### 2022-04-25 paper/main.tex
added paper

\documentclass{article}

\usepackage{arxiv}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsmath}
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{cleveref}       % smart cross-referencing
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{doi}
\usepackage{bm}             % bold fonts for math symbols
\usepackage[misc]{ifsym}    % collection of symbols
\usepackage{caption}
\usepackage{subcaption}
\usepackage{float}          % force figure positioning in text
\usepackage{chngcntr}       % counters per section
\usepackage{tikz}
\usepackage{pgfplots}
\usepgfplotslibrary{dateplot, groupplots}
\pgfplotsset{compat=newest}
\usepackage{plotpolicies}   % custom package to plot policy-executions

% default \cite to use parentheses
\renewcommand{\cite}[1]{\citep{#1}}

\newcommand{\foi}{\lambda}

%MDP
%state space
\newcommand{\mdpstatespace}{\mathcal{S}}
%action space
\newcommand{\mdpactionspace}{\mathcal{A}}
%mdp discount factor
\newcommand{\mdpdiscount}{\gamma}
%mdp policy
\newcommand{\mdppolicy}{\pi}
%mdpvaluefunction
\newcommand{\mdpvaluefunction}{V}
%vectorial value function
\newcommand{\momdpvaluefunction}{\mathbf{V}}
%mdpqfunction
\newcommand{\mdpqfunction}{Q}
%mdp q function estimate
\newcommand{\mdpqfunctionestimate}{\hat{Q}}
%mdp optimal q function
\newcommand{\mdpoptimalqfunction}{Q^{*}}
%q step size
\newcommand{\qstepsize}{\alpha}
%\advantageestimate
\newcommand{\advantageestimate}{\hat{A}}
%transition function
\newcommand{\mdptransition}{\mathcal{T}}
%reward fn
\newcommand{\mdprewardfn}{\mathcal{R}}
%vectorial reward fn
\newcommand{\momdprewardfn}{\bm{\mathcal{R}}}
%mdp state
\newcommand{\state}{s}
\newcommand{\mdpstate}{\mathbf{s}}
%mdp reward
\newcommand{\mdpreward}{\mathbf{r}}
\newcommand{\momdpreturn}{\mathbf{R}}
%mdp action
\newcommand{\action}{a}
%vectorial action
\newcommand{\mdpaction}{\mathbf{a}}
% the 6 social contact matrices
\newcommand{\chome}{C_{\text{home}}}
\newcommand{\cwork}{C_{\text{work}}}
\newcommand{\ctransport}{C_{\text{transport}}}
\newcommand{\cschool}{C_{\text{school}}}
\newcommand{\cleisure}{C_{\text{leisure}}}
\newcommand{\cother}{C_{\text{other}}}
% tentative momdp env name
\newcommand{\momdpname}{MOBelCov}
%other params
\newcommand{\agegroups}{K}

\title{Exploring the Pareto front of multi-objective COVID-19 mitigation policies using reinforcement learning}

% Here you can change the date presented in the paper title
%\date{September 9, 1985}
% Or remove it
\date{}

\author{Mathieu Reymond \\
	Vrije Universiteit Brussel\\
	Brussels, Belgium \\
	\texttt{mathieu.reymond@vub.be} \\
	\And
	Conor F. Hayes \\
	National University of Ireland Galway\\
	Galway, Ireland\\
% 	\texttt{c.hayes13@nuigalway.ie} \\
	\And
	Lander Willem \\
	University of Antwerp \\
	Antwerp, Belgium \\
	\And
	Roxana R\u{a}dulescu \\
	Vrije Universiteit Brussel\\
	Brussels, Belgium \\
	\And
	Steven Abrams \\
	Hasselt University \\
	Hasselt, Belgium \\
	\And
	Diederik M.\ Roijers \\
	HU University of Applied Sciences Utrecht \\
	Utrecht, the Netherlands \\
	\And
	Enda Howley \\
	National University of Ireland Galway\\
	Galway, Ireland\\
	\And
	Patrick Mannion \\
	National University of Ireland Galway\\
	Galway, Ireland\\
	\And
	Niel Hens \\
	Hasselt University \\
	Hasselt, Belgium \\
	\And
	Ann Now\'{e} \\
	Vrije Universiteit Brussel\\
	Brussels, Belgium \\
	\And
	Pieter Libin \\
	Vrije Universiteit Brussel\\
	Brussels, Belgium \\
% 	\AND
% 	Coauthor \\
% 	Affiliation \\
% 	Address \\
% 	\texttt{email} \\
}

% \author{Mathieu Reymond\inst{1}{\Letter} \and Conor F. Hayes \inst{4} \and Lander Willem  \inst{6}\and Roxana R\u{a}dulescu \inst{1} \and Steven Abrams \inst{3,7} \and Diederik M.\ Roijers\inst{1,5}\and Enda Howley\inst{4} \and Patrick Mannion\inst{4} \and Niel Hens \inst{3,6} \and Ann Now\'{e}\inst{1} \and Pieter Libin\inst{1,2,3}{\Letter}}

% \authorrunning{M. Reymond et al.}

% \institute{ 
% Artificial intelligence lab, Vrije Universiteit Brussel, Brussels, Belgium \email{\{pieter.libin,mathieu.reymond,diederik.roijers,ann.nowe\}@vub.be} 
% \and
% Rega institute, KU Leuven, Leuven, Belgium 
% \and  
% Data science institute, Hasselt University, Hasselt, Belgium
% \and
% National University of Ireland Galway, Ireland \email{\{c.hayes13\}@nuigalway.ie}
% \and
% HU University of Applied Sciences Utrecht, the Netherlands
% \and
% Centre for Health Economic Research and Modelling Infectious Diseases
% CHERMID, University of Antwerp, Belgium
% \and
% Global Health Institute, University of Antwerp, Belgium
% }

% Uncomment to override  the `A preprint' in the header
%\renewcommand{\headeright}{Technical Report}
%\renewcommand{\undertitle}{Technical Report}
\renewcommand{\shorttitle}{Exploring the Pareto front of multi-objective COVID-19 mitigation policies}

%%% Add PDF metadata to help others organize their library
%%% Once the PDF is generated, you can check the metadata with
%%% $ pdfinfo template.pdf
\hypersetup{
pdftitle={Exploring the Pareto front of multi-objective COVID-19 mitigation policies using reinforcement learning},
pdfauthor={Mathieu Reymond, Conor F.~Hayes, Lander Willem, Roxana R\u{a}dulescu, Steven Abrams , Diederik M.\ Roijers, Enda Howley, Patrick Mannion, Niel Hens, Ann Now\'{e}, Pieter Libin},
pdfkeywords={Multi-objective reinforcement learning, Epidemic control, COVID-19 epidemic models},
}

\begin{document}
\maketitle

\begin{abstract}
Infectious disease outbreaks can have a disruptive impact on public health and societal processes. As decision making in the context of epidemic mitigation is hard, reinforcement learning provides a methodology to automatically learn prevention strategies in combination with complex epidemic models. Current research focuses on optimizing policies with respect to a single objective, such as the pathogen's attack rate. However, as the mitigation of epidemics involves distinct, and possibly conflicting, criteria (i.a., prevalence, mortality, morbidity, cost), a multi-objective decision approach is warranted to learn balanced policies. To lift this decision-making process to real-world epidemic models, we apply deep multi-objective reinforcement learning and build upon a state-of-the-art algorithm, Pareto Conditioned Networks (PCN), to learn a set of solutions that approximates the Pareto front of the decision problem. We consider the first wave of the Belgian COVID-19 epidemic, which was mitigated by a lockdown, and study different deconfinement strategies, aiming to minimize both COVID-19 cases (i.e., infections and hospitalizations) and the societal burden that is induced by the applied mitigation measures. We contribute a multi-objective Markov decision process that encapsulates the stochastic compartment model that was used to inform policy makers during the COVID-19 epidemic. As these social mitigation measures are implemented in a continuous action space that modulates the contact matrix of the age-structured epidemic model, we extend PCN to this setting. We evaluate the solution set that PCN returns, and observe that it correctly learns to reduce the social burden whenever the hospitalization rates are sufficiently low. In this work, we thus demonstrate that multi-objective reinforcement learning is attainable in complex epidemiological models and provides essential insights to balance complex mitigation policies.
\end{abstract}

% keywords can be removed
\keywords{Multi-objective reinforcement learning  \and Epidemic control \and COVID-19 epidemic models}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
As shown by the COVID-19 pandemic, infectious disease outbreaks represent a paramount global challenge that should be tackled by prevention strategies. To this end, understanding the complex dynamics that underlie these epidemics is essential. Epidemiological transmission models allow us to capture and understand such dynamics and facilitate the study of prevention strategies through simulation. However, developing efficient mitigation strategies remains a challenging process, given the non-linear and complex nature of epidemics.

For this reason, reinforcement learning provides a methodology to automatically learn prevention strategies in combination with complex epidemic models~\cite{libin2020}.
Previous research typically focuses on optimising policies with respect to a single objective, such as the pathogen's attack rate, while the mitigation of epidemics is a problem that inherently covers distinct  and possibly conflicting criteria (i.a., prevalence, mortality, morbidity, cost). Therefore, optimizing on a single objective requires that  these distinct criteria are somehow aggregated into a single metric.
Generally, during learning, a decision maker will rely on an expert's assistance to manually design such a metric. However, a decision maker may
be unaware of all possible trade-offs between different solutions, which may leave the decision maker unaware of possible solutions that better reflect their preferences, potentially resulting in sub-optimal performance. Furthermore, manually designing such metrics is time consuming, costly and error-prone, as this non-intuitive process requires repetitive and tedious tuning to achieve the desired behaviour~\cite{hayes2021practical}. Moreover, taking a single objective approach has several other limiting factors~\cite{hayes2021practical,roijers2013survey}.

To alleviate the challenging process of defining the learning problem explicitly, we can directly learn optimal behaviours by explicitly taking a multi-objective approach. By assuming that a decision maker will always prefer solutions for which at least one objective improves, it is possible to learn a set of optimal solutions called the \emph{Pareto front}. This enables decision makers to review each solution on the Pareto front before making a decision, thereby being aware of the trade-offs that a solution may imply.

In this work, we investigate the use of \emph{multi-objective reinforcement learning} (MORL) to learn a set of solutions that approximate the Pareto front of multi-objective mitigation strategies. We consider the first wave of the Belgian COVID-19 epidemic, which was mitigated by a hard lockdown~\cite{willem2021impact}. When the incidence of confirmed cases were steadily dropping, epidemiological experts were asked to investigate strategies to exit from the stringent lockdown which was imposed.
Here, we consider the epidemiological model developed by \citet{abrams2021modelling} that was constructed to describe the Belgian COVID-19 epidemic, and was fitted to hospitalisation incidence data and serial sero-prevalence data.
This model constitutes a stochastic discrete-time age-structured compartmental model that simulates mitigation strategies by varying social distancing parameters concerning school, work and leisure contacts.  
Based on this model, we \emph{contribute \momdpname, a novel mutli-objective epidemiological environment}, in the form of a multi-objective Markov decision process (MOMDP). \momdpname\ encapsulates the epidemiological model developed by \citet{abrams2021modelling} to implement state transitions, with an action space that combines a proportional reduction of school, work and leisure contacts at each time step and defines a reward function based on two objectives: the attack rate (i.e., proportion of the population affected by the pathogen) and social burden.

To learn and explore the trade-offs between the attack rate and social burden we build upon a state-of-the-art MORL approach, namely Pareto Conditioned Networks (PCN)~\cite{reymond2022pcn}, which uses a single neural network to learn the policies that belong to the Pareto front.
As PCN is an algorithm designed for discrete action-spaces, we extend it towards continuous action-spaces to accommodate \momdpname's action-space. With this continuous action variant of PCN, we explore the Pareto front of multi-objective COVID-19 mitigation policies.

By evaluating the solution set of mitigation policies returned by PCN, we observe that PCN minimises the social burden in scenarios where hospitalization rates are sufficiently low. Therefore, in this work we illustrate that multi-objective reinforcement learning can be utilised to provide important insights surrounding the trade-offs between complex mitigation polices in real-world epidemiological models.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Background}
\subsection{Multi-Objective Reinforcement Learning}
Real-world decisions problems often have multiple and possibly conflicting objectives. Multi-objective reinforcement learning (MORL) can be used to find optimal solutions for sequential decision making problems with multiple objectives~\cite{hayes2021practical}. For MORL, problems are typically modelled as a multi-objective Markov decision process (MOMDP), i.e., a tuple, $\mathcal{M} = (\mdpstatespace, \mdpactionspace, \mdptransition, \mdpdiscount, \momdprewardfn)$, where $\mdpstatespace$, $\mdpactionspace$ are the state and action spaces respectively, $\mdptransition \colon \mdpstatespace \times \mdpactionspace \times \mdpstatespace  \to \left[ 0, 1 \right]$ is a probabilistic transition function, $\mdpdiscount$ is a discount factor determining the importance of future rewards and $\momdprewardfn \colon \mdpstatespace \times \mdpactionspace \times \mdpstatespace \to \mathbb{R}^n$ is an $n$-dimensional vector-valued immediate reward function, where $n$ denotes the number of objectives. For single-objective RL, i.e., when $n = 1$, the goal is to find the policy $\mdppolicy^{*}$ that maximizes the expected sum of discounted rewards, also called the expected return:
\begin{equation}
\mdppolicy^*= \arg\max_\mdppolicy \mathbb{E} \left [ \sum_{t=0}^h \mdpdiscount^t r_t |\ \mdppolicy, s_0 \right ],
\end{equation}
where $r_t = \mdprewardfn(\state_t, \action_t, \state_{t+1})$.
In contrast, in MORL, $n > 1$ which leads to vectorial returns. In this case, there can be policies for which, without any additional information, it is impossible to know if one is better than the other. For example, we cannot decide which policy between $\mdppolicy_1, \mdppolicy_2$ is optimal if they lead to expected returns (also called V-values) $\momdpvaluefunction^{\mdppolicy_1}=(0,1), \momdpvaluefunction^{\mdppolicy_2}=(1,0)$ respectively. We call these solutions \emph{non-dominated}, i.e., solutions for which it is impossible to improve an objective without hampering another. The set that contains all the non-dominated solutions of the decision problem is called the \emph{Pareto front} $\mathcal{F}$. Our goal is to find the set of policies that lead to all the V-values contained in the Pareto front $\Pi^* = \{\ \pi | \momdpvaluefunction^{\pi} \in \mathcal{F}\}$. In general, we call any set of V-values a \emph{solution set}. When a solution set contains only  non-dominated V-values, it is referred to as a \emph{coverage set}. In the case that no $\mdppolicy$ exists that has a $\momdpvaluefunction^{\pi}$ dominating any of the solutions in a coverage set, then this coverage set is the Pareto front.

\subsection{Multi-Objective Metrics}
Comparing the learned coverage sets of different algorithms is a non-trivial task, as one algorithm's output might dominate the other in some region of the objective-space, but be dominated in another. Intuitively, one would generally prefer the algorithm that covers a wider range of decision maker preferences. In this work we use several metrics to evaluate our algorithm's performance.

A widely used metric in the literature is called the \emph{hypervolume}~\cite{zitzler2003}. This metric evaluates the learned coverage set by computing its volume with respect to a fixed specified reference point. This metric is, by definition, the highest for the Pareto front, as no other possible solution can increase its volume (since they are all dominated). One disadvantage of the hypervolume is that it can be difficult to interpret. For example, when working in high-dimensional objective-spaces, adding or removing a single point can lead to wildly different hypervolume values, especially if the point lies close to an extremum of the space.

To alleviate these shortcomings, we additionally evaluate our work on a different metric called the $\varepsilon$-indicator $I_\varepsilon$~\cite{zitzler2003}, which measures how close a coverage set is to the Pareto front $\mathcal{F}$. Intuitively, $I_\varepsilon$ shows that any solution of $\mathcal{F}$ is \emph{at most} $\varepsilon$ better with respect to each objective $o$ than the closest solution of the evaluated coverage set:

\begin{equation}
\label{eq:epsilon-metric}
    I_{\varepsilon} = 
    \inf_{\varepsilon\in\mathbb{R}} 
    \{ 
    \forall\ \momdpvaluefunction^\pi\!{\in}\ \mathcal{F}, ~
    \exists\ \momdpvaluefunction^{\pi'}\!{\in}\ \hat{\Pi} :\ 
    || \mdpvaluefunction^\pi - \mdpvaluefunction_o^{\pi'} ||_{\infty} \le \varepsilon
    \}
\end{equation}

The main disadvantage of this metric is that we need the true Pareto front to compute it, which in the case of our MOMDP (see Section~\ref{sec:sars-cov2-momdp}) is unknown. To still gain insights from our learned policies, we approximate the true Pareto front using the non-dominated policies across all runs.

We note that the $\varepsilon$-indicator metric is quite pessimistic, as it measures worst-case performance~\cite{zintgraf2015}, e.g., it will still report bad performance as long as a single point of the Pareto front is not correctly modeled, even if all the other points are covered. As such, we also use the $I_{\varepsilon-mean}$ \cite{reymond2022pcn} which measures the \emph{average} $\varepsilon$ distance of the solutions in $\mathcal{F}$ with respect to the evaluated coverage set.

Figure~\ref{fig:paretofrontmetrics} shows a visual representation of the hypervolume and $\varepsilon$ metrics in two dimensions.

\begin{figure}[t!]
    \centering
    \begin{tikzpicture}
      \begin{axis}[legend pos=south west, xmin=-1, xmax=4.5, ymin=-0.5, ymax=4.5,ytick={0,1,2,3,4,5},scale only axis,width=0.4\textwidth]
      \addplot[only marks,mark=x,mark size=3] coordinates {(-0.5,0)};
      \addplot[only marks,mark=*] coordinates {(1,4)(2,2)(4,1)};
      \addplot[only marks,mark=o] coordinates {(0.5,3)(0.75,2.3)(2.3,1)(3.3,0.7)};
      % Hypervolume
      \fill [cyan,opacity=0.2] (-0.5,0) -- (-0.5,3) -- (0.5,3) -- (0.5,2.3) -- (0.75,2.3) --
      (0.75,1) -- (2.3,1) -- (2.3,0.7) -- (3.3,0.7) -- (3.3,0) -- cycle;
      % epsilons
      \draw [dashed, -|] (0.5, 3) -- node[right]{$\varepsilon_1$} (0.5,4);
      %\draw [dashed, -|] (0.75,2.3) -- node[below]{x} (2,2.3);
      \draw [dashed, -|] (2.3,1) -- node[right]{$\varepsilon_2$} (2.3,2);
      \draw [dashed, -|] (3.3,0.7) -- node[below]{$\varepsilon_3$} (4,0.7);
      \draw [opacity=0.2,line width=2,orange, dotted] (-0.5,0) -- (-0.5,4) -- (1,4) -- (1,2) --
      (2,2) -- (2,1) -- (4,1) -- (4,0) -- (-0.5,0);
      \end{axis}
    \end{tikzpicture}
    \caption{Example of a Pareto front (black dots) and a coverage set (white dots) in a 2-objective environment. The hypervolume metric (in light blue) measures the volume of all dominated solutions with respect to some reference point (cross). The $\varepsilon$ metrics first compute the maximum distance between each point in the Pareto front and its closest point in the coverage set ($\varepsilon_i$). We can then take their maximum value to compute the $I_\varepsilon$ metric, or their mean value to obtain the $I_{\varepsilon-mean}$ metric of the coverage set.}
    \label{fig:paretofrontmetrics}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{COVID-19 model and the \momdpname\  MOMDP}
\label{sec:sars-cov2-momdp}
\subsection{Stochastic compartment model for SARS-CoV-2}
\label{sec:sars-cov2-momdp_model}
As an environment to evaluate non-pharmaceutical interventions, we consider the adapted version of the SEIR compartmental model presented by Abrams et al., that was used to investigate exit strategies in Belgium after the first epidemic wave of SARS-CoV-2~\cite{abrams2021modelling}. This model constitutes a discrete-time stochastic model, that considers an age-structured population. This model generalises a standard SEIR model\footnote{A standard SEIR model divides the population into four different states, i.e., susceptible, exposed, infectious and recovered individuals.}, extended to capture the different stages of disease spread and history that are associated with SARS-CoV-2 (i.e., pre-symptomatic, asymptomatic, symptomatic with mild symptoms and symptomatic with severe symptoms) and to represent the stages associated with severe disease, i.e., hospitalisation, admission to the intensive care unit (ICU) and death. 
As we consider an age-structured population, we consider this extended SEIR structure for $\agegroups = 10$ age groups,  i.e., $[0-10), [10-20), [20-30), [30-40), [40-50), [50-60), [60-70), [70-80), [80-90), [90,\infty)$. Contacts of the different age-groups which impact the propagation rate of the epidemic are modeled using social contact matrices (SCMs). We define a SCM for 6 different social environments: $\chome, \cwork, \ctransport, \cschool, \cleisure, \cother$ for the home, work, transport, school, leisure, other environments respectively.
The model is described by a set of ordinary differential equations (ODEs), as described in SI.
By formulating this set of differential equations defined above as a chain-binomial process we can obtain stochastic trajectories from this model~\cite{abrams2021modelling}.

\subsection{Interventions strategies}
\label{sec:interventions}
In order to model different types of interventions, we follow \citet{abrams2021modelling}. Firstly, in order to consider distinct exit scenarios, we alter the SCMs to reflect a contact reduction in a particular age group. Secondly, we assume that compliance to the interventions is gradual and model this using a logistic compliance function (see details in SI).
We consider a contact reduction function that imposes a proportional reduction of work (including transport) $p_w$, school $p_s$ and leisure $p_l$ contacts, which is implemented as a linear combination of contact matrices:
\begin{equation}
\label{eq:scm}
    \hat{C} = \chome + p_w (\cwork + \ctransport) + p_s \cschool + p_l (\cleisure + \cother)
\end{equation}

\subsection{The \momdpname\ Environment}
\begin{figure}[t!]
    \centering
    \begin{tikzpicture}
     \draw (0.0pt, 0.0pt)node[fill=green!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](0){$\textbf{S}$};
     \draw (60.0pt, 0.0pt)node[fill=orange!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](1){$\textbf{E}$};
     \draw (120.0pt, 0.0pt)node[fill=red!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](2){$\textbf{I}^{presym}$};
     \draw (180.0pt, 10.0pt)node[fill=red!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](3){$\textbf{I}^{asym}$};
     \draw (170.0pt, -30.0pt)node[fill=red!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](4){$\textbf{I}^{mild}$};
     \draw (170.0pt, -70.0pt)node[fill=red!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](5){$\textbf{I}^{sev}$};
     \draw (230.0pt, -50.0pt)node[fill=blue!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](6){$\textbf{I}^{hosp}$};
      \draw (75.0pt, -70.0pt)node[fill=gray!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](10){$\textbf{H}^{new}$};
     \draw (230.0pt, -90.0pt)node[fill=blue!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](7){$\textbf{I}^{icu}$};
     \draw (300.0pt, -50.0pt)node[fill=orange!50, thick, minimum height=0.6cm,minimum width=0.6cm, draw](8){$\textbf{R}$};
     \draw (300.0pt, -90.0pt)node[fill=purple!50, thick, minimum height=0.6cm,minimum width=0.6cm, draw](9){$\textbf{D}$};

     %\path[->] (6) edge[dotted, bend right = 50]
     \path[->] (0) edge[] node[above]{$\boldsymbol{\lambda}(t)$} (1);
     \path[->] (1) edge[] node[above]{$\gamma$} (2);
     \path[->] (2) edge[] node[below]{$\textbf{p}\theta$} (3);
     \path[->] (2) edge[] node[near end, below left]{$(1-\textbf{p})\theta$} (4);
     \path[->] (4) edge[] node[left]{$\boldsymbol{\psi}$} (5);
     \path[->] (5) edge[] node[near end, below left]{$(1-\phi_{1})\boldsymbol{\omega}$} (7);
     \path[->] (5) edge[] node[above]{$\phi_{1}\boldsymbol{\omega}$} (6);
     \path[->] (7) edge[] node[near end, below right]{$\boldsymbol{\delta}_{4}$} (8);
     \path[->] (7) edge[] node[above]{$\boldsymbol{\tau}_{2}$} (9);
     \path[->] (6) edge[] node[above]{$\boldsymbol{\delta}_{3}$} (8);
     \path[->] (6) edge[] node[near start, below left]{$\boldsymbol{\tau}_{1}$} (9);
     \path[->] (3) edge[bend left = 30] node[above right]{$\delta_{1}$} (8);
     \path[->] (4) edge[bend left = 30] node[above left]{$\boldsymbol{\delta}_{2}$} (8);
      \path[->] (5) edge[] node[below]{$\textbf{I}^{sev} - \textbf{H}^{new}$} (10);
    \end{tikzpicture}

    \caption{Schematic diagram of the compartmental model for SARS-CoV-2 presented by \citet{abrams2021modelling} which is used to derive the MOMDP.}
    \label{fig:ode_mdel}
\end{figure}
In order to apply multi-objective reinforcement learning, we construct the \momdpname\ MOMDP based on the epidemiological model introduced in Section~\ref{sec:sars-cov2-momdp_model} and graphically depicted in Figure~\ref{fig:ode_mdel}.

\paragraph{State-space:} The state of the MOMDP consists of the state variables that make up the epidemiological model (see Figure~\ref{fig:ode_mdel}), to which we refer as $\mdpstate_m$, together with the social contact matrix $\hat{C}$ that is currently in place.
$\mdpstate_m$ directly corresponds to the aggregation of the state variables in the epidemiological model, i.e., a tuple,
\begin{equation}
    \{S_k, E_k, I_k^{presym}, I_k^{asym}, I_k^{mild}, I_k^{sev}, I_k^{hosp}, I_k^{icu}, H^{new}_k, D_k, R_k\},
\end{equation}
for each age group $k \in \{1, \ldots, \agegroups\}$, where $S$ encodes the members of the population who are susceptible to infection and $E$ encodes the members of the population who have been exposed to COVID-19. Moreover, $I^{presym}$, $I^{asym}$, $I^{mild}$, $I^{hosp}$, $I^{icu}$ are the members of the population infected with COVID-19 and are, respectively, presymptomatic, asymptomatic, have mild symptoms, are hospitalised, or are in the ICU. Finally, $H^{new}_k$ represents the number of newly hospitalised individuals in age group $k$.

Therefore, we define a state in \momdpname\ as follows:
\begin{equation}
    \mdpstate = \mdpstate_m \cup \hat{C}
\end{equation}

We paramaterise the model using the mean of the posteriors as reported by \citet{abrams2021modelling} (details in SI).

\paragraph{Action-space:} Our actions concern the installment of a social contact matrix $\hat{C}$, with a particular reduction (see Section~\ref{sec:interventions}). To this end, we use the proportional reduction parameters $p_w, p_s, p_l$ defined in Section~\ref{sec:interventions}. Thus, each $\mdpaction \in \mdpactionspace$ is a 3-dimensional continuous vector in $[0,1]^3$ (i.e., $\mdpaction = [p_{w}, p_{s}, p_{l}]$) which impacts the SCM according to Equation~\ref{eq:scm}.

\paragraph{Transition function:} The model defined by \citet{abrams2021modelling} utilises a model transition probability $M(\mdpstate_m' \mid \mdpstate_m, \hat{C})$ (see SI for details on $M$), where $\hat{C}$ the currently installed SCM, that progresses the epidemiological model in one timestep.
We use this function as the transition function in \momdpname.
For each timestep $t$, we simulate the model for one week, using $\hat{C}$ obtained from $\mdpaction_t$.

As mentioned above and detailed in the SI, the compartment model is described by a set of ODEs. To obtain a stochastic version, the ODEs can be formulated as a chain-binomial process. Based on these cases, we also create and experiment on two versions of the \momdpname\ environment: the (deterministic) \emph{ODE model} (i.e., a MOMDP with a deterministic transition function) and the (stochastic) \emph{Binomial model} (i.e., a MOMDP with a stochastic transition function).

\paragraph{Reward function:} We define a vector reward function which considers multiple objectives: attack rate (i.e., infections, hospitalisations) and the burden imposed by the interventions on the population.

The attack rate in terms of infections is defined as the difference in susceptibles from the current state and next state~\cite{libin2020}. Since this is a cost that needs to be minimized, we defined the corresponding reward function as the negative attack rate:
\begin{equation}
\mdprewardfn_{\text{ARI}}(\mdpstate,\mdpaction,\mdpstate') = -(\sum_{k=1}^{\agegroups}S_k(\mdpstate)-\sum_{k=1}^{\agegroups}S_k(\mdpstate')).
\end{equation}

The reward function to reduce the attack rate in terms of hospitalisations is simply defined as the negative number of new hospitalizations:
\begin{equation}
\mdprewardfn_{\text{ARH}}(\mdpstate,\mdpaction,\mdpstate') = -\sum_{k=1}^{\agegroups}H^\text{new}_k(\mdpstate)
\label{eqn:attack-rate-hosp}
\end{equation}

Finally, we use the missed contacts resulting from the intervention measures as a proxy for societal burden. To quantify missed contacts, we consider the original social contact matrix $C$ and the installed social contact matrix $\hat{C}$, and compute the difference $\hat{C} - C$. 
The resulting difference matrix quantifies the average frequency of contacts missed. To determine missed contacts for the entire population, we apply the calculated difference matrix to the population sizes of the respective age groups that are currently uninfected (i.e., susceptible and recovered individuals). Therefore we define the social burden reward function $\mdprewardfn_{\text{SB}}$, as follows:

\begin{equation}
    \mdprewardfn_{\text{SB}}(\mdpstate,\mdpaction,\mdpstate') = \sum_{i=1}^{\agegroups}\sum_{j=1}^{\agegroups}(\hat{C}-C)_{ij}S_i(\mdpstate)S_j(\mdpstate) + \sum_{i=1}^{\agegroups}\sum_{j=1}^{\agegroups}(\hat{C}-C)_{ij}R_i(\mdpstate)R_j(\mdpstate),
\end{equation}
where $S_k(\mdpstate)$ represents the number of susceptible individuals in age group $k$ in state $\mdpstate$ and $R_k$ represents the number of recovered individuals in age group $k$ in state $\mdpstate$. In Section~\ref{sec:experiments}, we optimize PCN on two different variants for the multi-objective reward function: $[\mdprewardfn_\text{ARH}, \mdprewardfn_\text{SB}]$ and $[\mdprewardfn_\text{ARI}, \mdprewardfn_\text{SB}]$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Pareto Conditioned Networks}
In multi-objective optimization, the set of optimal policies can grow exponentially in the number of objectives. Thus, recovering them all is an expensive process and requires an exhaustive exploration of the complete state space. To address this problem, we use Pareto Conditioned Networks~(PCN), a method that uses a single neural network to encompass all non-dominated
policies \cite{reymond2022pcn}. The key idea behind PCN is to use supervised learning techniques to improve the policy instead of resorting to temporal-difference learning. This eliminates the moving-target problem \cite{schmidhuber2019}, resulting in stable learning algorithm.

PCN uses a single neural network that takes a tuple $\langle \mdpstate, \hat{h}, \mathbf{\hat{R}}  \rangle$ as input. $\mathbf{\hat{R}}$ represents the \emph{desired return} of the decision maker, i.e. the return PCN should reach at the end of the episode. $\hat{h}$ denotes the \emph{desired horizon} that expresses the number of timesteps that should be executed before reaching $\mathbf{\hat{R}}$. At execution time, both $\hat{h}$ and $\mathbf{\hat{R}}$ are chosen by the decision maker at the start of the episode. Then, at every timestep, the desired horizon is updated according to the perceived reward $\mdpreward_t$, $\mathbf{\hat{R}} \leftarrow \mathbf{\hat{R}} - \mdpreward_t$ and the desired horizon is decreased by one, $\hat{h} \leftarrow \hat{h}-1$.

PCN's neural network has an output for each action $\action_i \in \mathcal{A}$. Each output represents the confidence the network has that, by taking the corresponding action in $\mdpstate$, $\mathbf{\hat{R}}$ will be reached in $\hat{h}$ timesteps. We can draw an analogy with a classification problem where the network should learn to classify $(\mdpstate, \hat{h}, \mathbf{\hat{R}})$ to its corresponding label $\action_i$.

Similar to classification, PCN requires a labeled dataset with training examples to learn a mapping from input to label. However, contrary to classification, the data present in the dataset is not fixed. Each time PCN executes a trajectory it creates a tuple $x=(\mdpstate_t, \momdpreturn_t, T-t)$, $y=\action_t$ for every transition of the trajectory and adds it to the dataset.

PCN collects new data for a fixed number of episodes, after which it re-trains the network with batch updates from the newly assembled dataset. This improves the policies induced by the network, which in turn enables the agent to gather superior data for the next training batch.

\subsection{PCN for continuous actions}
\label{sec:pcn-continuous}

PCN trains the network as a classification problem, where each class represents a different action. Transitions $x= \langle \mdpstate_t, h_t, \momdpreturn_t \rangle, y=\action_t$ are sampled from the dataset, and the ground-truth output $y$ is compared with the predicted output $\hat{y}=\pi(\mdpstate_t, h_t, \momdpreturn_t)$. The predictor (i.e., the policy) is then updated using the cross-entropy loss function:
\begin{equation}
    H = -\sum_{\action \in \mathcal{A}}{y_{\action} \log \pi(\action|\mdpstate_t,h_t,\momdpreturn_t)}
\end{equation}
where $y_\action = 1$ if $\action = \action_t$ and $y_\action = 0$ otherwise.

While the original PCN algorithm is designed for MOMDPs with discrete actions-spaces, the problem we tackle (see Section~\ref{sec:sars-cov2-momdp}) is defined in terms of a continuous action-space. We thus extend PCN for the continuous action-space setting.
First, we change the output of the neural network such that there is a single output value for each dimension of the action-space. Since the actions should be bound in the domain of possible actions ($[0,1]$ in the case of \momdpname, see Section~\ref{sec:sars-cov2-momdp}), we apply a tanh non-linearity function on this output.
Second, the problem becomes a regression problem instead of a classification problem, as the labeled dataset uses continuous labels $y = \mdpaction_t$ instead of categories. We thus use a Mean Squared Error (MSE) loss to update the policy:

\begin{equation}
    MSE = \frac{1}{|\mathcal{A}|}\sum_{a \in \mathcal{A}}{(\hat{y}_a - y_a)^2}
\end{equation}

Since learning the full set of Pareto-efficient policies $\Pi^*$ requires that the policies $\pi^* \in \Pi^*$ are deterministic stationary policies \cite{roijers2013survey}, we use the output $\mathbf{\hat{y}}$ as action at execution time. However, PCN improves its policy through exploration, by continuously updating its dataset with better trajectories. Thus, at training time, we follow \cite{lillicrap2015continuous} and use a stochastic policy by adding random noise sampled from a Normal distribution to the action: 
\begin{equation}
    \mdpaction_t = \pi(\mdpstate_t, h_t, \momdpreturn_t) + \eta s \text { with } s \sim \mathcal{N},
\end{equation}

where $\eta$ is a hyper-parameter defining the magnitude of noise to be added.

\subsection{Coping with stochastic transitions}

PCN trains its policy on a dataset that is collected by executing trajectories. It assumes that reenacting a transition from the dataset leads to the same episodic return. When the transition function $\mathcal{T}$ of the MOMDP is deterministic, the whole trajectory can be faithfully reenacted, which guarantees that we obtain the same return. Combined with the fact that PCN's policy is deterministic at execution time, conditioning the policy on a target episodic return is equivalent to conditioning it on the V-value $\mathbf{V}$.

However, when $\mdptransition$ is stochastic we lose this guarantee. We cope with this issue by adding small random noise to $\mathbb{R}_t$ when performing gradient descent, which reduces the risk of overfitting~\cite{zur2009noise}. Moreover, although the Binomial model of \momdpname\ is stochastic, the possible next-states resulting from a state-action pair are similar to each other. This allows PCN to compensate if $\mdpreward_t = \momdprewardfn(\mdpstate_t, \mdpaction_t, \mdpstate_{t+1})$ is somewhat worse than expected. This is confirmed by our experiments (see Section~\ref{sec:experiments}), where the coverage sets learned by PCN on the ODE model and on the Binomial model are similar.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Analysing COVID-19 deconfinement policies}
\label{sec:experiments}
Our goal is to use PCN to learn deconfinment strategies in the \momdpname\ environment. We aim to obtain policies that balance between the social burden experienced by the population and the epidemiological objective of minimising the attack rate. To this end we consider two cases for the vectorial reward functions $[\mdprewardfn_\text{ARH}, \mdprewardfn_\text{SB}]$ and $[\mdprewardfn_\text{ARI}, \mdprewardfn_\text{SB}]$, to learn and analyse policies under different targets with respect to the considered attack rate.

We now evaluate our extension of PCN for continuous action-spaces on both the ODE and Binomial models. Conform to \citet{abrams2021modelling}, the simulation starts on the 1st of March 2020, by seeding a number of infections in the population. Two weeks later, on the 14th of March, the Belgian government initiated a full lockdown of the country. This is implemented by fixing the actions $p_w, p_s, p_l$ to $0.2, 0, 0.1$ respectively. This lockdown ended on the 4th of May 2020, at which point the government decided on a multi-phase exit strategy to incrementally reduce teleworking, reopen schools and allow leisure activities, such as the re-opening of bars and the cultural sector. It is from this day onward that PCN aims to learn policies for diverse exit strategies, compromising between the total number of daily new hospitalizations and the total number of contacts lost as a proxy for social burden. The simulation lasts throughout the school holidays (from 01/07/2020 to 31/08/2020). Schools are closed during the school holidays, which is simulated by setting $p_s = 0$, regardless of the corresponding value outputted by the policy, i.e., during periods of school closure $p_s$ is ignored.

To evaluate the quality of the policies discovered by PCN, we compare it to a baseline. The baseline consists of a set of 100 fixed policies, that iterate over all the possible social restriction levels, with values ranging from 0 to 1. In other words, the fixed policies directly operate in a fine-grained manner on the whole contact reduction function $\hat{C}$. This allows us to obtain a strong baseline for potential exit strategies over the objective space. We note that while such fixed policies are a feasible approach, they do not scale well in terms of action and objective spaces and they will not be able to provide an adaptive restriction level, which is what we aim to provide using PCN.

All experiments are averaged over 5 runs. The hyper-parameters and the neural network architecture can be found in the SI.

\begin{figure}
    \centering
    \begin{tikzpicture}
        \pgfplotsset{
          scale only axis,
          height=6cm,
          xlabel={Cumulative number of daily new hospitalizations},
          ylabel={Cumulative lost contacts}
        }
      \begin{axis}[
        legend entries={{$\text{PCN}_\text{ARH}$ (ODE)}, {$\text{PCN}_\text{ARI}$ (ODE)}, {Fixed}},
        width=9.5cm
      ]
      %\addplot[only marks,draw opacity=0.5,mark=x,green] table [x=o_0, y=o_1, col sep=comma] {data/cs/cs_ode.csv};
      \addplot[only marks,draw opacity=0.5,mark=x,red] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/cs_ode.csv};
      \addplot[only marks,draw opacity=0.5,mark=o,blue] table [x=o_1, y=o_5, col sep=comma] {data/cs/ari/cs_ode.csv};
      %\addplot[only marks,mark=o,yellow] table [x=o_0, y=o_1, col sep=comma] {data/cs/cs_binomial.csv};
      %\addplot[only marks,draw opacity=0.5,mark=o,blue] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/cs_binomial.csv};
      %\addplot[only marks,draw opacity=0.5,mark=diamond,cyan] table [x=o_1, y=o_5, col sep=comma] {data/cs/ari/cs_binomial.csv};
      \addplot[only marks,mark=+] table [x=o_0, y=o_1, col sep=comma] {data/cs/cs_fixed.csv};
      \end{axis}
    \end{tikzpicture}
    \caption{The Pareto front of policies discovered by PCN using the ODE model (Binomial omitted for clarity, see SI), showing the different compromises between the number of hospitalizations and the number of lost contacts. Although $\text{PCN}_\text{ARI}$ was trained on the number of infections, it still shows a competitive coverage set with respect to hospitalizations.}
    \label{fig:pareto-front}
\end{figure}

\subsection{Learned coverage set}

We learn a coverage set that ranges from imposing minimal restrictions to enforcing many restrictions (see Figure~\ref{fig:policy-execution}). The coverage set is shown in Figure~\ref{fig:pareto-front}.

We notice that the coverage sets discovered by PCN almost completely dominate the coverage set of the baseline, showing that there are much better alternatives to the fixed policies. This is most visible in the compromising policies, where one has to carefully choose when to remove social restrictions while at the same time minimizing the impact on daily new hospitalizations. In these scenarios, PCN discovers policies that can drastically reduce the total number of new hospitalizations (e.g. more than $20000$) for the same social burden. Analysing the executions of such policies (see Figure~\ref{fig:policy-execution}, middle plot) shows a flattened hospitalization curve, with a gradual increase of social freedom during the school holidays such that the peak stays stable and gradually decreases over time.

Interestingly, we notice that the most restrictive policy (i.e. the one that prioritizes hospitalizations over social burden, see Figure~\ref{fig:policy-execution}, bottom plot) still starts to gradually increase $p_w$ and $p_l$ from the end of July onward. This is because by then, the epidemic has mostly faded out, and it is safe to reduce social restrictions. The timing of this reduction is important as reducing restrictions too soon can lead to a new wave. PCN learns the impact of its decisions over time, and correctly infers the timing at which restrictions can be safely lifted.

\begin{figure}
  \centering
  \input{fig_policy-execution}
  \caption{Selection of policies learned by PCN, from most restrictive in terms of social burden (top) to least restrictive (bottom).
  }
  \label{fig:policy-execution}
\end{figure}

\subsection{ODE versus Binomial models}
\label{sec:ode-vs-binomial}

Next, we compare the performance of PCN when trained on the ODE and Binomial models. The ODE model has the advantage of using a deterministic transition function, for which PCN is suited. However, taking into account stochasticity, using the Binomial model, is important as individual behavioural changes introduce substantial uncertainty in the course of the outbreak and require stochastic model evaluations to evaluate the impact of this uncertainty on the effectiveness of the intervention strategies. Regardless, the results displayed in Table~\ref{tab:coverage-set} show that the solution sets that PCN finds using the Binomial model are competitive with the ODE model, as indicated by their hypervolumes and $I_{\varepsilon-mean}$ metrics. We do note that the difference in $I_{\varepsilon}$ indicator is more significant, indicating that a few solutions discovered using the ODE model are not found when using the Binomial model, impacting worst-case performance. Overall, we see that our extension of PCN performs efficiently, even with some degree of stochasticity in the transition function.

\begin{table}[t]
  \centering
  \setlength{\tabcolsep}{0.5em} % for the horizontal padding
  {\renewcommand{\arraystretch}{1.2}% for the vertical padding
  \input{tab_coverage-set}
  }
  \caption{Evaluation metrics for the coverage sets comparing hospitalizations with social burden. In general training on the ODE results in slightly better coverage sets than on the Binomial model. Training on infections (ARI) still provides a competitive coverage set in terms of hospitalizations. All PCN coverage sets outperform the baseline.}
  \label{tab:coverage-set}
\end{table}

\subsection{$R_\text{ARH}$ versus on $R_\text{ARI}$}
\label{sec:arh-vs-ari}

We now assess the difference in coverage sets when optimizing on $\mdprewardfn_\text{ARH}$ versus $\mdprewardfn_\text{ARI}$. Although at a different scale, our experiments show that infections and hospitalizations are correlated. This is confirmed in Figure~\ref{fig:pareto-front}: the coverage set \emph{in terms of hopsitalizations} learned by $\text{PCN}_\text{ARI}$ is only slightly worse than the one learned by $\text{PCN}_\text{ARH}$, even though $\text{PCN}_\text{ARI}$ optimized \emph{on the infection attack rate}. This is expected, as during the initial phase of the epidemic, limited immunity was present in the population (few natural immunity and no vaccines), which induces a tight coupling between infection and hospitalisation cases. Moreover, the opposite also holds: both coverage sets are similar in terms of \emph{infections}. However, there is one notable exception. A number of policies learned by $\text{PCN}_\text{ARH}$ start removing social restrictions before the epidemic is rooted out, resulting in a rise of hospitalizations at the end of the summer holidays. Since, following Abrams et al., this is when the simulation stops, the total number of hospitalizations remains low. The number of infections, however, starts growing before this rise in hospitalizations and, since its growth is exponential, ends up high. We note that $\text{PCN}_\text{ARH}$ also learns alternative policies for similar $[\mdprewardfn_\text{ARH}, \mdprewardfn_\text{SB}]$ that do not exhibit this behavior. This shows that, even though a policy might lead to a return that matches the preferences of the decision maker, its execution might still lead to undesired behavior. We argue that the use of such expert systems for decision making should be paired with careful interpretation.

\subsection{Robustness of policy executions}
\label{sec:pcn-robustness}

The dataset of trajectories that PCN is trained on is pruned over time to keep only the most relevant trajectories. The returns of these trajectories are used in Figure~\ref{fig:pareto-front} to show the discovered coverage set. Each of these returns can be used as desired return for policy execution. We now assess the robustness of the executed policies, by comparing the averaged return obtained over multiple policy executions with the corresponding target return. We show that the executed policies reliably obtain returns that are similar to the desired return used to condition PCN.

Results are shown in Table~\ref{tab:pcn-robustness}. The $I_\varepsilon$ indicators shows that, over all policies, the decision maker will lose at worst $0.046$ and $0.068$ normalized returns in any of the objectives for the ODE, Binomial versions respectively. On average, it will lose $0.007$, $0.024$ normalized returns respectively, which corresponds to either a total of $3633, 1059$ more hospitalizations than expected or a total of $262, 76$ less contacts than expected.

\begin{table}[t]
    \centering
    \setlength{\tabcolsep}{0.5em} % for the horizontal padding
    {\renewcommand{\arraystretch}{1.2}% for the vertical padding
    \input{tab_pcn-robustness}
    }
    \caption{Comparing the difference in the desired return provided to PCN and the actual return PCN obtained when executing its policy. We see that, regardless of the setting, the learned policy faithfully receives a return similar to its desired return.}
    \label{tab:pcn-robustness}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related work}
Reinforcement learning (RL) has been used in conjunction with epidemiological models to learn policies to limit the spread of diseases and predict the effects of possible mitigation strategies \cite{probert2019context,ernst2006clinical}. For example, RL has been used extensively in modelling and controlling the spread of influenza \cite{das2008large,libin2020,libin2018bayesian}.

RL and Deep RL have been used extensively as a decision making aid to reduce the spread of COVID-19. For example, to learn effective mitigation strategies \cite{ohi2020exploring}, to learn efficacy of lockdown and travel restrictions \cite{kwak2021covid} and to limit the influx of asymptomatic travellers \cite{bastani2021efficient}.

Multi-objective methods have also been deployed to learn optimal strategies to mitigate the spread of COVID-19. Wan et al. \cite{wan2021multi} implement a model-based multi-objective policy search method and demonstrate their method on COVID-19 data from China. Given the method is model-based, a model of the transition function must be learned by sampling from the environment. The method proposed by Wan et al. \cite{wan2021multi} only considers a discrete action space which limits the application of their algorithm. Wan et al. \cite{wan2021multi} use linear weights to compute a set of Pareto optimal policies. However, methods which use linear weights can only learn policies on the convex-hull of the Pareto front \cite{vamplew2008limitations}, therefore the full Pareto front cannot be learned. It is important to note the method proposed by Kompella et al. \cite{kompella2020reinforcement} considers multiple objectives. However, the objectives are combined using a weighted sum with hand-tuned weights which are determined by the authors. The weighted sum is applied by the reward function and a single objective RL method is used to learn a single optimal policy.
In contrast to previous work, our approach makes no assumptions regarding the scalarisation function of the user and is able to discover Pareto fronts of arbitrary shape.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion and discussion}
Making decisions on how to mitigate epidemics has important ethical implications with respect to public health and societal burden. In this regard, it is crucial to approach this decision making from a balanced perspective, to which end we argue that multi-objective decision making is crucial. In this work, we establish a novel approach, i.e., an expert system, to study multi-faceted policies, and this approach shows great potential with respect to future epidemic control. We are aware of the ethical implications that expert systems have on the decision process and we make the disclaimer that all results based on, or derived from, the expert system that we propose should be carefully interpreted by experts in the field of public health, and in a much broader context of economics, well-being and education. We note that the work in this manuscript was conducted by a multi-disciplinary consortium that includes computer scientists and scientists with a background public health, epidemiology and bio-statistics.

In this work, we focus on the clinical outcomes of the intervention strategies and use the reduced contacts as proxy for the quality of life lost. This could be extended into more formal economic evaluations by assessing the health and monetary benefits and costs of different interventions for various stakeholders. The COVID-19 pandemic demonstrated the broad impact of infectious diseases on sectors outside health care. This stresses the need to capture a societal and thus multi-objective perspective in the decision making process on public health and health care interventions. Our learned policies confirm this, showing that focusing solely on reducing the number of hospitalizations results in taking drastic measures -- more than a thousand social interactions lost per person over the span of 4 months -- that may have a long-lasting impact on the population.

To conclude, we show that multi-objective reinforcement learning can be used to learn a wide set of high-quality policies on real-world problems, providing the decision maker with insightful and diverse alternatives, and showing the impact of taking extreme measures.

\bibliographystyle{unsrtnat}
\bibliography{references}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\appendix
\counterwithin{figure}{section}
\counterwithin{table}{section}
\counterwithin{equation}{section}

\section{Compartmental model for SARS-CoV-2}
In this work we utilise the compartmental model proposed by \citet{abrams2021modelling} and extend the model to a multi-objective Markov decision process (MOMDP). The MOMDP used in this work is outlined in the main paper. However, we describe the details of the underlying compartmental model used to model the spread of COVID-19 below.

\begin{figure}[h]
    \centering
     \begin{tikzpicture}
     \draw (0.0pt, 0.0pt)node[fill=green!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](0){$\textbf{S}$};
     \draw (60.0pt, 0.0pt)node[fill=orange!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](1){$\textbf{E}$};
     \draw (120.0pt, 0.0pt)node[fill=red!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](2){$\textbf{I}^{presym}$};
     \draw (180.0pt, 10.0pt)node[fill=red!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](3){$\textbf{I}^{asym}$};
     \draw (170.0pt, -30.0pt)node[fill=red!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](4){$\textbf{I}^{mild}$};
     \draw (170.0pt, -70.0pt)node[fill=red!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](5){$\textbf{I}^{sev}$};
     \draw (230.0pt, -50.0pt)node[fill=blue!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](6){$\textbf{I}^{hosp}$};
     \draw (230.0pt, -90.0pt)node[fill=blue!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](7){$\textbf{I}^{icu}$};
     \draw (300.0pt, -50.0pt)node[fill=orange!50, thick, minimum height=0.6cm,minimum width=0.6cm, draw](8){$\textbf{R}$};
     \draw (300.0pt, -90.0pt)node[fill=purple!50, thick, minimum height=0.6cm,minimum width=0.6cm, draw](9){$\textbf{D}$};

     %\path[->] (6) edge[dotted, bend right = 50]
     \path[->] (0) edge[] node[above]{$\boldsymbol{\lambda}(t)$} (1);
     \path[->] (1) edge[] node[above]{$\gamma$} (2);
     \path[->] (2) edge[] node[below]{$\textbf{p}\theta$} (3);
     \path[->] (2) edge[] node[near end, below left]{$(1-\textbf{p})\theta$} (4);
     \path[->] (4) edge[] node[left]{$\boldsymbol{\psi}$} (5);
     \path[->] (5) edge[] node[near end, below left]{$(1-\phi_{1})\boldsymbol{\omega}$} (7);
     \path[->] (5) edge[] node[above]{$\phi_{1}\boldsymbol{\omega}$} (6);
     \path[->] (7) edge[] node[near end, below right]{$\boldsymbol{\delta}_{4}$} (8);
     \path[->] (7) edge[] node[above]{$\boldsymbol{\tau}_{2}$} (9);
     \path[->] (6) edge[] node[above]{$\boldsymbol{\delta}_{3}$} (8);
     \path[->] (6) edge[] node[near start, below left]{$\boldsymbol{\tau}_{1}$} (9);
     \path[->] (3) edge[bend left = 30] node[above right]{$\delta_{1}$} (8);
     \path[->] (4) edge[bend left = 30] node[above left]{$\boldsymbol{\delta}_{2}$} (8);

    \end{tikzpicture}

    \caption{Schematic diagram of the compartmental model for SARS-CoV-2 presented by \citet{abrams2021modelling} which is used to derive the MOMDP.
    }
    \label{fig:ode_mdel}
\end{figure}

We utilise an adapted SEIR mathematical compartmental model proposed by Abrams et al. to contruct the MOBelCov MOMDP with deterministic and stochastic state transitions. In this model members of the population are susceptible to infection when they are in compartment $\textbf{S}$\footnote{Boldface vector notation is used to denote the multiple age groups for each compartment.}. If an individual comes into contact with an infections individual then they move to the exposed compartment, $\textbf{E}$, at a time specific rate, $\boldsymbol{\lambda} (t)$. After a period of time an exposed individual becomes infectious, where they move to the pre-symptomatic compartment, $\textbf{I}^{presym}$, at rate $\gamma$. Once infected, individuals develop mild symptoms, $\textbf{I}^{mild}$ , with probability $1 - \textbf{p}$ or do not develop any symptoms, $\textbf{I}^{asym}$, with probability $\textbf{p}$, where asymptomatic individuals recover, $\textbf{R}$, at rate $\delta_{1}$. Individuals who experience symptoms can suffer from a mild infection, $\textbf{I}^{mild}$, and recover at rate $\boldsymbol{\delta}_{2}$, or they suffer from a more severe infection, $\textbf{I}^{sev}$, at a rate $\boldsymbol{\psi}$. Individuals with a severe infection are then transferred to a hospital for treatment, $\textbf{I}^{hosp}$ with probability $\phi_{1}$. However, some individuals become critically ill and are transferred directly to the intensive care unit (ICU) with probability $1 - \psi_{1}$. Individuals in the hospital, $\textbf{I}^{hosp}$ and $\textbf{I}^{icu}$, recover at rate $\boldsymbol{\delta}_{3}$ or $\boldsymbol{\delta}_{4}$ and die at rate $\boldsymbol{\tau}_{3}$ and $\boldsymbol{\tau}_{4}$ respectively. Figure \ref{fig:ode_mdel} outlines the compartmental model defined by Abrams et al.~\cite{abrams2021modelling}.

\subsection{Deterministic Compartmental Model}
\label{sec:deterministic_comp_model}
The flows of the deterministic model are defined by a set of ordinary differential equations, which are outlined as follows:

\begin{align*}
\frac{d\textbf{S}(t)}{dt} & = -\lambda(t)(\textbf{S})(t), \\
%
\frac{d\textbf{E}(t)}{dt} & = \lambda(t)(\textbf{S})(t) - \gamma \textbf{E}(t), \\
%
\frac{d\textbf{I}^{presym}(t)}{dt} & = \gamma \textbf{E}(t) - \theta \textbf{I}^{presym}(t), \\
%
\frac{d\textbf{I}^{asym}(t)}{dt} & = \theta p \textbf{I}^{presym}(t) - \delta_{1} \textbf{I}^{asym}(t), \\
%
\frac{d\textbf{I}^{mild}(t)}{dt} & = \theta (1 - p) \textbf{I}^{presym}(t) - \{ \psi + \omega_{2} \} \textbf{I}^{mild}(t), \\
%
\frac{d\textbf{I}^{sev}(t)}{dt} & = \psi \textbf{I}^{mild}(t) - \omega \textbf{I}^{sev}(t), \\
%
\frac{d\textbf{I}^{hosp}(t)}{dt} & = \phi_{1} \omega \textbf{I}^{sev}(t) - \{ \delta_{3} + \tau_{1} \} \textbf{I}^{hosp}(t), \\
%
\frac{d\textbf{I}^{icu}(t)}{dt} & = (1 - \phi_1) \omega \textbf{I}^{sev}(t) - \{ \delta_{4} + \tau_{2} \} \textbf{I}^{icu}(t), \\
%
\frac{d\textbf{D}(t)}{dt} & = \tau_{1} \textbf{I}^{hosp}(t) - \tau_{2} \textbf{I}^{icu}(t), \\
%
\frac{d\textbf{R}(t)}{dt} & = \delta_{1} \textbf{I}^{asym}(t) + \delta_{2} \textbf{I}^{mild}(t) + \delta_{3} \textbf{I}^{hosp}(t) + \delta_{4} \textbf{I}^{icu}(t)
\end{align*}

where, for example, $\textbf{S} = (S_{1}(t), S_{2}(t), ..., S_{k}(t))^{T}$ is the vector representing the susceptible members of the population of each age group $k$ at time $t$.

In this set of ordinary differential equations, each state variable represents a vector over all age groups for a particular compartment at time $t$.
Infection dynamics are governed by an age-specific force of infection $\foi$:
\begin{equation}
    \foi(k,t) = \sum_{k'=1}^{K}\beta(k,k')I_{k'}(t),
\end{equation}
where $k$ is the respective age group of a total of K age groups, and $\beta(k,k')$ is the time-invariant transmission rate that encodes the average per capita rate at which an infectious individual in age group $k$ makes an effective contact with a susceptible individual in age group $k'$, per unit of time.

Under the social contact hypothesis \cite{wallinga2006using}, we have that:
\begin{equation}
\beta(k,k') = q \cdot C(k,k'),
\end{equation}
where q is a proportionality factor and the matrix $C$ denotes the social mixing behaviour within and between different age groups in the population, and is referred to as a social contact matrix. %(TODO: add appropriate reference).
Following \citet{abrams2021modelling}, we rely on distinct social contact matrices for symptomatic and asymptomatic individuals, respectively $C_{s}$ and $C_{a}$. Therefore it is possible to define the transmission rates for both symptomatic and asymptomatic individuals as follows:

\begin{equation}
\boldsymbol{\beta}_{s}(k, k') = q_{s} \cdot C_{s}(k, k'),    
\end{equation}
and
\begin{equation}
\boldsymbol{\beta}_{a}(k, k') = q_{a} \cdot C_{a}(k, k').  
\end{equation}
Moreover, the age-dependent force of infection can be defined as follows:
\begin{equation}
    \boldsymbol{\lambda}(t) = \boldsymbol{\beta}_{a} \times \{ \textbf{I}^{presym}(t) + \textbf{I}^{asym}(t) \} +
    \boldsymbol{\beta}_{s} \times \{ \textbf{I}^{mild}(t) + \textbf{I}^{sev}(t) \},
\end{equation}
where $\boldsymbol{\lambda}(t) = (\lambda(1, t), \lambda(1, t), ..., \lambda(K, t))$. For all further information about the different compartments and parameters please refer to the work of \citet{abrams2021modelling}

To create a version of MOBelCov with a deterministic transition function, $M$, we focus on the deterministic model proposed by \citet{abrams2021modelling}. Given that the transitions within the compartmental model are deterministic it is possible to utilise the highlighted model transitions for MOBelCov. Applying the contact matrix $\hat{C}$ to the model state $s_{m}$ progresses the model for one timestep and returns a new compartmental model state $s'_{m}$. Given $s_{m}$ and $\hat{C}$, the  deterministic compartmental model returns $s'_{m}$ with probability of $1$. Therefore, it is possible to use this process as a deterministic transition function, $M$, for MOBelCov.

\subsection{Stochastic Compartmental Model}
Intervening in the spread of the virus by, for example, reducing social contacts or government interventions introduces uncertainty in the further course of the outbreak. Therefore, to understand how this uncertainty affect the spread of the disease we introduce a stochastic component model which can model the uncertainty generated by interventions in social contacts.

By formulating the set of differential equations defined above, as a chain-binomial, we can obtain stochastic trajectories from this model \cite{bailey1975mathematical}. A chain-binomal model assumes a stochastic model where infected individuals are generated by some underlying probability distribution. For the stochastic model we consider a time interval $(t, t +h]$, where $h$ is defined as the length between two consecutive time points. Similar to \citet{abrams2021modelling}, in this work we set $h = \frac{1}{24}$. \citet{abrams2021modelling} define the set of differential equations as a chain binomial as follows:

\begin{align*}
\textbf{S}_{t+h} (k) & = \textbf{S}_{t} (k) - \textbf{E}_{new, t+h}(k),\\
%
\textbf{E}_{t+h}(k) & = \textbf{E}_{t}(k) + \textbf{E}_{new, t+h}(k) - \textbf{I}^{presym}_{new, t + h} (k),\\
%
\textbf{I}^{presym}_{t+h}(k) & = \textbf{I}^{presym}_{t}(k) + \textbf{I}^{presym}_{new, t + h}(k) - \textbf{I}^{asym}_{new, t + h}(k) - \textbf{I}^{mild}_{new, t + h}(k),\\
%
\textbf{I}^{asym}_{t+h}(k) & = \textbf{I}^{asym}_{t}(k) + \textbf{I}^{asym}_{new, t + h}(k) - \textbf{R}^{asym}_{new, t + h}(k),\\
%
\textbf{I}^{mild}_{t+h}(k) & = \textbf{I}^{mild}_{t}(k) + \textbf{I}^{mild}_{new, t + h}(k) - \textbf{I}^{sev}_{new, t + h}(k) - \textbf{R}^{mild}_{new, t + h}(k),\\
%
\textbf{I}^{sev}_{t+h}(k) & = \textbf{I}^{sev}_{t}(k) + \textbf{I}^{sev}_{new, t + h}(k) - \textbf{I}^{hosp}_{new, t + h}(k) - \textbf{I}^{icu}_{new, t + h}(k),\\
%
\textbf{I}^{hosp}_{t+h}(k) & = \textbf{I}^{hosp}_{t}(k) + \textbf{I}^{hosp}_{new, t + h}(k) - \textbf{D}^{hosp}_{new, t + h}(k) - \textbf{R}^{hosp}_{new, t + h}(k),\\
%
\textbf{I}^{icu}_{t+h}(k) & = \textbf{I}^{icu}_{t}(k) + \textbf{I}^{icu}_{new, t + h}(k) - \textbf{D}^{icu}_{new, t + h}(k) - \textbf{R}^{icu}_{new, t + h}(k),\\
%
\textbf{D}_{t+h}(k) & = \textbf{D}_{t}(k) + \textbf{D}^{hosp}_{new, t + h}(k) + \textbf{D}^{icu}_{new, t + h}(k),\\
%
\textbf{R}_{t+h}(k) & = \textbf{R}_{t}(k) + \textbf{R}^{asym}_{new, t + h}(k) + \textbf{R}^{mild}_{new, t + h}(k) + \textbf{R}^{hosp}_{new, t + h}(k) + \textbf{R}^{icu}_{new, t + h}(k)
\end{align*}

where,

\begin{align*}
\textbf{E}_{new, t+h} & \sim \textit{Binomial} \left( \textbf{S}_{t} (k), p^{*}_t (k) = 1 - \{ 1 -  p^{*}_t (k)\}^{\textbf{I}_{t}} \right), \\
%
p^{*}_t (k) & = 1 - exp \left[ -h \sum^{K}_{k' = 1} \beta_{asym}(k, k')\{ \textbf{I}^{asym}_{t}(k')\} + \beta_{sym}(k, k') \{\textbf{I}^{mild}_{t}(k')+\textbf{I}^{sev}_{t}(k')\} \right],\\
%
\textbf{I}^{presym}_{new, t+h} (k)&  \sim Binomial\left( \textbf{I}^{presym}_{t}(k), 1 - exp(-hp(k)\theta) \right),\\
%
\textbf{I}^{mild}_{new, t+h} (k) & \sim Binomial\left( \textbf{I}^{presym}_{t}(k), 1 - exp\left[-h \{1-p(k)\}\theta \right]) \right),\\
%
\textbf{I}^{sev}_{new, t+h} (k) & \sim Binomial\left( \textbf{I}^{mild}_{t}(k), 1 - exp\{-h\psi(k)\}) \right),\\
%
\textbf{I}^{hosp}_{new, t+h} (k) & \sim Binomial\left( \textbf{I}^{sev}_{t}(k), 1 - exp\{-h\phi_{1}(k)\omega(k)\}) \right),\\
%
\textbf{I}^{icu}_{new, t+h} (k) & \sim Binomial\left( \textbf{I}^{sev}_{t}(k), 1 - exp\left[-h \{1-\phi_{1}(k)\}\omega(k) \right] \right),\\
%
\textbf{D}^{hosp}_{new, t+h} (k) & \sim Binomial\left( \textbf{I}^{hosp}_{t}(k), 1 - exp\{-h\tau_{1}(k) \} \right),\\
%
\textbf{D}^{icu}_{new, t+h} (k) & \sim Binomial\left( \textbf{I}^{icu}_{t}(k), 1 - exp\{-h\tau_{2}(k) \} \right),\\
%
\textbf{R}^{asym}_{new, t+h} (k) & \sim Binomial\left( \textbf{I}^{asym}_{t}(k), 1 - exp\left(-h\delta_{2}(k) \right) \right), \\
%
\textbf{R}^{hosp}_{new, t+h} (k) & \sim Binomial\left( \textbf{I}^{hosp}_{t}(k), 1 - exp\{-h\delta_{3}(k) \} \right),\\
%
\textbf{R}^{icu}_{new, t+h} (k) & \sim Binomial\left( \textbf{I}^{icu}_{t}(k), 1 - exp\{-h\delta_{4}(k) \} \right).
\end{align*}

Given MOBelCov also calculates new hospitalisations, $\textbf{H}^{new}$, we define $\textbf{H}^{new}$ for the stochastic compartmental model as follows:
\[
\textbf{H}^{new}_{t+h}(k) = \textbf{H}^{new}_{t}(k) + \textbf{I}^{hosp}_{new, t+h} (k).
\]
For more details on this model and the chain-binomial representation of the differential equations, we refer the reader to the work of \citet{abrams2021modelling}.

To create a version of MOBelCov with a stochastic transition function, $M$, we utilise the stochastic compartmental model outlined above. Given the transitions within the compartmental model are derived by an underlying probability distribution it is possible to utilise the stochastic compartmental model transitions for MOBelCov. As previously outlined in Section \ref{sec:deterministic_comp_model} the contact matrix $\hat{C}$ applied the model state $s_{m}$ progresses the model and returns a new model state $s'_{m}$. Given the underlying model dynamics are governed in a probabilistic manner, the model returns $s'_{m}$ stochastically. Therefore, it is possible to use this process as a stochastic transition function, $M$, for MOBelCov.

\subsection{A Note on Model Parameters}

The model is parameterised using the mean of the posteriors as reported by \citet{abrams2021modelling}.

The population size for each of the considered age groups was taken from the Belgian statistical agency STATBEL\footnote{\url{https://statbel.fgov.be/nl/themas/bevolking/structuur-van-de-bevolking\#figures}}. To initialise the model, we used the number of confirmed cases until 13 March 2020 \cite{abrams2021modelling}, as reported by the Belgian agency for public health Sciensano\footnote{\url{https://epistat.wiv-isp.be/covid/}}.

\subsection{Modelling interventions}
\label{sec:interventions}
In order to model different types of interventions, we follow \citet{abrams2021modelling}. Firstly, 
to consider distinct exit scenarios, we alter the social contact matrices to reflect a contact reduction in a particular age group. Secondly, we assume that compliance to the interventions is gradual and model this using a logistic compliance function. We use the logistic compliance function in function of time $t$ proposed by Abrams et al.,
\begin{equation}
c(t,t_I) = \frac{\exp(\beta^*_0 + \beta^*_1(t-t_I))}{1+\exp(\beta^*_0 + \beta^*_1(t-t_I))},
\end{equation}
where $t_I$ indicates the time the intervention started \cite{abrams2021modelling}.
We initialise $\beta^*_1$ to the value estimated in by Abrams et al. and choose $\beta^*_0=-5$, as an intercept to have $c(t)=0$ for $t=0$, in correspondence with Figure~F2 in the Supplementary Information of \citet{abrams2021modelling}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Additional results}
\subsection{Learned coverage sets}

In Figure~\ref{fig:pareto-front-arh}, we display the coverage set with respect to the number of hospitalizations of the learned policies for both the ODE and Binomial variants, as the Binomial variants were omitted in the main paper for clarity. When trained on $\mathbf{R}_{\text{ARH}}$, the Binomial coverage set is on par with its ODE counterpart. On the other hand, the coverage sets when trained on $\mathbf{R}_{\text{ARI}}$ are slightly worse, as confirmed by the evaluation metrics in Table~\ref{tab:coverage-set}.

\begin{figure}[h]
    \centering
    \input{fig_arh}
    \caption{The Pareto front of policies discovered by PCN, showing the different compromises between the number of hospitalizations and the number of lost contacts. Although $\text{PCN}_\text{ARI}$ was trained on the number of infections, it still shows a competitive coverage set w.r.t. hospitalizations.}
    \label{fig:pareto-front-arh}
\end{figure}

Moreover, we also show the coverage sets with respect to the number of infections in Figure~\ref{fig:pareto-front-ari}. Interestingly, both variants trained on the Binomial model learn a coverage set that almost completely dominates the variants trained on the ODE model. This is due to the stochasticity of the Binomial model. PCN stores trajectories in its dataset, and keep a selection for training. Due to the stochasticity in the transition function, some trajectories might result in higher returns, even though the same policy was applied. Due to the high variability in infections, this difference might become significant. This is also why the variants using the Binomial model show a worse $I_{\varepsilon}$ and $I_{\varepsilon}-mean$ than their ODE counterpart: their desired return stems from a single trajectory, while we evaluate the policies over multiple trajectories. The resulting average return is then different than the desired return. Section~\ref{sec:arh-vs-ari} explains why some policies trained on $\mathbf{R}_{\text{ARH}}$ show suboptimal behavior with respect to the number of infections.

\begin{figure}[h]
    \centering
    \input{fig_ari}
    \caption{The Pareto front of policies discovered by PCN, showing the different compromises between the number of hospitalizations and the number of lost contacts. Although $\text{PCN}_\text{ARI}$ was trained on the number of infections, it still shows a competitive coverage set w.r.t. hospitalizations.}
    \label{fig:pareto-front-ari}
\end{figure}

\subsection{Policy executions}

PCN learns a coverage set containing more than 100 different policies. To gain a better insight about their behavior, and how they differ from each other, we plot executions of each learned policy in Figure~\ref{fig:policy-executions-ode}. The plots are displayed from the least restrictive policy in terms of social burden to the most restrictive one.

We notice that policies~60 to 140 display similar behavior. At first, they all completely restrict social contact, effectively continuing the lockdown. Afterwards, they progressively lessen these restrictions. What differs in these policies is the timing and the speed at which the restrictions are lessened. These policies correspond to the right-most part of the coverage set displayed in Figure~\ref{fig:pareto-front-arh}, which shows many alternative for the $[0, 40000]$ hospitalizations segment.

\begin{figure}
    \centering
    \plotPolicies{0}{19}
    \caption{Execution of policies 1 to 20.}
    \label{fig:policy-executions-ode}
\end{figure}
\begin{figure}\ContinuedFloat
    \centering
    \plotPolicies{20}{39}
    \caption{Execution of policies 21 to 40.}
    \label{fig:policy-executions-ode}
\end{figure}
\begin{figure}\ContinuedFloat
    \centering
    \plotPolicies{40}{59}
    \caption{Execution of policies 41 to 60.}
    \label{fig:policy-executions-ode}
\end{figure}
\begin{figure}\ContinuedFloat
    \centering
    \plotPolicies{60}{79}
    \caption{Execution of policies 61 to 80.}
    \label{fig:policy-executions-ode}
\end{figure}
\begin{figure}\ContinuedFloat
    \centering
    \plotPolicies{80}{99}
    \caption{Execution of policies 81 to 100.}
    \label{fig:policy-executions-ode}
\end{figure}
\begin{figure}\ContinuedFloat
    \centering
    \plotPolicies{100}{119}
    \caption{Execution of policies 101 to 120.}
    \label{fig:policy-executions-ode}
\end{figure}
\begin{figure}\ContinuedFloat
    \centering
    \plotPolicies{120}{139}
    \caption{Execution of policies 121 to 140.}
    \label{fig:policy-executions-ode}
\end{figure}

\subsection{Experiment parameters}

We used the same hyper-parameters across all experiments. Each experiment resulted in 5 independent trials. Finally, we performed a grid-search over possible hyper-parameter values. All hyper-parameters used and their possible values explored during grid-search are displayed in Table~\ref{tab:pcn-hyperparameters}.

\begin{table}[t]
    \centering
    \begin{tabular}{|c|c|c|}
        \hline
        hyper-parameter & value & grid-search \\
        \hline
        learning rate & $0.001$ &  \\
        total training timesteps & $300000$  & \\
        batch size & $256$ & $256, 1024$ \\
        model updates & $50$ & \\
        episodes between updates & $10$ & \\
        ER size (in episodes) & $1000$ & $400, 500, 1000$ \\
        initial random episodes & $200$ & $50, 200$ \\
        exploration noise & $0.1$ & $0, 0.1, 0.2$ \\
        desired return noise & $0.2$ & $0, 0.1, 0.2$ \\
        reward scaling & $[10000, 100]$ & \\
        \hline
    \end{tabular}
    \caption{The different hyper-parameters used by our extension of PCN. The right-most column also shows, when applicable, the different values tried during grid-search.}
    \label{tab:pcn-hyperparameters}
\end{table}

\begin{table}[h]
    \centering
    \input{tab_nn-architecture}
    \caption{The 4 different neural network architectures explored for our experiments. All the displayed results use the \texttt{conv1d-big} variant.}
    \label{tab:nn-architecture}
\end{table}

\subsection{Neural network architecture}

Next to the hyper-parameter search, we also performed a grid search over 4 different neural network architectures. All the architectures have the same structure. We use a compartment embedding $sc_{emb}$, a SCM embedding $sm_{emb}$ and a school-holidays embedding $sh_{emb}$ that take as inputs the compartment, the previous $p_w, p_s, p_l$ values (as they fully define the SCM $\hat{C}$) and a boolean flag for school holidays, respectively. All these embeddings have a same-sized embedding of 64, which are multiplied together to form the full state embedding. This state-embedding is used as input for another network, $s_{emb}$. Additionally, we use a common embedding $c_{emb}$ for the concatenation of the desired return and horizon. Finally, the results of $s_{emb}$ and $c_{emb}$ are multiplied together, before passing through a fully connected network $fc$ that has 3 outputs, one for $p_w, p_s, p_l$ respectively.

All the architectures of the different components are displayed in Table~\ref{tab:nn-architecture}. The variant used in all experiments is \texttt{conv1d-big}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related Work}
RL has also been used to learn effective mitigation strategies which limit the spread of COVID-19 \cite{ohi2020exploring}. Kwak et al. \cite{kwak2021covid} use deep RL, with an SIRD model, to learn efficacy of lockdown and travel restrictions in controlling the COVID-19 pandemic using data collected from various government organisations.  Bastani et al. \cite{bastani2021efficient} use a RL algorithm called Eva to limit the influx of asymptomatic travellers infected with SARS-CoV-2 to Greece. Kompella et al. \cite{kompella2020reinforcement} define a COVID-19 simulator and a methodology to learn mitigation policies that minimize the economic impact without overwhelming the hospital capacity. RL has also been used to learn effective COVID-19 vaccination distribution strategies \cite{beigi2021application,awasthi2020vacsim} and to create decision support systems for mitigating the spread of COVID-19.

\end{document}

### 2022-04-25 paper/tab_nn-architecture.tex
added paper

\begin{tabular}{|c|c|c|c|c|c|c|}
    \hline
        variant & $sc_{emb}$      & $sm_{emb}$   & $sh_{emb}$   & $s_{emb}$    & $c_{emb}$    & $fc$ \\
    \hline
                & conv1d(10,20)   & linear(3,64) & linear(1,64) & linear(64,64)& linear(3,64) & linear(64,64)\\
                & relu            & sigmoid      & sigmoid      & sigmoid      & sigmoid      & relu\\
    conv1d-small  & conv1d(20,20)   &              &              &              &              & linear(64,3)\\
                & relu            &              &              &         &         &      \\
                & linear(100,64)  &              &              &         &         &      \\
                & sigmoid         &              &              &         &         &      \\
    \hline
                & conv1d(10,20)   & linear(3,64) & linear(1,64) & linear(64,64)& linear(3,64) & linear(64,64)\\
                & relu            & relu         & relu         & relu         & sigmoid      & relu\\
    conv1d-big & conv1d(20,20)   & linear(64,64)& linear(64,64)&              &              & linear(64,3)\\
                & relu            & sigmoid      & sigmoid      &              &              &      \\
                & linear(100,64)  &          &          &         &         &      \\
                & sigmoid         &          &          &         &         &      \\
    \hline
                & linear(130,64)  & linear(3,64) & linear(1,64) & linear(64,64)& linear(3,64) & linear(64,64)\\
    dense-small & sigmoid         & sigmoid      & sigmoid      & sigmoid      & sigmoid      & relu\\
                &                 &              &              &              &              & linear(64,3)\\
    \hline
                & linear(130,64)  & linear(3,64) & linear(1,64) & linear(64,64)& linear(3,64) & linear(64,64)\\
    dense-big & relu            & relu         & relu         & relu         & sigmoid      & relu\\
                & linear(64,64)   & linear(64,64)& linear(64,64)&              &              & linear(64,3)\\
                & sigmoid         & sigmoid      & sigmoid      &              &              &      \\
    \hline
\end{tabular}

### 2022-07-27 main_pcn.py
change sb scaling

@@ -250,7 +250,7 @@ if __name__ == '__main__':
 
     env_type = 'ODE' if args.env == 'ode' else 'Binomial'
     n_evaluations = 1 if env_type == 'ODE' else 10
-    scale = np.array([800000, 10000, 50., 20, 50, 100])
+    scale = np.array([800000, 10000, 50., 20, 50, 90])
     ref_point = np.array([-15000000, -200000, -1000.0, -1000.0, -1000.0, -1000.0])/scale
     scaling_factor = torch.tensor([[1, 1, 1, 1, 1, 1, 0.1]]).to(device)
     max_return = np.array([0, 0, 0, 0, 0, 0])/scale

### 2022-07-27 pcn.py
add noise

@@ -327,7 +327,7 @@ def train(env,
         loss = []
         entropy = []
         for _ in range(n_model_updates):
-            l, lp = update_model(model, opt, experience_replay, batch_size=batch_size)
+            l, lp = update_model(model, opt, experience_replay, batch_size=batch_size, noise=noise)
             loss.append(l.detach().cpu().numpy())
             lp = lp.detach().cpu().numpy()
             ent = np.sum(-np.exp(lp)*lp)

### 2022-07-27 paper/fig_arh.tex

\begin{tikzpicture}
    \pgfplotsset{
        scale only axis,
        height=6cm,
        xlabel={total number of daily new hospitalizations},
        ylabel={total lost contacts}
    }
    \begin{axis}[
    legend entries={{$\text{PCN}_\text{ARH}$ (ODE)}, {$\text{PCN}_\text{ARI}$ (ODE)}, {$\text{PCN}_\text{ARH}$ (Binomial)}, {$\text{PCN}_\text{ARI}$ (Binomial)}, {Fixed}},
    width=9.5cm
    ]
    \addplot[only marks,draw opacity=0.5,mark=x,red] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/cs_ode.csv};
    \addplot[only marks,draw opacity=0.5,mark=o,blue] table [x=o_1, y=o_5, col sep=comma] {data/cs/ari/cs_ode.csv};
    \addplot[only marks,draw opacity=0.5,mark=triangle,orange] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/cs_binomial.csv};
    \addplot[only marks,draw opacity=0.5,mark=diamond,cyan] table [x=o_1, y=o_5, col sep=comma] {data/cs/ari/cs_binomial.csv};
    \addplot[only marks,mark=+] table [x=o_0, y=o_1, col sep=comma] {data/cs/cs_fixed.csv};
    \end{axis}
\end{tikzpicture}

### 2022-07-27 paper/fig_ari.tex

\begin{tikzpicture}
    \pgfplotsset{
      scale only axis,
      height=6cm,
      xlabel={total number of infections},
      ylabel={total lost contacts}
    }
  \begin{axis}[
    legend entries={{$\text{PCN}_\text{ARH}$ (ODE)}, {$\text{PCN}_\text{ARI}$ (ODE)}, {$\text{PCN}_\text{ARH}$ (Binomial)}, {$\text{PCN}_\text{ARI}$ (Binomial)}},
    %, {Fixed}},
    width=9.5cm
  ]
  \addplot[only marks,draw opacity=0.5,mark=x,red] table [x=o_0, y=o_5, col sep=comma] {data/cs/arh/cs_ode.csv};
  \addplot[only marks,draw opacity=0.5,mark=o,blue] table [x=o_0, y=o_5, col sep=comma] {data/cs/ari/cs_ode.csv};
  \addplot[only marks,draw opacity=0.5,mark=triangle,orange] table [x=o_0, y=o_5, col sep=comma] {data/cs/arh/cs_binomial.csv};
  \addplot[only marks,draw opacity=0.5,mark=diamond,cyan] table [x=o_0, y=o_5, col sep=comma] {data/cs/ari/cs_binomial.csv};
  %\addplot[only marks,mark=+] table [x=o_0, y=o_1, col sep=comma] {data/cs/cs_fixed.csv};
  \end{axis}
\end{tikzpicture}

### 2022-07-27 paper/fig_policy-execution.tex

\begin{tikzpicture}
\pgfplotsset{
      scale only axis,
      date coordinates in=x,
      xticklabel=\day/\month,
      height=2cm,
    }
    \begin{groupplot}[
    group style={
            % set how the plots should be organized
            group size=1 by 3,
            % only show ticklabels and axis labels on the bottom
            y descriptions at=edge left,
            x descriptions at=edge bottom,
            % set the `vertical sep' to zero
            horizontal sep=0pt,
            vertical sep=0pt
        },
      width=9.5cm,
      axis y line*=left,
      xlabel={time},
      ylabel={people},
    ]
    % GROUP WITH LEFT Y-AXIS: PEOPLE (HOSP, ICU, DEATHS)
    \nextgroupplot[scaled y ticks=false]
    \addplot[mark=none, cyan] table [x=dates, y=i_hosp_new, col sep=comma] {data/ode/policy_0.csv};
    \label{i_hosp_new}
    \addplot[mark=none, orange] table [x=dates, y=i_icu_new, col sep=comma] {data/ode/policy_0.csv};
    \label{i_icu_new}
    \addplot[mark=none, red] table [x=dates, y=d_new, col sep=comma] {data/ode/policy_0.csv};
    \label{d_new}
    \nextgroupplot[scaled y ticks=false]
    \addplot[mark=none, cyan] table [x=dates, y=i_hosp_new, col sep=comma] {data/ode/policy_30.csv};
    \addplot[mark=none, orange] table [x=dates, y=i_icu_new, col sep=comma] {data/ode/policy_30.csv};
    \addplot[mark=none, red] table [x=dates, y=d_new, col sep=comma] {data/ode/policy_30.csv};
    \nextgroupplot[scaled y ticks=false]
    \addplot[mark=none, cyan] table [x=dates, y=i_hosp_new, col sep=comma] {data/ode/policy_139.csv};
    \addplot[mark=none, orange] table [x=dates, y=i_icu_new, col sep=comma] {data/ode/policy_139.csv};
    \addplot[mark=none, red] table [x=dates, y=d_new, col sep=comma] {data/ode/policy_139.csv};
    \end{groupplot}
    % GROUP WITH RIGHT Y-AXIS: ACTIONS (P_W, P_S, P_L)
    \begin{groupplot}[
    group style={
            % set how the plots should be organized
            group size=1 by 3,
            % only show ticklabels and axis labels on the bottom
            y descriptions at=edge right,
            x descriptions at=edge bottom,
            % set the `vertical sep' to zero
            horizontal sep=0pt,
            vertical sep=0pt
        },
      width=9.5cm,
      axis y line*=right,
      axis x line=none,
      ymax=1.1,
      ylabel={proportion},
      legend style={at={(0.5,1)},anchor=south},
    %   legend pos=outer north east,
      legend columns=3, 
      legend cell align=left,
    ]
    \nextgroupplot[scaled y ticks=false]
    \addplot[mark=none, dashed, blue] table [x=dates, y=p_w, col sep=comma] {data/ode/policy_0.csv};
    \label{p_w}
    \addplot[mark=none, dashed, magenta] table [x=dates, y=p_s, col sep=comma] {data/ode/policy_0.csv};
    \label{p_s}
    \addplot[mark=none, dashed, darkgray] table [x=dates, y=p_l, col sep=comma] {data/ode/policy_0.csv};
    \label{p_l}
    \addlegendimage{/pgfplots/refstyle=i_hosp_new}\addlegendentry{$p_w$}
    \addlegendimage{/pgfplots/refstyle=i_icu_new}\addlegendentry{$p_l$}
    \addlegendimage{/pgfplots/refstyle=d_new}\addlegendentry{$p_s$}
    \addlegendimage{/pgfplots/refstyle=p_w}\addlegendentry{{daily new hosp}}
    \addlegendimage{/pgfplots/refstyle=p_s}\addlegendentry{{daily new ICU}}
    \addlegendimage{/pgfplots/refstyle=p_l}\addlegendentry{{daily deaths}}
    \nextgroupplot[scaled y ticks=false]
    \addplot[mark=none, dashed, blue] table [x=dates, y=p_w, col sep=comma] {data/ode/policy_30.csv};
    \addplot[mark=none, dashed, magenta] table [x=dates, y=p_s, col sep=comma] {data/ode/policy_30.csv};
    \addplot[mark=none, dashed, darkgray] table [x=dates, y=p_l, col sep=comma] {data/ode/policy_30.csv};
    \nextgroupplot[scaled y ticks=false]
    \addplot[mark=none, dashed, blue] table [x=dates, y=p_w, col sep=comma] {data/ode/policy_139.csv};
    \addplot[mark=none, dashed, magenta] table [x=dates, y=p_s, col sep=comma] {data/ode/policy_139.csv};
    \addplot[mark=none, dashed, darkgray] table [x=dates, y=p_l, col sep=comma] {data/ode/policy_139.csv};
    \end{groupplot}
  \end{tikzpicture}

### 2023-01-19 paper/fig_budget.tex
update paper

\begin{tikzpicture}
    \pgfplotsset{
        scale only axis,
        height=6cm,
        xlabel={total number of daily new hospitalizations},
        ylabel={total lost contacts}
    }
    \begin{axis}[
    legend entries={{$\text{PCN}_\text{ARH}$ (ODE)}, {Budget 2}, {Budget 3}, {Budget 4}, {Budget 5}, {Budget None}},
    width=9.5cm
    ]
    \addplot[only marks,draw opacity=0.5,mark=x,red] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/cs_ode.csv};
    \addplot[only marks,draw opacity=0.5,mark=o,blue] table [x=o_1, y=o_5, col sep=comma] {data/cs/budget/budget_2.csv};
    \addplot[only marks,draw opacity=0.5,mark=square,pink] table [x=o_1, y=o_5, col sep=comma] {data/cs/budget/budget_3.csv};
    \addplot[only marks,draw opacity=0.5,mark=triangle,orange] table [x=o_1, y=o_5, col sep=comma] {data/cs/budget/budget_4.csv};
    \addplot[only marks,draw opacity=0.5,mark=diamond,cyan] table [x=o_1, y=o_5, col sep=comma] {data/cs/budget/budget_5.csv};
    \addplot[only marks,mark=+] table [x=o_1, y=o_5, col sep=comma] {data/cs/budget/budget_None.csv};
    \end{axis}
\end{tikzpicture}

### 2023-01-19 paper/fig_budget_binomial.tex
update paper

\begin{tikzpicture}
    \pgfplotsset{
        scale only axis,
        height=6cm,
        xlabel={total number of daily new hospitalizations},
        ylabel={total lost contacts}
    }
    \begin{axis}[
    legend entries={{$\text{PCN}_\text{ARH}$ (Binomial)}, {Budget 2}, {Budget 3}, {Budget 4}, {Budget 5}, {Budget None}},
    width=9.5cm
    ]
    \addplot[only marks,draw opacity=0.5,mark=x,red] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/cs_binomial.csv};
    \addplot[only marks,draw opacity=0.5,mark=o,blue] table [x=o_1, y=o_5, col sep=comma] {data/cs/budget/binomial/budget_2.csv};
    \addplot[only marks,draw opacity=0.5,mark=square,pink] table [x=o_1, y=o_5, col sep=comma] {data/cs/budget/binomial/budget_3.csv};
    \addplot[only marks,draw opacity=0.5,mark=triangle,orange] table [x=o_1, y=o_5, col sep=comma] {data/cs/budget/binomial/budget_4.csv};
    \addplot[only marks,draw opacity=0.5,mark=diamond,cyan] table [x=o_1, y=o_5, col sep=comma] {data/cs/budget/binomial/budget_5.csv};
    \addplot[only marks,mark=+] table [x=o_1, y=o_5, col sep=comma] {data/cs/budget/binomial/budget_None.csv};
    \end{axis}
\end{tikzpicture}

### 2023-01-19 paper/main.tex
update paper

@@
 \usepackage{booktabs}       % professional-quality tables
-\usepackage{amsmath}
+\usepackage{amsmath,bm}
 \usepackage{amsfonts}       % blackboard math symbols
@@
 \newcommand{\foi}{\lambda}
+\newcommand\added[1]{\textcolor{blue}{#1}}
 
@@
 
-\paragraph{State-space:} The state of the MOMDP consists of the state variables that make up the epidemiological model (see Figure~\ref{fig:ode_mdel}), to which we refer as $\mdpstate_m$, together with the social contact matrix $\hat{C}$ that is currently in place.
+\paragraph{Action-space:} Our actions concern the installment of a social contact matrix $\hat{C}$, with a particular reduction (see Section~\ref{sec:interventions}).
+To this end, we use the proportional reduction parameters $p_w, p_s, p_l$ defined in Section~\ref{sec:interventions}.
+Thus, each $\mdpaction \in \mdpactionspace$ is a 3-dimensional continuous vector in $[0,1]^3$ (i.e., $\mdpaction = [p_{w}, p_{s}, p_{l}]$) which impacts the SCM according to Equation~\ref{eq:scm}.
+
+\paragraph{Transition function:} The model defined by \citet{abrams2021modelling} utilises a model transition probability $M(\mdpstate_m' \mid \mdpstate_m, \hat{C})$ (see SI for details on $M$), where $\hat{C}$ the currently installed SCM, that progresses the epidemiological model in one timestep.
+We use this function as the transition function in \momdpname.
+For each timestep $t$, we simulate the model for one week, using $\hat{C}$ obtained from $\mdpaction_t$.
+
+As mentioned above and detailed in the SI, the compartment model is described by a set of ODEs.
+To obtain a stochastic version, the ODEs can be formulated as a chain-binomial process.
+Based on these cases, we also create and experiment on two versions of the \momdpname\ environment: the (deterministic) \emph{ODE model} (i.e., a MOMDP with a deterministic transition function) and the (stochastic) \emph{Binomial model} (i.e., a MOMDP with a stochastic transition function).
+
+\added{In a classical MDP, executing an action $\action_t$ in any state $\state_t$ leads to a next state $\state_{t+1}$ according to the transition function $\mdptransition$.
+At every timestep, the policy is free to choose the action to perform.
+In our case, this corresponds to a different restriction $[p_{w}, p_{s}, p_{l}]$ every week.
+However, imposing new social restrictions every week does not seem very realistic.
+It can be frustrating to having to cope with the ever-changing rules.
+In case of the exit strategy after the lockdown of the first COVID wave in Belgium, the Belgian government executed a progressive plan that consisted of 3 different stages.
+Each stage resulted in the progressive lessening of social restrictions.}
+
+\added{In order to simulate a more realistic multi-stage exit strategy, we incorporate a \emph{budget per restriction-change} on the MOMDP.
+Concretely, if the social restriction proposed by the policy is different from the current one in place, we consider this a next stage and reduce our restriction-change budget by one.
+When the budget reaches 0, we consider the exit strategy to be final, and impose the current social restriction for the remainder of the episode.
+Finally, we can simulate a no-limit budget setting by setting the budget to the length of the episode.}
+
+\paragraph{State-space:} The state of the MOMDP \added{is a tuple that consists of 3 elements.
+The first element, $\mdpstate_m$, are the state-variables that make up the epidemiological model (see Figure~\ref{fig:ode_mdel}).
+They are required to simulate the underlying epidemiological model for a full week.}
+
 $\mdpstate_m$ directly corresponds to the aggregation of the state variables in the epidemiological model, i.e., a tuple,
@@
 
+We parameterize the \added{epidemiological} model using the mean of the posteriors as reported by \citet{abrams2021modelling} (details in SI).
+
+\added{The second element of the tuple consists of the social contact matrix $\hat{C}$ that is currently in place.
+The reason to incorporate it in the state-space is two-fold.
+First, \citet{abrams2021modelling} define a compliance function, simulating the time people need to get used to the new rules set in place.
+As such, during the simulated week, there is a gradual shift from the current $\hat{C}$ to the next one.
+We thus require to know the current $\hat{C}$.}
+
+\added{Secondly, we require the current $\hat{C}$ to determine if we enter a new stage of the multi-stage exit strategy -- and thus consume part of the restriction-change budget -- or not.}
+
+\added{The third and last element of the tuple consists of the restriction-change budget $b$.
+We note that we incorporate a separate budget per action-dimension, so $p_{w}, p_{s}$ and $p_{l}$ each have their own budget, $b = [b_w, b_s, b_l]$.
+It is thus possible that, at timestep $t$, the budget for one of the dimensions is reduced but not the others.}
+
 Therefore, we define a state in \momdpname\ as follows:
 \begin{equation}
-    \mdpstate = \mdpstate_m \cup \hat{C}
-\end{equation}
-
-We paramaterise the model using the mean of the posteriors as reported by \citet{abrams2021modelling} (details in SI).
-
-\paragraph{Action-space:} Our actions concern the installment of a social contact matrix $\hat{C}$, with a particular reduction (see Section~\ref{sec:interventions}).
-To this end, we use the proportional reduction parameters $p_w, p_s, p_l$ defined in Section~\ref{sec:interventions}.
-Thus, each $\mdpaction \in \mdpactionspace$ is a 3-dimensional continuous vector in $[0,1]^3$ (i.e., $\mdpaction = [p_{w}, p_{s}, p_{l}]$) which impacts the SCM according to Equation~\ref{eq:scm}.
-
-\paragraph{Transition function:} The model defined by \citet{abrams2021modelling} utilises a model transition probability $M(\mdpstate_m' \mid \mdpstate_m, \hat{C})$ (see SI for details on $M$), where $\hat{C}$ the currently installed SCM, that progresses the epidemiological model in one timestep.
-We use this function as the transition function in \momdpname.
-For each timestep $t$, we simulate the model for one week, using $\hat{C}$ obtained from $\mdpaction_t$.
-
-As mentioned above and detailed in the SI, the compartment model is described by a set of ODEs.
-To obtain a stochastic version, the ODEs can be formulated as a chain-binomial process.
-Based on these cases, we also create and experiment on two versions of the \momdpname\ environment: the (deterministic) \emph{ODE model} (i.e., a MOMDP with a deterministic transition function) and the (stochastic) \emph{Binomial model} (i.e., a MOMDP with a stochastic transition function).
+    \mdpstate = \mdpstate_m \cup \hat{C} \cup b
+\end{equation}
 
@@
 Thus, recovering them all is an expensive process and requires an exhaustive exploration of the complete state space.
-To address this problem, we use Pareto Conditioned Networks~(PCN), a method that uses a single neural network to encompass all non-dominated
-policies \cite{reymond2022pcn}.
+To address this problem, we use Pareto Conditioned Networks~(PCN), a method that uses a single neural network to encompass all non-dominated policies \cite{reymond2022pcn}.
 The key idea behind PCN is to use supervised learning techniques to improve the policy instead of resorting to temporal-difference learning.
-This eliminates the moving-target problem \cite{schmidhuber2019}, resulting in stable learning algorithm.
+% This eliminates the moving-target problem \cite{schmidhuber2019}, resulting in stable learning algorithm.
+
+\added{TODO REPHRASE, FROM PCN PAPER: Using neural networks as function approximators in RL comes with many challenges.
+One of them is that the target (e.g., the optimal action of the policy) is not known in advance --- as opposed to classical supervised learning where the ground-truth target is provided.
+As the behavior of the agent improves over time, the action used as target can change, often leading to hard-to-tune and brittle learners \cite{mnih2015,fu2019}.}
+
+\added{Instead of trying to continuously improve the policy by learning actions that should lead to the highest cumulative reward, PCN flips the problem, by learning actions that should lead to any \emph{desired} cumulative reward (be it high or low).
+In this way, all past trajectories can be reused for supervision, since their returns are known, as well as the actions needed to reach said returns.
+We can thus train a policy that, conditioned on a desired return, provides the optimal action to reach said return.
+By leveraging the generalization properties of neural networks, we can accumulate incrementally better experience by conditioning on increasingly higher reward-goals}
 
@@
 However, contrary to classification, the data present in the dataset is not fixed.
-Each time PCN executes a trajectory it creates a tuple $x=(\mdpstate_t, \momdpreturn_t, T-t)$, $y=\action_t$ for every transition of the trajectory and adds it to the dataset.
-
-PCN collects new data for a fixed number of episodes, after which it re-trains the network with batch updates from the newly assembled dataset.
-This improves the policies induced by the network, which in turn enables the agent to gather superior data for the next training batch.
-
-\subsection{PCN for continuous actions}
+\added{PCN collects data from the trajectories experienced while exploring the environment.
+Thus, the dataset improves over time, as we collect better and better trajectories.}
+
+\added{Concretely, the PCN algorithm is a loop that repeats 3 main steps:
+\begin{enumerate}
+    \item Using the current policy to interact with the environment and generate trajectories
+    \item Update and prune the dataset to incorporate these new trajectories and remove the least interesting trajectories in terms of returns
+    \item Update the policy by training it on the updated dataset
+\end{enumerate}
+}
+
+\subsection{Building and updating the dataset}
+\label{sec:pcn-building-dataset}
+
+\added{Given a trajectory composed of $T$ transitions $(s_0, a_0, \bm{r}_0, s_1), \dots, (s_{T-1}, a_{T-1}, \bm{r}_{T-1}, s_T)$ we can compute, for each transition at timestep $t$, the future discounted return $\bm{R}_t = \sum_{i=t}^T \gamma^i\bm{r}_i$ and the leftover horizon $h_t = T-t$.
+Since for this trajectory executing action $a_t$ in state $s_t$ resulted in return $\mathbf{R}_t$ in $h_t$ timesteps, we add a datapoint with input $\langle s, \hat{h}, \mathbf{\hat{R}}\rangle = \langle s_t, h_t, \mathbf{R}_t \rangle$ and output $a=a_t$ to the dataset.
+In other words, when the observed return corresponds to the desired return in that state, then $a_t$ is the optimal action to take.}
+
+\added{Since we have a fix-sized dataset, adding this trajectory means we need to remove another one.
+Our aim is to keep trajectories that span different parts of the objective space, such that we can explore as many types of trade-offs as possible.
+However, to improve the trade-offs we currently have in our dataset, we need to only keep solutions that are close to our current estimate of the Pareto front.
+Thus, we assign a priority for each trajectory in our dataset that combine two metrics: the \emph{crowding distance}, which measures the distance of a solution with its closest neighbors, and the \emph{l2-norm} with the closest non-dominated solution, which indicates how close the solution is from our current estimate of the Pareto front.}
+
+\added{The dataset is then updated by only keeping the trajectories with the top-$N$ priorities.}
+
+\subsection{Generating trajectories with the current policy}
+\label{sec:pcn-policy-exploration}
+
+\added{In the previous section, we explain how PCN updates the dataset given a trajectory. In this section, we explain how PCN produces said trajectory.}
+
+\added{It is unrealistic to expect PCN to reliably produce trajectories with high-valued desired returns when it has only been trained on datapoints originating from random trajectories.
+Rather, we can expect PCN to produce trajectories with returns in the range of the ones from the current training data.
+Therefore, if we obtain trajectories reaching high returns, PCN will be able to confidently return high-return policies.}
+
+\added{PCN leverages the fact that, due to the generalization capabilities of neural networks, the policies obtained from the network will still be reliable even if the desired return is marginally higher than what is present in the training data.
+In fact, they will perform similar actions to those in the training data, but lead to a higher return.
+Thus, we incrementally condition the network on better and better returns, in order to obtain trajectories that extend the boundaries of PCN's current coverage set.}
+
+\added{More precisely, we randomly select a non-dominated return $\mathbf{R}_{nd}$ and its corresponding horizon $\hat{h}$ from the dataset.
+By randomly picking a non-dominated return from the entire coverage set we ensure equal chance of improvement to each part of the objective space.
+To generate trajectories that improve upon $\mathbf{R}_{nd}$, we randomly select an objective $o$ and increase $R_{nd,o}$ by a sample from the uniform distribution $U(0, \sigma_o)$, where $\sigma_o$ is the standard deviation for the selected objective, using all non-dominated returns from the trajectories in the dataset.
+This then becomes our desired return $\bm{\hat{R}}$.
+By restricting the improvement to at most $\sigma_o$, $\mathbf{\hat{R}}$ stays in the range of possible achievable returns and, by only modifying one objective at a time, the changes to the network's input compared to the training data are kept at a minimum.}
+
+\subsection{Training the network for continuous actions}
 \label{sec:pcn-continuous}
@@
     \centering
+    \input{fig_budget}
+    \caption{The Pareto front of policies discovered by PCN, showing the different compromises between the number of hospitalizations and the number of lost contacts.
+Although $\text{PCN}_\text{ARI}$ was trained on the number of infections, it still shows a competitive coverage set w.r.t.
+hospitalizations.}
+    \label{fig:pareto-front-arh}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \input{fig_budget_binomial}
+    \caption{The Pareto front of policies discovered by PCN, showing the different compromises between the number of hospitalizations and the number of lost contacts.
+Although $\text{PCN}_\text{ARI}$ was trained on the number of infections, it still shows a competitive coverage set w.r.t.
+hospitalizations.}
+    \label{fig:pareto-front-arh}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
     \input{fig_arh}
@@
     \label{fig:pareto-front-arh}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+\begin{tikzpicture}
+    \pgfplotsset{
+        scale only axis,
+        height=6cm,
+        xlabel={total number of daily new hospitalizations},
+        ylabel={total lost contacts}
+    }
+    \begin{axis}[
+    legend entries={{Budget 2 (ODE)}, {Budget 2 (Binomial)}, {Fixed}},
+    width=9.5cm
+    ]
+    \addplot[only marks,draw opacity=0.5,mark=x,red] table [x=o_1, y=o_5, col sep=comma] {data/cs/budget/budget_2.csv};
+    \addplot[only marks,draw opacity=0.5,mark=o,blue] table [x=o_1, y=o_5, col sep=comma] {data/cs/budget/binomial/budget_2.csv};
+    \addplot[only marks,mark=+] table [x=o_0, y=o_1, col sep=comma] {data/cs/cs_fixed.csv};
+    \end{axis}
+\end{tikzpicture}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+\begin{tikzpicture}
+    \pgfplotsset{
+        scale only axis,
+        height=6cm,
+        xlabel={total number of daily new hospitalizations},
+        ylabel={total lost contacts}
+    }
+    \begin{axis}[
+    legend entries={{Budget 3 (ODE)}, {Budget 3 (Binomial)}, {Fixed}},
+    width=9.5cm
+    ]
+    \addplot[only marks,draw opacity=0.5,mark=x,red] table [x=o_1, y=o_5, col sep=comma] {data/cs/budget/budget_3.csv};
+    \addplot[only marks,draw opacity=0.5,mark=o,blue] table [x=o_1, y=o_5, col sep=comma] {data/cs/budget/binomial/budget_3.csv};
+    \addplot[only marks,mark=+] table [x=o_0, y=o_1, col sep=comma] {data/cs/cs_fixed.csv};
+    \end{axis}
+\end{tikzpicture}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+\begin{tikzpicture}
+    \pgfplotsset{
+        scale only axis,
+        height=6cm,
+        xlabel={total number of daily new hospitalizations},
+        ylabel={total lost contacts}
+    }
+    \begin{axis}[
+    legend entries={{Budget 4 (ODE)}, {Budget 4 (Binomial)}, {Fixed}},
+    width=9.5cm
+    ]
+    \addplot[only marks,draw opacity=0.5,mark=x,red] table [x=o_1, y=o_5, col sep=comma] {data/cs/budget/budget_4.csv};
+    \addplot[only marks,draw opacity=0.5,mark=o,blue] table [x=o_1, y=o_5, col sep=comma] {data/cs/budget/binomial/budget_4.csv};
+    \addplot[only marks,mark=+] table [x=o_0, y=o_1, col sep=comma] {data/cs/cs_fixed.csv};
+    \end{axis}
+\end{tikzpicture}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+\begin{tikzpicture}
+    \pgfplotsset{
+        scale only axis,
+        height=6cm,
+        xlabel={total number of daily new hospitalizations},
+        ylabel={total lost contacts}
+    }
+    \begin{axis}[
+    legend entries={{Budget 5 (ODE)}, {Budget 5 (Binomial)}, {Fixed}},
+    width=9.5cm
+    ]
+    \addplot[only marks,draw opacity=0.5,mark=x,red] table [x=o_1, y=o_5, col sep=comma] {data/cs/budget/budget_5.csv};
+    \addplot[only marks,draw opacity=0.5,mark=o,blue] table [x=o_1, y=o_5, col sep=comma] {data/cs/budget/binomial/budget_5.csv};
+    \addplot[only marks,mark=+] table [x=o_0, y=o_1, col sep=comma] {data/cs/cs_fixed.csv};
+    \end{axis}
+\end{tikzpicture}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+\begin{tikzpicture}
+    \pgfplotsset{
+        scale only axis,
+        height=6cm,
+        xlabel={total number of daily new hospitalizations},
+        ylabel={total lost contacts}
+    }
+    \begin{axis}[
+    legend entries={{Budget N (ODE)}, {Budget N (Binomial)}, {Fixed}},
+    width=9.5cm
+    ]
+    \addplot[only marks,draw opacity=0.5,mark=x,red] table [x=o_1, y=o_5, col sep=comma] {data/cs/budget/budget_None.csv};
+    \addplot[only marks,draw opacity=0.5,mark=o,blue] table [x=o_1, y=o_5, col sep=comma] {data/cs/budget/binomial/budget_None.csv};
+    \addplot[only marks,mark=+] table [x=o_0, y=o_1, col sep=comma] {data/cs/cs_fixed.csv};
+    \end{axis}
+\end{tikzpicture}
 \end{figure}

### 2023-01-19 distil_policy.py

@@ -1,0 +1,520 @@
+import pickle
+import pathlib
+
+from numpy.random import beta
+import torch
+from torch.utils.data import TensorDataset, DataLoader
+from dataclasses import dataclass
+from SoftDecisionTree.sdt.model import SoftDecisionTree
+import wandb
+import os
+import time
+import json
+import matplotlib.pyplot as plt
+import numpy as np
+from copy import deepcopy
+from plotly.subplots import make_subplots
+import plotly.graph_objects as go
+
+# with open('/tmp/tables.pkl', 'rb') as f:
+#     tables = pickle.load(f)
+# with open('/tmp/configs.pkl', 'rb') as f:
+#     configs = pickle.load(f)
+
+batch_size = 32
+epochs = 5000
+
+BUDGET_IDX = 0
+STATE_IDX = 1
+EVENT_IDX = 2
+
+@dataclass
+class Transition(object):
+    observation: np.ndarray
+    action: int
+    reward: float
+    next_observation: np.ndarray
+    terminal: bool
+
+@dataclass
+class AugmentedTransition(object):
+    observation: np.ndarray
+    action: int
+    reward: float
+    next_observation: np.ndarray
+    terminal: bool
+    t_r: np.ndarray
+    t_h: int
+
+
+@dataclass
+class SDTARGS(object):
+    depth : int
+    lmbda : float
+    input_dim : int
+    output_dim: int
+    log_interval: int
+    save : str
+    tensorboard : bool
+    device: str
+
+
+def download_policy_logs(run_id):
+    path = f'mreymond/pcn-covid/runs/{run_id}'
+    api = wandb.Api(timeout=30)
+    run = api.run(path=path)
+    # get artifacts from run
+    run_dict = {}
+    for name in ('coverage_set', 'nd_coverage_set', 'executions', 'execution-transitions'):
+        artifact_path = f'mreymond/{run.project}/run-{run.id}-{name}:latest'
+        print(f'getting: {artifact_path}')
+        # download artifact
+        try:
+            file_path = api.artifact(artifact_path).file(f'/tmp')
+            if file_path[-4:] == '.pkl':
+                # load the downloaded file
+                with open(file_path, 'rb') as f:
+                    table = pickle.load(f)
+            else:
+                # load the downloaded file
+                with open(file_path, 'r') as f:
+                    table = json.load(f)
+            os.remove(file_path)
+            run_dict[name] = table
+            time.sleep(3)
+        except Exception as e:
+            print(f'skipping... {e}')
[…]

### 2023-01-19 eval_pcn.py

@@ -16,7 +16,37 @@ def run_episode(env, model, desired_return, desired_horizon, max_return):
     done = False
     while not done:
         action = choose_action(model, obs, desired_return, desired_horizon, eval=True)
-        n_obs, reward, done, _ = env.step(action)
+        n_obs, reward, done, info = env.step(action)
+        if 'action' in info:
+            action = info['action']
+
+        transitions.append(Transition(
+            observation=obs[1],
+            action=env.action(action),
+            reward=np.float32(reward).copy(),
+            next_observation=n_obs[1],
+            terminal=done
+        ))
+
+        obs = n_obs
+        # clip desired return, to return-upper-bound, 
+        # to avoid negative returns giving impossible desired returns
+        # reward = np.array((reward[1], reward[2]))
+        desired_return = np.clip(desired_return-reward, None, max_return, dtype=np.float32)
+        # clip desired horizon to avoid negative horizons
+        desired_horizon = np.float32(max(desired_horizon-1, 1.))
+    return transitions
+
+
+def run_fixed_episode(env, actions, desired_return, desired_horizon, max_return):
+    transitions = []
+    obs = env.reset()
+    done = False; step = 0
+    while not done:
+        action = actions[step]
+        n_obs, reward, done, info = env.step(action)
+        if 'action' in info:
+            action = info['action']
 
         transitions.append(Transition(
             observation=obs[0],
@@ -27,6 +57,7 @@ def run_episode(env, model, desired_return, desired_horizon, max_return):
         ))
 
         obs = n_obs
+        step += 1
         # clip desired return, to return-upper-bound, 
         # to avoid negative returns giving impossible desired returns
         # reward = np.array((reward[1], reward[2]))
@@ -59,12 +90,13 @@ def plot_episode(transitions, alpha=1.):
     ax.plot(i_hosp_new, alpha=alpha, label='hosp', color='blue')
     ax.plot(i_icu_new,  alpha=alpha, label='icu', color='green')
     ax.plot(i_hosp_new+i_icu_new, label='hosp+icu',  alpha=alpha, color='orange')
+    ax.plot(d_new, alpha=alpha, label='deaths', color='red')
[…]

### 2023-01-19 experiment_plots.py

@@ -12,6 +12,17 @@ import pandas as pd
 pre = sys.argv[1]
 
 
+def baseline_runs(logdir):
+    with open(Path(logdir) / 'returns.pkl', 'rb') as f:
+        import pickle
+        pf = pickle.load(f)
+        pf = np.mean(pf, axis=1)
+        pf = non_dominated(pf)
+    return [{
+        'pareto_front': pf,
+        'hypervolume': []
+    }]
+
 def udrl_runs(logdir):
     runs = []
     for i, path in enumerate(Path(logdir).rglob('log.h5')):
@@ -19,6 +30,7 @@ def udrl_runs(logdir):
         with h5py.File(path, 'r') as logfile:
             pf = logfile['train/leaves/ndarray']
             try:
+                raise KeyError()
                 hv = logfile['train/hypervolume']
             except KeyError:
                 # recompute hypervolume
@@ -33,7 +45,8 @@ def udrl_runs(logdir):
                     'walkroom8':np.array([-10.0]*8),
                     'walkroom9':np.array([-10.0]*9),
                     'ode':np.array([-50000, -2000.0])/np.array([10000, 100.]),
-                    'binomial':np.array([-50000, -2000.0])/np.array([10000, 100.])
+                    'binomial':np.array([-50000, -2000.0])/np.array([10000, 100.]),
+                    # 'binomial':np.array([-50000, -2000.0, -2000.0, -2000.0])/np.array([10000, 50., 50, 50])
                 }
                 for rp in ref_points.keys():
                     if rp in logdir:
@@ -49,11 +62,11 @@ def udrl_runs(logdir):
                 steps = logfile['train/leaves/step']
                 hv = np.stack((steps, hv), axis=1)
             # best points found over training
-            bp = non_dominated(np.concatenate(pf, axis=0))
+            # bp = non_dominated(np.concatenate(pf, axis=0))
             run = {
                 'pareto_front': pf[-1],
                 'hypervolume': hv[:],
-                'best_points': bp
+                # 'best_points': bp
             }
             runs.append(run)
     return runs
@@ -103,18 +116,23 @@ def show_pareto_front(all_runs):
             plt.gca().scatter(*[pf[:,j] for j in range(pf.shape[1])], c=weights, vmin=-1, vmax=len(runs)+1, cmap=cmaps[k])
     plt.show()
 
-
+from itertools import combinations
[…]

### 2023-01-31 paper/fig/arh.tex
plot for different budgets

\begin{tikzpicture}
\pgfplotsset{
    scale only axis,
    % height=6cm,
    xlabel={total daily new hospitalizations},
    ylabel={total lost contacts},
    width=0.45\textwidth
}
\begin{groupplot}[
    group style={
        group size=2 by 1,
        y descriptions at=edge left,
        vertical sep=0pt,
        horizontal sep=0pt,
    },
    legend entries={{$\text{PCN}_{b=2}$}, {$\text{PCN}_{b=3}$}, {$\text{PCN}_{b=4}$}, {$\text{PCN}_{b=5}$}, {$\text{PCN}$}, {Fixed}},
]
\nextgroupplot[mark size=1pt,draw opacity=1]
        \addplot[only marks,mark=o,pc1] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/ode/budget_2.csv};
        \addplot[only marks,mark=o,pc2] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/ode/budget_3.csv};
        \addplot[only marks,mark=o,pc3] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/ode/budget_4.csv};
        \addplot[only marks,mark=o,pc4] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/ode/budget_5.csv};
        \addplot[only marks,mark=o,pc5] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/ode/budget_None.csv};
        \addplot[only marks,mark=+] table [x=o_0, y=o_1, col sep=comma] {data/cs/cs_fixed.csv};
\nextgroupplot[mark size=1pt]
        \addplot[only marks,draw opacity=1,fill opacity=0.5,mark=o,pc1] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/binomial/budget_2.csv};
        \addplot[only marks,draw opacity=1,fill opacity=0.5,mark=o,pc2] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/binomial/budget_3.csv};
        \addplot[only marks,draw opacity=1,fill opacity=0.5,mark=o,pc3] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/binomial/budget_4.csv};
        \addplot[only marks,draw opacity=1,fill opacity=0.5,mark=o,pc4] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/binomial/budget_5.csv};
        \addplot[only marks,draw opacity=1,fill opacity=0.5,mark=o,pc5] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/binomial/budget_None.csv};
        \addplot[only marks,mark=+] table [x=o_0, y=o_1, col sep=comma] {data/cs/cs_fixed.csv};

\end{groupplot}
\end{tikzpicture}

### 2023-01-31 paper/main.tex
plot for different budgets
remove previous plot

@@
 \usepackage{pgfplots}
+\usepackage{comment}
 \usepgfplotslibrary{dateplot, groupplots}
@@
 \newcommand{\agegroups}{K}
+
+% color scheme for plots: plot-color-i
+\definecolor{pc1}{RGB}{76, 120, 168}
+\definecolor{pc2}{RGB}{245, 133, 24}
+\definecolor{pc3}{RGB}{228, 87, 88}
+\definecolor{pc4}{RGB}{114, 183, 178}
+\definecolor{pc5}{RGB}{84, 162, 75}
+\definecolor{pc6}{RGB}{238, 202, 59}
+\definecolor{pc7}{RGB}{178, 121, 162}
 
@@
     \centering
-    \begin{tikzpicture}
-        \pgfplotsset{
-          scale only axis,
-          height=6cm,
-          xlabel={Cumulative number of daily new hospitalizations},
-          ylabel={Cumulative lost contacts}
-        }
-      \begin{axis}[
-        legend entries={{$\text{PCN}_\text{ARH}$ (ODE)}, {$\text{PCN}_\text{ARI}$ (ODE)}, {Fixed}},
-        width=9.5cm
-      ]
-      %\addplot[only marks,draw opacity=0.5,mark=x,green] table [x=o_0, y=o_1, col sep=comma] {data/cs/cs_ode.csv};
-      \addplot[only marks,draw opacity=0.5,mark=x,red] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/cs_ode.csv};
-      \addplot[only marks,draw opacity=0.5,mark=o,blue] table [x=o_1, y=o_5, col sep=comma] {data/cs/ari/cs_ode.csv};
-      %\addplot[only marks,mark=o,yellow] table [x=o_0, y=o_1, col sep=comma] {data/cs/cs_binomial.csv};
-      %\addplot[only marks,draw opacity=0.5,mark=o,blue] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/cs_binomial.csv};
-      %\addplot[only marks,draw opacity=0.5,mark=diamond,cyan] table [x=o_1, y=o_5, col sep=comma] {data/cs/ari/cs_binomial.csv};
-      \addplot[only marks,mark=+] table [x=o_0, y=o_1, col sep=comma] {data/cs/cs_fixed.csv};
-      \end{axis}
-    \end{tikzpicture}
+    \input{fig/arh.tex}
     \caption{The Pareto front of policies discovered by PCN using the ODE model (Binomial omitted for clarity, see SI), showing the different compromises between the number of hospitalizations and the number of lost contacts.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+\begin{comment}
 \appendix
@@
 RL has also been used to learn effective COVID-19 vaccination distribution strategies \cite{beigi2021application,awasthi2020vacsim} and to create decision support systems for mitigating the spread of COVID-19.
-
+\end{comment}
 \end{document}

### 2023-01-31 paper/fig/all_budgets.tex
replace policy-execution plots with budget plots

\begin{tikzpicture}
    \pgfplotsset{
          scale only axis,
          date coordinates in=x,
          xticklabel=\day/\month,
          height=2cm,
        }
        \begin{groupplot}[
        group style={
                % set how the plots should be organized
                group size=1 by 5,
                % only show ticklabels and axis labels on the bottom
                y descriptions at=edge left,
                x descriptions at=edge bottom,
                % set the `vertical sep' to zero
                horizontal sep=0pt,
                vertical sep=0pt
            },
          width=9.5cm,
          axis y line*=left,
          xlabel={time},
          ylabel={people},
        ]
        % GROUP WITH LEFT Y-AXIS: PEOPLE (HOSP, ICU, DEATHS)
        \nextgroupplot[scaled y ticks=false]
        \addplot[mark=none, cyan] table [x=dates, y=i_hosp_new, col sep=comma] {data/transitions/arh/binomial/budget_2/policies-transitions/72/run_0.csv};
        \label{i_hosp_new}
        \addplot[mark=none, orange] table [x=dates, y=i_icu_new, col sep=comma] {data/transitions/arh/binomial/budget_2/policies-transitions/72/run_0.csv};
        \label{i_icu_new}
        \addplot[mark=none, red] table [x=dates, y=d_new, col sep=comma] {data/transitions/arh/binomial/budget_2/policies-transitions/72/run_0.csv};
        \label{d_new}
        \nextgroupplot[scaled y ticks=false]
        \addplot[mark=none, cyan] table [x=dates, y=i_hosp_new, col sep=comma] {data/transitions/arh/binomial/budget_3/policies-transitions/36/run_0.csv};
        \addplot[mark=none, orange] table [x=dates, y=i_icu_new, col sep=comma] {data/transitions/arh/binomial/budget_3/policies-transitions/36/run_0.csv};
        \addplot[mark=none, red] table [x=dates, y=d_new, col sep=comma] {data/transitions/arh/binomial/budget_3/policies-transitions/36/run_0.csv};
        \nextgroupplot[scaled y ticks=false]
        \addplot[mark=none, cyan] table [x=dates, y=i_hosp_new, col sep=comma] {data/transitions/arh/binomial/budget_4/policies-transitions/25/run_0.csv};
        \addplot[mark=none, orange] table [x=dates, y=i_icu_new, col sep=comma] {data/transitions/arh/binomial/budget_4/policies-transitions/25/run_0.csv};
        \addplot[mark=none, red] table [x=dates, y=d_new, col sep=comma] {data/transitions/arh/binomial/budget_4/policies-transitions/25/run_0.csv};
        \nextgroupplot[scaled y ticks=false]
        \addplot[mark=none, cyan] table [x=dates, y=i_hosp_new, col sep=comma] {data/transitions/arh/binomial/budget_5/policies-transitions/41/run_0.csv};
        \addplot[mark=none, orange] table [x=dates, y=i_icu_new, col sep=comma] {data/transitions/arh/binomial/budget_5/policies-transitions/41/run_0.csv};
        \addplot[mark=none, red] table [x=dates, y=d_new, col sep=comma] {data/transitions/arh/binomial/budget_5/policies-transitions/41/run_0.csv};
        \nextgroupplot[scaled y ticks=false]
        \addplot[mark=none, cyan] table [x=dates, y=i_hosp_new, col sep=comma] {data/transitions/arh/binomial/budget_None/policies-transitions/48/run_0.csv};
        \addplot[mark=none, orange] table [x=dates, y=i_icu_new, col sep=comma] {data/transitions/arh/binomial/budget_None/policies-transitions/48/run_0.csv};
        \addplot[mark=none, red] table [x=dates, y=d_new, col sep=comma] {data/transitions/arh/binomial/budget_None/policies-transitions/48/run_0.csv};
        \end{groupplot}
        % GROUP WITH RIGHT Y-AXIS: ACTIONS (P_W, P_S, P_L)
        \begin{groupplot}[
        group style={
                % set how the plots should be organized
                group size=1 by 5,
                % only show ticklabels and axis labels on the bottom
                y descriptions at=edge right,
                x descriptions at=edge bottom,
                % set the `vertical sep' to zero
                horizontal sep=0pt,
                vertical sep=0pt
            },
          width=9.5cm,
          axis y line*=right,
          axis x line=none,
          ymax=1.1,
          ylabel={proportion},
          legend style={at={(0.5,1)},anchor=south},
        %   legend pos=outer north east,
          legend columns=3, 
          legend cell align=left,
        ]
        \nextgroupplot[scaled y ticks=false]
        \addplot[mark=none, dashed, blue] table [x=dates, y=p_w, col sep=comma] {data/transitions/arh/binomial/budget_2/policies-transitions/72/run_0.csv};
        \label{p_w}
        \addplot[mark=none, dashed, magenta] table [x=dates, y=p_s, col sep=comma] {data/transitions/arh/binomial/budget_2/policies-transitions/72/run_0.csv};
        \label{p_s}
        \addplot[mark=none, dashed, darkgray] table [x=dates, y=p_l, col sep=comma] {data/transitions/arh/binomial/budget_2/policies-transitions/72/run_0.csv};
        \label{p_l}
        \addlegendimage{/pgfplots/refstyle=i_hosp_new}\addlegendentry{$p_w$}
        \addlegendimage{/pgfplots/refstyle=i_icu_new}\addlegendentry{$p_l$}
        \addlegendimage{/pgfplots/refstyle=d_new}\addlegendentry{$p_s$}
        \addlegendimage{/pgfplots/refstyle=p_w}\addlegendentry{{daily new hosp}}
        \addlegendimage{/pgfplots/refstyle=p_s}\addlegendentry{{daily new ICU}}
        \addlegendimage{/pgfplots/refstyle=p_l}\addlegendentry{{daily deaths}}
        \nextgroupplot[scaled y ticks=false]
        \addplot[mark=none, dashed, blue] table [x=dates, y=p_w, col sep=comma] {data/transitions/arh/binomial/budget_3/policies-transitions/36/run_0.csv};
        \addplot[mark=none, dashed, magenta] table [x=dates, y=p_s, col sep=comma] {data/transitions/arh/binomial/budget_3/policies-transitions/36/run_0.csv};
        \addplot[mark=none, dashed, darkgray] table [x=dates, y=p_l, col sep=comma] {data/transitions/arh/binomial/budget_3/policies-transitions/36/run_0.csv};
        \nextgroupplot[scaled y ticks=false]
        \addplot[mark=none, dashed, blue] table [x=dates, y=p_w, col sep=comma] {data/transitions/arh/binomial/budget_4/policies-transitions/25/run_0.csv};
        \addplot[mark=none, dashed, magenta] table [x=dates, y=p_s, col sep=comma] {data/transitions/arh/binomial/budget_4/policies-transitions/25/run_0.csv};
        \addplot[mark=none, dashed, darkgray] table [x=dates, y=p_l, col sep=comma] {data/transitions/arh/binomial/budget_4/policies-transitions/25/run_0.csv};
        \nextgroupplot[scaled y ticks=false]
        \addplot[mark=none, dashed, blue] table [x=dates, y=p_w, col sep=comma] {data/transitions/arh/binomial/budget_5/policies-transitions/41/run_0.csv};
        \addplot[mark=none, dashed, magenta] table [x=dates, y=p_s, col sep=comma] {data/transitions/arh/binomial/budget_5/policies-transitions/41/run_0.csv};
        \addplot[mark=none, dashed, darkgray] table [x=dates, y=p_l, col sep=comma] {data/transitions/arh/binomial/budget_5/policies-transitions/41/run_0.csv};
        \nextgroupplot[scaled y ticks=false]
        \addplot[mark=none, dashed, blue] table [x=dates, y=p_w, col sep=comma] {data/transitions/arh/binomial/budget_None/policies-transitions/48/run_0.csv};
        \addplot[mark=none, dashed, magenta] table [x=dates, y=p_s, col sep=comma] {data/transitions/arh/binomial/budget_None/policies-transitions/48/run_0.csv};
        \addplot[mark=none, dashed, darkgray] table [x=dates, y=p_l, col sep=comma] {data/transitions/arh/binomial/budget_None/policies-transitions/48/run_0.csv};
        \end{groupplot}
      \end{tikzpicture}

### 2023-01-31 paper/fig/budget_5.tex
replace policy-execution plots with budget plots

\begin{tikzpicture}
    \pgfplotsset{
          scale only axis,
          date coordinates in=x,
          xticklabel=\day/\month,
          height=2cm,
        }
        \begin{groupplot}[
        group style={
                % set how the plots should be organized
                group size=1 by 3,
                % only show ticklabels and axis labels on the bottom
                y descriptions at=edge left,
                x descriptions at=edge bottom,
                % set the `vertical sep' to zero
                horizontal sep=0pt,
                vertical sep=0pt
            },
          width=9.5cm,
          axis y line*=left,
          xlabel={time},
          ylabel={people},
        ]
        % GROUP WITH LEFT Y-AXIS: PEOPLE (HOSP, ICU, DEATHS)
        \nextgroupplot[scaled y ticks=false]
        \addplot[mark=none, cyan] table [x=dates, y=i_hosp_new, col sep=comma] {data/transitions/arh/binomial/budget_5/policies-transitions/0/run_0.csv};
        \label{i_hosp_new}
        \addplot[mark=none, orange] table [x=dates, y=i_icu_new, col sep=comma] {data/transitions/arh/binomial/budget_5/policies-transitions//0/run_0.csv};
        \label{i_icu_new}
        \addplot[mark=none, red] table [x=dates, y=d_new, col sep=comma] {data/transitions/arh/binomial/budget_5/policies-transitions//0/run_0.csv};
        \label{d_new}
        \nextgroupplot[scaled y ticks=false]
        \addplot[mark=none, cyan] table [x=dates, y=i_hosp_new, col sep=comma] {data/transitions/arh/binomial/budget_5/policies-transitions/45/run_0.csv};
        \addplot[mark=none, orange] table [x=dates, y=i_icu_new, col sep=comma] {data/transitions/arh/binomial/budget_5/policies-transitions/45/run_0.csv};
        \addplot[mark=none, red] table [x=dates, y=d_new, col sep=comma] {data/transitions/arh/binomial/budget_5/policies-transitions/45/run_0.csv};
        \nextgroupplot[scaled y ticks=false]
        \addplot[mark=none, cyan] table [x=dates, y=i_hosp_new, col sep=comma] {data/transitions/arh/binomial/budget_5/policies-transitions/114/run_0.csv};
        \addplot[mark=none, orange] table [x=dates, y=i_icu_new, col sep=comma] {data/transitions/arh/binomial/budget_5/policies-transitions/114/run_0.csv};
        \addplot[mark=none, red] table [x=dates, y=d_new, col sep=comma] {data/transitions/arh/binomial/budget_5/policies-transitions/114/run_0.csv};
        \end{groupplot}
        % GROUP WITH RIGHT Y-AXIS: ACTIONS (P_W, P_S, P_L)
        \begin{groupplot}[
        group style={
                % set how the plots should be organized
                group size=1 by 3,
                % only show ticklabels and axis labels on the bottom
                y descriptions at=edge right,
                x descriptions at=edge bottom,
                % set the `vertical sep' to zero
                horizontal sep=0pt,
                vertical sep=0pt
            },
          width=9.5cm,
          axis y line*=right,
          axis x line=none,
          ymax=1.1,
          ylabel={proportion},
          legend style={at={(0.5,1)},anchor=south},
        %   legend pos=outer north east,
          legend columns=3, 
          legend cell align=left,
        ]
        \nextgroupplot[scaled y ticks=false]
        \addplot[mark=none, dashed, blue] table [x=dates, y=p_w, col sep=comma] {data/transitions/arh/binomial/budget_5/policies-transitions//0/run_0.csv};
        \label{p_w}
        \addplot[mark=none, dashed, magenta] table [x=dates, y=p_s, col sep=comma] {data/transitions/arh/binomial/budget_5/policies-transitions//0/run_0.csv};
        \label{p_s}
        \addplot[mark=none, dashed, darkgray] table [x=dates, y=p_l, col sep=comma] {data/transitions/arh/binomial/budget_5/policies-transitions//0/run_0.csv};
        \label{p_l}
        \addlegendimage{/pgfplots/refstyle=i_hosp_new}\addlegendentry{$p_w$}
        \addlegendimage{/pgfplots/refstyle=i_icu_new}\addlegendentry{$p_l$}
        \addlegendimage{/pgfplots/refstyle=d_new}\addlegendentry{$p_s$}
        \addlegendimage{/pgfplots/refstyle=p_w}\addlegendentry{{daily new hosp}}
        \addlegendimage{/pgfplots/refstyle=p_s}\addlegendentry{{daily new ICU}}
        \addlegendimage{/pgfplots/refstyle=p_l}\addlegendentry{{daily deaths}}
        \nextgroupplot[scaled y ticks=false]
        \addplot[mark=none, dashed, blue] table [x=dates, y=p_w, col sep=comma] {data/transitions/arh/binomial/budget_5/policies-transitions/45/run_0.csv};
        \addplot[mark=none, dashed, magenta] table [x=dates, y=p_s, col sep=comma] {data/transitions/arh/binomial/budget_5/policies-transitions/45/run_0.csv};
        \addplot[mark=none, dashed, darkgray] table [x=dates, y=p_l, col sep=comma] {data/transitions/arh/binomial/budget_5/policies-transitions/45/run_0.csv};
        \nextgroupplot[scaled y ticks=false]
        \addplot[mark=none, dashed, blue] table [x=dates, y=p_w, col sep=comma] {data/transitions/arh/binomial/budget_5/policies-transitions/114/run_0.csv};
        \addplot[mark=none, dashed, magenta] table [x=dates, y=p_s, col sep=comma] {data/transitions/arh/binomial/budget_5/policies-transitions/114/run_0.csv};
        \addplot[mark=none, dashed, darkgray] table [x=dates, y=p_l, col sep=comma] {data/transitions/arh/binomial/budget_5/policies-transitions/114/run_0.csv};
        \end{groupplot}
      \end{tikzpicture}

### 2023-02-01 paper/main.tex
replace policy-execution plots with budget plots
processed pieters comments (introduction)

@@
 %mdp reward
-\newcommand{\mdpreward}{\mathbf{r}}
+\newcommand{\mdpreward}{r}
+\newcommand{\momdpreward}{\mathbf{r}}
 \newcommand{\momdpreturn}{\mathbf{R}}
@@
 \section{Introduction}
-As shown by the COVID-19 pandemic, infectious disease outbreaks represent a paramount global challenge that should be tackled by prevention strategies.
+As shown by the COVID-19 pandemic, infectious disease outbreaks represent a major global challenge.
 To this end, understanding the complex dynamics that underlie these epidemics is essential.
@@
 
-For this reason, reinforcement learning provides a methodology to automatically learn prevention strategies in combination with complex epidemic models~\cite{libin2020}.
-Previous research typically focuses on optimising policies with respect to a single objective, such as the pathogen's attack rate, while the mitigation of epidemics is a problem that inherently covers distinct  and possibly conflicting criteria (i.a., prevalence, mortality, morbidity, cost).
+To address these challenges, reinforcement learning provides a methodology to automatically learn mitigation strategies in combination with complex epidemic models~\cite{libin2020}.
+Previous research focused on optimising policies with respect to a single objective, such as the pathogen's attack rate, while the mitigation of epidemics is a problem that inherently covers distinct  and possibly conflicting criteria (i.a., prevalence, mortality, morbidity, cost).
 Therefore, optimizing on a single objective requires that  these distinct criteria are somehow aggregated into a single metric.
@@
 This model constitutes a stochastic discrete-time age-structured compartmental model that simulates mitigation strategies by varying social distancing parameters concerning school, work and leisure contacts.
-Based on this model, we \emph{contribute \momdpname, a novel mutli-objective epidemiological environment}, in the form of a multi-objective Markov decision process (MOMDP).
-\momdpname\ encapsulates the epidemiological model developed by \citet{abrams2021modelling} to implement state transitions, with an action space that combines a proportional reduction of school, work and leisure contacts at each time step and defines a reward function based on two objectives: the attack rate (i.e., proportion of the population affected by the pathogen) and social burden.
-
-To learn and explore the trade-offs between the attack rate and social burden we build upon a state-of-the-art MORL approach, namely Pareto Conditioned Networks (PCN)~\cite{reymond2022pcn}, which uses a single neural network to learn the policies that belong to the Pareto front.
+Based on this model, we contribute \momdpname, a novel mutli-objective epidemiological environment, in the form of a multi-objective Markov decision process (MOMDP) \cite{roijers2013survey}.
+\momdpname\ encapsulates the epidemiological model developed by \citet{abrams2021modelling} to implement state transitions, with an action space that combines a proportional reduction of school, work and leisure contacts at each time step, Furthermore, it defines a reward function based on two objectives: the attack rate (i.e., proportion of the population affected by the pathogen) and the social burden that is induced by the mitigation measures.
+
+To learn and explore the trade-offs between the attack rate and social burden we present a state-of-the-art MORL approach, Pareto Conditioned Networks (PCN)~\cite{reymond2022pcn}.
+PCN uses a single neural network to learn the policies that belong to the Pareto front.
 As PCN is an algorithm designed for discrete action-spaces, we extend it towards continuous action-spaces to accommodate \momdpname's action-space.
@@
 \subsection{Multi-Objective Reinforcement Learning}
-Real-world decisions problems often have multiple and possibly conflicting objectives.
+Many real-world decisions problems often have multiple and possibly conflicting objectives.
 Multi-objective reinforcement learning (MORL) can be used to find optimal solutions for sequential decision making problems with multiple objectives~\cite{hayes2021practical}.
-For MORL, problems are typically modelled as a multi-objective Markov decision process (MOMDP), i.e., a tuple, $\mathcal{M} = (\mdpstatespace, \mdpactionspace, \mdptransition, \mdpdiscount, \momdprewardfn)$, where $\mdpstatespace$, $\mdpactionspace$ are the state and action spaces respectively, $\mdptransition \colon \mdpstatespace \times \mdpactionspace \times \mdpstatespace  \to \left[ 0, 1 \right]$ is a probabilistic transition function, $\mdpdiscount$ is a discount factor determining the importance of future rewards and $\momdprewardfn \colon \mdpstatespace \times \mdpactionspace \times \mdpstatespace \to \mathbb{R}^n$ is an $n$-dimensional vector-valued immediate reward function, where $n$ denotes the number of objectives.
-For single-objective RL, i.e., when $n = 1$, the goal is to find the policy $\mdppolicy^{*}$ that maximizes the expected sum of discounted rewards, also called the expected return:
-\begin{equation}
-\mdppolicy^*= \arg\max_\mdppolicy \mathbb{E} \left [ \sum_{t=0}^h \mdpdiscount^t r_t |\ \mdppolicy, s_0 \right ],
-\end{equation}
-where $r_t = \mdprewardfn(\state_t, \action_t, \state_{t+1})$.
+For MORL, problems are typically modelled as a multi-objective Markov decision process (MOMDP), i.e., a tuple, $\mathcal{M} = (\mdpstatespace, \mdpactionspace, \mdptransition, \mdpdiscount, \momdprewardfn)$, where $\mdpstatespace$, $\mdpactionspace$ denote the state and action spaces respectively, $\mdptransition \colon \mdpstatespace \times \mdpactionspace \times \mdpstatespace  \to \left[ 0, 1 \right]$ denotes a probabilistic transition function, $\mdpdiscount$ is a discount factor determining the importance of future rewards and $\momdprewardfn \colon \mdpstatespace \times \mdpactionspace \times \mdpstatespace \to \mathbb{R}^n$ is an $n$-dimensional vector-valued immediate reward function, where $n$ corresponds to the number of objectives.
+
+While MOMDPs define the actions an agent can take, it does not define which actions are taken in each state.
+We call this the agent's \emph{policy}.
+This policy dictates the agent's behavior.
+More formally, a policy $\mdppolicy$ is a probabilistic function $\mdpstatespace \times \mdpactionspace \to \left[ 0, 1 \right]$ determining the probability, for each state $s \in \mdpstatespace$, to take an action $a \in \mdpactionspace$.
+We measure the performance of $\mdppolicy$ through the expected sum of discounted rewards it achieves.
+This is denoted as the $\momdpvaluefunction$-value:
+
+\begin{equation}
+    \momdpvaluefunction^\mdppolicy = \mathbb{E} \left [ \sum_{t=0}^h \mdpdiscount^t \momdpreward_t |\ \mdppolicy, s_0 \right ],
+\end{equation}
+%
+where $\momdpreward_t = \mdprewardfn(\state_t, \action_t, \state_{t+1})$.
+
+For single-objective RL, i.e., when $n = 1$, the goal is to find the policy $\mdppolicy^{*}$ that maximizes the $\mdpvaluefunction$-value:
+\begin{equation}
+\mdppolicy^*= \arg\max_\mdppolicy \mdpvaluefunction^\mdppolicy.
+\end{equation}
+
 In contrast, in MORL, $n > 1$ which leads to vectorial returns.
@@
 At execution time, both $\hat{h}$ and $\mathbf{\hat{R}}$ are chosen by the decision maker at the start of the episode.
-Then, at every timestep, the desired horizon is updated according to the perceived reward $\mdpreward_t$, $\mathbf{\hat{R}} \leftarrow \mathbf{\hat{R}} - \mdpreward_t$ and the desired horizon is decreased by one, $\hat{h} \leftarrow \hat{h}-1$.
+Then, at every timestep, the desired horizon is updated according to the perceived reward $\momdpreward_t$, $\mathbf{\hat{R}} \leftarrow \mathbf{\hat{R}} - \momdpreward_t$ and the desired horizon is decreased by one, $\hat{h} \leftarrow \hat{h}-1$.
 
@@
 Moreover, although the Binomial model of \momdpname\ is stochastic, the possible next-states resulting from a state-action pair are similar to each other.
-This allows PCN to compensate if $\mdpreward_t = \momdprewardfn(\mdpstate_t, \mdpaction_t, \mdpstate_{t+1})$ is somewhat worse than expected.
+This allows PCN to compensate if $\momdpreward_t = \momdprewardfn(\mdpstate_t, \mdpaction_t, \mdpstate_{t+1})$ is somewhat worse than expected.
 This is confirmed by our experiments (see Section~\ref{sec:experiments}), where the coverage sets learned by PCN on the ODE model and on the Binomial model are similar.
@@
     \centering
-  \input{fig_policy-execution}
+    \input{fig/budget_5}
+    \caption{Selection of policies learned by PCN, from most restrictive in terms of social burden (top) to least restrictive (bottom).
+    }
+    \label{fig:policy-execution}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \input{fig/all_budgets}
     \caption{Selection of policies learned by PCN, from most restrictive in terms of social burden (top) to least restrictive (bottom).

### 2023-02-02 paper/main.tex
processed pieters comments (up to 3.2)

@@
 \subsection{Multi-Objective Metrics}
-Comparing the learned coverage sets of different algorithms is a non-trivial task, as one algorithm's output might dominate the other in some region of the objective-space, but be dominated in another.
+Comparing the learned coverage sets produced by different algorithms is a non-trivial task, as one algorithm's output might dominate the other in some region of the objective-space, but be dominated in another.
 Intuitively, one would generally prefer the algorithm that covers a wider range of decision maker preferences.
@@
 This metric evaluates the learned coverage set by computing its volume with respect to a fixed specified reference point.
-This metric is, by definition, the highest for the Pareto front, as no other possible solution can increase its volume (since they are all dominated).
+The reference point is taken as a lower bound on the achievable returns so that the volumes are always positive.
+By definition, the solutions contained in the Pareto front dominate all other possible solutions.
+Thus, no other solution can further increase the volume under the Pareto front.
+This means that the hypervolume is the highest for the Pareto front.
 One disadvantage of the hypervolume is that it can be difficult to interpret.
-For example, when working in high-dimensional objective-spaces, adding or removing a single point can lead to wildly different hypervolume values, especially if the point lies close to an extremum of the space.
-
-To alleviate these shortcomings, we additionally evaluate our work on a different metric called the $\varepsilon$-indicator $I_\varepsilon$~\cite{zitzler2003}, which measures how close a coverage set is to the Pareto front $\mathcal{F}$.
+For example, when working in high-dimensional objective-spaces, adding or removing a single point can drastically change hypervolume values, especially if the point lies close to an extremum of said space.
+
+To counterpoise these shortcomings, we additionally evaluate our work on a different metric called the $\varepsilon$-indicator $I_\varepsilon$~\cite{zitzler2003}, which measures how close a coverage set is to the Pareto front $\mathcal{F}$.
 Intuitively, $I_\varepsilon$ shows that any solution of $\mathcal{F}$ is \emph{at most} $\varepsilon$ better with respect to each objective $o$ than the closest solution of the evaluated coverage set:
@@
     \exists\ \momdpvaluefunction^{\pi'}\!{\in}\ \hat{\Pi} :\ 
-    || \mdpvaluefunction^\pi - \mdpvaluefunction_o^{\pi'} ||_{\infty} \le \varepsilon
-    \}
-\end{equation}
+    || \momdpvaluefunction^\pi - \momdpvaluefunction^{\pi'} ||_{\infty} \le \varepsilon
+    \},
+\end{equation}
+%
+where $||.||_{\infty}$, the L-infinity norm, returns the magnitude of the largest entry of a vector.
 
@@
 
-We note that the $\varepsilon$-indicator metric is quite pessimistic, as it measures worst-case performance~\cite{zintgraf2015}, e.g., it will still report bad performance as long as a single point of the Pareto front is not correctly modeled, even if all the other points are covered.
+We note that the $\varepsilon$-indicator metric is quite pessimistic, as it measures worst-case performance~\cite{zintgraf2015}, i.e., e.g., it will still report bad performance as long as a single point of the Pareto front is not correctly modeled, even if all the other points are covered.
 As such, we also use the $I_{\varepsilon-mean}$ \cite{reymond2022pcn} which measures the \emph{average} $\varepsilon$ distance of the solutions in $\mathcal{F}$ with respect to the evaluated coverage set.
@@
 This model generalises a standard SEIR model\footnote{A standard SEIR model divides the population into four different states, i.e., susceptible, exposed, infectious and recovered individuals.}, extended to capture the different stages of disease spread and history that are associated with SARS-CoV-2 (i.e., pre-symptomatic, asymptomatic, symptomatic with mild symptoms and symptomatic with severe symptoms) and to represent the stages associated with severe disease, i.e., hospitalisation, admission to the intensive care unit (ICU) and death.
-As we consider an age-structured population, we consider this extended SEIR structure for $\agegroups = 10$ age groups,  i.e., $[0-10), [10-20), [20-30), [30-40), [40-50), [50-60), [60-70), [70-80), [80-90), [90,\infty)$.
-Contacts of the different age-groups which impact the propagation rate of the epidemic are modeled using social contact matrices (SCMs).
+As we consider an age-structured population, we consider this extended SEIR structure for $\agegroups = 10$ age groups,  i.e., $[0-10), [10-20), [20-30), [30-40), [40-50), [50-60), [60-70), [70-80), [80-90), [90,100+)$.
+Contacts of the different age-groups, which impact the propagation rate of the epidemic, are modeled using social contact matrices (SCMs).
 We define a SCM for 6 different social environments: $\chome, \cwork, \ctransport, \cschool, \cleisure, \cother$ for the home, work, transport, school, leisure, other environments respectively.
 The model is described by a set of ordinary differential equations (ODEs), as described in SI.
-By formulating this set of differential equations defined above as a chain-binomial process we can obtain stochastic trajectories from this model~\cite{abrams2021modelling}.
+By formulating this set of differential equations a chain-binomial process (as detailed in SI) we can obtain stochastic trajectories from this model~\cite{abrams2021modelling}.
 

### 2023-02-02 paper/main.tex
processed pieters comments (up to 4.)

@@
 \newcommand{\agegroups}{K}
+\newcommand{\budget}{\bm{b}}
 
@@
 In order to model different types of interventions, we follow \citet{abrams2021modelling}.
-Firstly, in order to consider distinct exit scenarios, we alter the SCMs to reflect a contact reduction in a particular age group.
+Firstly, in order to consider distinct exit scenarios, we modulate the SCMs to reflect a contact reduction in a particular age group.
 Secondly, we assume that compliance to the interventions is gradual and model this using a logistic compliance function (see details in SI).
-We consider a contact reduction function that imposes a proportional reduction of work (including transport) $p_w$, school $p_s$ and leisure $p_l$ contacts, which is implemented as a linear combination of contact matrices:
+We consider a contact reduction function that imposes a proportional reduction of work (including transport) $p_w$, school $p_s$ and leisure $p_l$ contacts, which is implemented as a linear combination of SCMs:
 \begin{equation}
@@
 We use this function as the transition function in \momdpname.
-For each timestep $t$, we simulate the model for one week, using $\hat{C}$ obtained from $\mdpaction_t$.
-
-As mentioned above and detailed in the SI, the compartment model is described by a set of ODEs.
+
+\added{In a classical MDP, executing an action $\action_t$ in any state $\state_t$ leads to a next state $\state_{t+1}$ according to the transition function $\mdptransition$.
+At every timestep $t$, the policy is free to choose the action to perform.
+In our case, this potentially results in a different restriction $[p_{w}, p_{s}, p_{l}]$ every week.
+However, we argue that in the context of mitigation policies, consistency is key, and policies that impose changes too frequently will be hard to follow.}
+
+\added{In order to learn realistic and consistent mitigation policies, we incorporate a \emph{budget} on the number of times a policy can change its actions until the terminal state of the MOMDP.
+Concretely, when the action changes, i.e., if the social restriction proposed by the policy is different from the current one in place, we reduce the budget by one.
+We only allow action changes as long as there is budget left.
+We note that we can simulate a no-limit budget setting by setting the budget to the length of the episode.}
+
+Finally, for each timestep $t$, our transition function $\mdptransition$ uses the model transition probability $M$ to simulate the model for one week, using $\hat{C}$ obtained from $\mdpaction_t$.
+
+% As mentioned above and detailed in the SI, the compartment model is described by a set of ODEs.
 To obtain a stochastic version, the ODEs can be formulated as a chain-binomial process.
@@
 
-\added{In a classical MDP, executing an action $\action_t$ in any state $\state_t$ leads to a next state $\state_{t+1}$ according to the transition function $\mdptransition$.
-At every timestep, the policy is free to choose the action to perform.
-In our case, this corresponds to a different restriction $[p_{w}, p_{s}, p_{l}]$ every week.
-However, imposing new social restrictions every week does not seem very realistic.
-It can be frustrating to having to cope with the ever-changing rules.
-In case of the exit strategy after the lockdown of the first COVID wave in Belgium, the Belgian government executed a progressive plan that consisted of 3 different stages.
-Each stage resulted in the progressive lessening of social restrictions.}
-
-\added{In order to simulate a more realistic multi-stage exit strategy, we incorporate a \emph{budget per restriction-change} on the MOMDP.
-Concretely, if the social restriction proposed by the policy is different from the current one in place, we consider this a next stage and reduce our restriction-change budget by one.
-When the budget reaches 0, we consider the exit strategy to be final, and impose the current social restriction for the remainder of the episode.
-Finally, we can simulate a no-limit budget setting by setting the budget to the length of the episode.}
-
-\paragraph{State-space:} The state of the MOMDP \added{is a tuple that consists of 3 elements.
-The first element, $\mdpstate_m$, are the state-variables that make up the epidemiological model (see Figure~\ref{fig:ode_mdel}).
-They are required to simulate the underlying epidemiological model for a full week.}
-
-$\mdpstate_m$ directly corresponds to the aggregation of the state variables in the epidemiological model, i.e., a tuple,
+\paragraph{State-space:} The state of the MOMDP \added{concerns a 3-tuple.
+The first element,} $\mdpstate_m$, directly corresponds to the aggregation of the state variables in the epidemiological model, i.e., a tuple,
 \begin{equation}
@@
 \end{equation}
+%
 for each age group $k \in \{1, \ldots, \agegroups\}$, where $S$ encodes the members of the population who are susceptible to infection and $E$ encodes the members of the population who have been exposed to COVID-19.
@@
 
-We parameterize the \added{epidemiological} model using the mean of the posteriors as reported by \citet{abrams2021modelling} (details in SI).
+We parameterize the \added{epidemiological} model using the mean of the posteriors as specified by \citet{abrams2021modelling} (details in SI).
 
@@
 
-\added{Secondly, we require the current $\hat{C}$ to determine if we enter a new stage of the multi-stage exit strategy -- and thus consume part of the restriction-change budget -- or not.}
-
-\added{The third and last element of the tuple consists of the restriction-change budget $b$.
-We note that we incorporate a separate budget per action-dimension, so $p_{w}, p_{s}$ and $p_{l}$ each have their own budget, $b = [b_w, b_s, b_l]$.
-It is thus possible that, at timestep $t$, the budget for one of the dimensions is reduced but not the others.}
+\added{Secondly, we require the current $\hat{C}$ to determine whether the action changes the social restrictions in place -- and thus consume part of the budget -- or not.}
+
+\added{The third element of the tuple concerns of the budget $\budget$.
+We note that we incorporate a distinct budget per action-dimension, so $p_{w}, p_{s}$ and $p_{l}$ each have their own budget, rendering $\budget$ a vector $\budget = [b_w, b_s, b_l]$.
+As such, it is possible that, at timestep $t$, the budget for one of the proportional reductions is reduced but not the others.}
 
@@
 
-\paragraph{Reward function:} We define a vector reward function which considers multiple objectives: attack rate (i.e., infections, hospitalisations) and the burden imposed by the interventions on the population.
+\paragraph{Reward function:} We define a vectorial reward function which considers multiple objectives: attack rate (i.e., infections, hospitalisations) and the social burden imposed by the interventions on the population.
 
@@
 
-The reward function to reduce the attack rate in terms of hospitalisations is simply defined as the negative number of new hospitalizations:
-\begin{equation}
-\mdprewardfn_{\text{ARH}}(\mdpstate,\mdpaction,\mdpstate') = -\sum_{k=1}^{\agegroups}H^\text{new}_k(\mdpstate)
+The reward function to reduce the attack rate in terms of hospitalisations is defined as the negative number of new hospitalizations:
+\begin{equation}
+\mdprewardfn_{\text{ARH}}(\mdpstate,\mdpaction,\mdpstate') = -\sum_{k=1}^{\agegroups}H^\text{new}_k(\mdpstate).
 \label{eqn:attack-rate-hosp}
@@
 Finally, we use the missed contacts resulting from the intervention measures as a proxy for societal burden.
-To quantify missed contacts, we consider the original social contact matrix $C$ and the installed social contact matrix $\hat{C}$, and compute the difference $\hat{C} - C$.
+To quantify missed contacts, we consider the original social contact matrix $C$ and the installed social contact matrix $\hat{C}$ to compute the difference $\hat{C} - C$.
 The resulting difference matrix quantifies the average frequency of contacts missed.
-To determine missed contacts for the entire population, we apply the calculated difference matrix to the population sizes of the respective age groups that are currently uninfected (i.e., susceptible and recovered individuals).
+To determine missed contacts for the entire population, we apply the difference matrix to the population sizes of the respective age groups that are currently uninfected (i.e., susceptible and recovered individuals).
 Therefore we define the social burden reward function $\mdprewardfn_{\text{SB}}$, as follows:

### 2023-02-02 paper/main.tex
processed pieters comments (section 4.)

@@
 \added{PCN collects data from the trajectories experienced while exploring the environment.
-Thus, the dataset improves over time, as we collect better and better trajectories.}
-
-\added{Concretely, the PCN algorithm is a loop that repeats 3 main steps:
+Thus, the dataset improves over time, as we collect increasingly better trajectories.}
+
+\added{Concretely, the PCN algorithm is a loop that repeatedly executes 3 main steps:
 \begin{enumerate}
-    \item Using the current policy to interact with the environment and generate trajectories
-    \item Update and prune the dataset to incorporate these new trajectories and remove the least interesting trajectories in terms of returns
-    \item Update the policy by training it on the updated dataset
+    \item Use the current policy to interact with the environment and generate trajectories (see Section~\ref{sec:pcn-policy-exploration}).
+    \item Update the dataset to incorporate these new trajectories and prun the least interesting trajectories in terms of returns (see Section~\ref{sec:pcn-building-dataset}).
+    \item Update the policy by training it on the updated dataset (see Section~\ref{sec:pcn-continuous}).
 \end{enumerate}

### 2023-02-07 paper/fig/arh_binomial.tex
impact of budgets on coverage set

--- 2023-01-31 paper/fig/arh.tex
@@
         % height=6cm,
-    xlabel={total daily new hospitalizations},
-    ylabel={total lost contacts},
-    width=0.45\textwidth
+        xlabel={total number of daily new hospitalizations},
+        ylabel={total lost contacts}
     }
-\begin{groupplot}[
-    group style={
-        group size=2 by 1,
-        y descriptions at=edge left,
-        vertical sep=0pt,
-        horizontal sep=0pt,
-    },
+    \begin{axis}[
     legend entries={{$\text{PCN}_{b=2}$}, {$\text{PCN}_{b=3}$}, {$\text{PCN}_{b=4}$}, {$\text{PCN}_{b=5}$}, {$\text{PCN}$}, {Fixed}},
+    % width=9.5cm,
+    mark size=1pt,
     ]
-\nextgroupplot[mark size=1pt,draw opacity=1]
-        \addplot[only marks,mark=o,pc1] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/ode/budget_2.csv};
-        \addplot[only marks,mark=o,pc2] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/ode/budget_3.csv};
-        \addplot[only marks,mark=o,pc3] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/ode/budget_4.csv};
-        \addplot[only marks,mark=o,pc4] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/ode/budget_5.csv};
-        \addplot[only marks,mark=o,pc5] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/ode/budget_None.csv};
-        \addplot[only marks,mark=+] table [x=o_0, y=o_1, col sep=comma] {data/cs/cs_fixed.csv};
-\nextgroupplot[mark size=1pt]
     \addplot[only marks,draw opacity=1,fill opacity=0.5,mark=o,pc1] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/binomial/budget_2.csv};
@@
     \addplot[only marks,mark=+] table [x=o_0, y=o_1, col sep=comma] {data/cs/cs_fixed.csv};
-
-\end{groupplot}
+    \end{axis}
 \end{tikzpicture}

### 2023-02-07 paper/main.tex
impact of budgets on coverage set

@@
 \newcommand\added[1]{\textcolor{blue}{#1}}
+\newcommand\todo[1]{\textcolor{red}{TODO: #1}}
 
@@
 Our goal is to use PCN to learn deconfinment strategies in the \momdpname\ environment.
-We aim to obtain policies that balance between the social burden experienced by the population and the epidemiological objective of minimising the attack rate.
+We aim to learn policies that balance between the epidemiological objective of minimising the attack rate and the social burden experienced by the population.
 To this end we consider two cases for the vectorial reward functions $[\mdprewardfn_\text{ARH}, \mdprewardfn_\text{SB}]$ and $[\mdprewardfn_\text{ARI}, \mdprewardfn_\text{SB}]$, to learn and analyse policies under different targets with respect to the considered attack rate.
 
-We now evaluate our extension of PCN for continuous action-spaces on both the ODE and Binomial models.
+To conduct this analysis, we evaluate our extension of PCN for continuous action-spaces on both the Binomial model.
 Conform to \citet{abrams2021modelling}, the simulation starts on the 1st of March 2020, by seeding a number of infections in the population.
@@
 
+% https://nl.wikipedia.org/wiki/Maatregelen_tijdens_de_coronacrisis_in_Belgi%C3%AB
+We draw the analogy with the multi-phase exit strategy established by the Belgian government and the restriction on the number of action-changes imposed by the \momdpname's budget $\budget$.
+Indeed, on the 11th of May 2020, exactly one week after the end of the lockdown, stores and certain types of jobs were allowed to reopen, under strict conditions.
+This corresponds to altering $p_w$ in our MOMDP.
+One week later, on 18th of May, primary and secondary schools reopened for limited sized class-groups, and the cultural sector reopened partially.
+This is equivalent to increasing $p_s$ and $p_l$.
+Further changes of restrictions occured on the 8th of June, 1st, 9th and 25th of July.
+Thus, we argue that, with a limited budget, we can achieve realistic and implementable policies.
+In our experiments, we consider budgets of 2 to 5, as these closely relate to the number of changes that occured until the end of the summer holidays of 2020.
+As an upper bound, we also consider a no-limit budget setting.
+
 To evaluate the quality of the policies discovered by PCN, we compare it to a baseline.
 The baseline consists of a set of 100 fixed policies, that iterate over all the possible social restriction levels, with values ranging from 0 to 1.
+Concretely, each policy uses a fixed proportional reduction $p_w=p_l=p_s=u, u\sim\mathcal{U}(0, 1)$ throughout the episode.
 In other words, the fixed policies directly operate in a fine-grained manner on the whole contact reduction function $\hat{C}$.
@@
     \centering
-    \input{fig/arh.tex}
-    \caption{The Pareto front of policies discovered by PCN using the ODE model (Binomial omitted for clarity, see SI), showing the different compromises between the number of hospitalizations and the number of lost contacts.
-Although $\text{PCN}_\text{ARI}$ was trained on the number of infections, it still shows a competitive coverage set with respect to hospitalizations.}
-    \label{fig:pareto-front}
+    \input{fig/arh_binomial.tex}
+    \caption{The Pareto front of policies discovered by PCN using the Binomial model, showing the different compromises between the number of hospitalizations and the number of lost contacts.
+As the budget increases, so does the coverage set learnt by PCN.
+Changes are most noticeable in the less restrictive trade-offs in terms of social burden.}
+    \label{fig:pf-binomal-arh-budgets}
 \end{figure}
@@
 We learn a coverage set that ranges from imposing minimal restrictions to enforcing many restrictions (see Figure~\ref{fig:policy-execution}).
-The coverage set is shown in Figure~\ref{fig:pareto-front}.
-
-We notice that the coverage sets discovered by PCN almost completely dominate the coverage set of the baseline, showing that there are much better alternatives to the fixed policies.
+The coverage set is shown in Figure~\ref{fig:pf-binomal-arh-budgets}.
+
+Regardless of the imposed budget, we notice that the coverage sets discovered by PCN almost completely dominate the coverage set of the baseline, demonstrating that there are better alternatives to the fixed policies.
 This is most visible in the compromising policies, where one has to carefully choose when to remove social restrictions while at the same time minimizing the impact on daily new hospitalizations.
@@
 
+\subsection{Impact of budgets on the coverage set}
+
+Looking at the coverage sets displayed in Figure~\ref{fig:pf-binomal-arh-budgets}, we notice that the budget impacts the coverage set that has been learnt.
+In general, an increase of budget results in an increasingly better coverage set.
+This is to be expected, as a higher budget give the agent more freedom to change its actions as the epidemic progresses.
+However, we also notice that this increase of coverage sets is not necessarily gradual.
+Between the run with a budget of 2 and budget of 3, the difference in coverage set seems marginal.
+\todo{insight}.
+
+Moreover, we observe that the improvement concentrates around the less restrictive policies in terms of social burden.
+We believe that this regio contains the most complex policies, as these try to maintain as much social freedom as possible, while containing the number of hospitalisations.
+We believe that, in these cases, the timing of the actions greatly correlate with the timing and duration of the peak of the epidemic, and a higher budget allows for more fine-grained control about managing this timing.
+To confirm this, we select, for each budget setting, the solution where the difference in performance is most noticeable, corresponding to the solutions with a total number of hospitalisations around $80000$.
+We plot the execution of the corresponding policies in Figure~\ref{fig:pe-binomial-arh-budgets} and analyse the difference they make in terms of social burden.
+First, we notice that the lower-budget policies are unable to lessen the social restrictions past the peak of the epidemic.
+In contrast, the setting with no budget restrictions finely controls the restrictions as the epidemic progresses, completely removing restrictions by the end of the wave.
+We note how the policy with a budget of 5 ressembles the execution of the one without restrictions but that, due to its budget, is unable to progressively lessen the restrictions and instead resorts to a halfway compromise.
+\todo{corresponding SB numbers}
+
+Compared to this specific regio of the coverage set, the difference in performance between different budget settings seems marginal around the extremas. \todo{explain}
+
+Finally, we observe that, while the extremas deliver similar trade-offs regardless of the chosen budget, the results differ for the setting without budget restrictions.
+Indeed, in this setting, PCN does not seem to learn the most and least restrictive policies, even though there are no constraints on the action-set.
+As explained in Section~\ref{sec:pcn-policy-exploration}, PCN searches for increasingly better solutions using a stochastic policy.
+Thus, at every timestep, the action can change compared to the previous one.
+Continuously outputting the same action (e.g., no social restrictions) becomes a complicated task.
+In comparison, for the settings with a limited budget, the action stays the same as the previous one once the budget has been spent.
+It is thus easier to discover the most extreme policies.
+Thus, in the specific case where we impose no budget, we believe the freedom of action actually hinders PCN's search.
+
 \begin{figure}
@@
     \input{fig/all_budgets}
-    \caption{Selection of policies learned by PCN, from most restrictive in terms of social burden (top) to least restrictive (bottom).
+    \caption{Execution of the policies attaining a number of hospitalisations around $80000$, for different budgets.
+From top to bottom are displayed the policy executions of runs with budget 2,3,4,5 and no-limit, respectively.
+We notice that the lower-budget policies are unable to lessen the social restrictions past the peak.
+The setting with no budget restrictions finely controls the restrictions as the epidemic progresses, completely removing restrictions by the end of the wave.
+Finally, there is no consensus on which social environment to restrict most, some policies providing similar restrictions for $p_w$ and $p_l$, while others putting harsher restrictions on one social environment than the other.
     }
-    \label{fig:policy-execution}
+    \label{fig:pe-binomial-arh-budgets}
 \end{figure}

### 2023-02-07 paper/fig/arh_binomial_both.tex
interpolated plot

\begin{tikzpicture}
    \pgfplotsset{
        scale only axis,
        % height=6cm,
        xlabel={total daily new hospitalizations},
        ylabel={total lost contacts},
        width=0.45\textwidth
    }
    \begin{groupplot}[
        group style={
            group size=2 by 1,
            y descriptions at=edge left,
            vertical sep=0pt,
            horizontal sep=0pt,
        },
        legend entries={{$\text{PCN}_{b=2}$}, {$\text{PCN}_{b=3}$}, {$\text{PCN}_{b=4}$}, {$\text{PCN}_{b=5}$}, {$\text{PCN}$}, {Fixed}},
    ]
    % interpolated plot
    \nextgroupplot[draw opacity=1]
    % mean line
    \foreach \b/\c in {2/pc1, 3/pc2, 4/pc3, 5/pc4, {None}/pc5}
    {
        \edef\temp{
            \noexpand\addplot[fill opacity=0.5,\c] table [x=mu_o_1, y=mu_o_5, col sep=comma] {data/cs/arh/binomial/interpolated/budget_\b.csv};
        }
        \temp
    }
    % std fillbetween, has to be after means for legend order
    \foreach \b/\c in {2/pc1, 3/pc2, 4/pc3, 5/pc4, {None}/pc5}
    {
        \edef\temp{
            \noexpand\addplot[name path=upper\b, draw=none] table [x=mu_o_1, y expr=\noexpand\thisrow{mu_o_5}+\noexpand\thisrow{std_o_5}, col sep=comma] {data/cs/arh/binomial/interpolated/budget_\b.csv};
            \noexpand\addplot[name path=lower\b, draw=none] table [x=mu_o_1, y expr=\noexpand\thisrow{mu_o_5}-\noexpand\thisrow{std_o_5}, col sep=comma] {data/cs/arh/binomial/interpolated/budget_\b.csv};
            \noexpand\addplot[\c!15] fill between[of={upper\b} and {lower\b}];
        }
        \temp
    }
    \addplot[only marks,mark=+] table [x=o_0, y=o_1, col sep=comma] {data/cs/cs_fixed.csv};
    % best performing run
    \nextgroupplot[mark size=1pt,draw opacity=1]
    \foreach \b/\c in {2/pc1, 3/pc2, 4/pc3, 5/pc4, {None}/pc5}
    {
        \edef\temp{
            \noexpand\addplot[only marks,mark=o,fill opacity=0.5,\c] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/binomial/budget_\b.csv};
        }
        \temp
    }
    \addplot[only marks,mark=+] table [x=o_0, y=o_1, col sep=comma] {data/cs/cs_fixed.csv};
    \end{groupplot}
\end{tikzpicture}

### 2023-02-08 paper/main.tex
interpolated plot
metric per budget
arh and ari table
remove ode vs binomial section

@@
 \usepackage{comment}
-\usepgfplotslibrary{dateplot, groupplots}
+\usepgfplotslibrary{dateplot, groupplots, fillbetween}
 \pgfplotsset{compat=newest}
@@
 \subsection{Coping with stochastic transitions}
+\label{sec:pcn-stochastic}
 
@@
 To conduct this analysis, we evaluate our extension of PCN for continuous action-spaces on both the Binomial model.
+\todo{reason for binomial}.
+As explained in Section~\ref{sec:pcn-stochastic}, we extend PCN for environments with stochastic transitions.
+While all results in this Section concern the Binomial model, we observe similar performance on the ODE model, showing that our extension does resolve the stochasticity issue.
+These results are available in the SI.
+
 Conform to \citet{abrams2021modelling}, the simulation starts on the 1st of March 2020, by seeding a number of infections in the population.
@@
     \centering
-    \input{fig/arh_binomial.tex}
+    \input{fig/arh_binomial_both}
     \caption{The Pareto front of policies discovered by PCN using the Binomial model, showing the different compromises between the number of hospitalizations and the number of lost contacts.
@@
 
-\subsection{ODE versus Binomial models}
-\label{sec:ode-vs-binomial}
-
-Next, we compare the performance of PCN when trained on the ODE and Binomial models.
-The ODE model has the advantage of using a deterministic transition function, for which PCN is suited.
-However, taking into account stochasticity, using the Binomial model, is important as individual behavioural changes introduce substantial uncertainty in the course of the outbreak and require stochastic model evaluations to evaluate the impact of this uncertainty on the effectiveness of the intervention strategies.
-Regardless, the results displayed in Table~\ref{tab:coverage-set} show that the solution sets that PCN finds using the Binomial model are competitive with the ODE model, as indicated by their hypervolumes and $I_{\varepsilon-mean}$ metrics.
-We do note that the difference in $I_{\varepsilon}$ indicator is more significant, indicating that a few solutions discovered using the ODE model are not found when using the Binomial model, impacting worst-case performance.
-Overall, we see that our extension of PCN performs efficiently, even with some degree of stochasticity in the transition function.
-
 \begin{table}[t]
@@
   {\renewcommand{\arraystretch}{1.2}% for the vertical padding
-  \input{tab_coverage-set}
+  \input{tab/arh/binomial}
   }

### 2023-02-08 paper/main.tex
update arh vs ari setction

@@
 We now assess the difference in coverage sets when optimizing on $\mdprewardfn_\text{ARH}$ versus $\mdprewardfn_\text{ARI}$.
-Although at a different scale, our experiments show that infections and hospitalizations are correlated.
-This is confirmed in Figure~\ref{fig:pareto-front}: the coverage set \emph{in terms of hopsitalizations} learned by $\text{PCN}_\text{ARI}$ is only slightly worse than the one learned by $\text{PCN}_\text{ARH}$, even though $\text{PCN}_\text{ARI}$ optimized \emph{on the infection attack rate}.
+Although these reward functions have a different scale (there are more infected persons than hospitalised ones), our experiments show that infections and hospitalizations are correlated.
 This is expected, as during the initial phase of the epidemic, limited immunity was present in the population (few natural immunity and no vaccines), which induces a tight coupling between infection and hospitalisation cases.
-Moreover, the opposite also holds: both coverage sets are similar in terms of \emph{infections}.
-However, there is one notable exception.
-A number of policies learned by $\text{PCN}_\text{ARH}$ start removing social restrictions before the epidemic is rooted out, resulting in a rise of hospitalizations at the end of the summer holidays.
-Since, following Abrams et al., this is when the simulation stops, the total number of hospitalizations remains low.
-The number of infections, however, starts growing before this rise in hospitalizations and, since its growth is exponential, ends up high.
-We note that $\text{PCN}_\text{ARH}$ also learns alternative policies for similar $[\mdprewardfn_\text{ARH}, \mdprewardfn_\text{SB}]$ that do not exhibit this behavior.
-This shows that, even though a policy might lead to a return that matches the preferences of the decision maker, its execution might still lead to undesired behavior.
-We argue that the use of such expert systems for decision making should be paired with careful interpretation.
+This is confirmed in Table~\ref{tab:coverage-set}.
+In this table, we show the different performance metrics (hypervolume, $I_\varepsilon$, $I_{\varepsilon-mean}$) with respect to the objectives $[\mdprewardfn_\text{ARH}, \mdprewardfn_\text{SB}]$.
+However, the left-side of the table shows the experiments where these objectives are also used as optimization criterions, while the right-side shows the experiments using $[\mdprewardfn_\text{ARI}, \mdprewardfn_\text{SB}]$ as optimization criterions.
+
+Even with the $\mdprewardfn_\text{ARI}$, the increased budget shows an increase in hypervolume in terms of hospitalisations.
+Moreover, those hypervolumes are close to the ones trained on $\mdprewardfn_\text{ARH}$, inidicating that their coverage sets are similar.
+However, regardless of the imposed budget, the hypervolumes are slightly worse.
+This is to be expected, since those experiments are not directly optimising on $\mdprewardfn_\text{ARH}$.
+We make similar conclusions for $I_\varepsilon$.
+For budgets 2, 3 and 5, the difference between the worst-performing policy for the $\mdprewardfn_\text{ARI}$ variant and the $\mdprewardfn_\text{ARH}$ is less than $0.01$, indicating less than $1\%$ difference in return values between the two variants when comparing their worst-performing policy.
+As an exception, we notice that PCN without budget restrictions results in better performance across every metric for the $\mdprewardfn_\text{ARI}$ variant.
+Still, due to the high standard deviation of the no-budget, $\mdprewardfn_\text{ARH}$ setting, we do not believe this difference is meaningfull.
+\todo{some kind of meaningfull conclusion}
 

### 2023-02-09 paper/main.tex
update pcn robustness section, table

@@
 
+\subsection{Learned coverage set}
+
 \begin{figure}
@@
     \label{fig:pf-binomal-arh-budgets}
-\end{figure}
-
-\subsection{Learned coverage set}
-
-We learn a coverage set that ranges from imposing minimal restrictions to enforcing many restrictions (see Figure~\ref{fig:policy-execution}).
-The coverage set is shown in Figure~\ref{fig:pf-binomal-arh-budgets}.
-
-Regardless of the imposed budget, we notice that the coverage sets discovered by PCN almost completely dominate the coverage set of the baseline, demonstrating that there are better alternatives to the fixed policies.
-This is most visible in the compromising policies, where one has to carefully choose when to remove social restrictions while at the same time minimizing the impact on daily new hospitalizations.
-In these scenarios, PCN discovers policies that can drastically reduce the total number of new hospitalizations (e.g.
-more than $20000$) for the same social burden.
-Analysing the executions of such policies (see Figure~\ref{fig:policy-execution}, middle plot) shows a flattened hospitalization curve, with a gradual increase of social freedom during the school holidays such that the peak stays stable and gradually decreases over time.
-
-Interestingly, we notice that the most restrictive policy (i.e.
-the one that prioritizes hospitalizations over social burden, see Figure~\ref{fig:policy-execution}, bottom plot) still starts to gradually increase $p_w$ and $p_l$ from the end of July onward.
-This is because by then, the epidemic has mostly faded out, and it is safe to reduce social restrictions.
-The timing of this reduction is important as reducing restrictions too soon can lead to a new wave.
-PCN learns the impact of its decisions over time, and correctly infers the timing at which restrictions can be safely lifted.
-
-\begin{figure}
-    \centering
-    \input{fig/budget_5}
-    \caption{Selection of policies learned by PCN, from most restrictive in terms of social burden (top) to least restrictive (bottom).
-    }
-    \label{fig:policy-execution}
-\end{figure}
-
-\subsection{Impact of budgets on the coverage set}
-
-Looking at the coverage sets displayed in Figure~\ref{fig:pf-binomal-arh-budgets}, we notice that the budget impacts the coverage set that has been learnt.
-In general, an increase of budget results in an increasingly better coverage set.
-This is to be expected, as a higher budget give the agent more freedom to change its actions as the epidemic progresses.
-However, we also notice that this increase of coverage sets is not necessarily gradual.
-Between the run with a budget of 2 and budget of 3, the difference in coverage set seems marginal.
-\todo{insight}.
-
-Moreover, we observe that the improvement concentrates around the less restrictive policies in terms of social burden.
-We believe that this regio contains the most complex policies, as these try to maintain as much social freedom as possible, while containing the number of hospitalisations.
-We believe that, in these cases, the timing of the actions greatly correlate with the timing and duration of the peak of the epidemic, and a higher budget allows for more fine-grained control about managing this timing.
-To confirm this, we select, for each budget setting, the solution where the difference in performance is most noticeable, corresponding to the solutions with a total number of hospitalisations around $80000$.
-We plot the execution of the corresponding policies in Figure~\ref{fig:pe-binomial-arh-budgets} and analyse the difference they make in terms of social burden.
-First, we notice that the lower-budget policies are unable to lessen the social restrictions past the peak of the epidemic.
-In contrast, the setting with no budget restrictions finely controls the restrictions as the epidemic progresses, completely removing restrictions by the end of the wave.
-We note how the policy with a budget of 5 ressembles the execution of the one without restrictions but that, due to its budget, is unable to progressively lessen the restrictions and instead resorts to a halfway compromise.
-\todo{corresponding SB numbers}
-
-Compared to this specific regio of the coverage set, the difference in performance between different budget settings seems marginal around the extremas. \todo{explain}
-
-Finally, we observe that, while the extremas deliver similar trade-offs regardless of the chosen budget, the results differ for the setting without budget restrictions.
-Indeed, in this setting, PCN does not seem to learn the most and least restrictive policies, even though there are no constraints on the action-set.
-As explained in Section~\ref{sec:pcn-policy-exploration}, PCN searches for increasingly better solutions using a stochastic policy.
-Thus, at every timestep, the action can change compared to the previous one.
-Continuously outputting the same action (e.g., no social restrictions) becomes a complicated task.
-In comparison, for the settings with a limited budget, the action stays the same as the previous one once the budget has been spent.
-It is thus easier to discover the most extreme policies.
-Thus, in the specific case where we impose no budget, we believe the freedom of action actually hinders PCN's search.
-
-\begin{figure}
-    \centering
-    \input{fig/all_budgets}
-    \caption{Execution of the policies attaining a number of hospitalisations around $80000$, for different budgets.
-From top to bottom are displayed the policy executions of runs with budget 2,3,4,5 and no-limit, respectively.
-We notice that the lower-budget policies are unable to lessen the social restrictions past the peak.
-The setting with no budget restrictions finely controls the restrictions as the epidemic progresses, completely removing restrictions by the end of the wave.
-Finally, there is no consensus on which social environment to restrict most, some policies providing similar restrictions for $p_w$ and $p_l$, while others putting harsher restrictions on one social environment than the other.
-    }
-    \label{fig:pe-binomial-arh-budgets}
 \end{figure}
@@
 
+We learn a coverage set that ranges from imposing minimal restrictions to enforcing many restrictions (see Figure~\ref{fig:policy-execution}).
+The coverage set is shown in Figure~\ref{fig:pf-binomal-arh-budgets}.
+
+Regardless of the imposed budget, we notice that the coverage sets discovered by PCN almost completely dominate the coverage set of the baseline, demonstrating that there are better alternatives to the fixed policies.
+This is most visible in the compromising policies, where one has to carefully choose when to remove social restrictions while at the same time minimizing the impact on daily new hospitalizations.
+In these scenarios, PCN discovers policies that can drastically reduce the total number of new hospitalizations (e.g.
+more than $20000$) for the same social burden.
+Analysing the executions of such policies (see Figure~\ref{fig:policy-execution}, middle plot) shows a flattened hospitalization curve, with a gradual increase of social freedom during the school holidays such that the peak stays stable and gradually decreases over time.
+
+Interestingly, we notice that the most restrictive policy (i.e.
+the one that prioritizes hospitalizations over social burden, see Figure~\ref{fig:policy-execution}, bottom plot) still starts to gradually increase $p_w$ and $p_l$ from the end of July onward.
+This is because by then, the epidemic has mostly faded out, and it is safe to reduce social restrictions.
+The timing of this reduction is important as reducing restrictions too soon can lead to a new wave.
+PCN learns the impact of its decisions over time, and correctly infers the timing at which restrictions can be safely lifted.
+
+\subsection{Impact of budgets on the coverage set}
+
+\begin{figure}
+    \centering
+    \input{fig/budget_5}
+    \caption{Selection of policies learned by PCN, from most restrictive in terms of social burden (top) to least restrictive (bottom).
+    }
+    \label{fig:policy-execution}
+\end{figure}
+
+Looking at the coverage sets displayed in Figure~\ref{fig:pf-binomal-arh-budgets}, we notice that the budget impacts the coverage set that has been learnt.
+In general, an increase of budget results in an increasingly better coverage set.
+This is to be expected, as a higher budget give the agent more freedom to change its actions as the epidemic progresses.
+However, we also notice that this increase of coverage sets is not necessarily gradual.
+Between the run with a budget of 2 and budget of 3, the difference in coverage set seems marginal.
+\todo{insight}.
+
+Moreover, we observe that the improvement concentrates around the less restrictive policies in terms of social burden.
+We believe that this regio contains the most complex policies, as these try to maintain as much social freedom as possible, while containing the number of hospitalisations.
+We believe that, in these cases, the timing of the actions greatly correlate with the timing and duration of the peak of the epidemic, and a higher budget allows for more fine-grained control about managing this timing.
+To confirm this, we select, for each budget setting, the solution where the difference in performance is most noticeable, corresponding to the solutions with a total number of hospitalisations around $80000$.
+We plot the execution of the corresponding policies in Figure~\ref{fig:pe-binomial-arh-budgets} and analyse the difference they make in terms of social burden.
+First, we notice that the lower-budget policies are unable to lessen the social restrictions past the peak of the epidemic.
+In contrast, the setting with no budget restrictions finely controls the restrictions as the epidemic progresses, completely removing restrictions by the end of the wave.
+We note how the policy with a budget of 5 ressembles the execution of the one without restrictions but that, due to its budget, is unable to progressively lessen the restrictions and instead resorts to a halfway compromise.
+\todo{corresponding SB numbers}
+
+Compared to this specific regio of the coverage set, the difference in performance between different budget settings seems marginal around the extremas. \todo{explain}
+
+Finally, we observe that, while the extremas deliver similar trade-offs regardless of the chosen budget, the results differ for the setting without budget restrictions.
+Indeed, in this setting, PCN does not seem to learn the most and least restrictive policies, even though there are no constraints on the action-set.
+As explained in Section~\ref{sec:pcn-policy-exploration}, PCN searches for increasingly better solutions using a stochastic policy.
+Thus, at every timestep, the action can change compared to the previous one.
+Continuously outputting the same action (e.g., no social restrictions) becomes a complicated task.
+In comparison, for the settings with a limited budget, the action stays the same as the previous one once the budget has been spent.
+It is thus easier to discover the most extreme policies.
+Thus, in the specific case where we impose no budget, we believe the freedom of action actually hinders PCN's search.
+
+\begin{figure}
+    \centering
+    \input{fig/all_budgets}
+    \caption{Execution of the policies attaining a number of hospitalisations around $80000$, for different budgets.
+From top to bottom are displayed the policy executions of runs with budget 2,3,4,5 and no-limit, respectively.
+We notice that the lower-budget policies are unable to lessen the social restrictions past the peak.
+The setting with no budget restrictions finely controls the restrictions as the epidemic progresses, completely removing restrictions by the end of the wave.
+Finally, there is no consensus on which social environment to restrict most, some policies providing similar restrictions for $p_w$ and $p_l$, while others putting harsher restrictions on one social environment than the other.
+    }
+    \label{fig:pe-binomial-arh-budgets}
+\end{figure}
+
 \subsection{$R_\text{ARH}$ versus on $R_\text{ARI}$}
@@
 
-The dataset of trajectories that PCN is trained on is pruned over time to keep only the most relevant trajectories.
-The returns of these trajectories are used in Figure~\ref{fig:pareto-front} to show the discovered coverage set.
-Each of these returns can be used as desired return for policy execution.
-We now assess the robustness of the executed policies, by comparing the averaged return obtained over multiple policy executions with the corresponding target return.
-We show that the executed policies reliably obtain returns that are similar to the desired return used to condition PCN.
-
-Results are shown in Table~\ref{tab:pcn-robustness}.
-The $I_\varepsilon$ indicators shows that, over all policies, the decision maker will lose at worst $0.046$ and $0.068$ normalized returns in any of the objectives for the ODE, Binomial versions respectively.
-On average, it will lose $0.007$, $0.024$ normalized returns respectively, which corresponds to either a total of $3633, 1059$ more hospitalizations than expected or a total of $262, 76$ less contacts than expected.
-
-\begin{table}[t]
+\begin{table}
     \centering
@@
     {\renewcommand{\arraystretch}{1.2}% for the vertical padding
-    \input{tab_pcn-robustness}
+    \input{tab/arh/arh_binomial_robustness}
     }
@@
 \end{table}
+
+The dataset of trajectories that PCN is trained on is pruned over time to keep only the most relevant trajectories.
+The returns of these trajectories are used in Figure~\ref{fig:pf-binomal-arh-budgets} to show the discovered coverage set.
+Each of these returns can be used as desired return for policy execution.
+We now assess the robustness of the executed policies, by comparing the return obtained after executing the policy with the corresponding target return.
+For each run, we execute the policy $10$ times and compute the $I_\varepsilon$ and $I_{\varepsilon-mean}$ metrics with respect to the coverage set learned during the run.
+We show that the executed policies reliably obtain returns that are similar to the desired return used to condition PCN.
+
+Results are shown in Table~\ref{tab:pcn-robustness}.
+The $I_\varepsilon$ indicators shows that, regardless of the budget, the decision maker will lose at worst a $0.050$ normalized return in any of the objectives.
+On average, it will lose $0.010$ returns.
+Moreover, we emphasize that the learned coverage set contains the non-dominated returns across the whole training procedure.
+Since the Binomial model is stochasticity, for multiple executions of the same policy, the executions that remains in the coverage set is the one with advantageous transitions due to this stochasticity.
+Thus, we expect our policy-executions to obtain slightly worse results than the target-solution selected from the coverage set.
+In light of this, we conclude that the policies trained by PCN are reliable and produce returns as close as possible from their chosen target.
 

### 2023-02-09 paper/main.tex
remove \added

@@
 \newcommand{\foi}{\lambda}
-\newcommand\added[1]{\textcolor{blue}{#1}}
 \newcommand\todo[1]{\textcolor{red}{TODO: #1}}
@@
 
-\added{In a classical MDP, executing an action $\action_t$ in any state $\state_t$ leads to a next state $\state_{t+1}$ according to the transition function $\mdptransition$.
+In a classical MDP, executing an action $\action_t$ in any state $\state_t$ leads to a next state $\state_{t+1}$ according to the transition function $\mdptransition$.
 At every timestep $t$, the policy is free to choose the action to perform.
 In our case, this potentially results in a different restriction $[p_{w}, p_{s}, p_{l}]$ every week.
-However, we argue that in the context of mitigation policies, consistency is key, and policies that impose changes too frequently will be hard to follow.}
-
-\added{In order to learn realistic and consistent mitigation policies, we incorporate a \emph{budget} on the number of times a policy can change its actions until the terminal state of the MOMDP.
+However, we argue that in the context of mitigation policies, consistency is key, and policies that impose changes too frequently will be hard to follow.
+
+In order to learn realistic and consistent mitigation policies, we incorporate a \emph{budget} on the number of times a policy can change its actions until the terminal state of the MOMDP.
 Concretely, when the action changes, i.e., if the social restriction proposed by the policy is different from the current one in place, we reduce the budget by one.
 We only allow action changes as long as there is budget left.
-We note that we can simulate a no-limit budget setting by setting the budget to the length of the episode.}
+We note that we can simulate a no-limit budget setting by setting the budget to the length of the episode.
 
@@
 
-\paragraph{State-space:} The state of the MOMDP \added{concerns a 3-tuple.
-The first element,} $\mdpstate_m$, directly corresponds to the aggregation of the state variables in the epidemiological model, i.e., a tuple,
+\paragraph{State-space:} The state of the MOMDP concerns a 3-tuple.
+The first element, $\mdpstate_m$, directly corresponds to the aggregation of the state variables in the epidemiological model, i.e., a tuple,
 \begin{equation}
@@
 
-We parameterize the \added{epidemiological} model using the mean of the posteriors as specified by \citet{abrams2021modelling} (details in SI).
-
-\added{The second element of the tuple consists of the social contact matrix $\hat{C}$ that is currently in place.
+We parameterize the epidemiological model using the mean of the posteriors as specified by \citet{abrams2021modelling} (details in SI).
+
+The second element of the tuple consists of the social contact matrix $\hat{C}$ that is currently in place.
 The reason to incorporate it in the state-space is two-fold.
@@
 As such, during the simulated week, there is a gradual shift from the current $\hat{C}$ to the next one.
-We thus require to know the current $\hat{C}$.}
-
-\added{Secondly, we require the current $\hat{C}$ to determine whether the action changes the social restrictions in place -- and thus consume part of the budget -- or not.}
-
-\added{The third element of the tuple concerns of the budget $\budget$.
+We thus require to know the current $\hat{C}$.
+
+Secondly, we require the current $\hat{C}$ to determine whether the action changes the social restrictions in place -- and thus consume part of the budget -- or not.
+
+The third element of the tuple concerns of the budget $\budget$.
 We note that we incorporate a distinct budget per action-dimension, so $p_{w}, p_{s}$ and $p_{l}$ each have their own budget, rendering $\budget$ a vector $\budget = [b_w, b_s, b_l]$.
-As such, it is possible that, at timestep $t$, the budget for one of the proportional reductions is reduced but not the others.}
+As such, it is possible that, at timestep $t$, the budget for one of the proportional reductions is reduced but not the others.
 
@@
 
-\added{TODO REPHRASE, FROM PCN PAPER: Using neural networks as function approximators in RL comes with many challenges.
+\todo{TODO REPHRASE, FROM PCN PAPER} Using neural networks as function approximators in RL comes with many challenges.
 One of them is that the target (e.g., the optimal action of the policy) is not known in advance --- as opposed to classical supervised learning where the ground-truth target is provided.
-As the behavior of the agent improves over time, the action used as target can change, often leading to hard-to-tune and brittle learners \cite{mnih2015,fu2019}.}
-
-\added{Instead of trying to continuously improve the policy by learning actions that should lead to the highest cumulative reward, PCN flips the problem, by learning actions that should lead to any \emph{desired} cumulative reward (be it high or low).
+As the behavior of the agent improves over time, the action used as target can change, often leading to hard-to-tune and brittle learners \cite{mnih2015,fu2019}.
+
+Instead of trying to continuously improve the policy by learning actions that should lead to the highest cumulative reward, PCN flips the problem, by learning actions that should lead to any \emph{desired} cumulative reward (be it high or low).
 In this way, all past trajectories can be reused for supervision, since their returns are known, as well as the actions needed to reach said returns.
 We can thus train a policy that, conditioned on a desired return, provides the optimal action to reach said return.
-By leveraging the generalization properties of neural networks, we can accumulate incrementally better experience by conditioning on increasingly higher reward-goals}
+By leveraging the generalization properties of neural networks, we can accumulate incrementally better experience by conditioning on increasingly higher reward-goals
 
@@
 However, contrary to classification, the data present in the dataset is not fixed.
-\added{PCN collects data from the trajectories experienced while exploring the environment.
-Thus, the dataset improves over time, as we collect increasingly better trajectories.}
-
-\added{Concretely, the PCN algorithm is a loop that repeatedly executes 3 main steps:
+PCN collects data from the trajectories experienced while exploring the environment.
+Thus, the dataset improves over time, as we collect increasingly better trajectories.
+
+Concretely, the PCN algorithm is a loop that repeatedly executes 3 main steps:
 \begin{enumerate}
@@
 \end{enumerate}
-}
 
@@
 
-\added{Given a trajectory composed of $T$ transitions $(s_0, a_0, \bm{r}_0, s_1), \dots, (s_{T-1}, a_{T-1}, \bm{r}_{T-1}, s_T)$ we can compute, for each transition at timestep $t$, the future discounted return $\bm{R}_t = \sum_{i=t}^T \gamma^i\bm{r}_i$ and the leftover horizon $h_t = T-t$.
+Given a trajectory composed of $T$ transitions $(s_0, a_0, \bm{r}_0, s_1), \dots, (s_{T-1}, a_{T-1}, \bm{r}_{T-1}, s_T)$ we can compute, for each transition at timestep $t$, the future discounted return $\bm{R}_t = \sum_{i=t}^T \gamma^i\bm{r}_i$ and the leftover horizon $h_t = T-t$.
 Since for this trajectory executing action $a_t$ in state $s_t$ resulted in return $\mathbf{R}_t$ in $h_t$ timesteps, we add a datapoint with input $\langle s, \hat{h}, \mathbf{\hat{R}}\rangle = \langle s_t, h_t, \mathbf{R}_t \rangle$ and output $a=a_t$ to the dataset.
-In other words, when the observed return corresponds to the desired return in that state, then $a_t$ is the optimal action to take.}
-
-\added{Since we have a fix-sized dataset, adding this trajectory means we need to remove another one.
+In other words, when the observed return corresponds to the desired return in that state, then $a_t$ is the optimal action to take.
+
+Since we have a fix-sized dataset, adding this trajectory means we need to remove another one.
 Our aim is to keep trajectories that span different parts of the objective space, such that we can explore as many types of trade-offs as possible.
 However, to improve the trade-offs we currently have in our dataset, we need to only keep solutions that are close to our current estimate of the Pareto front.
-Thus, we assign a priority for each trajectory in our dataset that combine two metrics: the \emph{crowding distance}, which measures the distance of a solution with its closest neighbors, and the \emph{l2-norm} with the closest non-dominated solution, which indicates how close the solution is from our current estimate of the Pareto front.}
-
-\added{The dataset is then updated by only keeping the trajectories with the top-$N$ priorities.}
+Thus, we assign a priority for each trajectory in our dataset that combine two metrics: the \emph{crowding distance}, which measures the distance of a solution with its closest neighbors, and the \emph{l2-norm} with the closest non-dominated solution, which indicates how close the solution is from our current estimate of the Pareto front.
+
+The dataset is then updated by only keeping the trajectories with the top-$N$ priorities.
 
@@
 
-\added{In the previous section, we explain how PCN updates the dataset given a trajectory. In this section, we explain how PCN produces said trajectory.}
-
-\added{It is unrealistic to expect PCN to reliably produce trajectories with high-valued desired returns when it has only been trained on datapoints originating from random trajectories.
+In the previous section, we explain how PCN updates the dataset given a trajectory. In this section, we explain how PCN produces said trajectory.
+
+It is unrealistic to expect PCN to reliably produce trajectories with high-valued desired returns when it has only been trained on datapoints originating from random trajectories.
 Rather, we can expect PCN to produce trajectories with returns in the range of the ones from the current training data.
-Therefore, if we obtain trajectories reaching high returns, PCN will be able to confidently return high-return policies.}
-
-\added{PCN leverages the fact that, due to the generalization capabilities of neural networks, the policies obtained from the network will still be reliable even if the desired return is marginally higher than what is present in the training data.
+Therefore, if we obtain trajectories reaching high returns, PCN will be able to confidently return high-return policies.
+
+PCN leverages the fact that, due to the generalization capabilities of neural networks, the policies obtained from the network will still be reliable even if the desired return is marginally higher than what is present in the training data.
 In fact, they will perform similar actions to those in the training data, but lead to a higher return.
-Thus, we incrementally condition the network on better and better returns, in order to obtain trajectories that extend the boundaries of PCN's current coverage set.}
-
-\added{More precisely, we randomly select a non-dominated return $\mathbf{R}_{nd}$ and its corresponding horizon $\hat{h}$ from the dataset.
+Thus, we incrementally condition the network on better and better returns, in order to obtain trajectories that extend the boundaries of PCN's current coverage set.
+
+More precisely, we randomly select a non-dominated return $\mathbf{R}_{nd}$ and its corresponding horizon $\hat{h}$ from the dataset.
 By randomly picking a non-dominated return from the entire coverage set we ensure equal chance of improvement to each part of the objective space.
@@
 This then becomes our desired return $\bm{\hat{R}}$.
-By restricting the improvement to at most $\sigma_o$, $\mathbf{\hat{R}}$ stays in the range of possible achievable returns and, by only modifying one objective at a time, the changes to the network's input compared to the training data are kept at a minimum.}
+By restricting the improvement to at most $\sigma_o$, $\mathbf{\hat{R}}$ stays in the range of possible achievable returns and, by only modifying one objective at a time, the changes to the network's input compared to the training data are kept at a minimum.
 
@@
 Still, due to the high standard deviation of the no-budget, $\mdprewardfn_\text{ARH}$ setting, we do not believe this difference is meaningfull.
-\todo{some kind of meaningfull conclusion}
+\todo{some kind of meaningful conclusion}
 

### 2023-02-09 paper/main.tex
update figure captions

@@
     \caption{Evaluation metrics for the coverage sets comparing hospitalizations with social burden.
-In general training on the ODE results in slightly better coverage sets than on the Binomial model.
+In general, an increase of budget results in a better coverage set.
 Training on infections (ARI) still provides a competitive coverage set in terms of hospitalizations.
@@
 The coverage set is shown in Figure~\ref{fig:pf-binomal-arh-budgets}.
+On the right, we display the coverage set of the best-performing run in terms of hypervolume, for each budget setting.
+On the left, we show an interpolated average of the coverage sets learned by the different runs.
 
@@
     \input{fig/budget_5}
-    \caption{Selection of policies learned by PCN, from most restrictive in terms of social burden (top) to least restrictive (bottom).
+    \caption{Selection of policies learned by $\text{PCN}_{b=5}$, from most restrictive in terms of social burden (top) to least restrictive (bottom).
     }

### 2023-02-09 paper/fig/arh_binomial_ode.tex
SI: ode vs binomial

\begin{tikzpicture}
    \pgfplotsset{
        scale only axis,
        % height=4cm,
        xlabel={total daily new hospitalizations},
        ylabel={total lost contacts},
        width=0.45\textwidth,
        every axis title/.append style={at={(0.45,0.0)}}
    }
    \begin{groupplot}[
        group style={
            group size=2 by 3,
            y descriptions at=edge left,
            x descriptions at=edge bottom,
            vertical sep=0pt,
            horizontal sep=0pt,
        },
        % legend entries={{$\text{PCN}_{b=2}$}, {$\text{PCN}_{b=3}$}, {$\text{PCN}_{b=4}$}, {$\text{PCN}_{b=5}$}, {$\text{PCN}$}, {Fixed}},
    ]
    % interpolated plot for ODE
    \nextgroupplot[mark size=1pt,draw opacity=1,legend entries={{$\text{PCN}_{b=2}$}, {$\text{PCN}_{b=3}$}, {$\text{PCN}_{b=4}$}, {$\text{PCN}_{b=5}$}, {$\text{PCN}$}, {Fixed}},title={ODE model, different budgets}]
    % mean line
    \foreach \b/\c in {2/pc1, 3/pc2, 4/pc3, 5/pc4, {None}/pc5}
    {
        \edef\temp{
            \noexpand\addplot[fill opacity=0.5,\c] table [x=mu_o_1, y=mu_o_5, col sep=comma] {data/cs/arh/ode/interpolated/budget_\b.csv};
        }
        \temp
    }
    % std fillbetween, has to be after means for legend order
    \foreach \b/\c in {2/pc1, 3/pc2, 4/pc3, 5/pc4, {None}/pc5}
    {
        \edef\temp{
            \noexpand\addplot[name path=upper\b, draw=none] table [x=mu_o_1, y expr=\noexpand\thisrow{mu_o_5}+\noexpand\thisrow{std_o_5}, col sep=comma] {data/cs/arh/ode/interpolated/budget_\b.csv};
            \noexpand\addplot[name path=lower\b, draw=none] table [x=mu_o_1, y expr=\noexpand\thisrow{mu_o_5}-\noexpand\thisrow{std_o_5}, col sep=comma] {data/cs/arh/ode/interpolated/budget_\b.csv};
            \noexpand\addplot[\c!15] fill between[of={upper\b} and {lower\b}];
        }
        \temp
    }
    \addplot[only marks,mark=+] table [x=o_0, y=o_1, col sep=comma] {data/cs/cs_fixed.csv};
    % ODE vs Binomial per budget
    \pgfplotsforeachungrouped \b in {2, 3, 4, 5, {None}}
    {
        % mean plots
        \edef\temp{
            \noexpand\nextgroupplot[draw opacity=1,legend entries={Binomial, ODE},title={$\text{PCN}_{b=\b}$}]
            \noexpand\addplot[fill opacity=0.5,pc1] table [x=mu_o_1, y=mu_o_5, col sep=comma] {data/cs/arh/binomial/interpolated/budget_\b.csv};
            \noexpand\addplot[fill opacity=0.5,pc2] table [x=mu_o_1, y=mu_o_5, col sep=comma] {data/cs/arh/ode/interpolated/budget_\b.csv};
            % std plots
            \noexpand\addplot[name path=binomialupper\b, draw=none] table [x=mu_o_1, y expr=\noexpand\thisrow{mu_o_5}+\noexpand\thisrow{std_o_5}, col sep=comma] {data/cs/arh/binomial/interpolated/budget_\b.csv};
            \noexpand\addplot[name path=binomiallower\b, draw=none] table [x=mu_o_1, y expr=\noexpand\thisrow{mu_o_5}-\noexpand\thisrow{std_o_5}, col sep=comma] {data/cs/arh/binomial/interpolated/budget_\b.csv};
            \noexpand\addplot[pc1!15] fill between[of={binomialupper\b} and {binomiallower\b}];

            \noexpand\addplot[name path=odeupper\b, draw=none] table [x=mu_o_1, y expr=\noexpand\thisrow{mu_o_5}+\noexpand\thisrow{std_o_5}, col sep=comma] {data/cs/arh/ode/interpolated/budget_\b.csv};
            \noexpand\addplot[name path=odelower\b, draw=none] table [x=mu_o_1, y expr=\noexpand\thisrow{mu_o_5}-\noexpand\thisrow{std_o_5}, col sep=comma] {data/cs/arh/ode/interpolated/budget_\b.csv};
            \noexpand\addplot[pc2!15] fill between[of={odeupper\b} and {odelower\b}];
        }
        \temp
    }
    \end{groupplot}
\end{tikzpicture}

### 2023-02-09 paper/main.tex
SI: ode vs binomial

@@
 The $I_\varepsilon$ indicators shows that, regardless of the budget, the decision maker will lose at worst a $0.050$ normalized return in any of the objectives.
-On average, it will lose $0.010$ returns.
+On average, it will lose $0.010$ normalized returns.
 Moreover, we emphasize that the learned coverage set contains the non-dominated returns across the whole training procedure.
@@
 \section{Additional results}
+
+\subsection{Comparison of coverage sets learned on ODE and Binomial models}
+\subsection{sec:cs-ode-vs-binomial}
+
+\begin{figure}
+    \centering
+    \input{fig/arh_binomial_ode}
+    \caption{ODE vs Binomial}
+    \label{fig:pf-binomial-ode}
+\end{figure}
+
+The coverage sets displayed in Figure~\ref{fig:pf-binomal-arh-budgets} correspond to PCN trained on the Binomial model.
+To show that PCN copes with the stochasticity of this model, we compare these coverage sets with the ones learned on the ODE model.
+
+Results are shown in Figure~\ref{fig:pf-binomial-ode}.
+First, in a similar fashion as Figure~\ref{fig:pf-binomal-arh-budgets}, we show the interpolated average coverage set for different budget setting, with PCN trained on the ODE model.
+We observe similar trends as for the Binomial model.
+Next, for each budget setting, we compare the interpolated average coverage set of the Binomial setting with the ODE setting.
+We observe similar coverage set, regardless of the budget setting, indicating that PCN is able to cope with stochastic transitions.
+
 \subsection{Learned coverage sets}

### 2023-03-21 paper/main.tex
pieter comments introduction

@@
 \section{Introduction}
-As shown by the COVID-19 pandemic, infectious disease outbreaks represent a major global challenge.
+Infectious disease outbreaks represent a major global challenge \todo{ref}.
 To this end, understanding the complex dynamics that underlie these epidemics is essential.
@@
 Therefore, optimizing on a single objective requires that  these distinct criteria are somehow aggregated into a single metric.
-Generally, during learning, a decision maker will rely on an expert's assistance to manually design such a metric. However, a decision maker may
-be unaware of all possible trade-offs between different solutions, which may leave the decision maker unaware of possible solutions that better reflect their preferences, potentially resulting in sub-optimal performance.
-Furthermore, manually designing such metrics is time consuming, costly and error-prone, as this non-intuitive process requires repetitive and tedious tuning to achieve the desired behaviour~\cite{hayes2021practical}.
-Moreover, taking a single objective approach has several other limiting factors~\cite{hayes2021practical,roijers2013survey}.
-
-To alleviate the challenging process of defining the learning problem explicitly, we can directly learn optimal behaviours by explicitly taking a multi-objective approach.
-By assuming that a decision maker will always prefer solutions for which at least one objective improves, it is possible to learn a set of optimal solutions called the \emph{Pareto front}.
-This enables decision makers to review each solution on the Pareto front before making a decision, thereby being aware of the trade-offs that a solution may imply.
-
-In this work, we investigate the use of \emph{multi-objective reinforcement learning} (MORL) to learn a set of solutions that approximate the Pareto front of multi-objective mitigation strategies.
-We consider the first wave of the Belgian COVID-19 epidemic, which was mitigated by a hard lockdown~\cite{willem2021impact}.
-When the incidence of confirmed cases were steadily dropping, epidemiological experts were asked to investigate strategies to exit from the stringent lockdown which was imposed.
-Here, we consider the epidemiological model developed by \citet{abrams2021modelling} that was constructed to describe the Belgian COVID-19 epidemic, and was fitted to hospitalisation incidence data and serial sero-prevalence data.
+Typically, a decision maker will rely on an expert's assistance to manually design such a metric.
+However, this is a challenging process, as the decision maker is generally not aware of all possible trade-offs, and as such this may result in a metric that exhibits sub-optimal performance.
+Furthermore, manually designing such metrics is time consuming, costly and error-prone, as this non-intuitive process requires repetitive and tedious tuning to achieve the desired behaviour~\cite{roijers2013survey}.
+Moreover, taking a single objective approach has several other limiting factors.
+For example, it reduces the explainability of the learned solution, as we cannot compare the learned behavior with alternatives.
+Moreover, the preferences of the decision maker might change over time, in which case the optimization process needs to be done again~\cite{hayes2021practical}.
+
+This challenging process can be circumvented by taking an explicit multi-objective approach that aims to learn the different trade-offs regarding the considered criteria.
+By assuming that a decision maker will always prefer solutions for which at least one objective improves, it is possible to learn a set of optimal solutions referred to as the \emph{Pareto front}.
+This enables decision makers to review each solution on the Pareto front before making a decision, thereby being aware of the trade-offs that each solution implies.
+
+In this work, we investigate the use of \emph{multi-objective reinforcement learning} (MORL) to learn a set of solutions that approximate the Pareto front of multi-objective epidemic mitigation strategies.
+We consider the first wave of the Belgian COVID-19 epidemic, which was mitigated by a strict lockdown~\cite{willem2021impact}.
+When the incidence of confirmed cases was steadily decreasing, epidemiological experts were tasked to investigate deconfinement strategies, to reduce the severe contact and mobility reductions.
+Here, we consider an epidemiological model that was constructed to describe the Belgian COVID-19 epidemic, and was fitted to hospitalisation incidence data and serial sero-prevalence data~\citet{abrams2021modelling}.
 This model constitutes a stochastic discrete-time age-structured compartmental model that simulates mitigation strategies by varying social distancing parameters concerning school, work and leisure contacts.
-Based on this model, we contribute \momdpname, a novel mutli-objective epidemiological environment, in the form of a multi-objective Markov decision process (MOMDP) \cite{roijers2013survey}.
+Based on this model, we contribute a novel mutli-objective epidemiological reinforcement learning environment~(\momdpname), in the form of a multi-objective Markov decision process (MOMDP) \cite{roijers2013survey}.
 \momdpname\ encapsulates the epidemiological model developed by \citet{abrams2021modelling} to implement state transitions, with an action space that combines a proportional reduction of school, work and leisure contacts at each time step, Furthermore, it defines a reward function based on two objectives: the attack rate (i.e., proportion of the population affected by the pathogen) and the social burden that is induced by the mitigation measures.
 
-To learn and explore the trade-offs between the attack rate and social burden we present a state-of-the-art MORL approach, Pareto Conditioned Networks (PCN)~\cite{reymond2022pcn}.
+To learn and explore the trade-offs between the attack rate and social burden we present a state-of-the-art MORL approach based on Pareto Conditioned Networks (PCN)~\cite{reymond2022pcn}.
 PCN uses a single neural network to learn the policies that belong to the Pareto front.
@@
 
-By evaluating the solution set of mitigation policies returned by PCN, we observe that PCN minimises the social burden in scenarios where hospitalization rates are sufficiently low.
+By evaluating the solution set of mitigation policies learned by PCN, we observe that PCN minimises the social burden in scenarios where hospitalization rates are sufficiently low \todo{update (challenges)}.
 Therefore, in this work we illustrate that multi-objective reinforcement learning can be utilised to provide important insights surrounding the trade-offs between complex mitigation polices in real-world epidemiological models.

### 2023-03-23 paper/main.tex
partial pieter comments

@@
 
+\newcommand{\ltuple}{\langle}
+\newcommand{\rtuple}{\rangle}
+
 % color scheme for plots: plot-color-i
@@
 \section{Background}
+In the following, we use bold notation $\momdpvaluefunction$ to denote a vectorial value, while non-bold notation $\mdpvaluefunction$ for its scalar counterpart.
+
 \subsection{Multi-Objective Reinforcement Learning}
-Many real-world decisions problems often have multiple and possibly conflicting objectives.
+Real-world decisions problems typically consider multiple and possibly conflicting objectives.
 Multi-objective reinforcement learning (MORL) can be used to find optimal solutions for sequential decision making problems with multiple objectives~\cite{hayes2021practical}.
-For MORL, problems are typically modelled as a multi-objective Markov decision process (MOMDP), i.e., a tuple, $\mathcal{M} = (\mdpstatespace, \mdpactionspace, \mdptransition, \mdpdiscount, \momdprewardfn)$, where $\mdpstatespace$, $\mdpactionspace$ denote the state and action spaces respectively, $\mdptransition \colon \mdpstatespace \times \mdpactionspace \times \mdpstatespace  \to \left[ 0, 1 \right]$ denotes a probabilistic transition function, $\mdpdiscount$ is a discount factor determining the importance of future rewards and $\momdprewardfn \colon \mdpstatespace \times \mdpactionspace \times \mdpstatespace \to \mathbb{R}^n$ is an $n$-dimensional vector-valued immediate reward function, where $n$ corresponds to the number of objectives.
-
-While MOMDPs define the actions an agent can take, it does not define which actions are taken in each state.
-We call this the agent's \emph{policy}.
-This policy dictates the agent's behavior.
-More formally, a policy $\mdppolicy$ is a probabilistic function $\mdpstatespace \times \mdpactionspace \to \left[ 0, 1 \right]$ determining the probability, for each state $s \in \mdpstatespace$, to take an action $a \in \mdpactionspace$.
-We measure the performance of $\mdppolicy$ through the expected sum of discounted rewards it achieves.
-This is denoted as the $\momdpvaluefunction$-value:
-
-\begin{equation}
-    \momdpvaluefunction^\mdppolicy = \mathbb{E} \left [ \sum_{t=0}^h \mdpdiscount^t \momdpreward_t |\ \mdppolicy, s_0 \right ],
-\end{equation}
-%
-where $\momdpreward_t = \mdprewardfn(\state_t, \action_t, \state_{t+1})$.
-
-For single-objective RL, i.e., when $n = 1$, the goal is to find the policy $\mdppolicy^{*}$ that maximizes the $\mdpvaluefunction$-value:
+Multi-objective sequential problems are typically modelled as a multi-objective Markov decision process (MOMDP), i.e., a tuple, $\mathcal{M} = \ltuple\mdpstatespace, \mdpactionspace, \mdptransition, \mdpdiscount, \momdprewardfn\rtuple$, where $\mdpstatespace$, $\mdpactionspace$ denote the state and action spaces respectively, $\mdptransition \colon \mdpstatespace \times \mdpactionspace \times \mdpstatespace  \to \left[ 0, 1 \right]$ denotes a probabilistic transition function, $\mdpdiscount$ is a discount factor determining the importance of future rewards and $\momdprewardfn \colon \mdpstatespace \times \mdpactionspace \times \mdpstatespace \to \mathbb{R}^n$ is an $n$-dimensional vector-valued immediate reward function, where $n$ corresponds to the number of objectives.
+
+While MOMDPs define the actions an agent can take, it does not define the agent's behavior, i.e., which actions are taken in each state.
+Given this MOMDP, an agent follows a policy $\mdppolicy$, that expresses the probability to take action $\action \in \mdpactionspace$ when in state $\state \in \mdpstatespace$: $\mdpstatespace \times \mdpactionspace \to \left[ 0, 1 \right]$.
+We measure the performance of $\mdppolicy$ through the expected sum of discounted rewards (denoted the \emph{Value}) it achieves from the initial state $\state_0$ until the end of the decision problem:
+
+\begin{equation}
+    \momdpvaluefunction^\mdppolicy = \mathbb{E} \left [ \sum_{t=0}^h \mdpdiscount^t \momdpreward_t |\ \mdppolicy, \state_0 \right ],
+\end{equation}
+%
+where $\momdpreward_t$ corresponds to the multi-objective reward observed at time $t$, by following policy $\mdppolicy$, given by the reward function $\mdprewardfn(\state_t, \action_t, \state_{t+1})$, where $\state_t$ is the current state, $\action_t$ corresponds to the action that was taken, and $\state_{t+1}$ signifies the state reached by taking action $\action_t$.
+
+For single-objective RL, where $n = 1$, the goal is to find the policy $\mdppolicy^{*}$ that maximizes the $\mdpvaluefunction$-value:
 \begin{equation}
@@
 
-In contrast, in MORL, $n > 1$ which leads to vectorial returns.
+In contrast, for MORL, $n > 1$ which leads to vectorial returns.
 In this case, there can be policies for which, without any additional information, it is impossible to know if one is better than the other.
-For example, we cannot decide which policy between $\mdppolicy_1, \mdppolicy_2$ is optimal if they lead to expected returns (also called V-values) $\momdpvaluefunction^{\mdppolicy_1}=(0,1), \momdpvaluefunction^{\mdppolicy_2}=(1,0)$ respectively.
+For example, it is impossible to decide which policy between $\mdppolicy_1, \mdppolicy_2$ is optimal if both policies lead to expected returns $\momdpvaluefunction^{\mdppolicy_1}=(0,1), \momdpvaluefunction^{\mdppolicy_2}=(1,0)$ respectively.
 We call these solutions \emph{non-dominated}, i.e., solutions for which it is impossible to improve an objective without hampering another.
 The set that contains all the non-dominated solutions of the decision problem is called the \emph{Pareto front} $\mathcal{F}$.
-Our goal is to find the set of policies that lead to all the V-values contained in the Pareto front $\Pi^* = \{\ \pi | \momdpvaluefunction^{\pi} \in \mathcal{F}\}$.
-In general, we call any set of V-values a \emph{solution set}.
-When a solution set contains only  non-dominated V-values, it is referred to as a \emph{coverage set}.
+Our goal is to find the set of policies that lead to all the $\mdpvaluefunction$-values contained in the Pareto front $\Pi^* = \{\ \pi | \momdpvaluefunction^{\pi} \in \mathcal{F}\}$.
+In general, we call any set of $\mdpvaluefunction$-values a \emph{solution set}.
+When a solution set contains only  non-dominated $\mdpvaluefunction$-values, it is referred to as a \emph{coverage set}.
 In the case that no $\mdppolicy$ exists that has a $\momdpvaluefunction^{\pi}$ dominating any of the solutions in a coverage set, then this coverage set is the Pareto front.
@@
 A widely used metric in the literature is called the \emph{hypervolume}~\cite{zitzler2003}.
-This metric evaluates the learned coverage set by computing its volume with respect to a fixed specified reference point.
+This metric evaluates the learned coverage set by computing its volume with respect to a fixed reference point.
 The reference point is taken as a lower bound on the achievable returns so that the volumes are always positive.
@@
 This means that the hypervolume is the highest for the Pareto front.
-One disadvantage of the hypervolume is that it can be difficult to interpret.
+One drawback of the hypervolume is that it can be difficult to interpret.
 For example, when working in high-dimensional objective-spaces, adding or removing a single point can drastically change hypervolume values, especially if the point lies close to an extremum of said space.
 
-To counterpoise these shortcomings, we additionally evaluate our work on a different metric called the $\varepsilon$-indicator $I_\varepsilon$~\cite{zitzler2003}, which measures how close a coverage set is to the Pareto front $\mathcal{F}$.
+To counterpoise these shortcomings, we evaluate our work using an additional metric called the $\varepsilon$-indicator $I_\varepsilon$~\cite{zitzler2003}, which measures how close a coverage set is to the Pareto front $\mathcal{F}$.
 Intuitively, $I_\varepsilon$ shows that any solution of $\mathcal{F}$ is \emph{at most} $\varepsilon$ better with respect to each objective $o$ than the closest solution of the evaluated coverage set:
@@
 
-The main disadvantage of this metric is that we need the true Pareto front to compute it, which in the case of our MOMDP (see Section~\ref{sec:sars-cov2-momdp}) is unknown.
+The main disadvantage of this metric is that we need the true Pareto front to compute it, which is unknown for our MOMDP (see Section~\ref{sec:sars-cov2-momdp}).
 To still gain insights from our learned policies, we approximate the true Pareto front using the non-dominated policies across all runs.
 
-We note that the $\varepsilon$-indicator metric is quite pessimistic, as it measures worst-case performance~\cite{zintgraf2015}, i.e., e.g., it will still report bad performance as long as a single point of the Pareto front is not correctly modeled, even if all the other points are covered.
+We note that the $\varepsilon$-indicator metric is quite pessimistic, as it measures worst-case performance~\cite{zintgraf2015}, i.e., it will still report low performance as long as a single point of the Pareto front is incorrectly modeled, even if all the other points are covered.
 As such, we also use the $I_{\varepsilon-mean}$ \cite{reymond2022pcn} which measures the \emph{average} $\varepsilon$ distance of the solutions in $\mathcal{F}$ with respect to the evaluated coverage set.
@@
 The hypervolume metric (in light blue) measures the volume of all dominated solutions with respect to some reference point (cross).
+The reference point is taken as a lower bound on the achievable returns, which is why it can differ from the origin.
 The $\varepsilon$ metrics first compute the maximum distance between each point in the Pareto front and its closest point in the coverage set ($\varepsilon_i$).
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{COVID-19 model and the \momdpname\  MOMDP}
+\section{COVID-19 model and \momdpname\  MOMDP}
 \label{sec:sars-cov2-momdp}
@@
 \label{sec:sars-cov2-momdp_model}
-As an environment to evaluate non-pharmaceutical interventions, we consider the adapted version of the SEIR compartmental model presented by Abrams et al., that was used to investigate exit strategies in Belgium after the first epidemic wave of SARS-CoV-2~\cite{abrams2021modelling}.
-This model constitutes a discrete-time stochastic model, that considers an age-structured population.
-This model generalises a standard SEIR model\footnote{A standard SEIR model divides the population into four different states, i.e., susceptible, exposed, infectious and recovered individuals.}, extended to capture the different stages of disease spread and history that are associated with SARS-CoV-2 (i.e., pre-symptomatic, asymptomatic, symptomatic with mild symptoms and symptomatic with severe symptoms) and to represent the stages associated with severe disease, i.e., hospitalisation, admission to the intensive care unit (ICU) and death.
+To evaluate non-pharmaceutical interventions, we consider the compartmental model presented by Abrams et al., that was used to investigate exit strategies in Belgium after the first epidemic wave of SARS-CoV-2~\cite{abrams2021modelling}.
+This model concerns a discrete-time stochastic model, that considers an age-structured population.
+The model generalises a standard SEIR model\footnote{A standard SEIR model divides the population into four different states, i.e., susceptible, exposed, infectious and recovered individuals.}, extended to capture the different stages of disease spread and history that are associated with SARS-CoV-2 (i.e., pre-symptomatic, asymptomatic, symptomatic with mild symptoms and symptomatic with severe symptoms) and to represent the stages associated with severe disease, i.e., hospitalisation, admission to the intensive care unit (ICU) and death.
 As we consider an age-structured population, we consider this extended SEIR structure for $\agegroups = 10$ age groups,  i.e., $[0-10), [10-20), [20-30), [30-40), [40-50), [50-60), [60-70), [70-80), [80-90), [90,100+)$.
-Contacts of the different age-groups, which impact the propagation rate of the epidemic, are modeled using social contact matrices (SCMs).
-We define a SCM for 6 different social environments: $\chome, \cwork, \ctransport, \cschool, \cleisure, \cother$ for the home, work, transport, school, leisure, other environments respectively.
+Contacts of the different age-groups, which impact the propagation rate of the epidemic, are modeled using social contact matrices.
+We define a social contact matrix for 6 different social environments: $\chome, \cwork, \ctransport, \cschool, \cleisure, \cother$, for the home, work, transport, school, leisure, other environments respectively.
 The model is described by a set of ordinary differential equations (ODEs), as described in SI.
@@
 \label{sec:interventions}
-In order to model different types of interventions, we follow \citet{abrams2021modelling}.
-Firstly, in order to consider distinct exit scenarios, we modulate the SCMs to reflect a contact reduction in a particular age group.
+To model different types of interventions, we follow the contact reduction scheme presented by \citet{abrams2021modelling}.
+Firstly, to consider distinct exit scenarios, we modulate the social contact matrices to reflect a contact reduction in a particular age group.
+
+We consider a contact reduction function that imposes a proportional reduction of work (including transport) $p_w$, school $p_s$ and leisure $p_l$ contacts, which is implemented as a linear combination of social contact matrices:
+\begin{equation}
+\label{eq:scm}
+    \hat{C}(p_w, p_s, p_l) = \chome + p_w (\cwork + \ctransport) + p_s \cschool + p_l (\cleisure + \cother)
+\end{equation}
+
+We denote $\hat{C}_t$ the social contact matrix at timestep $t$, resulting from the reduction function $\hat{C}$.
 Secondly, we assume that compliance to the interventions is gradual and model this using a logistic compliance function (see details in SI).
-We consider a contact reduction function that imposes a proportional reduction of work (including transport) $p_w$, school $p_s$ and leisure $p_l$ contacts, which is implemented as a linear combination of SCMs:
-\begin{equation}
-\label{eq:scm}
-    \hat{C} = \chome + p_w (\cwork + \ctransport) + p_s \cschool + p_l (\cleisure + \cother)
-\end{equation}
 
@@
 In order to apply multi-objective reinforcement learning, we construct the \momdpname\ MOMDP based on the epidemiological model introduced in Section~\ref{sec:sars-cov2-momdp_model} and graphically depicted in Figure~\ref{fig:ode_mdel}.
-
-\paragraph{Action-space:} Our actions concern the installment of a social contact matrix $\hat{C}$, with a particular reduction (see Section~\ref{sec:interventions}).
+Moreover, we consider a finite-horizon setting where we simulate the compartment model for a fixed number of weeks.
+
+\paragraph{Action-space:} Our actions concern the installment of a social contact matrix with a particular reduction resulting from the reduction function $\hat{C}$ (see Section~\ref{sec:interventions}).
 To this end, we use the proportional reduction parameters $p_w, p_s, p_l$ defined in Section~\ref{sec:interventions}.
-Thus, each $\mdpaction \in \mdpactionspace$ is a 3-dimensional continuous vector in $[0,1]^3$ (i.e., $\mdpaction = [p_{w}, p_{s}, p_{l}]$) which impacts the SCM according to Equation~\ref{eq:scm}.
-
-\paragraph{Transition function:} The model defined by \citet{abrams2021modelling} utilises a model transition probability $M(\mdpstate_m' \mid \mdpstate_m, \hat{C})$ (see SI for details on $M$), where $\hat{C}$ the currently installed SCM, that progresses the epidemiological model in one timestep.
+Thus, each $\mdpaction \in \mdpactionspace$ is a 3-dimensional continuous vector in $[0,1]^3$ (i.e., $\mdpaction = [p_{w}, p_{s}, p_{l}]$) which impacts the social contact matrix according to Equation~\ref{eq:scm}.
+
+\paragraph{Transition function:} The model defined by \citet{abrams2021modelling} utilises a model transition probability $M(\mdpstate_m' \mid \mdpstate_m, \hat{C}, p_{w}, p_{s}, p_{l})$ (see SI for details on $M$), where $\hat{C}(p_{w}, p_{s}, p_{l})$ returns the currently installed social contact matrix, that progresses the epidemiological model in one timestep.
 We use this function as the transition function in \momdpname.
@@
 In a classical MDP, executing an action $\action_t$ in any state $\state_t$ leads to a next state $\state_{t+1}$ according to the transition function $\mdptransition$.
-At every timestep $t$, the policy is free to choose the action to perform.
+At every timestep $t$, the agent is free to choose the action to perform.
 In our case, this potentially results in a different restriction $[p_{w}, p_{s}, p_{l}]$ every week.
-However, we argue that in the context of mitigation policies, consistency is key, and policies that impose changes too frequently will be hard to follow.
-
-In order to learn realistic and consistent mitigation policies, we incorporate a \emph{budget} on the number of times a policy can change its actions until the terminal state of the MOMDP.
-Concretely, when the action changes, i.e., if the social restriction proposed by the policy is different from the current one in place, we reduce the budget by one.
+However, we argue that in the context of mitigation policies, consistency is important and policies that impose changes too frequently will be hard to adhere to.
+
+In order to learn realistic and consistent mitigation policies, we introduce a \emph{budget} regarding the number of times a policy can change its actions until the terminal state of the MOMDP is reached.
+Concretely, when the action changes, i.e., if the social restriction proposed by the policy is different from the one that is currently in place, we reduce the budget by one.
 We only allow action changes as long as there is budget left.
-We note that we can simulate a no-limit budget setting by setting the budget to the length of the episode.
-
-Finally, for each timestep $t$, our transition function $\mdptransition$ uses the model transition probability $M$ to simulate the model for one week, using $\hat{C}$ obtained from $\mdpaction_t$.
+We note that we can mimic a no-limit budget setting by choosing a budget that corresponds to the horizon of the environment.
+
+Finally, for each timestep $t$, our transition function $\mdptransition$ uses the model transition probability $M$ to simulate the model for one week, using $\hat{C}_t$ obtained from $\mdpaction_t$.
 
@@
 \begin{equation}
-    \{S_k, E_k, I_k^{presym}, I_k^{asym}, I_k^{mild}, I_k^{sev}, I_k^{hosp}, I_k^{icu}, H^{new}_k, D_k, R_k\},
+    \ltuple S_k, E_k, I_k^{presym}, I_k^{asym}, I_k^{mild}, I_k^{sev}, I_k^{hosp}, I_k^{icu}, H^{new}_k, D_k, R_k \rtuple ,
 \end{equation}
@@
 for each age group $k \in \{1, \ldots, \agegroups\}$, where $S$ encodes the members of the population who are susceptible to infection and $E$ encodes the members of the population who have been exposed to COVID-19.
-Moreover, $I^{presym}$, $I^{asym}$, $I^{mild}$, $I^{hosp}$, $I^{icu}$ are the members of the population infected with COVID-19 and are, respectively, presymptomatic, asymptomatic, have mild symptoms, are hospitalised, or are in the ICU.
+Moreover, $I^{presym}$, $I^{asym}$, $I^{mild}$, $I^{hosp}$, $I^{icu}$ denote the members of the population infected with COVID-19 and are, respectively, presymptomatic, asymptomatic, mildly symptomatic, hospitalised, or in the ICU.
 Finally, $H^{new}_k$ represents the number of newly hospitalised individuals in age group $k$.
@@
 
-The second element of the tuple consists of the social contact matrix $\hat{C}$ that is currently in place.
+The second element of the tuple consists of the social contact matrix $\hat{C}_t$ that is currently in place.
 The reason to incorporate it in the state-space is two-fold.
 First, \citet{abrams2021modelling} define a compliance function, simulating the time people need to get used to the new rules set in place.
-As such, during the simulated week, there is a gradual shift from the current $\hat{C}$ to the next one.
-We thus require to know the current $\hat{C}$.
-
-Secondly, we require the current $\hat{C}$ to determine whether the action changes the social restrictions in place -- and thus consume part of the budget -- or not.
-
-The third element of the tuple concerns of the budget $\budget$.
-We note that we incorporate a distinct budget per action-dimension, so $p_{w}, p_{s}$ and $p_{l}$ each have their own budget, rendering $\budget$ a vector $\budget = [b_w, b_s, b_l]$.
+As such, during the simulated week, there is a gradual shift from the current $\hat{C}_t$ to the new social contact matrix, $\hat{C}_{t+1}$.
+As such, we need to maintain the current $\hat{C}_t$, to maintain a Markovian environment.
+
+Secondly, we require the current $\hat{C}_t$ to determine whether the action changes the social restrictions in place, and thus consume part of the budget.
+
+The third element of the tuple concerns the budget $\budget$.
+We incorporate a distinct budget per action-dimension, so $p_{w}, p_{s}$ and $p_{l}$ each have their own budget, resulting in a vector $\budget = [b_w, b_s, b_l]$.
 As such, it is possible that, at timestep $t$, the budget for one of the proportional reductions is reduced but not the others.
@@
 
-\paragraph{Reward function:} We define a vectorial reward function which considers multiple objectives: attack rate (i.e., infections, hospitalisations) and the social burden imposed by the interventions on the population.
-
-The attack rate in terms of infections is defined as the difference in susceptibles from the current state and next state~\cite{libin2020}.
+\paragraph{Reward function:} We define a vectorial reward function which considers multiple objectives: attack rate (i.e., infections and hospitalisations) and the social burden imposed by the interventions on the population.
+
+The attack rate in terms of infections is defined as the difference in susceptibles from the current state to the next state~\cite{libin2020}.
 Since this is a cost that needs to be minimized, we defined the corresponding reward function as the negative attack rate:
@@
 To determine missed contacts for the entire population, we apply the difference matrix to the population sizes of the respective age groups that are currently uninfected (i.e., susceptible and recovered individuals).
-Therefore we define the social burden reward function $\mdprewardfn_{\text{SB}}$, as follows:
+Formally, we define the social burden reward function $\mdprewardfn_{\text{SB}}$, as follows:
 
@@
 where $S_k(\mdpstate)$ represents the number of susceptible individuals in age group $k$ in state $\mdpstate$ and $R_k$ represents the number of recovered individuals in age group $k$ in state $\mdpstate$.
-In Section~\ref{sec:experiments}, we optimize PCN on two different variants for the multi-objective reward function: $[\mdprewardfn_\text{ARH}, \mdprewardfn_\text{SB}]$ and $[\mdprewardfn_\text{ARI}, \mdprewardfn_\text{SB}]$.
+In Section~\ref{sec:experiments}, we optimize PCN on two different variants for the multi-objective reward function: $[\mdprewardfn_\text{ARH}, \mdprewardfn_\text{SB}]$ and $[\mdprewardfn_\text{ARI}, \mdprewardfn_\text{SB}]$, to study the impact of these distinct attack rate quantities.
 
@@
 In multi-objective optimization, the set of optimal policies can grow exponentially in the number of objectives.
-Thus, recovering them all is an expensive process and requires an exhaustive exploration of the complete state space.
+Thus, recovering them all is a computationally expensive process and requires an exhaustive exploration of the complete state space.
 To address this problem, we use Pareto Conditioned Networks~(PCN), a method that uses a single neural network to encompass all non-dominated policies \cite{reymond2022pcn}.
@@
 
-\todo{TODO REPHRASE, FROM PCN PAPER} Using neural networks as function approximators in RL comes with many challenges.
-One of them is that the target (e.g., the optimal action of the policy) is not known in advance --- as opposed to classical supervised learning where the ground-truth target is provided.
+One major challenge of using neural networks as function approximators in RL is that the target (e.g., the optimal action of the policy) is not known in advance, as opposed to classical supervised learning where the ground-truth target is provided.
 As the behavior of the agent improves over time, the action used as target can change, often leading to hard-to-tune and brittle learners \cite{mnih2015,fu2019}.
@@
 
-PCN uses a single neural network that takes a tuple $\langle \mdpstate, \hat{h}, \mathbf{\hat{R}}  \rangle$ as input.
+Instead of using a neural network that takes a state $\mdpstate$ as input, PCN uses a single neural network that takes a tuple $\ltuple \mdpstate, \hat{h}, \mathbf{\hat{R}} \rtuple$ as input.
 $\mathbf{\hat{R}}$ represents the \emph{desired return} of the decision maker, i.e.
@@
 Each output represents the confidence the network has that, by taking the corresponding action in $\mdpstate$, $\mathbf{\hat{R}}$ will be reached in $\hat{h}$ timesteps.
-We can draw an analogy with a classification problem where the network should learn to classify $(\mdpstate, \hat{h}, \mathbf{\hat{R}})$ to its corresponding label $\action_i$.
+We can draw an analogy with a classification problem where the network should learn to classify $\ltuple \mdpstate, \hat{h}, \mathbf{\hat{R}}\rtuple$ to its corresponding label $\action_i$.
 
@@
 
-Concretely, the PCN algorithm is a loop that repeatedly executes 3 main steps:
+Concretely, the PCN algorithm implements an iterative process that repeatedly executes 3 main steps:
 \begin{enumerate}
     \item Use the current policy to interact with the environment and generate trajectories (see Section~\ref{sec:pcn-policy-exploration}).
-    \item Update the dataset to incorporate these new trajectories and prun the least interesting trajectories in terms of returns (see Section~\ref{sec:pcn-building-dataset}).
+Initially, use a random policy to fill the dataset with random trajectories.
+    \item Update the dataset to incorporate these new trajectories and prune the least interesting trajectories in terms of returns (see Section~\ref{sec:pcn-building-dataset}).
     \item Update the policy by training it on the updated dataset (see Section~\ref{sec:pcn-continuous}).
@@
 
-Given a trajectory composed of $T$ transitions $(s_0, a_0, \bm{r}_0, s_1), \dots, (s_{T-1}, a_{T-1}, \bm{r}_{T-1}, s_T)$ we can compute, for each transition at timestep $t$, the future discounted return $\bm{R}_t = \sum_{i=t}^T \gamma^i\bm{r}_i$ and the leftover horizon $h_t = T-t$.
+Given a trajectory composed of $T$ transitions $\ltuple s_0, a_0, \bm{r}_0, s_1\rtuple, \dots, \ltuple s_{T-1}, a_{T-1}, \bm{r}_{T-1}, s_T\rtuple$ we can compute, for each transition at timestep $t$, the future discounted return $\bm{R}_t = \sum_{i=t}^T \gamma^i\bm{r}_i$ and the leftover horizon $h_t = T-t$.
 Since for this trajectory executing action $a_t$ in state $s_t$ resulted in return $\mathbf{R}_t$ in $h_t$ timesteps, we add a datapoint with input $\langle s, \hat{h}, \mathbf{\hat{R}}\rangle = \langle s_t, h_t, \mathbf{R}_t \rangle$ and output $a=a_t$ to the dataset.
@@
 However, to improve the trade-offs we currently have in our dataset, we need to only keep solutions that are close to our current estimate of the Pareto front.
-Thus, we assign a priority for each trajectory in our dataset that combine two metrics: the \emph{crowding distance}, which measures the distance of a solution with its closest neighbors, and the \emph{l2-norm} with the closest non-dominated solution, which indicates how close the solution is from our current estimate of the Pareto front.
-
-The dataset is then updated by only keeping the trajectories with the top-$N$ priorities.
+Thus, we assign a priority for each trajectory in our dataset that combines two metrics: the \emph{crowding distance}, which measures the distance of a solution with its closest neighbors, and the \emph{$l_2$-norm} with the closest non-dominated solution, which indicates how close the solution is from our current estimate of the Pareto front.
+
+The dataset is then updated by only keeping the trajectories with the top-$N$ priorities, where $N$ is the size of the dataset.
 
@@
 It is unrealistic to expect PCN to reliably produce trajectories with high-valued desired returns when it has only been trained on datapoints originating from random trajectories.
-Rather, we can expect PCN to produce trajectories with returns in the range of the ones from the current training data.
+Rather, we expect PCN to produce trajectories with returns in the range of the ones from the current training data.
 Therefore, if we obtain trajectories reaching high returns, PCN will be able to confidently return high-return policies.
@@
 In fact, they will perform similar actions to those in the training data, but lead to a higher return.
-Thus, we incrementally condition the network on better and better returns, in order to obtain trajectories that extend the boundaries of PCN's current coverage set.
+Thus, we incrementally condition the network on increasingly higher returns, to obtain trajectories that extend the boundaries of PCN's current coverage set.
 
@@
 By randomly picking a non-dominated return from the entire coverage set we ensure equal chance of improvement to each part of the objective space.
-To generate trajectories that improve upon $\mathbf{R}_{nd}$, we randomly select an objective $o$ and increase $R_{nd,o}$ by a sample from the uniform distribution $U(0, \sigma_o)$, where $\sigma_o$ is the standard deviation for the selected objective, using all non-dominated returns from the trajectories in the dataset.
+To generate trajectories that improve upon $\mathbf{R}_{nd}$, we randomly select an objective $o$ and increase $R_{nd,o}$ by a sample from the uniform distribution $\mathcal{U}(0, \sigma_o)$, where $\sigma_o$ is the standard deviation for the selected objective, using all non-dominated returns from the trajectories in the dataset.
 This then becomes our desired return $\bm{\hat{R}}$.
@@
 Transitions $x= \langle \mdpstate_t, h_t, \momdpreturn_t \rangle, y=\action_t$ are sampled from the dataset, and the ground-truth output $y$ is compared with the predicted output $\hat{y}=\pi(\mdpstate_t, h_t, \momdpreturn_t)$.
-The predictor (i.e., the policy) is then updated using the cross-entropy loss function:
+The predictor (i.e., the policy) is then updated using the cross-entropy loss function \cite{shore1980axiomatic}:
 \begin{equation}
@@
 While the original PCN algorithm is designed for MOMDPs with discrete actions-spaces, the problem we tackle (see Section~\ref{sec:sars-cov2-momdp}) is defined in terms of a continuous action-space.
-We thus extend PCN for the continuous action-space setting.
-First, we change the output of the neural network such that there is a single output value for each dimension of the action-space.
-Since the actions should be bound in the domain of possible actions ($[0,1]$ in the case of \momdpname, see Section~\ref{sec:sars-cov2-momdp}), we apply a tanh non-linearity function on this output.
-Second, the problem becomes a regression problem instead of a classification problem, as the labeled dataset uses continuous labels $y = \mdpaction_t$ instead of categories.
+We thus extend PCN to the continuous action-space setting.
+We change the output of the neural network such that there is a single output value for each dimension of the action-space.
+Since the actions should be bound in the domain of possible actions ($[0,1]$ in the case of \momdpname, see Section~\ref{sec:sars-cov2-momdp}), we apply a tanh non-linearity function on each output.
+As such, we have a regression problem instead of a classification problem, as the labeled dataset now uses continuous labels $y = \mdpaction_t$ instead of categories.
 We thus use a Mean Squared Error (MSE) loss to update the policy:
@@
 However, PCN improves its policy through exploration, by continuously updating its dataset with better trajectories.
-Thus, at training time, we follow \cite{lillicrap2015continuous} and use a stochastic policy by adding random noise sampled from a Normal distribution to the action:
+Thus, at training time, we use a stochastic policy by adding random noise to the action \cite{lillicrap2015continuous}:
 \begin{equation}
@@
 
-where $\eta$ is a hyper-parameter defining the magnitude of noise to be added.
+where $\mathcal{N}$ is the standard Normal distribution and $\eta$ is a hyper-parameter defining the magnitude of noise to be added.
 
@@
 It assumes that reenacting a transition from the dataset leads to the same episodic return.
-When the transition function $\mathcal{T}$ of the MOMDP is deterministic, the whole trajectory can be faithfully reenacted, which guarantees that we obtain the same return.
+When the transition function $\mathcal{T}$ of the MOMDP is deterministic, the whole trajectory can be faithfully reenacted, which guarantees the same return.
 Combined with the fact that PCN's policy is deterministic at execution time, conditioning the policy on a target episodic return is equivalent to conditioning it on the V-value $\mathbf{V}$.
 
-However, when $\mdptransition$ is stochastic we lose this guarantee.
-We cope with this issue by adding small random noise to $\mathbb{R}_t$ when performing gradient descent, which reduces the risk of overfitting~\cite{zur2009noise}.
-Moreover, although the Binomial model of \momdpname\ is stochastic, the possible next-states resulting from a state-action pair are similar to each other.
+However, when $\mdptransition$ is stochastic this can no longer be guaranteed.
+To mitigate this, we add small random noise to $\mathbb{R}_t$ when performing gradient descent, which reduces the risk of overfitting~\cite{zur2009noise}.
+Moreover, while the Binomial \momdpname\ model is stochastic, the possible next-states resulting from a state-action pair are similar to each other.
 This allows PCN to compensate if $\momdpreward_t = \momdprewardfn(\mdpstate_t, \mdpaction_t, \mdpstate_{t+1})$ is somewhat worse than expected.
-This is confirmed by our experiments (see Section~\ref{sec:experiments}), where the coverage sets learned by PCN on the ODE model and on the Binomial model are similar.
+This is confirmed in our experiments (see Section~\ref{sec:experiments}), where the coverage sets learned by PCN on the ODE model and on the Binomial model are similar.
 

### 2023-03-23 paper/main.tex
include ode into main

@@
 % the 6 social contact matrices
+\newcommand{\ctotal}{C}
 \newcommand{\chome}{C_{\text{home}}}
@@
 %other params
+\newcommand{\proportionalityfactor}{q}
+\newcommand{\agegroup}{k}
 \newcommand{\agegroups}{K}
@@
 The model generalises a standard SEIR model\footnote{A standard SEIR model divides the population into four different states, i.e., susceptible, exposed, infectious and recovered individuals.}, extended to capture the different stages of disease spread and history that are associated with SARS-CoV-2 (i.e., pre-symptomatic, asymptomatic, symptomatic with mild symptoms and symptomatic with severe symptoms) and to represent the stages associated with severe disease, i.e., hospitalisation, admission to the intensive care unit (ICU) and death.
+
+A visual representation of the model is depicted in Fig~\ref{fig:ode_mdel}. It shows the different compartments, as well as the flow rates at which individuals move from one compartment to another.
+
+These flow rates are defined by a set of ordinary differential equations, which are outlined as follows:
+
+\begin{align*}
+\frac{d\textbf{S}(t)}{dt} & = -\lambda(t)(\textbf{S})(t), \\
+%
+\frac{d\textbf{E}(t)}{dt} & = \lambda(t)(\textbf{S})(t) - \gamma \textbf{E}(t), \\
+%
+\frac{d\textbf{I}^{presym}(t)}{dt} & = \gamma \textbf{E}(t) - \theta \textbf{I}^{presym}(t), \\
+%
+\frac{d\textbf{I}^{asym}(t)}{dt} & = \theta p \textbf{I}^{presym}(t) - \delta_{1} \textbf{I}^{asym}(t), \\
+%
+\frac{d\textbf{I}^{mild}(t)}{dt} & = \theta (1 - p) \textbf{I}^{presym}(t) - \{ \psi + \omega_{2} \} \textbf{I}^{mild}(t), \\
+%
+\frac{d\textbf{I}^{sev}(t)}{dt} & = \psi \textbf{I}^{mild}(t) - \omega \textbf{I}^{sev}(t), \\
+%
+\frac{d\textbf{I}^{hosp}(t)}{dt} & = \phi_{1} \omega \textbf{I}^{sev}(t) - \{ \delta_{3} + \tau_{1} \} \textbf{I}^{hosp}(t), \\
+%
+\frac{d\textbf{I}^{icu}(t)}{dt} & = (1 - \phi_1) \omega \textbf{I}^{sev}(t) - \{ \delta_{4} + \tau_{2} \} \textbf{I}^{icu}(t), \\
+%
+\frac{d\textbf{D}(t)}{dt} & = \tau_{1} \textbf{I}^{hosp}(t) - \tau_{2} \textbf{I}^{icu}(t), \\
+%
+\frac{d\textbf{R}(t)}{dt} & = \delta_{1} \textbf{I}^{asym}(t) + \delta_{2} \textbf{I}^{mild}(t) + \delta_{3} \textbf{I}^{hosp}(t) + \delta_{4} \textbf{I}^{icu}(t)
+\end{align*}
+%
+where, for example, $\textbf{S} = (S_{1}(t), S_{2}(t), ..., S_{\agegroup}(t))^{T}$ is the vector representing the susceptible members of the population of each age group $\agegroup$ at time $t$.
+
+In this set of ordinary differential equations, each state variable represents a vector over all age groups for a particular compartment at time $t$.
+Infection dynamics are governed by an age-specific force of infection $\foi$:
+
+\begin{equation}
+    \foi(\agegroup,t) = \sum_{\agegroup'=1}^{\agegroups}\beta(\agegroup,\agegroup')I_{\agegroup'}(t),
+\end{equation}
+%
+where $\agegroup$ is the respective age group of a total of $\agegroups$ age groups, and $\beta(\agegroup,\agegroup')$ is the time-invariant transmission rate that encodes the average per capita rate at which an infectious individual in age group $\agegroup$ makes an effective contact with a susceptible individual in age group $\agegroup'$, per unit of time.
+
 As we consider an age-structured population, we consider this extended SEIR structure for $\agegroups = 10$ age groups,  i.e., $[0-10), [10-20), [20-30), [30-40), [40-50), [50-60), [60-70), [70-80), [80-90), [90,100+)$.
@@
 We define a social contact matrix for 6 different social environments: $\chome, \cwork, \ctransport, \cschool, \cleisure, \cother$, for the home, work, transport, school, leisure, other environments respectively.
-The model is described by a set of ordinary differential equations (ODEs), as described in SI.
-By formulating this set of differential equations a chain-binomial process (as detailed in SI) we can obtain stochastic trajectories from this model~\cite{abrams2021modelling}.
+The social contact matrix across all social environments is defined as:
+
+\begin{equation}
+    \ctotal = \chome + \cwork + \ctransport + \cschool + \cleisure + \cother
+\end{equation}
+
+Under the social contact hypothesis \cite{wallinga2006using}, we have that:
+%
+\begin{equation}
+\beta(\agegroup,\agegroup') = \proportionalityfactor \cdot \ctotal(\agegroup,\agegroup'),
+\end{equation}
+%
+where $\proportionalityfactor$ is a proportionality factor.
+
+Following \citet{abrams2021modelling}, we rely on distinct social contact matrices for symptomatic and asymptomatic individuals, respectively $\ctotal_{s}$ and $\ctotal_{a}$.
+Therefore, it is possible to define the transmission rates for both symptomatic and asymptomatic individuals as follows:
+
+\begin{equation}
+\boldsymbol{\beta}_{s}(\agegroup, \agegroup) = \proportionalityfactor_{s} \cdot C_{s}(\agegroup, \agegroup'),    
+\end{equation}
+and
+\begin{equation}
+\boldsymbol{\beta}_{a}(\agegroup, \agegroup') = \proportionalityfactor_{a} \cdot C_{a}(\agegroup, \agegroup').  
+\end{equation}
+
+Moreover, the age-dependent force of infection can be defined as follows:
+
+\begin{equation}
+    \boldsymbol{\lambda}(t) = \boldsymbol{\beta}_{a} \times \{ \textbf{I}^{presym}(t) + \textbf{I}^{asym}(t) \} +
+    \boldsymbol{\beta}_{s} \times \{ \textbf{I}^{mild}(t) + \textbf{I}^{sev}(t) \},
+\end{equation}
+%
+where $\boldsymbol{\lambda}(t) = (\lambda(1, t), \lambda(1, t), ..., \lambda(\agegroups, t))$.
+For all further information about the different compartments and parameters please refer to the work of \citet{abrams2021modelling}.
+
+By formulating the set of differential equations defined above, as a chain-binomial, we can obtain stochastic trajectories from this model \cite{bailey1975mathematical}.
+A chain-binomial model assumes a stochastic model where infected individuals are generated by some underlying probability distribution.
+For the stochastic model we consider a time interval $(t, t +h]$, where $h$ is defined as the length between two consecutive time points.
+Similar to \citet{abrams2021modelling}, in this work we set $h = \frac{1}{24}$.
+
+\todo{epi friends, please check} Time- and age-dependent behavioral changes introduce substantial uncertainty regarding the outcome of an outbreak.
+To evaluate intervention strategies that modulate such behavioral changes, the use of stochastic epidemiological models is warranted \cite{abrams2021modelling}.
+Moreover, the effect of stochasticity is most pronounced when studying the initial growth of an epidemic \cite{britton2009epidemic}, be it during its emergence, or when implementing deconfinement strategies.
+
+\todo{This is why we use the binomial model in our analysis.}
 
@@
 To mitigate this, we add small random noise to $\mathbb{R}_t$ when performing gradient descent, which reduces the risk of overfitting~\cite{zur2009noise}.
-Moreover, while the Binomial \momdpname\ model is stochastic, the possible next-states resulting from a state-action pair are similar to each other.
-This allows PCN to compensate if $\momdpreward_t = \momdprewardfn(\mdpstate_t, \mdpaction_t, \mdpstate_{t+1})$ is somewhat worse than expected.
-This is confirmed in our experiments (see Section~\ref{sec:experiments}), where the coverage sets learned by PCN on the ODE model and on the Binomial model are similar.
+Moreover, while the \momdpname\ model is stochastic, the variation is entirely due to the sampling of the binomial distributions in the binomial-chain.
+While this variation accumulates over time, the time window we consider for each timestep (i.e., one week) is short enough that the accumulation stays limited.
+Thus, the possible next-states resulting from a state-action pair are similar to each other.
+This allows PCN to compensate if $\momdpreward_t = \momdprewardfn(\mdpstate_t, \mdpaction_t, \mdpstate_{t+1})$ is worse than expected.
+
+Although we use a stochastic model to cope with the uncertainty of the outcome of the outbreak, it is possible to deterministically evaluate the set of ordinary differential equations that define the model.
+We assess the validity of our approach by executing PCN on this deterministic variant, and observing similar performance as with the \momdpname\ model.
+\todo{Results are available in SI}.
 

### 2023-03-24 paper/main.tex
pieter comments

@@
 Our goal is to use PCN to learn deconfinment strategies in the \momdpname\ environment.
-We aim to learn policies that balance between the epidemiological objective of minimising the attack rate and the social burden experienced by the population.
-To this end we consider two cases for the vectorial reward functions $[\mdprewardfn_\text{ARH}, \mdprewardfn_\text{SB}]$ and $[\mdprewardfn_\text{ARI}, \mdprewardfn_\text{SB}]$, to learn and analyse policies under different targets with respect to the considered attack rate.
-
-To conduct this analysis, we evaluate our extension of PCN for continuous action-spaces on both the Binomial model.
-\todo{reason for binomial}.
+We aim to learn policies that balance between the epidemiological objective of minimising the attack rate (i.e., $\mdprewardfn_\text{ARH}$ for hospitalisation and $\mdprewardfn_\text{ARI}$ for infection) and the social burden (i.e., $\mdprewardfn_\text{SB}$) experienced by the population population due to the implement mitigation measures.
+To this end, we consider two cases for the vectorial reward functions $[\mdprewardfn_\text{ARH}, \mdprewardfn_\text{SB}]$ and $[\mdprewardfn_\text{ARI}, \mdprewardfn_\text{SB}]$, to learn and analyse policies under different targets with respect to the considered attack rate.
+
+To conduct this analysis, we apply our extension of PCN for continuous action-spaces on the \momdpname\ model.
 As explained in Section~\ref{sec:pcn-stochastic}, we extend PCN for environments with stochastic transitions.
-While all results in this Section concern the Binomial model, we observe similar performance on the ODE model, showing that our extension does resolve the stochasticity issue.
-These results are available in the SI.
 
 Conform to \citet{abrams2021modelling}, the simulation starts on the 1st of March 2020, by seeding a number of infections in the population.
-Two weeks later, on the 14th of March, the Belgian government initiated a full lockdown of the country.
+Two weeks later, on the 14th of March, the Belgian government initiated a full lockdown.
 This is implemented by fixing the actions $p_w, p_s, p_l$ to $0.2, 0, 0.1$ respectively.
 This lockdown ended on the 4th of May 2020, at which point the government decided on a multi-phase exit strategy to incrementally reduce teleworking, reopen schools and allow leisure activities, such as the re-opening of bars and the cultural sector.
-It is from this day onward that PCN aims to learn policies for diverse exit strategies, compromising between the total number of daily new hospitalizations and the total number of contacts lost as a proxy for social burden.
+It is from this day onward that PCN aims to learn policies for diverse exit strategies, compromising between the total number of daily hospitalizations and the total number of contacts lost as a proxy for social burden.
 The simulation lasts throughout the school holidays (from 01/07/2020 to 31/08/2020).
@@
 
-To evaluate the quality of the policies discovered by PCN, we compare it to a baseline.
-The baseline consists of a set of 100 fixed policies, that iterate over all the possible social restriction levels, with values ranging from 0 to 1.
+To evaluate the quality of the policies learned by PCN, we compare PCN to a baseline.
+This baseline consists of a set of 100 fixed policies, that iterate over all the possible social restriction levels, with values ranging between 0 and 1.
 Concretely, each policy uses a fixed proportional reduction $p_w=p_l=p_s=u, u\sim\mathcal{U}(0, 1)$ throughout the episode.
@@
     \input{fig/arh_binomial_both}
-    \caption{The Pareto front of policies discovered by PCN using the Binomial model, showing the different compromises between the number of hospitalizations and the number of lost contacts.
-As the budget increases, so does the coverage set learnt by PCN.
+    \caption{The Pareto front of policies discovered by PCN using \momdpname, showing the different compromises between the number of hospitalizations and the number of lost contacts.
+On the right, we show, for each budget setting (colored, subscript indicates budget) the coverage set learned by the best performing run.
+On the left, we show an interpolated average of the coverage sets learned by the different runs, with the shaded regions corresponding to the standard deviation.
+For comparison, the basline is displayed on both plots (in black).
+As the budget increases, so does the size of the coverage set learnt by PCN.
 Changes are most noticeable in the less restrictive trade-offs in terms of social burden.}
@@
 
-We learn a coverage set that ranges from imposing minimal restrictions to enforcing many restrictions (see Figure~\ref{fig:policy-execution}).
-The coverage set is shown in Figure~\ref{fig:pf-binomal-arh-budgets}.
-On the right, we display the coverage set of the best-performing run in terms of hypervolume, for each budget setting.
+We learn a coverage set (see Figure~\ref{fig:pf-binomal-arh-budgets}) that ranges from imposing minimal restrictions to enforcing many restrictions.
+In Figure~\ref{fig:pf-binomal-arh-budgets}, we display on the right the coverage set of the best-performing run in terms of hypervolume, for each budget setting.
 On the left, we show an interpolated average of the coverage sets learned by the different runs.
@@
 Regardless of the imposed budget, we notice that the coverage sets discovered by PCN almost completely dominate the coverage set of the baseline, demonstrating that there are better alternatives to the fixed policies.
-This is most visible in the compromising policies, where one has to carefully choose when to remove social restrictions while at the same time minimizing the impact on daily new hospitalizations.
-In these scenarios, PCN discovers policies that can drastically reduce the total number of new hospitalizations (e.g.
-more than $20000$) for the same social burden.
-Analysing the executions of such policies (see Figure~\ref{fig:policy-execution}, middle plot) shows a flattened hospitalization curve, with a gradual increase of social freedom during the school holidays such that the peak stays stable and gradually decreases over time.
+This is most evident in the compromising policies, where one has to carefully choose when to remove social restrictions while at the same time minimizing the impact on daily new hospitalizations.
+In these scenarios, PCN learns policies that drastically reduce the total number of new hospitalizations (e.g., more than $20000$) for the same social burden.
+We analyse the executions of such policies in Figure~\ref{fig:policy-execution} (middle plot), that shows a flattened hospitalization curve, with a gradual increase of social freedom during the school holidays such that the curve of the epidemic is flattened and gradually decreases over time.
 
@@
 
+Finally, the top plot shows that, without imposing social restrictions, the number of hospitalisations peaks on the 15th of June.
+By the beginning of July, the epidemic has spread out over the majority of the population \todo{herd immunity?}, and the number of admissions at the hospital has been reduced to a fraction of the number of hospitalisations at the peak.
+Thus, without social restrictions, we do not take advantage of the natural decrease of social contacts due to school holidays, as the majority of the population has already been infected before the start of the holidays.
+\todo{mention the slow decrease of $p_l$ during the summer?
+This is simply due to PCN not having found better}
+
 \subsection{Impact of budgets on the coverage set}
@@
     \caption{Selection of policies learned by $\text{PCN}_{b=5}$, from most restrictive in terms of social burden (top) to least restrictive (bottom).
+The x-axis represents the time, starting from the end of the lockdown, on the 4th of May, until the end of the school holidays, on the 1st of September.
+Since the lockdown is simulated before the start of the exit strategy, the start-state differs for each episode (i.e., the hospital already contains infected individuals).
+The left y-axis represents the number of individuals affected by the epidemic.
+The full-lined plots represent the number of individuals admitted into the hospital, ICU and deceased between the last timestep and the current one.
+The plot showing the newly deceased individuals closely relates to the ICU admissions.
+The right y-axis represents the proportional reduction in effect, with 1 meaning a business-as-usual policy, and 0 meaning a complete suppression of social contacts.
+The dotted-lined plots represent the proportional reductions for the work, leasure and school environments.
+We note that the school reduction automatically goes to 0 at the start of the school holidays.
     }
@@
 
-Looking at the coverage sets displayed in Figure~\ref{fig:pf-binomal-arh-budgets}, we notice that the budget impacts the coverage set that has been learnt.
-In general, an increase of budget results in an increasingly better coverage set.
+Figure~\ref{fig:pf-binomal-arh-budgets} demonstrates that the budget impacts the learnt coverage set.
+In general, an increase of budget is associated with an increasingly better coverage set, as policies learned using a higher budget dominate the ones learned with a lower budget.
 This is to be expected, as a higher budget give the agent more freedom to change its actions as the epidemic progresses.
@@
 
-Moreover, we observe that the improvement concentrates around the less restrictive policies in terms of social burden.
-We believe that this regio contains the most complex policies, as these try to maintain as much social freedom as possible, while containing the number of hospitalisations.
-We believe that, in these cases, the timing of the actions greatly correlate with the timing and duration of the peak of the epidemic, and a higher budget allows for more fine-grained control about managing this timing.
+Moreover, we observe that the difference is concentrated around the less restrictive policies in terms of social burden.
+We hypothesise that this region contains the most complex policies, as these try to maintain as much social freedom as possible, while containing the number of hospitalisations.
+In these cases, the timing of the actions greatly correlate with the timing and duration of the peak of the epidemic, and a higher budget allows for more fine-grained control about managing this timing.
 To confirm this, we select, for each budget setting, the solution where the difference in performance is most noticeable, corresponding to the solutions with a total number of hospitalisations around $80000$.
-We plot the execution of the corresponding policies in Figure~\ref{fig:pe-binomial-arh-budgets} and analyse the difference they make in terms of social burden.
-First, we notice that the lower-budget policies are unable to lessen the social restrictions past the peak of the epidemic.
+We plot the execution of the corresponding policies in Figure~\ref{fig:pe-binomial-arh-budgets} and analyse their impact in terms of social burden.
+First, we observe that the lower-budget policies are unable to reduce the social restrictions past the peak of the epidemic.
 In contrast, the setting with no budget restrictions finely controls the restrictions as the epidemic progresses, completely removing restrictions by the end of the wave.
-We note how the policy with a budget of 5 ressembles the execution of the one without restrictions but that, due to its budget, is unable to progressively lessen the restrictions and instead resorts to a halfway compromise.
-\todo{corresponding SB numbers}
-
-Compared to this specific regio of the coverage set, the difference in performance between different budget settings seems marginal around the extremas. \todo{explain}
-
-Finally, we observe that, while the extremas deliver similar trade-offs regardless of the chosen budget, the results differ for the setting without budget restrictions.
-Indeed, in this setting, PCN does not seem to learn the most and least restrictive policies, even though there are no constraints on the action-set.
+Second, we note that the policy with a budget of 5 ressembles the execution of the one without restrictions.
+However, due to its budget, the policy is unable to progressively reduce the restrictions and instead resorts to a halfway compromise.
+\todo{corresponding SB numbers.} Compared to this specific region of the coverage set, the difference in performance between different budget settings seems marginal around the extremas.
+At the extremas, the policies are less complex (e.g., business-as-usual, resulting in the same action executed throughout the episode) and are thus less impacted by the budget restrictions.
+
+Finally, we observe that, while the extremas deliver similar trade-offs for any of the chosen budgets, these trade-offs differ for the setting without budget restrictions.
+Indeed, in this setting, PCN does not learn the most extreme policies with respect to restrictions, even though there are no constraints on the action-set.
 As explained in Section~\ref{sec:pcn-policy-exploration}, PCN searches for increasingly better solutions using a stochastic policy.
@@
 In comparison, for the settings with a limited budget, the action stays the same as the previous one once the budget has been spent.
-It is thus easier to discover the most extreme policies.
-Thus, in the specific case where we impose no budget, we believe the freedom of action actually hinders PCN's search.
+As such, it is easier to learn the most extreme policies.
+Thus, in the specific case where we impose no budget, the freedom of action actually hinders PCN's search, for certain regions of the rweard-space.
 
@@
     \caption{Execution of the policies attaining a number of hospitalisations around $80000$, for different budgets.
-From top to bottom are displayed the policy executions of runs with budget 2,3,4,5 and no-limit, respectively.
-We notice that the lower-budget policies are unable to lessen the social restrictions past the peak.
-The setting with no budget restrictions finely controls the restrictions as the epidemic progresses, completely removing restrictions by the end of the wave.
-Finally, there is no consensus on which social environment to restrict most, some policies providing similar restrictions for $p_w$ and $p_l$, while others putting harsher restrictions on one social environment than the other.
+From top to bottom we display the policy executions with budget 2,3,4,5 and no-limit, respectively.
+We notice that the lower-budget policies are unable to reduce the social restrictions past the peak.
+The setting without budget restrictions finely controls the restrictions as the epidemic progresses, completely removing restrictions by the end of the wave.
+Finally, there is no consensus on which social environment to restrict most: certain policies provide similar restrictions for $p_w$ and $p_l$, while others impose harsher restrictions on one social environment than the other.
     }
@@
 
-We now assess the difference in coverage sets when optimizing on $\mdprewardfn_\text{ARH}$ versus $\mdprewardfn_\text{ARI}$.
+Next, we assess the difference in coverage sets when optimizing on $\mdprewardfn_\text{ARH}$ versus $\mdprewardfn_\text{ARI}$.
 Although these reward functions have a different scale (there are more infected persons than hospitalised ones), our experiments show that infections and hospitalizations are correlated.
-This is expected, as during the initial phase of the epidemic, limited immunity was present in the population (few natural immunity and no vaccines), which induces a tight coupling between infection and hospitalisation cases.
+This is expected, as during the initial phase of the epidemic, limited immunity was present in the population (i.e., limited natural immunity and no vaccines), which induces a tight coupling between infection and hospitalisation cases.
 This is confirmed in Table~\ref{tab:coverage-set}.
 In this table, we show the different performance metrics (hypervolume, $I_\varepsilon$, $I_{\varepsilon-mean}$) with respect to the objectives $[\mdprewardfn_\text{ARH}, \mdprewardfn_\text{SB}]$.
-However, the left-side of the table shows the experiments where these objectives are also used as optimization criterions, while the right-side shows the experiments using $[\mdprewardfn_\text{ARI}, \mdprewardfn_\text{SB}]$ as optimization criterions.
+The table is split in two parts.
+The left-side shows the different performance metrics, for PCN using $[\mdprewardfn_\text{ARH}, \mdprewardfn_\text{SB}]$ as optimization criterions.
+The right-side shows the same performance metrics, but with PCN using $[\mdprewardfn_\text{ARI}, \mdprewardfn_\text{SB}]$ as optimization criterions.
 
 Even with the $\mdprewardfn_\text{ARI}$, the increased budget shows an increase in hypervolume in terms of hospitalisations.
-Moreover, those hypervolumes are close to the ones trained on $\mdprewardfn_\text{ARH}$, inidicating that their coverage sets are similar.
+Moreover, those hypervolumes are close to the ones trained on $\mdprewardfn_\text{ARH}$, indicating that their coverage sets are similar.
 However, regardless of the imposed budget, the hypervolumes are slightly worse.
@@
 The dataset of trajectories that PCN is trained on is pruned over time to keep only the most relevant trajectories.
-The returns of these trajectories are used in Figure~\ref{fig:pf-binomal-arh-budgets} to show the discovered coverage set.
+The returns of these trajectories are used in Figure~\ref{fig:pf-binomal-arh-budgets} to show the learned coverage set.
 Each of these returns can be used as desired return for policy execution.
@@
 The $I_\varepsilon$ indicators shows that, regardless of the budget, the decision maker will lose at worst a $0.050$ normalized return in any of the objectives.
-On average, it will lose $0.010$ normalized returns.
-Moreover, we emphasize that the learned coverage set contains the non-dominated returns across the whole training procedure.
-Since the Binomial model is stochasticity, for multiple executions of the same policy, the executions that remains in the coverage set is the one with advantageous transitions due to this stochasticity.
-Thus, we expect our policy-executions to obtain slightly worse results than the target-solution selected from the coverage set.
-In light of this, we conclude that the policies trained by PCN are reliable and produce returns as close as possible from their chosen target.
+On average, it will lose $0.010$ normalized returns, i.e., on average, the return obtained by executing a policy will either result in an additional $1441$ hospitalisations than expected, or result in $12$ additional social contacts lost.
+Moreover, we emphasize that the learned coverage set contains the non-dominated returns encountered over the whole training procedure.
+Since the \momdpname\ model is stochastic, for multiple executions of the same policy, the executions kept in the coverage sets are the ones for which the samples from the binomial-chain resulted in a better progression of the epidemic than average.
+Thus, we expect our policy-executions to be close to the target selected from the coverage set, but not on exactly target.
+Based on this analysis, we conclude that the policies trained by PCN are robust and produce returns as close as possible from their chosen target.
 
@@
 \section{Related work}
-Reinforcement learning (RL) has been used in conjunction with epidemiological models to learn policies to limit the spread of diseases and predict the effects of possible mitigation strategies \cite{probert2019context,ernst2006clinical}.
-For example, RL has been used extensively in modelling and controlling the spread of influenza \cite{das2008large,libin2020,libin2018bayesian}.
+Reinforcement learning (RL) has been used in conjunction with epidemiological models to learn policies to limit the spread of diseases and predict the effects of possible mitigation strategies \cite{probert2019context}.
+For example, RL has been used to model the mitigation of influenza \cite{das2008large,libin2020,libin2018bayesian}.
 
@@
 \cite{wan2021multi} implement a model-based multi-objective policy search method and demonstrate their method on COVID-19 data from China.
-Given the method is model-based, a model of the transition function must be learned by sampling from the environment.
+Given that this method is model-based, a model of the transition function must be learned by sampling from the environment.
 The method proposed by Wan et al.
@@
 However, methods which use linear weights can only learn policies on the convex-hull of the Pareto front \cite{vamplew2008limitations}, therefore the full Pareto front cannot be learned.
-It is important to note the method proposed by Kompella et al.
+We note that the method proposed by Kompella et al.
 \cite{kompella2020reinforcement} considers multiple objectives.
@@
 We are aware of the ethical implications that expert systems have on the decision process and we make the disclaimer that all results based on, or derived from, the expert system that we propose should be carefully interpreted by experts in the field of public health, and in a much broader context of economics, well-being and education.
-We note that the work in this manuscript was conducted by a multi-disciplinary consortium that includes computer scientists and scientists with a background public health, epidemiology and bio-statistics.
+We note that the work in this manuscript was conducted by a interdisciplinary consortium that includes computer scientists and scientists with a background public health, epidemiology and bio-statistics.
 

### 2023-03-31 paper/fig/arh_binomial_both.tex
missing data
pieter comments

@@
         },
-        legend entries={{$\text{PCN}_{b=2}$}, {$\text{PCN}_{b=3}$}, {$\text{PCN}_{b=4}$}, {$\text{PCN}_{b=5}$}, {$\text{PCN}$}, {Fixed}},
+        legend entries={{$\text{PCN}_{b=2}$}, {$\text{PCN}_{b=3}$}, {$\text{PCN}_{b=4}$}, {$\text{PCN}_{b=5}$}, {$\text{PCN}_{b=\infty}$}, {Fixed}},
     ]
     % interpolated plot
-    \nextgroupplot[draw opacity=1]
+    \nextgroupplot[mark size=1pt,draw opacity=1]
     % mean line

### 2023-03-31 paper/main.tex
pieter comments

@@
 \usepackage{tikz}
-\usepackage{pgfplots}
+\usepackage{pgfplots,etoolbox}
 \usepackage{comment}
+\usetikzlibrary{external}
 \usepgfplotslibrary{dateplot, groupplots, fillbetween}
 \pgfplotsset{compat=newest}
+\tikzexternalize[prefix=cache/]
+
 \usepackage{plotpolicies}   % custom package to plot policy-executions
@@
 
-The main disadvantage of this metric is that we need the true Pareto front to compute it, which is unknown for our MOMDP (see Section~\ref{sec:sars-cov2-momdp}).
+The main disadvantage of this metric is that we need the true Pareto front to compute it, which is unknown for our MOMDP (see Sec.~\ref{sec:sars-cov2-momdp}).
 To still gain insights from our learned policies, we approximate the true Pareto front using the non-dominated policies across all runs.
@@
 
-Figure~\ref{fig:paretofrontmetrics} shows a visual representation of the hypervolume and $\varepsilon$ metrics in two dimensions.
+Fig.~\ref{fig:paretofrontmetrics} shows a visual representation of the hypervolume and $\varepsilon$ metrics in two dimensions.
 
@@
 
-A visual representation of the model is depicted in Fig~\ref{fig:ode_mdel}. It shows the different compartments, as well as the flow rates at which individuals move from one compartment to another.
+A visual representation of the model is depicted in Fig~\ref{fig:compartment_model}.
+It shows the different compartments, as well as the flow rates at which individuals move from one compartment to another.
 
@@
 We denote $\hat{C}_t$ the social contact matrix at timestep $t$, resulting from the reduction function $\hat{C}$.
-Secondly, we assume that compliance to the interventions is gradual and model this using a logistic compliance function (see details in SI).
+Secondly, we assume that compliance to the interventions is gradual and model this using a logistic compliance function (see details in Sec.~\ref{sec:compliance-function}).
 
@@
     \caption{Schematic diagram of the compartmental model for SARS-CoV-2 presented by \citet{abrams2021modelling} which is used to derive the MOMDP.}
-    \label{fig:ode_mdel}
+    \label{fig:compartment_model}
 \end{figure}
-In order to apply multi-objective reinforcement learning, we construct the \momdpname\ MOMDP based on the epidemiological model introduced in Section~\ref{sec:sars-cov2-momdp_model} and graphically depicted in Figure~\ref{fig:ode_mdel}.
+In order to apply multi-objective reinforcement learning, we construct the \momdpname\ MOMDP based on the epidemiological model introduced in Sec.~\ref{sec:sars-cov2-momdp_model} and graphically depicted in Fig.~\ref{fig:compartment_model}.
 Moreover, we consider a finite-horizon setting where we simulate the compartment model for a fixed number of weeks.
 
-\paragraph{Action-space:} Our actions concern the installment of a social contact matrix with a particular reduction resulting from the reduction function $\hat{C}$ (see Section~\ref{sec:interventions}).
-To this end, we use the proportional reduction parameters $p_w, p_s, p_l$ defined in Section~\ref{sec:interventions}.
-Thus, each $\mdpaction \in \mdpactionspace$ is a 3-dimensional continuous vector in $[0,1]^3$ (i.e., $\mdpaction = [p_{w}, p_{s}, p_{l}]$) which impacts the social contact matrix according to Equation~\ref{eq:scm}.
-
-\paragraph{Transition function:} The model defined by \citet{abrams2021modelling} utilises a model transition probability $M(\mdpstate_m' \mid \mdpstate_m, \hat{C}, p_{w}, p_{s}, p_{l})$ (see SI for details on $M$), where $\hat{C}(p_{w}, p_{s}, p_{l})$ returns the currently installed social contact matrix, that progresses the epidemiological model in one timestep.
+\paragraph{Action-space:} Our actions concern the installment of a social contact matrix with a particular reduction resulting from the reduction function $\hat{C}$ (see Sec.~\ref{sec:interventions}).
+To this end, we use the proportional reduction parameters $p_w, p_s, p_l$ defined in Sec.~\ref{sec:interventions}.
+Thus, each $\mdpaction \in \mdpactionspace$ is a 3-dimensional continuous vector in $[0,1]^3$ (i.e., $\mdpaction = [p_{w}, p_{s}, p_{l}]$) which impacts the social contact matrix according to Eq.~\ref{eq:scm}.
+
+\paragraph{Transition function:} The model defined by \citet{abrams2021modelling} utilises a model transition probability $M(\mdpstate_m' \mid \mdpstate_m, \hat{C}, p_{w}, p_{s}, p_{l})$ (see Sec.~\ref{sec:binomial-model} for details on $M$), where $\hat{C}(p_{w}, p_{s}, p_{l})$ returns the currently installed social contact matrix, that progresses the epidemiological model in one timestep.
 We use this function as the transition function in \momdpname.
@@
 
-We parameterize the epidemiological model using the mean of the posteriors as specified by \citet{abrams2021modelling} (details in SI).
+We parameterize the epidemiological model using the mean of the posteriors as specified by \citet{abrams2021modelling} (details in Sec.~\ref{sec:compartment-model-parameters}).
 
@@
 where $S_k(\mdpstate)$ represents the number of susceptible individuals in age group $k$ in state $\mdpstate$ and $R_k$ represents the number of recovered individuals in age group $k$ in state $\mdpstate$.
-In Section~\ref{sec:experiments}, we optimize PCN on two different variants for the multi-objective reward function: $[\mdprewardfn_\text{ARH}, \mdprewardfn_\text{SB}]$ and $[\mdprewardfn_\text{ARI}, \mdprewardfn_\text{SB}]$, to study the impact of these distinct attack rate quantities.
+In Sec.~\ref{sec:experiments}, we optimize PCN on two different variants for the multi-objective reward function: $[\mdprewardfn_\text{ARH}, \mdprewardfn_\text{SB}]$ and $[\mdprewardfn_\text{ARI}, \mdprewardfn_\text{SB}]$, to study the impact of these distinct attack rate quantities.
 
@@
 \begin{enumerate}
-    \item Use the current policy to interact with the environment and generate trajectories (see Section~\ref{sec:pcn-policy-exploration}).
+    \item Use the current policy to interact with the environment and generate trajectories (see Sec.~\ref{sec:pcn-policy-exploration}).
 Initially, use a random policy to fill the dataset with random trajectories.
-    \item Update the dataset to incorporate these new trajectories and prune the least interesting trajectories in terms of returns (see Section~\ref{sec:pcn-building-dataset}).
-    \item Update the policy by training it on the updated dataset (see Section~\ref{sec:pcn-continuous}).
+    \item Update the dataset to incorporate these new trajectories and prune the least interesting trajectories in terms of returns (see Sec.~\ref{sec:pcn-building-dataset}).
+    \item Update the policy by training it on the updated dataset (see Sec.~\ref{sec:pcn-continuous}).
 \end{enumerate}
@@
 
-While the original PCN algorithm is designed for MOMDPs with discrete actions-spaces, the problem we tackle (see Section~\ref{sec:sars-cov2-momdp}) is defined in terms of a continuous action-space.
+While the original PCN algorithm is designed for MOMDPs with discrete actions-spaces, the problem we tackle (see Sec.~\ref{sec:sars-cov2-momdp}) is defined in terms of a continuous action-space.
 We thus extend PCN to the continuous action-space setting.
 We change the output of the neural network such that there is a single output value for each dimension of the action-space.
-Since the actions should be bound in the domain of possible actions ($[0,1]$ in the case of \momdpname, see Section~\ref{sec:sars-cov2-momdp}), we apply a tanh non-linearity function on each output.
+Since the actions should be bound in the domain of possible actions ($[0,1]$ in the case of \momdpname, see Sec.~\ref{sec:sars-cov2-momdp}), we apply a tanh non-linearity function on each output.
 As such, we have a regression problem instead of a classification problem, as the labeled dataset now uses continuous labels $y = \mdpaction_t$ instead of categories.
@@
 We assess the validity of our approach by executing PCN on this deterministic variant, and observing similar performance as with the \momdpname\ model.
-\todo{Results are available in SI}.
+We show the results in Sec.~\ref{sec:cs-ode-vs-binomial}.
 
@@
 To conduct this analysis, we apply our extension of PCN for continuous action-spaces on the \momdpname\ model.
-As explained in Section~\ref{sec:pcn-stochastic}, we extend PCN for environments with stochastic transitions.
+As explained in Sec.~\ref{sec:pcn-stochastic}, we extend PCN for environments with stochastic transitions.
 
@@
 
-All experiments are averaged over 5 runs. The hyper-parameters and the neural network architecture can be found in the SI.
+All experiments are averaged over 10 runs.
+The hyper-parameters and the neural network architecture can be found in Sec.s~\ref{sec:pcn-hyperparameters} and Sec.~\ref{sec:pcn-architecture}, respectively.
 
@@
 Changes are most noticeable in the less restrictive trade-offs in terms of social burden.}
-    \label{fig:pf-binomal-arh-budgets}
+    \label{fig:pf-binomial-arh-budgets}
 \end{figure}
@@
 
-We learn a coverage set (see Figure~\ref{fig:pf-binomal-arh-budgets}) that ranges from imposing minimal restrictions to enforcing many restrictions.
-In Figure~\ref{fig:pf-binomal-arh-budgets}, we display on the right the coverage set of the best-performing run in terms of hypervolume, for each budget setting.
+We learn a coverage set (see Fig.~\ref{fig:pf-binomial-arh-budgets}) that ranges from imposing minimal restrictions to enforcing many restrictions.
+In Fig.~\ref{fig:pf-binomial-arh-budgets}, we display on the right the coverage set of the best-performing run in terms of hypervolume, for each budget setting.
 On the left, we show an interpolated average of the coverage sets learned by the different runs.
@@
 In these scenarios, PCN learns policies that drastically reduce the total number of new hospitalizations (e.g., more than $20000$) for the same social burden.
-We analyse the executions of such policies in Figure~\ref{fig:policy-execution} (middle plot), that shows a flattened hospitalization curve, with a gradual increase of social freedom during the school holidays such that the curve of the epidemic is flattened and gradually decreases over time.
+We analyse the executions of such policies in Fig.~\ref{fig:policy-execution} (middle plot), that shows a flattened hospitalization curve, with a gradual increase of social freedom during the school holidays such that the curve of the epidemic is flattened and gradually decreases over time.
 
 Interestingly, we notice that the most restrictive policy (i.e.
-the one that prioritizes hospitalizations over social burden, see Figure~\ref{fig:policy-execution}, bottom plot) still starts to gradually increase $p_w$ and $p_l$ from the end of July onward.
+the one that prioritizes hospitalizations over social burden, see Fig.~\ref{fig:policy-execution}, bottom plot) still starts to gradually increase $p_w$ and $p_l$ from the end of July onward.
 This is because by then, the epidemic has mostly faded out, and it is safe to reduce social restrictions.
@@
 
-Figure~\ref{fig:pf-binomal-arh-budgets} demonstrates that the budget impacts the learnt coverage set.
+Fig.~\ref{fig:pf-binomial-arh-budgets} demonstrates that the budget impacts the learnt coverage set.
 In general, an increase of budget is associated with an increasingly better coverage set, as policies learned using a higher budget dominate the ones learned with a lower budget.
@@
 However, we also notice that this increase of coverage sets is not necessarily gradual.
-Between the run with a budget of 2 and budget of 3, the difference in coverage set seems marginal.
+% Between the run with a budget of 2 and budget of 3, the difference in coverage set seems marginal.
 \todo{insight}.
@@
 In these cases, the timing of the actions greatly correlate with the timing and duration of the peak of the epidemic, and a higher budget allows for more fine-grained control about managing this timing.
-To confirm this, we select, for each budget setting, the solution where the difference in performance is most noticeable, corresponding to the solutions with a total number of hospitalisations around $80000$.
-We plot the execution of the corresponding policies in Figure~\ref{fig:pe-binomial-arh-budgets} and analyse their impact in terms of social burden.
+To confirm this, we select, for each budget setting, the solution where the difference in performance is most noticeable, corresponding to the solutions with a total number of hospitalisations around $80000$ (which is in the middle of the range of possible hospitalisations, as can be seen in Fig.~\ref{fig:pf-binomial-arh-budgets}).
+We plot the execution of the corresponding policies in Fig.~\ref{fig:pe-binomial-arh-budgets} and analyse their impact in terms of social burden.
 First, we observe that the lower-budget policies are unable to reduce the social restrictions past the peak of the epidemic.
@@
 However, due to its budget, the policy is unable to progressively reduce the restrictions and instead resorts to a halfway compromise.
-\todo{corresponding SB numbers.} Compared to this specific region of the coverage set, the difference in performance between different budget settings seems marginal around the extremas.
+Compared to this specific region of the coverage set, the difference in performance between different budget settings seems marginal around the extremas.
 At the extremas, the policies are less complex (e.g., business-as-usual, resulting in the same action executed throughout the episode) and are thus less impacted by the budget restrictions.
@@
 Indeed, in this setting, PCN does not learn the most extreme policies with respect to restrictions, even though there are no constraints on the action-set.
-As explained in Section~\ref{sec:pcn-policy-exploration}, PCN searches for increasingly better solutions using a stochastic policy.
+As explained in Sec.~\ref{sec:pcn-policy-exploration}, PCN searches for increasingly better solutions using a stochastic policy.
 Thus, at every timestep, the action can change compared to the previous one.
@@
 As an exception, we notice that PCN without budget restrictions results in better performance across every metric for the $\mdprewardfn_\text{ARI}$ variant.
-Still, due to the high standard deviation of the no-budget, $\mdprewardfn_\text{ARH}$ setting, we do not believe this difference is meaningfull.
+Still, due to the high standard deviation of the no-budget, $\mdprewardfn_\text{ARH}$ setting, we do not believe this difference is meaningful.
 \todo{some kind of meaningful conclusion}
@@
 The dataset of trajectories that PCN is trained on is pruned over time to keep only the most relevant trajectories.
-The returns of these trajectories are used in Figure~\ref{fig:pf-binomal-arh-budgets} to show the learned coverage set.
+The returns of these trajectories are used in Fig.~\ref{fig:pf-binomial-arh-budgets} to show the learned coverage set.
 Each of these returns can be used as desired return for policy execution.
@@
 
+\todo{Although we use an age-structured compartment model, with social contact matrices to model social interactions, this remains a model that evaluates the progression of the pandemic as an aggregated process over the population.
+Individual-based models could provide more detailed and localized policies, potentially further improving the quality of possible trade-offs, and would provide an interesting avenue for future work.
+However, due to the computational cost of simulating such models, and the number of interactions required by reinforcement learning in general, this remains a challenging problem.}
+
+\todo{Moreover, although we are able to interpret and analyse the obtained policies and their corresponding trade-offs, as we can plot the Pareto front for 2 objectives, this approach cannot be used for problems with more objectives.
+Defining interpretation and visualization tools for many-objective problems is necessary if we want to include additional objectives for our problem.}
+
+\todo{Additionally, PCN is able to cope with model stochasticity, because the stochasticity is limited.
+In settings where the stochasticity is more pronounced, e.g., designing policies to control the initial outbreak of an epidemic, further methodological extensions are warranted.}
+
+\todo{Finally, while we have focused on social burden as a trade-off for hospitalisations, our approach can be used for related problems, involving different compromises, such as balancing the number of lost schooldays with respect to new infections in schools \cite{torneri2021controlling}, contact tracing effort versus its effect \cite{willem2021impact}, the impact of antivirals on the epidemic versus resistance to antivirals \cite{torneri2020prospect}, or the cost of universal testing versus its impact on the epidemic \cite{libin2021assessing}.}
+
 To conclude, we show that multi-objective reinforcement learning can be used to learn a wide set of high-quality policies on real-world problems, providing the decision maker with insightful and diverse alternatives, and showing the impact of taking extreme measures.
+\todo{This can be done for realistic settings, where we budget the number of changes of social reductions, such that the policies are practical and can be deployed in real-life scenarios.}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{comment}
 \appendix
@@
 
-\section{Compartmental model for SARS-CoV-2}
+\section{Stochastic Compartmental Model}
+\label{sec:binomial-model}
 In this work we utilise the compartmental model proposed by \citet{abrams2021modelling} and extend the model to a multi-objective Markov decision process (MOMDP).
-The MOMDP used in this work is outlined in the main paper.
-However, we describe the details of the underlying compartmental model used to model the spread of COVID-19 below.
-
-\begin{figure}[h]
-    \centering
-     \begin{tikzpicture}
-     \draw (0.0pt, 0.0pt)node[fill=green!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](0){$\textbf{S}$};
-     \draw (60.0pt, 0.0pt)node[fill=orange!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](1){$\textbf{E}$};
-     \draw (120.0pt, 0.0pt)node[fill=red!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](2){$\textbf{I}^{presym}$};
-     \draw (180.0pt, 10.0pt)node[fill=red!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](3){$\textbf{I}^{asym}$};
-     \draw (170.0pt, -30.0pt)node[fill=red!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](4){$\textbf{I}^{mild}$};
-     \draw (170.0pt, -70.0pt)node[fill=red!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](5){$\textbf{I}^{sev}$};
-     \draw (230.0pt, -50.0pt)node[fill=blue!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](6){$\textbf{I}^{hosp}$};
-     \draw (230.0pt, -90.0pt)node[fill=blue!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](7){$\textbf{I}^{icu}$};
-     \draw (300.0pt, -50.0pt)node[fill=orange!50, thick, minimum height=0.6cm,minimum width=0.6cm, draw](8){$\textbf{R}$};
-     \draw (300.0pt, -90.0pt)node[fill=purple!50, thick, minimum height=0.6cm,minimum width=0.6cm, draw](9){$\textbf{D}$};
-
-     %\path[->] (6) edge[dotted, bend right = 50]
-     \path[->] (0) edge[] node[above]{$\boldsymbol{\lambda}(t)$} (1);
-     \path[->] (1) edge[] node[above]{$\gamma$} (2);
-     \path[->] (2) edge[] node[below]{$\textbf{p}\theta$} (3);
-     \path[->] (2) edge[] node[near end, below left]{$(1-\textbf{p})\theta$} (4);
-     \path[->] (4) edge[] node[left]{$\boldsymbol{\psi}$} (5);
-     \path[->] (5) edge[] node[near end, below left]{$(1-\phi_{1})\boldsymbol{\omega}$} (7);
-     \path[->] (5) edge[] node[above]{$\phi_{1}\boldsymbol{\omega}$} (6);
-     \path[->] (7) edge[] node[near end, below right]{$\boldsymbol{\delta}_{4}$} (8);
-     \path[->] (7) edge[] node[above]{$\boldsymbol{\tau}_{2}$} (9);
-     \path[->] (6) edge[] node[above]{$\boldsymbol{\delta}_{3}$} (8);
-     \path[->] (6) edge[] node[near start, below left]{$\boldsymbol{\tau}_{1}$} (9);
-     \path[->] (3) edge[bend left = 30] node[above right]{$\delta_{1}$} (8);
-     \path[->] (4) edge[bend left = 30] node[above left]{$\boldsymbol{\delta}_{2}$} (8);
-
-    \end{tikzpicture}
-
-    \caption{Schematic diagram of the compartmental model for SARS-CoV-2 presented by \citet{abrams2021modelling} which is used to derive the MOMDP.
-    }
-    \label{fig:ode_mdel}
-\end{figure}
-
-We utilise an adapted SEIR mathematical compartmental model proposed by Abrams et al.
-to contruct the MOBelCov MOMDP with deterministic and stochastic state transitions.
-In this model members of the population are susceptible to infection when they are in compartment $\textbf{S}$\footnote{Boldface vector notation is used to denote the multiple age groups for each compartment.}.
-If an individual comes into contact with an infections individual then they move to the exposed compartment, $\textbf{E}$, at a time specific rate, $\boldsymbol{\lambda} (t)$.
-After a period of time an exposed individual becomes infectious, where they move to the pre-symptomatic compartment, $\textbf{I}^{presym}$, at rate $\gamma$.
-Once infected, individuals develop mild symptoms, $\textbf{I}^{mild}$ , with probability $1 - \textbf{p}$ or do not develop any symptoms, $\textbf{I}^{asym}$, with probability $\textbf{p}$, where asymptomatic individuals recover, $\textbf{R}$, at rate $\delta_{1}$.
-Individuals who experience symptoms can suffer from a mild infection, $\textbf{I}^{mild}$, and recover at rate $\boldsymbol{\delta}_{2}$, or they suffer from a more severe infection, $\textbf{I}^{sev}$, at a rate $\boldsymbol{\psi}$.
-Individuals with a severe infection are then transferred to a hospital for treatment, $\textbf{I}^{hosp}$ with probability $\phi_{1}$.
-However, some individuals become critically ill and are transferred directly to the intensive care unit (ICU) with probability $1 - \psi_{1}$.
-Individuals in the hospital, $\textbf{I}^{hosp}$ and $\textbf{I}^{icu}$, recover at rate $\boldsymbol{\delta}_{3}$ or $\boldsymbol{\delta}_{4}$ and die at rate $\boldsymbol{\tau}_{3}$ and $\boldsymbol{\tau}_{4}$ respectively.
-Figure \ref{fig:ode_mdel} outlines the compartmental model defined by Abrams et al.~\cite{abrams2021modelling}.
-
-\subsection{Deterministic Compartmental Model}
-\label{sec:deterministic_comp_model}
-The flows of the deterministic model are defined by a set of ordinary differential equations, which are outlined as follows:
-
-\begin{align*}
-\frac{d\textbf{S}(t)}{dt} & = -\lambda(t)(\textbf{S})(t), \\
-%
-\frac{d\textbf{E}(t)}{dt} & = \lambda(t)(\textbf{S})(t) - \gamma \textbf{E}(t), \\
-%
-\frac{d\textbf{I}^{presym}(t)}{dt} & = \gamma \textbf{E}(t) - \theta \textbf{I}^{presym}(t), \\
-%
-\frac{d\textbf{I}^{asym}(t)}{dt} & = \theta p \textbf{I}^{presym}(t) - \delta_{1} \textbf{I}^{asym}(t), \\
-%
-\frac{d\textbf{I}^{mild}(t)}{dt} & = \theta (1 - p) \textbf{I}^{presym}(t) - \{ \psi + \omega_{2} \} \textbf{I}^{mild}(t), \\
-%
-\frac{d\textbf{I}^{sev}(t)}{dt} & = \psi \textbf{I}^{mild}(t) - \omega \textbf{I}^{sev}(t), \\
-%
-\frac{d\textbf{I}^{hosp}(t)}{dt} & = \phi_{1} \omega \textbf{I}^{sev}(t) - \{ \delta_{3} + \tau_{1} \} \textbf{I}^{hosp}(t), \\
-%
-\frac{d\textbf{I}^{icu}(t)}{dt} & = (1 - \phi_1) \omega \textbf{I}^{sev}(t) - \{ \delta_{4} + \tau_{2} \} \textbf{I}^{icu}(t), \\
-%
-\frac{d\textbf{D}(t)}{dt} & = \tau_{1} \textbf{I}^{hosp}(t) - \tau_{2} \textbf{I}^{icu}(t), \\
-%
-\frac{d\textbf{R}(t)}{dt} & = \delta_{1} \textbf{I}^{asym}(t) + \delta_{2} \textbf{I}^{mild}(t) + \delta_{3} \textbf{I}^{hosp}(t) + \delta_{4} \textbf{I}^{icu}(t)
-\end{align*}
-
-where, for example, $\textbf{S} = (S_{1}(t), S_{2}(t), ..., S_{k}(t))^{T}$ is the vector representing the susceptible members of the population of each age group $k$ at time $t$.
-
-In this set of ordinary differential equations, each state variable represents a vector over all age groups for a particular compartment at time $t$.
-Infection dynamics are governed by an age-specific force of infection $\foi$:
-\begin{equation}
-    \foi(k,t) = \sum_{k'=1}^{K}\beta(k,k')I_{k'}(t),
-\end{equation}
-where $k$ is the respective age group of a total of K age groups, and $\beta(k,k')$ is the time-invariant transmission rate that encodes the average per capita rate at which an infectious individual in age group $k$ makes an effective contact with a susceptible individual in age group $k'$, per unit of time.
-
-Under the social contact hypothesis \cite{wallinga2006using}, we have that:
-\begin{equation}
-\beta(k,k') = q \cdot C(k,k'),
-\end{equation}
-where q is a proportionality factor and the matrix $C$ denotes the social mixing behaviour within and between different age groups in the population, and is referred to as a social contact matrix.
-%(TODO: add appropriate reference).
-Following \citet{abrams2021modelling}, we rely on distinct social contact matrices for symptomatic and asymptomatic individuals, respectively $C_{s}$ and $C_{a}$.
-Therefore it is possible to define the transmission rates for both symptomatic and asymptomatic individuals as follows:
-
-\begin{equation}
-\boldsymbol{\beta}_{s}(k, k') = q_{s} \cdot C_{s}(k, k'),    
-\end{equation}
-and
-\begin{equation}
-\boldsymbol{\beta}_{a}(k, k') = q_{a} \cdot C_{a}(k, k').  
-\end{equation}
-Moreover, the age-dependent force of infection can be defined as follows:
-\begin{equation}
-    \boldsymbol{\lambda}(t) = \boldsymbol{\beta}_{a} \times \{ \textbf{I}^{presym}(t) + \textbf{I}^{asym}(t) \} +
-    \boldsymbol{\beta}_{s} \times \{ \textbf{I}^{mild}(t) + \textbf{I}^{sev}(t) \},
-\end{equation}
-where $\boldsymbol{\lambda}(t) = (\lambda(1, t), \lambda(1, t), ..., \lambda(K, t))$.
-For all further information about the different compartments and parameters please refer to the work of \citet{abrams2021modelling}
-
-To create a version of MOBelCov with a deterministic transition function, $M$, we focus on the deterministic model proposed by \citet{abrams2021modelling}.
-Given that the transitions within the compartmental model are deterministic it is possible to utilise the highlighted model transitions for MOBelCov.
-Applying the contact matrix $\hat{C}$ to the model state $s_{m}$ progresses the model for one timestep and returns a new compartmental model state $s'_{m}$.
-Given $s_{m}$ and $\hat{C}$, the  deterministic compartmental model returns $s'_{m}$ with probability of $1$.
-Therefore, it is possible to use this process as a deterministic transition function, $M$, for MOBelCov.
-
-\subsection{Stochastic Compartmental Model}
+Fig.~\ref{fig:compartment_model} shows a visual depiction of the compartment model.
+The flows of the deterministic model are defined by a set of ordienary differential equations, outlined in Sec.~\ref{sec:sars-cov2-momdp_model}.
+
 Intervening in the spread of the virus by, for example, reducing social contacts or government interventions introduces uncertainty in the further course of the outbreak.
@@
 
-By formulating the set of differential equations defined above, as a chain-binomial, we can obtain stochastic trajectories from this model \cite{bailey1975mathematical}.
-A chain-binomal model assumes a stochastic model where infected individuals are generated by some underlying probability distribution.
+By formulating the set of differential equations defined in Sec.~\ref{sec:sars-cov2-momdp_model}, as a chain-binomial, we can obtain stochastic trajectories from this model \cite{bailey1975mathematical}.
+A chain-binomial model assumes a stochastic model where infected individuals are generated by some underlying probability distribution.
 For the stochastic model we consider a time interval $(t, t +h]$, where $h$ is defined as the length between two consecutive time points.
@@
 
-Given MOBelCov also calculates new hospitalisations, $\textbf{H}^{new}$, we define $\textbf{H}^{new}$ for the stochastic compartmental model as follows:
+Given \momdpname\ also calculates new hospitalisations, $\textbf{H}^{new}$, we define $\textbf{H}^{new}$ for the stochastic compartmental model as follows:
 \[
@@
 
-To create a version of MOBelCov with a stochastic transition function, $M$, we utilise the stochastic compartmental model outlined above.
-Given the transitions within the compartmental model are derived by an underlying probability distribution it is possible to utilise the stochastic compartmental model transitions for MOBelCov.
-As previously outlined in Section \ref{sec:deterministic_comp_model} the contact matrix $\hat{C}$ applied the model state $s_{m}$ progresses the model and returns a new model state $s'_{m}$.
+To create a version of \momdpname\ with a stochastic transition function, $M$, we utilise the stochastic compartmental model outlined above.
+Given the transitions within the compartmental model are derived by an underlying probability distribution it is possible to utilise the stochastic compartmental model transitions for \momdpname\.
+As previously outlined in Sec.
+\ref{sec:deterministic_comp_model} the contact matrix $\hat{C}$ applied the model state $s_{m}$ progresses the model and returns a new model state $s'_{m}$.
 Given the underlying model dynamics are governed in a probabilistic manner, the model returns $s'_{m}$ stochastically.
-Therefore, it is possible to use this process as a stochastic transition function, $M$, for MOBelCov.
+Therefore, it is possible to use this process as a stochastic transition function, $M$, for \momdpname.
 
 \subsection{A Note on Model Parameters}
+\label{sec:compartment-model-parameters}
 
@@
 \subsection{Modelling interventions}
-\label{sec:interventions}
-In order to model different types of interventions, we follow \citet{abrams2021modelling}. Firstly, 
-to consider distinct exit scenarios, we alter the social contact matrices to reflect a contact reduction in a particular age group.
+\label{sec:compliance-function}
+In order to model different types of interventions, we follow \citet{abrams2021modelling}.
+Firstly, to consider distinct exit scenarios, we alter the social contact matrices to reflect a contact reduction in a particular age group.
 Secondly, we assume that compliance to the interventions is gradual and model this using a logistic compliance function.
-We use the logistic compliance function in function of time $t$ proposed by Abrams et al.,
+We use the logistic compliance function in function of time $t$,
+%
 \begin{equation}
@@
 \end{equation}
-where $t_I$ indicates the time the intervention started \cite{abrams2021modelling}.
+%
+where $t_I$ indicates the time the intervention started.
 We initialise $\beta^*_1$ to the value estimated in by Abrams et al.
-and choose $\beta^*_0=-5$, as an intercept to have $c(t)=0$ for $t=0$, in correspondence with Figure~F2 in the Supplementary Information of \citet{abrams2021modelling}.
+and choose $\beta^*_0=-5$, as an intercept to have $c(t)=0$ for $t=0$, in correspondence with Fig.~F2 in the Supplementary Information of \citet{abrams2021modelling}.
 
@@
 \subsection{Comparison of coverage sets learned on ODE and Binomial models}
-\subsection{sec:cs-ode-vs-binomial}
-
-\begin{figure}
+\label{sec:cs-ode-vs-binomial}
+
+\begin{figure}[t!]
     \centering
     \input{fig/arh_binomial_ode}
-    \caption{ODE vs Binomial}
+    \caption{Comparison of learned coverage sets when PCN interacts with the ODE (i.e., deterministic) variant of the compartment model.
+We observe that, regardless of the chosen budget, PCN learns a similar coverage set on both variants of the model, indicating that it is able to cope with the stochasticity present in the Binomial variant.}
     \label{fig:pf-binomial-ode}
@@
 
-The coverage sets displayed in Figure~\ref{fig:pf-binomal-arh-budgets} correspond to PCN trained on the Binomial model.
-To show that PCN copes with the stochasticity of this model, we compare these coverage sets with the ones learned on the ODE model.
-
-Results are shown in Figure~\ref{fig:pf-binomial-ode}.
-First, in a similar fashion as Figure~\ref{fig:pf-binomal-arh-budgets}, we show the interpolated average coverage set for different budget setting, with PCN trained on the ODE model.
+The coverage sets displayed in Fig.~\ref{fig:pf-binomial-arh-budgets} correspond to PCN trained on the \momdpname\ model, which is stochastic.
+To show that PCN copes with the stochasticity of this model, we compare these coverage sets with the ones learned on the ODE model, which is deterministic.
+
+Results are shown in Fig.~\ref{fig:pf-binomial-ode}.
+First, in a similar fashion as Fig.~\ref{fig:pf-binomial-arh-budgets}, we show the interpolated average coverage set for different budget setting, with PCN trained on the ODE model.
 We observe similar trends as for the Binomial model.
@@
 
-\subsection{Learned coverage sets}
-
-In Figure~\ref{fig:pareto-front-arh}, we display the coverage set with respect to the number of hospitalizations of the learned policies for both the ODE and Binomial variants, as the Binomial variants were omitted in the main paper for clarity.
-When trained on $\mathbf{R}_{\text{ARH}}$, the Binomial coverage set is on par with its ODE counterpart.
-On the other hand, the coverage sets when trained on $\mathbf{R}_{\text{ARI}}$ are slightly worse, as confirmed by the evaluation metrics in Table~\ref{tab:coverage-set}.
-
-\begin{figure}[h]
-    \centering
-    \input{fig_budget}
-    \caption{The Pareto front of policies discovered by PCN, showing the different compromises between the number of hospitalizations and the number of lost contacts.
-Although $\text{PCN}_\text{ARI}$ was trained on the number of infections, it still shows a competitive coverage set w.r.t.
-hospitalizations.}
-    \label{fig:pareto-front-arh}
+\subsection{Comparison of coverage sets learned on $\mathbf{R}_{\text{ARH}}$ and $\mathbf{R}_{\text{ARI}}$}
+\label{sec:cs-arh-vs-ari}
+
+In Sec.~\ref{sec:arh-vs-ari}, we show that we can learn all trade-offs between social burden and hospitalisations, while using the attack rate over infections as reward function.
+
+Results are shown in Fig.~\ref{fig:pf-binomial-arh-ari}.
+We observe that the learned coverage sets are similar, regardless of the budget setting.
+Still, the coverage set of when trained on $\mathbf{R}_{\text{ARI}}$ is systematically dominated by the one when trained on $\mathbf{R}_{\text{ARH}}$.
+\todo{Please check: is this actually true?} While hospitalisations and infections are highly correlated, they differ in terms of age-groups.
+Older age-groups are more susceptible to be hospitalised after being infected, but they do not form the majority of the population.
+For trade-offs where infections and social burden need to be balanced, the proportional reductions target different social environments than for trade-offs balancing hospitalisations and social burden.
+For example, the work environment is majorly comprised of individuals with a more robust immune system, reducing the social contact in this environment greatly affects the number of infections, but has a lesser impact on the number of hospitalisations.
+
+\begin{figure}[t!]
+    \centering
+    \input{fig/arh_binomial_ari}
+    \caption{Comparison of learned coverage sets when PCN learns using the attack rate of infections $\mathbf{R}_{\text{ARI}}$.}
+    \label{fig:pf-binomial-arh-ari}
 \end{figure}
 
-\begin{figure}[h]
-    \centering
-    \input{fig_budget_binomial}
-    \caption{The Pareto front of policies discovered by PCN, showing the different compromises between the number of hospitalizations and the number of lost contacts.
-Although $\text{PCN}_\text{ARI}$ was trained on the number of infections, it still shows a competitive coverage set w.r.t.
-hospitalizations.}
-    \label{fig:pareto-front-arh}
+\subsection{Policy executions}
+
+Depending on the budget, PCN learns a coverage set containing more than 150 different policies.
+To gain a better insight about their behavior, and how they differ from each other, we plot executions of each learned policy in Fig.s~\ref{fig:policy-executions-b2-0}-\ref{fig:policy-executions-b2-140}.
+The plots are displayed from the least restrictive policy in terms of social burden to the most restrictive one.
+Since the \momdpname\ model is stochastic, we show 10 executions of the same policy, on each plot.
+
+\foreach \i/\j in {0/19, 20/39, 40/59, 60/79, 80/99, 100/119, 120/139, 140/153}{
+    \begin{figure}
+        \centering
+        \plotPolicies{\i}{\j}{2}
+        \caption{Execution of policies \i\ to \j, with a budget of 2.}
+        \label{fig:policy-executions-b2-\i}
     \end{figure}
-
-\begin{figure}[h]
-    \centering
-    \input{fig_arh}
-    \caption{The Pareto front of policies discovered by PCN, showing the different compromises between the number of hospitalizations and the number of lost contacts.
-Although $\text{PCN}_\text{ARI}$ was trained on the number of infections, it still shows a competitive coverage set w.r.t.
-hospitalizations.}
-    \label{fig:pareto-front-arh}
+}
+
+\foreach \i/\j in {0/19, 20/39, 40/59, 60/79, 80/99, 100/119, 120/139, 140/159, 160/160}{
+    \begin{figure}
+        \centering
+        \plotPolicies{\i}{\j}{3}
+        \caption{Execution of policies \i\ to \j, with a budget of 3.}
+        \label{fig:policy-executions-b3-\i}
     \end{figure}
-
-\begin{figure}[h]
-    \centering
-\begin{tikzpicture}
-    \pgfplotsset{
-        scale only axis,
-        height=6cm,
-        xlabel={total number of daily new hospitalizations},
-        ylabel={total lost contacts}
 }
-    \begin{axis}[
-    legend entries={{Budget 2 (ODE)}, {Budget 2 (Binomial)}, {Fixed}},
-    width=9.5cm
-    ]
-    \addplot[only marks,draw opacity=0.5,mark=x,red] table [x=o_1, y=o_5, col sep=comma] {data/cs/budget/budget_2.csv};
-    \addplot[only marks,draw opacity=0.5,mark=o,blue] table [x=o_1, y=o_5, col sep=comma] {data/cs/budget/binomial/budget_2.csv};
-    \addplot[only marks,mark=+] table [x=o_0, y=o_1, col sep=comma] {data/cs/cs_fixed.csv};
-    \end{axis}
-\end{tikzpicture}
+
+\foreach \i/\j in {0/19, 20/39, 40/59, 60/79, 80/93}{
+    \begin{figure}
+        \centering
+        \plotPolicies{\i}{\j}{4}
+        \caption{Execution of policies \i\ to \j, with a budget of 4.}
+        \label{fig:policy-executions-b4-\i}
     \end{figure}
-
-\begin{figure}[h]
-    \centering
-\begin{tikzpicture}
-    \pgfplotsset{
-        scale only axis,
-        height=6cm,
-        xlabel={total number of daily new hospitalizations},
-        ylabel={total lost contacts}
 }
-    \begin{axis}[
-    legend entries={{Budget 3 (ODE)}, {Budget 3 (Binomial)}, {Fixed}},
-    width=9.5cm
-    ]
-    \addplot[only marks,draw opacity=0.5,mark=x,red] table [x=o_1, y=o_5, col sep=comma] {data/cs/budget/budget_3.csv};
-    \addplot[only marks,draw opacity=0.5,mark=o,blue] table [x=o_1, y=o_5, col sep=comma] {data/cs/budget/binomial/budget_3.csv};
-    \addplot[only marks,mark=+] table [x=o_0, y=o_1, col sep=comma] {data/cs/cs_fixed.csv};
-    \end{axis}
-\end{tikzpicture}
+
+\foreach \i/\j in {0/19, 20/39, 40/59, 60/79, 80/99, 100/114}{
+    \begin{figure}
+        \centering
+        \plotPolicies{\i}{\j}{5}
+        \caption{Execution of policies \i\ to \j, with a budget of 5.}
+        \label{fig:policy-executions-b5-\i}
     \end{figure}
-
-\begin{figure}[h]
-    \centering
-\begin{tikzpicture}
-    \pgfplotsset{
-        scale only axis,
-        height=6cm,
-        xlabel={total number of daily new hospitalizations},
-        ylabel={total lost contacts}
 }
-    \begin{axis}[
-    legend entries={{Budget 4 (ODE)}, {Budget 4 (Binomial)}, {Fixed}},
-    width=9.5cm
-    ]
-    \addplot[only marks,draw opacity=0.5,mark=x,red] table [x=o_1, y=o_5, col sep=comma] {data/cs/budget/budget_4.csv};
-    \addplot[only marks,draw opacity=0.5,mark=o,blue] table [x=o_1, y=o_5, col sep=comma] {data/cs/budget/binomial/budget_4.csv};
-    \addplot[only marks,mark=+] table [x=o_0, y=o_1, col sep=comma] {data/cs/cs_fixed.csv};
-    \end{axis}
-\end{tikzpicture}
+
+\foreach \i/\j in {0/19, 20/39, 40/59, 60/79, 80/99, 100/116}{
+    \begin{figure}
+        \centering
+        \plotPolicies{\i}{\j}{None}
+        \caption{Execution of policies \i\ to \j, with no budget restriction ($b=\infty$).}
+        \label{fig:policy-executions-binf-\i}
     \end{figure}
-
-\begin{figure}[h]
-    \centering
-\begin{tikzpicture}
-    \pgfplotsset{
-        scale only axis,
-        height=6cm,
-        xlabel={total number of daily new hospitalizations},
-        ylabel={total lost contacts}
 }
-    \begin{axis}[
-    legend entries={{Budget 5 (ODE)}, {Budget 5 (Binomial)}, {Fixed}},
-    width=9.5cm
-    ]
-    \addplot[only marks,draw opacity=0.5,mark=x,red] table [x=o_1, y=o_5, col sep=comma] {data/cs/budget/budget_5.csv};
-    \addplot[only marks,draw opacity=0.5,mark=o,blue] table [x=o_1, y=o_5, col sep=comma] {data/cs/budget/binomial/budget_5.csv};
-    \addplot[only marks,mark=+] table [x=o_0, y=o_1, col sep=comma] {data/cs/cs_fixed.csv};
-    \end{axis}
-\end{tikzpicture}
-\end{figure}
-
-\begin{figure}[h]
-    \centering
-\begin{tikzpicture}
-    \pgfplotsset{
-        scale only axis,
-        height=6cm,
-        xlabel={total number of daily new hospitalizations},
-        ylabel={total lost contacts}
-    }
-    \begin{axis}[
-    legend entries={{Budget N (ODE)}, {Budget N (Binomial)}, {Fixed}},
-    width=9.5cm
-    ]
-    \addplot[only marks,draw opacity=0.5,mark=x,red] table [x=o_1, y=o_5, col sep=comma] {data/cs/budget/budget_None.csv};
-    \addplot[only marks,draw opacity=0.5,mark=o,blue] table [x=o_1, y=o_5, col sep=comma] {data/cs/budget/binomial/budget_None.csv};
-    \addplot[only marks,mark=+] table [x=o_0, y=o_1, col sep=comma] {data/cs/cs_fixed.csv};
-    \end{axis}
-\end{tikzpicture}
-\end{figure}
-
-Moreover, we also show the coverage sets with respect to the number of infections in Figure~\ref{fig:pareto-front-ari}.
-Interestingly, both variants trained on the Binomial model learn a coverage set that almost completely dominates the variants trained on the ODE model.
-This is due to the stochasticity of the Binomial model.
-PCN stores trajectories in its dataset, and keep a selection for training.
-Due to the stochasticity in the transition function, some trajectories might result in higher returns, even though the same policy was applied.
-Due to the high variability in infections, this difference might become significant.
-This is also why the variants using the Binomial model show a worse $I_{\varepsilon}$ and $I_{\varepsilon}-mean$ than their ODE counterpart: their desired return stems from a single trajectory, while we evaluate the policies over multiple trajectories.
-The resulting average return is then different than the desired return.
-Section~\ref{sec:arh-vs-ari} explains why some policies trained on $\mathbf{R}_{\text{ARH}}$ show suboptimal behavior with respect to the number of infections.
-
-\begin{figure}[h]
-    \centering
-    \input{fig_ari}
-    \caption{The Pareto front of policies discovered by PCN, showing the different compromises between the number of hospitalizations and the number of lost contacts.
-Although $\text{PCN}_\text{ARI}$ was trained on the number of infections, it still shows a competitive coverage set w.r.t.
-hospitalizations.}
-    \label{fig:pareto-front-ari}
-\end{figure}
-
-\subsection{Policy executions}
-
-PCN learns a coverage set containing more than 100 different policies.
-To gain a better insight about their behavior, and how they differ from each other, we plot executions of each learned policy in Figure~\ref{fig:policy-executions-ode}.
-The plots are displayed from the least restrictive policy in terms of social burden to the most restrictive one.
-
-We notice that policies~60 to 140 display similar behavior.
-At first, they all completely restrict social contact, effectively continuing the lockdown.
-Afterwards, they progressively lessen these restrictions.
-What differs in these policies is the timing and the speed at which the restrictions are lessened.
-These policies correspond to the right-most part of the coverage set displayed in Figure~\ref{fig:pareto-front-arh}, which shows many alternative for the $[0, 40000]$ hospitalizations segment.
-
-\begin{figure}
-    \centering
-    \plotPolicies{0}{19}
-    \caption{Execution of policies 1 to 20.}
-    \label{fig:policy-executions-ode}
-\end{figure}
-\begin{figure}\ContinuedFloat
-    \centering
-    \plotPolicies{20}{39}
-    \caption{Execution of policies 21 to 40.}
-    \label{fig:policy-executions-ode}
-\end{figure}
-\begin{figure}\ContinuedFloat
-    \centering
-    \plotPolicies{40}{59}
-    \caption{Execution of policies 41 to 60.}
-    \label{fig:policy-executions-ode}
-\end{figure}
-\begin{figure}\ContinuedFloat
-    \centering
-    \plotPolicies{60}{79}
-    \caption{Execution of policies 61 to 80.}
-    \label{fig:policy-executions-ode}
-\end{figure}
-\begin{figure}\ContinuedFloat
-    \centering
-    \plotPolicies{80}{99}
-    \caption{Execution of policies 81 to 100.}
-    \label{fig:policy-executions-ode}
-\end{figure}
-\begin{figure}\ContinuedFloat
-    \centering
-    \plotPolicies{100}{119}
-    \caption{Execution of policies 101 to 120.}
-    \label{fig:policy-executions-ode}
-\end{figure}
-\begin{figure}\ContinuedFloat
-    \centering
-    \plotPolicies{120}{139}
-    \caption{Execution of policies 121 to 140.}
-    \label{fig:policy-executions-ode}
-\end{figure}
 
 \subsection{Experiment parameters}
+\label{sec:pcn-hyperparameters}
 
 We used the same hyper-parameters across all experiments.
-Each experiment resulted in 5 independent trials.
+Each experiment resulted in 10 independent trials.
 Finally, we performed a grid-search over possible hyper-parameter values.
@@
         exploration noise & $0.1$ & $0, 0.1, 0.2$ \\
-        desired return noise & $0.2$ & $0, 0.1, 0.2$ \\
+        desired return noise & $0.05$ & $0, 0.05, 0.1, 0.2$ \\
         reward scaling & $[10000, 100]$ & \\
@@
     \input{tab_nn-architecture}
-    \caption{The 4 different neural network architectures explored for our experiments. All the displayed results use the \texttt{conv1d-big} variant.}
+    \caption{The 4 different neural network architectures explored for our experiments. All the displayed results use the \texttt{dense-big} variant.}
     \label{tab:nn-architecture}
@@
 \subsection{Neural network architecture}
+\label{sec:pcn-architecture}
 
@@
 All the architectures have the same structure.
-We use a compartment embedding $sc_{emb}$, a SCM embedding $sm_{emb}$ and a school-holidays embedding $sh_{emb}$ that take as inputs the compartment, the previous $p_w, p_s, p_l$ values (as they fully define the SCM $\hat{C}$) and a boolean flag for school holidays, respectively.
+We use a compartment embedding $sc_{emb}$, a social contact matrix embedding $sm_{emb}$ and a school-holidays embedding $sh_{emb}$ that take as inputs the compartment, the previous $p_w, p_s, p_l$ values (as they fully define the SCM $\hat{C}$) and a boolean flag for school holidays, respectively.
 All these embeddings have a same-sized embedding of 64, which are multiplied together to form the full state embedding.
@@
 
-All the architectures of the different components are displayed in Table~\ref{tab:nn-architecture}. The variant used in all experiments is \texttt{conv1d-big}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Related Work}
-RL has also been used to learn effective mitigation strategies which limit the spread of COVID-19 \cite{ohi2020exploring}.
-Kwak et al.
-\cite{kwak2021covid} use deep RL, with an SIRD model, to learn efficacy of lockdown and travel restrictions in controlling the COVID-19 pandemic using data collected from various government organisations.
-Bastani et al.
-\cite{bastani2021efficient} use a RL algorithm called Eva to limit the influx of asymptomatic travellers infected with SARS-CoV-2 to Greece.
-Kompella et al.
-\cite{kompella2020reinforcement} define a COVID-19 simulator and a methodology to learn mitigation policies that minimize the economic impact without overwhelming the hospital capacity.
-RL has also been used to learn effective COVID-19 vaccination distribution strategies \cite{beigi2021application,awasthi2020vacsim} and to create decision support systems for mitigating the spread of COVID-19.
-\end{comment}
+All the architectures of the different components are displayed in Table~\ref{tab:nn-architecture}. The variant used in all experiments is \texttt{dense-big}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \end{document}

### 2023-03-31 paper/main.tex
pieter comments

@@
 \newcommand{\foi}{\lambda}
-\newcommand\todo[1]{\textcolor{red}{TODO: #1}}
+\newcommand\todo[1]{\textcolor{red}{(TODO: #1)}}
 
@@
 %mdp state
-\newcommand{\state}{s}
 \newcommand{\mdpstate}{\mathbf{s}}
@@
 	Brussels, Belgium \\
+    \texttt{pieter.libin@vub.be} \\
 % 	\AND
@@
 \section{Introduction}
-Infectious disease outbreaks represent a major global challenge \todo{ref}.
+Infectious disease outbreaks represent a major global challenge \cite{miranda2022tale}.
 To this end, understanding the complex dynamics that underlie these epidemics is essential.
@@
 Typically, a decision maker will rely on an expert's assistance to manually design such a metric.
-However, this is a challenging process, as the decision maker is generally not aware of all possible trade-offs, and as such this may result in a metric that exhibits sub-optimal performance.
-Furthermore, manually designing such metrics is time consuming, costly and error-prone, as this non-intuitive process requires repetitive and tedious tuning to achieve the desired behaviour~\cite{roijers2013survey}.
+However, this is a challenging process, as the decision maker is generally not aware of all possible trade-offs, and as such this may result in a metric that exhibits suboptimal performance.
+Furthermore, manually designing such metrics is time-consuming, costly and error-prone, as this non-intuitive process requires repetitive and tedious tuning to achieve the desired behaviour~\cite{roijers2013survey}.
 Moreover, taking a single objective approach has several other limiting factors.
@@
 This challenging process can be circumvented by taking an explicit multi-objective approach that aims to learn the different trade-offs regarding the considered criteria.
-By assuming that a decision maker will always prefer solutions for which at least one objective improves, it is possible to learn a set of optimal solutions referred to as the \emph{Pareto front}.
+By assuming that a decision maker will always prefer solutions for which at least one objective improves, it is possible to learn a set of optimal solutions referred to as the \emph{Pareto front}~\cite{hayes2021practical}.
 This enables decision makers to review each solution on the Pareto front before making a decision, thereby being aware of the trade-offs that each solution implies.
@@
 We consider the first wave of the Belgian COVID-19 epidemic, which was mitigated by a strict lockdown~\cite{willem2021impact}.
-When the incidence of confirmed cases was steadily decreasing, epidemiological experts were tasked to investigate deconfinement strategies, to reduce the severe contact and mobility reductions.
-Here, we consider an epidemiological model that was constructed to describe the Belgian COVID-19 epidemic, and was fitted to hospitalisation incidence data and serial sero-prevalence data~\citet{abrams2021modelling}.
+When the incidence of confirmed cases was steadily decreasing, epidemiological experts were tasked to investigate deconfinement strategies, to reduce the severe social contact and mobility restrictions.
+Here, we consider an epidemiological model that was constructed to describe the Belgian COVID-19 epidemic and was fitted to hospitalisation incidence data and serial sero-prevalence data~\cite{abrams2021modelling}.
 This model constitutes a stochastic discrete-time age-structured compartmental model that simulates mitigation strategies by varying social distancing parameters concerning school, work and leisure contacts.
@@
 
-To learn and explore the trade-offs between the attack rate and social burden we present a state-of-the-art MORL approach based on Pareto Conditioned Networks (PCN)~\cite{reymond2022pcn}.
+To learn and explore the trade-offs between the attack rate and social burden we use a state-of-the-art MORL approach based on Pareto Conditioned Networks (PCN)~\cite{reymond2022pcn}.
 PCN uses a single neural network to learn the policies that belong to the Pareto front.
@@
 
-By evaluating the solution set of mitigation policies learned by PCN, we observe that PCN minimises the social burden in scenarios where hospitalization rates are sufficiently low \todo{update (challenges)}.
-Therefore, in this work we illustrate that multi-objective reinforcement learning can be utilised to provide important insights surrounding the trade-offs between complex mitigation polices in real-world epidemiological models.
+By evaluating the solution set of mitigation policies learned by PCN, we observe that PCN minimises the social burden in scenarios where hospitalization rates are sufficiently low.
+Therefore, in this work we illustrate that multi-objective reinforcement learning can provide important insights concerning the trade-offs between complex mitigation polices in real-world epidemiological models.
 
@@
 \section{Background}
-In the following, we use bold notation $\momdpvaluefunction$ to denote a vectorial value, while non-bold notation $\mdpvaluefunction$ for its scalar counterpart.
+In the following, we use bold notation to denote variables with a vectorial value, while non-bold notation for its scalar counterpart.
 
@@
 While MOMDPs define the actions an agent can take, it does not define the agent's behavior, i.e., which actions are taken in each state.
-Given this MOMDP, an agent follows a policy $\mdppolicy$, that expresses the probability to take action $\action \in \mdpactionspace$ when in state $\state \in \mdpstatespace$: $\mdpstatespace \times \mdpactionspace \to \left[ 0, 1 \right]$.
-We measure the performance of $\mdppolicy$ through the expected sum of discounted rewards (denoted the \emph{Value}) it achieves from the initial state $\state_0$ until the end of the decision problem:
-
-\begin{equation}
-    \momdpvaluefunction^\mdppolicy = \mathbb{E} \left [ \sum_{t=0}^h \mdpdiscount^t \momdpreward_t |\ \mdppolicy, \state_0 \right ],
-\end{equation}
-%
-where $\momdpreward_t$ corresponds to the multi-objective reward observed at time $t$, by following policy $\mdppolicy$, given by the reward function $\mdprewardfn(\state_t, \action_t, \state_{t+1})$, where $\state_t$ is the current state, $\action_t$ corresponds to the action that was taken, and $\state_{t+1}$ signifies the state reached by taking action $\action_t$.
-
-For single-objective RL, where $n = 1$, the goal is to find the policy $\mdppolicy^{*}$ that maximizes the $\mdpvaluefunction$-value:
+Given this MOMDP, an agent follows a policy $\mdppolicy$, that expresses the probability to take action $\action \in \mdpactionspace$ when in state $\mdpstate \in \mdpstatespace$: $\mdpstatespace \times \mdpactionspace \to \left[ 0, 1 \right]$.
+We measure the performance of $\mdppolicy$ through the expected sum of discounted rewards (denoted the \emph{Value}) it achieves from the initial state $\mdpstate_0$ until the end of the decision problem:
+
+\begin{equation}
+    \momdpvaluefunction^\mdppolicy = \mathbb{E} \left [ \sum_{t=0}^h \mdpdiscount^t \momdpreward_t |\ \mdppolicy, \mdpstate_0 \right ],
+\end{equation}
+%
+where $\momdpreward_t$ corresponds to the multi-objective reward observed at time $t$, by following policy $\mdppolicy$, given by the reward function $\mdprewardfn(\mdpstate_t, \action_t, \mdpstate_{t+1})$, where $\mdpstate_t$ is the current state, $\action_t$ corresponds to the action that was taken, and $\mdpstate_{t+1}$ signifies the state reached by taking action $\action_t$.
+
+On the one hand, for single-objective RL, where $n = 1$, the goal is to find the policy $\mdppolicy^{*}$ that maximizes the $\mdpvaluefunction$-value:
 \begin{equation}
@@
 
-In contrast, for MORL, $n > 1$ which leads to vectorial returns.
+On the other hand, for MORL, $n > 1$ which leads to vectorial returns.
 In this case, there can be policies for which, without any additional information, it is impossible to know if one is better than the other.
@@
 Intuitively, one would generally prefer the algorithm that covers a wider range of decision maker preferences.
-In this work we use several metrics to evaluate our algorithm's performance.
+%In this work we use several metrics to evaluate our algorithm's performance.
 
@@
 %
-where $||.||_{\infty}$, the L-infinity norm, returns the magnitude of the largest entry of a vector.
+where $||.||_{\infty}$, the L-infinity norm, is defined as the magnitude of the largest entry of a vector.
 
@@
 We note that the $\varepsilon$-indicator metric is quite pessimistic, as it measures worst-case performance~\cite{zintgraf2015}, i.e., it will still report low performance as long as a single point of the Pareto front is incorrectly modeled, even if all the other points are covered.
-As such, we also use the $I_{\varepsilon-mean}$ \cite{reymond2022pcn} which measures the \emph{average} $\varepsilon$ distance of the solutions in $\mathcal{F}$ with respect to the evaluated coverage set.
+As such, we also use the $I_{\varepsilon-mean}$ which measures the \emph{average} $\varepsilon$ distance of the solutions in $\mathcal{F}$ with respect to the evaluated coverage set~\cite{reymond2022pcn}.
 
@@
     \caption{Example of a Pareto front (black dots) and a coverage set (white dots) in a 2-objective environment.
-The hypervolume metric (in light blue) measures the volume of all dominated solutions with respect to some reference point (cross).
+The hypervolume metric (in light blue) measures the volume of all dominated solutions with respect to a reference point (cross).
 The reference point is taken as a lower bound on the achievable returns, which is why it can differ from the origin.
@@
 
-A visual representation of the model is depicted in Fig~\ref{fig:compartment_model}.
-It shows the different compartments, as well as the flow rates at which individuals move from one compartment to another.
+A visual representation of the model is depicted in Fig.~\ref{fig:compartment_model}. It shows the different compartments, as well as the flow rates at which individuals move between compartments.
 
@@
 %
-where, for example, $\textbf{S} = (S_{1}(t), S_{2}(t), ..., S_{\agegroup}(t))^{T}$ is the vector representing the susceptible members of the population of each age group $\agegroup$ at time $t$.
 
 In this set of ordinary differential equations, each state variable represents a vector over all age groups for a particular compartment at time $t$.
+For example, $\textbf{S} = (S_{1}(t), S_{2}(t), ..., S_{\agegroup}(t))^{T}$ is the vector representing the susceptible members of the population of each age group $\agegroup$ at time $t$.
 Infection dynamics are governed by an age-specific force of infection $\foi$:
@@
 %
-where $\agegroup$ is the respective age group of a total of $\agegroups$ age groups, and $\beta(\agegroup,\agegroup')$ is the time-invariant transmission rate that encodes the average per capita rate at which an infectious individual in age group $\agegroup$ makes an effective contact with a susceptible individual in age group $\agegroup'$, per unit of time.
+where $\agegroups$ is the total number of age groups, and $\beta(\agegroup,\agegroup')$ is the time-invariant transmission rate that encodes the average per capita rate at which an infectious individual in age group $\agegroup$ makes an effective contact with a susceptible individual in age group $\agegroup'$, per unit of time.
 
@@
 Following \citet{abrams2021modelling}, we rely on distinct social contact matrices for symptomatic and asymptomatic individuals, respectively $\ctotal_{s}$ and $\ctotal_{a}$.
-Therefore, it is possible to define the transmission rates for both symptomatic and asymptomatic individuals as follows:
+Therefore, we define the transmission rates for both symptomatic and asymptomatic individuals as follows:
 
@@
 
-Moreover, the age-dependent force of infection can be defined as follows:
+The age-dependent force of infection can be defined as follows:
 
@@
 %
-where $\boldsymbol{\lambda}(t) = (\lambda(1, t), \lambda(1, t), ..., \lambda(\agegroups, t))$.
-For all further information about the different compartments and parameters please refer to the work of \citet{abrams2021modelling}.
-
-By formulating the set of differential equations defined above, as a chain-binomial, we can obtain stochastic trajectories from this model \cite{bailey1975mathematical}.
-A chain-binomial model assumes a stochastic model where infected individuals are generated by some underlying probability distribution.
-For the stochastic model we consider a time interval $(t, t +h]$, where $h$ is defined as the length between two consecutive time points.
-Similar to \citet{abrams2021modelling}, in this work we set $h = \frac{1}{24}$.
+where $\boldsymbol{\lambda}(t) = (\lambda(1, t), \lambda(2, t), ..., \lambda(\agegroups, t))$.
+For all further information about the different compartments and parameters we refer the reader to the work of \citet{abrams2021modelling}.
 
@@
 
-\todo{This is why we use the binomial model in our analysis.}
+By formulating the set of differential equations defined above, as a chain-binomial, we can obtain stochastic trajectories from this model \cite{bailey1975mathematical}.
+A chain-binomial model assumes a stochastic model where infected individuals are generated by some underlying probability distribution.
+For this stochastic model we consider a time interval $(t, t +h]$, where $h$ is defined as the length between two consecutive time points.
+In this work we set $h = \frac{1}{240}$.
 
@@
 We denote $\hat{C}_t$ the social contact matrix at timestep $t$, resulting from the reduction function $\hat{C}$.
-Secondly, we assume that compliance to the interventions is gradual and model this using a logistic compliance function (see details in Sec.~\ref{sec:compliance-function}).
+Secondly, we assume that compliance to the interventions is gradual and model this using a logistic compliance function (see details in Sec.~\ref{sec:compliance-function} of the Appendix).
 
@@
 
-\paragraph{Transition function:} The model defined by \citet{abrams2021modelling} utilises a model transition probability $M(\mdpstate_m' \mid \mdpstate_m, \hat{C}, p_{w}, p_{s}, p_{l})$ (see Sec.~\ref{sec:binomial-model} for details on $M$), where $\hat{C}(p_{w}, p_{s}, p_{l})$ returns the currently installed social contact matrix, that progresses the epidemiological model in one timestep.
+\paragraph{Transition function:} The model defined by \citet{abrams2021modelling} utilises a model transition probability $M$ (see Sec.~\ref{sec:binomial-model} of the Appendix for details on $M$), that progresses the epidemiological model in one timestep based on the currently installed social contact matrix $\hat{C}(p_{w}, p_{s}, p_{l})$.
 We use this function as the transition function in \momdpname.
 
-In a classical MDP, executing an action $\action_t$ in any state $\state_t$ leads to a next state $\state_{t+1}$ according to the transition function $\mdptransition$.
+In a classical MDP, executing an action $\mdpaction_t$ in any state $\mdpstate_t$ leads to a next state $\mdpstate_{t+1}$ according to the transition function $\mdptransition$.
 At every timestep $t$, the agent is free to choose the action to perform.
@@
 
-We parameterize the epidemiological model using the mean of the posteriors as specified by \citet{abrams2021modelling} (details in Sec.~\ref{sec:compartment-model-parameters}).
+We parameterize the epidemiological model using the mean of the posteriors as specified by \citet{abrams2021modelling} (details in Sec.~\ref{sec:compartment-model-parameters} of the Appendix).
 
@@
 As such, during the simulated week, there is a gradual shift from the current $\hat{C}_t$ to the new social contact matrix, $\hat{C}_{t+1}$.
-As such, we need to maintain the current $\hat{C}_t$, to maintain a Markovian environment.
-
+Thus, we need to include the current $\hat{C}_t$, to maintain a Markovian environment.
 Secondly, we require the current $\hat{C}_t$ to determine whether the action changes the social restrictions in place, and thus consume part of the budget.
@@
 \begin{equation}
-    \mdpstate = \mdpstate_m \cup \hat{C} \cup b
+    \mdpstate = \mdpstate_m \cup \hat{C} \cup \budget
 \end{equation}
@@
 We can thus train a policy that, conditioned on a desired return, provides the optimal action to reach said return.
-By leveraging the generalization properties of neural networks, we can accumulate incrementally better experience by conditioning on increasingly higher reward-goals
+By leveraging the generalization properties of neural networks, we can accumulate increasingly better experience by conditioning on increasingly higher reward-goals.
 
 Instead of using a neural network that takes a state $\mdpstate$ as input, PCN uses a single neural network that takes a tuple $\ltuple \mdpstate, \hat{h}, \mathbf{\hat{R}} \rtuple$ as input.
-$\mathbf{\hat{R}}$ represents the \emph{desired return} of the decision maker, i.e.
-the return PCN should reach at the end of the episode.
+$\mathbf{\hat{R}}$ represents the \emph{desired return} of the decision maker, i.e., the return PCN should reach at the end of the episode.
 $\hat{h}$ denotes the \emph{desired horizon} that expresses the number of timesteps that should be executed before reaching $\mathbf{\hat{R}}$.
@@
     \item Use the current policy to interact with the environment and generate trajectories (see Sec.~\ref{sec:pcn-policy-exploration}).
-Initially, use a random policy to fill the dataset with random trajectories.
+Initially, a random policy is used to fill the dataset with random trajectories.
     \item Update the dataset to incorporate these new trajectories and prune the least interesting trajectories in terms of returns (see Sec.~\ref{sec:pcn-building-dataset}).
@@
 
-Given a trajectory composed of $T$ transitions $\ltuple s_0, a_0, \bm{r}_0, s_1\rtuple, \dots, \ltuple s_{T-1}, a_{T-1}, \bm{r}_{T-1}, s_T\rtuple$ we can compute, for each transition at timestep $t$, the future discounted return $\bm{R}_t = \sum_{i=t}^T \gamma^i\bm{r}_i$ and the leftover horizon $h_t = T-t$.
+Given a trajectory composed of $T$ transitions $\ltuple \mdpstate_0, \action_0, \bm{r}_0, \mdpstate_1\rtuple, \dots, \ltuple \mdpstate_{T-1}, \action_{T-1}, \bm{r}_{T-1}, \mdpstate_T\rtuple$ we can compute, for each transition at timestep $t$, the future discounted return $\bm{R}_t = \sum_{i=t}^T \gamma^i\bm{r}_i$ and the leftover horizon $h_t = T-t$.
 Since for this trajectory executing action $a_t$ in state $s_t$ resulted in return $\mathbf{R}_t$ in $h_t$ timesteps, we add a datapoint with input $\langle s, \hat{h}, \mathbf{\hat{R}}\rangle = \langle s_t, h_t, \mathbf{R}_t \rangle$ and output $a=a_t$ to the dataset.
@@
 
-Since we have a fix-sized dataset, adding this trajectory means we need to remove another one.
+So that our neural network is trained on relevant transitions only, we keep a fix-sized dataset.
+Thus, adding a trajectory means we need to remove another one.
 Our aim is to keep trajectories that span different parts of the objective space, such that we can explore as many types of trade-offs as possible.
 However, to improve the trade-offs we currently have in our dataset, we need to only keep solutions that are close to our current estimate of the Pareto front.
-Thus, we assign a priority for each trajectory in our dataset that combines two metrics: the \emph{crowding distance}, which measures the distance of a solution with its closest neighbors, and the \emph{$l_2$-norm} with the closest non-dominated solution, which indicates how close the solution is from our current estimate of the Pareto front.
+Thus, we assign a priority for each trajectory in our dataset that combines two metrics: the \emph{crowding distance}~\cite{deb2000}, which measures the distance of a solution with its closest neighbors, and the \emph{$\ell^2$-norm} with the closest non-dominated solution, which indicates how close the solution is from our current estimate of the Pareto front.
 
@@
 More precisely, we randomly select a non-dominated return $\mathbf{R}_{nd}$ and its corresponding horizon $\hat{h}$ from the dataset.
-By randomly picking a non-dominated return from the entire coverage set we ensure equal chance of improvement to each part of the objective space.
+By randomly picking a non-dominated return from the entire coverage set we ensure equal probability to improve each part of the objective space.
 To generate trajectories that improve upon $\mathbf{R}_{nd}$, we randomly select an objective $o$ and increase $R_{nd,o}$ by a sample from the uniform distribution $\mathcal{U}(0, \sigma_o)$, where $\sigma_o$ is the standard deviation for the selected objective, using all non-dominated returns from the trajectories in the dataset.
 This then becomes our desired return $\bm{\hat{R}}$.
-By restricting the improvement to at most $\sigma_o$, $\mathbf{\hat{R}}$ stays in the range of possible achievable returns and, by only modifying one objective at a time, the changes to the network's input compared to the training data are kept at a minimum.
+By restricting the improvement to at most $\sigma_o$, $\mathbf{\hat{R}}$ stays in the range of achievable returns.
+By only modifying one objective at a time, the changes to the network's input compared to the training data are kept at a minimum.
 
@@
 We change the output of the neural network such that there is a single output value for each dimension of the action-space.
-Since the actions should be bound in the domain of possible actions ($[0,1]$ in the case of \momdpname, see Sec.~\ref{sec:sars-cov2-momdp}), we apply a tanh non-linearity function on each output.
+Since the actions should be bound in the domain of possible actions ($[0,1]$ in the case of \momdpname, see Sec.~\ref{sec:sars-cov2-momdp}), we apply a sigmoid non-linearity function on each output, as the output of the sigmoid function is bound in $[0,1]$.
 As such, we have a regression problem instead of a classification problem, as the labeled dataset now uses continuous labels $y = \mdpaction_t$ instead of categories.
@@
 
-Since learning the full set of Pareto-efficient policies $\Pi^*$ requires that the policies $\pi^* \in \Pi^*$ are deterministic stationary policies \cite{roijers2013survey}, we use the output $\mathbf{\hat{y}}$ as action at execution time.
+Learning the full set of Pareto-efficient policies $\Pi^*$ requires that the policies $\pi^* \in \Pi^*$ are deterministic stationary policies.
+If stochastic policies are permitted, the set of optimal policies corresponds to the convex part of the Pareto front~\cite{roijers2013survey}.
+However, we argue that, in the context of mitigation policies, deterministic policies are required, as the population needs to be informed in advance of imposed measures.
+Thus, we use the output $\mathbf{\hat{y}}$, which is deterministic, as action at execution time.
 However, PCN improves its policy through exploration, by continuously updating its dataset with better trajectories.
@@
 However, when $\mdptransition$ is stochastic this can no longer be guaranteed.
-To mitigate this, we add small random noise to $\mathbb{R}_t$ when performing gradient descent, which reduces the risk of overfitting~\cite{zur2009noise}.
+To mitigate this, we add a limited amount of random noise to $\mathbf{R}_t$ when performing gradient descent, which reduces the risk of overfitting~\cite{zur2009noise}.
 Moreover, while the \momdpname\ model is stochastic, the variation is entirely due to the sampling of the binomial distributions in the binomial-chain.
-While this variation accumulates over time, the time window we consider for each timestep (i.e., one week) is short enough that the accumulation stays limited.
+While this variation accumulates over time, the time window we consider for each timestep (i.e., one week) is short enough that the accumulation remains bounded.
 Thus, the possible next-states resulting from a state-action pair are similar to each other.
@@
 We assess the validity of our approach by executing PCN on this deterministic variant, and observing similar performance as with the \momdpname\ model.
-We show the results in Sec.~\ref{sec:cs-ode-vs-binomial}.
+We show the results in Sec.~\ref{sec:cs-ode-vs-binomial} of the Appendix.
 
@@
 It is from this day onward that PCN aims to learn policies for diverse exit strategies, compromising between the total number of daily hospitalizations and the total number of contacts lost as a proxy for social burden.
-The simulation lasts throughout the school holidays (from 01/07/2020 to 31/08/2020).
+The simulation lasts throughout the summer school holidays, from 01/07/2020 to 31/08/2020.
 Schools are closed during the school holidays, which is simulated by setting $p_s = 0$, regardless of the corresponding value outputted by the policy, i.e., during periods of school closure $p_s$ is ignored.
@@
 We draw the analogy with the multi-phase exit strategy established by the Belgian government and the restriction on the number of action-changes imposed by the \momdpname's budget $\budget$.
-Indeed, on the 11th of May 2020, exactly one week after the end of the lockdown, stores and certain types of jobs were allowed to reopen, under strict conditions.
+Indeed, on the 11th of May 2020, exactly one week after the end of the lockdown, stores and certain companies were allowed to reopen, under strict conditions.
 This corresponds to altering $p_w$ in our MOMDP.
@@
 Further changes of restrictions occured on the 8th of June, 1st, 9th and 25th of July.
-Thus, we argue that, with a limited budget, we can achieve realistic and implementable policies.
+Thus, we argue that, with a limited budget, we can achieve realistic policies.
 In our experiments, we consider budgets of 2 to 5, as these closely relate to the number of changes that occured until the end of the summer holidays of 2020.
@@
 All experiments are averaged over 10 runs.
-The hyper-parameters and the neural network architecture can be found in Sec.s~\ref{sec:pcn-hyperparameters} and Sec.~\ref{sec:pcn-architecture}, respectively.
+The hyper-parameters and the neural network architecture can be found in Sec.s~\ref{sec:pcn-hyperparameters} and Sec.~\ref{sec:pcn-architecture} of the Appendix, respectively.
 
@@
 
-Interestingly, we notice that the most restrictive policy (i.e.
-the one that prioritizes hospitalizations over social burden, see Fig.~\ref{fig:policy-execution}, bottom plot) still starts to gradually increase $p_w$ and $p_l$ from the end of July onward.
+Interestingly, we notice that the most restrictive policy (i.e., the one that prioritizes hospitalizations over social burden, see Fig.~\ref{fig:policy-execution}, bottom plot) still starts to gradually increase $p_w$ and $p_l$ from the end of July onward.
 This is because by then, the epidemic has mostly faded out, and it is safe to reduce social restrictions.
@@
 Finally, the top plot shows that, without imposing social restrictions, the number of hospitalisations peaks on the 15th of June.
-By the beginning of July, the epidemic has spread out over the majority of the population \todo{herd immunity?}, and the number of admissions at the hospital has been reduced to a fraction of the number of hospitalisations at the peak.
-Thus, without social restrictions, we do not take advantage of the natural decrease of social contacts due to school holidays, as the majority of the population has already been infected before the start of the holidays.
-\todo{mention the slow decrease of $p_l$ during the summer?
-This is simply due to PCN not having found better}
+By the beginning of July, the epidemic has spread out over the majority of the population \todo{epi friends: can we mention here that we reached herd immunity?}, and the number of admissions at the hospital has been reduced to a fraction of the number of hospitalisations at the peak.
+Thus, without social restrictions, we do not take advantage of the natural decrease of social contacts due to school holidays, as a significant proportion of the population has already been infected before the start of the holidays.
 
@@
 In general, an increase of budget is associated with an increasingly better coverage set, as policies learned using a higher budget dominate the ones learned with a lower budget.
-This is to be expected, as a higher budget give the agent more freedom to change its actions as the epidemic progresses.
-However, we also notice that this increase of coverage sets is not necessarily gradual.
+This is to be expected, as a higher budget gives the agent more freedom to change its actions as the epidemic progresses.
+%However, we also notice that this increase of coverage sets is not necessarily gradual.
 % Between the run with a budget of 2 and budget of 3, the difference in coverage set seems marginal.
@@
 Moreover, we observe that the difference is concentrated around the less restrictive policies in terms of social burden.
-We hypothesise that this region contains the most complex policies, as these try to maintain as much social freedom as possible, while containing the number of hospitalisations.
-In these cases, the timing of the actions greatly correlate with the timing and duration of the peak of the epidemic, and a higher budget allows for more fine-grained control about managing this timing.
+We postulate that this region contains the most complex policies, as these try to maintain as much social freedom as possible, while containing the number of hospitalisations.
+In these cases, the timing of the actions coincide with the timing and duration of the peak of the epidemic, and a higher budget allows for more fine-grained control to manage this timing.
 To confirm this, we select, for each budget setting, the solution where the difference in performance is most noticeable, corresponding to the solutions with a total number of hospitalisations around $80000$ (which is in the middle of the range of possible hospitalisations, as can be seen in Fig.~\ref{fig:pf-binomial-arh-budgets}).
@@
 First, we observe that the lower-budget policies are unable to reduce the social restrictions past the peak of the epidemic.
-In contrast, the setting with no budget restrictions finely controls the restrictions as the epidemic progresses, completely removing restrictions by the end of the wave.
-Second, we note that the policy with a budget of 5 ressembles the execution of the one without restrictions.
+In contrast, the setting with no budget restrictions meticulously controls the restrictions as the epidemic progresses, completely removing restrictions by the end of the wave.
+Second, we note that the policy with a budget of 5 resembles the execution of the one without restrictions.
 However, due to its budget, the policy is unable to progressively reduce the restrictions and instead resorts to a halfway compromise.
@@
 As such, it is easier to learn the most extreme policies.
-Thus, in the specific case where we impose no budget, the freedom of action actually hinders PCN's search, for certain regions of the rweard-space.
+Thus, in the specific case where we have an unlimited budget, the freedom of action actually hinders PCN's search, for certain regions of the reward-space.
 
@@
 Next, we assess the difference in coverage sets when optimizing on $\mdprewardfn_\text{ARH}$ versus $\mdprewardfn_\text{ARI}$.
-Although these reward functions have a different scale (there are more infected persons than hospitalised ones), our experiments show that infections and hospitalizations are correlated.
+Although these reward functions have a different scale (there are more infected persons than hospitalised ones), our experiments show that infections and hospitalizations are tightly correlated.
 This is expected, as during the initial phase of the epidemic, limited immunity was present in the population (i.e., limited natural immunity and no vaccines), which induces a tight coupling between infection and hospitalisation cases.
@@
 However, regardless of the imposed budget, the hypervolumes are slightly worse.
-This is to be expected, since those experiments are not directly optimising on $\mdprewardfn_\text{ARH}$.
-We make similar conclusions for $I_\varepsilon$.
-For budgets 2, 3 and 5, the difference between the worst-performing policy for the $\mdprewardfn_\text{ARI}$ variant and the $\mdprewardfn_\text{ARH}$ is less than $0.01$, indicating less than $1\%$ difference in return values between the two variants when comparing their worst-performing policy.
+This is to be expected, since those experiments are not directly optimised on $\mdprewardfn_\text{ARH}$.
+We draw a similar conclusion for $I_\varepsilon$: for budgets 2, 3 and 5, the difference between the worst-performing policy for the $\mdprewardfn_\text{ARI}$ variant and the $\mdprewardfn_\text{ARH}$ is less than $0.01$, indicating less than $1\%$ difference in return values between the two variants when comparing their worst-performing policy.
 As an exception, we notice that PCN without budget restrictions results in better performance across every metric for the $\mdprewardfn_\text{ARI}$ variant.
-Still, due to the high standard deviation of the no-budget, $\mdprewardfn_\text{ARH}$ setting, we do not believe this difference is meaningful.
-\todo{some kind of meaningful conclusion}
+Still, due to the high standard deviation of the unlimited budget, $\mdprewardfn_\text{ARH}$ setting, we do not believe this difference is meaningful.
+Thus, we could optimise on the attack rate of hospitalisations with $\mdprewardfn_\text{ARI}$.
+As there is a 2 week delay for hospitalisations, this would facilitate learning policies to react to unexpected changes earlier than using $\mdprewardfn_\text{ARH}$, given that a good proxy to the actual number of infections was available (e.g., due to a scale up of PCR testing, as was the case after the first lockdown).
 
@@
 The dataset of trajectories that PCN is trained on is pruned over time to keep only the most relevant trajectories.
-The returns of these trajectories are used in Fig.~\ref{fig:pf-binomial-arh-budgets} to show the learned coverage set.
+The returns of these trajectories are used in Fig.~\ref{fig:pf-binomial-arh-budgets} to visualize the learned coverage set.
 Each of these returns can be used as desired return for policy execution.
@@
 Since the \momdpname\ model is stochastic, for multiple executions of the same policy, the executions kept in the coverage sets are the ones for which the samples from the binomial-chain resulted in a better progression of the epidemic than average.
-Thus, we expect our policy-executions to be close to the target selected from the coverage set, but not on exactly target.
+Thus, we expect our policy-executions to be close to the target selected from the coverage set, but not exactly on target.
 Based on this analysis, we conclude that the policies trained by PCN are robust and produce returns as close as possible from their chosen target.
@@
 \section{Related work}
-Reinforcement learning (RL) has been used in conjunction with epidemiological models to learn policies to limit the spread of diseases and predict the effects of possible mitigation strategies \cite{probert2019context}.
-For example, RL has been used to model the mitigation of influenza \cite{das2008large,libin2020,libin2018bayesian}.
+Reinforcement learning (RL) has been used in conjunction with epidemiological models to learn policies to limit the spread of diseases and predict the effects of possible mitigation strategies \cite{probert2019context,libin2018bayesian,libin2020}.
 
@@
 Making decisions on how to mitigate epidemics has important ethical implications with respect to public health and societal burden.
-In this regard, it is crucial to approach this decision making from a balanced perspective, to which end we argue that multi-objective decision making is crucial.
-In this work, we establish a novel approach, i.e., an expert system, to study multi-faceted policies, and this approach shows great potential with respect to future epidemic control.
-We are aware of the ethical implications that expert systems have on the decision process and we make the disclaimer that all results based on, or derived from, the expert system that we propose should be carefully interpreted by experts in the field of public health, and in a much broader context of economics, well-being and education.
+In this regard, it is crucial to approach this decision making from a balanced perspective, to which end we argue that multi-objective decision making is essential.
+In this work, we establish a novel approach, i.e., an expert system, to study multi-faceted policies, and this approach shows great potential to study future epidemic mitigation policies.
+We are aware of the ethical implications that expert systems have on the decision process and we make the disclaimer that all results based on the expert system that we propose should be carefully interpreted by experts in the field of public health, and in a broader context that encompasses health economics, well-being and education \todo{epi friends, other things to add here?
+To make sure we are exhaustive}.
 We note that the work in this manuscript was conducted by a interdisciplinary consortium that includes computer scientists and scientists with a background public health, epidemiology and bio-statistics.
 
-In this work, we focus on the clinical outcomes of the intervention strategies and use the reduced contacts as proxy for the quality of life lost.
-This could be extended into more formal economic evaluations by assessing the health and monetary benefits and costs of different interventions for various stakeholders.
-The COVID-19 pandemic demonstrated the broad impact of infectious diseases on sectors outside health care.
+In this work, we focus on the clinical outcomes of intervention strategies and use the reduced contacts as proxy for social burden.
+This could be extended into more formal health economic evaluations, by designing reward functions that explicitely consider distinct health economic principles.
+The COVID-19 pandemic demonstrates the broad impact of infectious diseases on sectors other than health care.
 This stresses the need to capture a societal and thus multi-objective perspective in the decision making process on public health and health care interventions.
@@
 
-\todo{Although we use an age-structured compartment model, with social contact matrices to model social interactions, this remains a model that evaluates the progression of the pandemic as an aggregated process over the population.
+Although we use an age-structured compartment model, with social contact matrices to model social interactions, this remains a model that evaluates the progression of the pandemic as an aggregated process over the population.
 Individual-based models could provide more detailed and localized policies, potentially further improving the quality of possible trade-offs, and would provide an interesting avenue for future work.
-However, due to the computational cost of simulating such models, and the number of interactions required by reinforcement learning in general, this remains a challenging problem.}
-
-\todo{Moreover, although we are able to interpret and analyse the obtained policies and their corresponding trade-offs, as we can plot the Pareto front for 2 objectives, this approach cannot be used for problems with more objectives.
-Defining interpretation and visualization tools for many-objective problems is necessary if we want to include additional objectives for our problem.}
-
-\todo{Additionally, PCN is able to cope with model stochasticity, because the stochasticity is limited.
-In settings where the stochasticity is more pronounced, e.g., designing policies to control the initial outbreak of an epidemic, further methodological extensions are warranted.}
-
-\todo{Finally, while we have focused on social burden as a trade-off for hospitalisations, our approach can be used for related problems, involving different compromises, such as balancing the number of lost schooldays with respect to new infections in schools \cite{torneri2021controlling}, contact tracing effort versus its effect \cite{willem2021impact}, the impact of antivirals on the epidemic versus resistance to antivirals \cite{torneri2020prospect}, or the cost of universal testing versus its impact on the epidemic \cite{libin2021assessing}.}
-
-To conclude, we show that multi-objective reinforcement learning can be used to learn a wide set of high-quality policies on real-world problems, providing the decision maker with insightful and diverse alternatives, and showing the impact of taking extreme measures.
-\todo{This can be done for realistic settings, where we budget the number of changes of social reductions, such that the policies are practical and can be deployed in real-life scenarios.}
+However, due to the computational cost of simulating such models, and the number of interactions required by reinforcement learning in general, this remains a challenging problem, that will require fundamental research to improve the sample efficiency of  multi-objective reinforcement learning algorithms.
+
+While we are able to interpret and analyse the obtained policies and their corresponding trade-offs, as we can plot the Pareto front for two objectives, this approach cannot be used for problems with more objectives, which will be necessary to cover reward functions that cover distinct health economic principles.
+To facilitate this kind of research, new algorithms are necessary to enable reinforcement learning in many-objective contexts and to interprete the learnt policies.
+
+In this work, PCN is able to cope with model stochasticity, as the stochasticity is limited.
+In settings where the stochasticity is more pronounced, e.g., when designing policies to control the initial outbreak of an epidemic, further methodological extensions are warranted.
+
+Finally, in this work we studied policies that aim to balance social burden and hospitalizations.
+Yet, the methodology that we propose shows promise to address a wide variety of public health challenges, such as balancing the number of lost schooldays with respect to the attack rate of infections in schools \cite{torneri2021controlling}, contact tracing effort compared to the impact of such policies \cite{willem2021impact}, the impact of antivirals on the epidemic while balancing the likelihood for resistance mutations to emerge \cite{torneri2020prospect}, and to balance the cost of universal testing and its impact on an emerging epidemic \cite{libin2021assessing}.
+
+To conclude, we show that multi-objective reinforcement learning can be used to learn a wide set of high-quality policies on real-world problems, providing the decision maker with insightful and diverse alternatives, and showing the impact of extreme measures.
+Furthermore, we show that action budgets can act as a regulariser that facilitates learning realistic policies that can be easily conveyed to decision makers.
 
@@
 To create a version of \momdpname\ with a stochastic transition function, $M$, we utilise the stochastic compartmental model outlined above.
-Given the transitions within the compartmental model are derived by an underlying probability distribution it is possible to utilise the stochastic compartmental model transitions for \momdpname\.
-As previously outlined in Sec.
-\ref{sec:deterministic_comp_model} the contact matrix $\hat{C}$ applied the model state $s_{m}$ progresses the model and returns a new model state $s'_{m}$.
+Given the transitions within the compartmental model are derived by an underlying probability distribution it is possible to utilise the stochastic compartmental model transitions for \momdpname.
+As previously outlined in Sec.~\ref{sec:sars-cov2-momdp_model} the contact matrix $\hat{C}$ applied the model state $s_{m}$ progresses the model and returns a new model state $s'_{m}$.
 Given the underlying model dynamics are governed in a probabilistic manner, the model returns $s'_{m}$ stochastically.

### 2023-10-19 paper/fig/arh_binomial_ari.tex
update figures

\begin{tikzpicture}
    \pgfplotsset{
        scale only axis,
        % height=4cm,
        xlabel={total daily new hospitalizations},
        ylabel={total lost contacts},
        width=0.45\textwidth,
        every axis title/.append style={at={(0.45,0.0)}},
        x dir= reverse,
        y dir= reverse,
    }
    \begin{groupplot}[
        group style={
            group size=2 by 3,
            y descriptions at=edge left,
            x descriptions at=edge bottom,
            vertical sep=0pt,
            horizontal sep=0pt,
        },
        % legend entries={{$\text{PCN}_{b=2}$}, {$\text{PCN}_{b=3}$}, {$\text{PCN}_{b=4}$}, {$\text{PCN}_{b=5}$}, {$\text{PCN}$}, {Fixed}},
    ]
    % interpolated plot for ARI
    \nextgroupplot[mark size=1pt,draw opacity=1,legend entries={{$\text{PCN}_{b=2}$}, {$\text{PCN}_{b=3}$}, {$\text{PCN}_{b=4}$}, {$\text{PCN}_{b=5}$}, {$\text{PCN}$}, {Fixed}},title={$\mdprewardfn_\text{ARI}$, different budgets}]
    % mean line
    \foreach \b/\c in {2/pc1, 3/pc2, 4/pc3, 5/pc4, {None}/pc5}
    {
        \edef\temp{
            \noexpand\addplot[fill opacity=0.5,\c] table [x expr=-1*\noexpand\thisrow{mu_o_1}, y expr=-1*\noexpand\thisrow{mu_o_5}, col sep=comma] {data/cs/ari/binomial/interpolated/budget_\b.csv};
        }
        \temp
    }
    % std fillbetween, has to be after means for legend order
    \foreach \b/\c in {2/pc1, 3/pc2, 4/pc3, 5/pc4, {None}/pc5}
    {
        \edef\temp{
            \noexpand\addplot[name path=upper\b, draw=none] table [x expr=-1*\noexpand\thisrow{mu_o_1}, y expr=-1*\noexpand\thisrow{mu_o_5}+\noexpand\thisrow{std_o_5}, col sep=comma] {data/cs/ari/binomial/interpolated/budget_\b.csv};
            \noexpand\addplot[name path=lower\b, draw=none] table [x expr=-1*\noexpand\thisrow{mu_o_1}, y expr=-1*\noexpand\thisrow{mu_o_5}-\noexpand\thisrow{std_o_5}, col sep=comma] {data/cs/ari/binomial/interpolated/budget_\b.csv};
            \noexpand\addplot[\c!15] fill between[of={upper\b} and {lower\b}];
        }
        \temp
    }
    \addplot[only marks,mark=+] table [x expr=-1*\thisrow{o_0}, y expr=-1*\thisrow{o_1}, col sep=comma] {data/cs/cs_fixed.csv};
    % ARI vs ARH per budget
    \pgfplotsforeachungrouped \b in {2, 3, 4, 5, {None}}
    {
        % mean plots
        \edef\temp{
            \noexpand\nextgroupplot[draw opacity=1,legend entries={$R_{\text{ARH}}$, $R_{\text{ARI}}$},title={$\text{PCN}_{b=\b}$}]
            \noexpand\addplot[fill opacity=0.5,pc1] table [x expr=-1*\noexpand\thisrow{mu_o_1}, y expr=-1*\noexpand\thisrow{mu_o_5}, col sep=comma] {data/cs/arh/binomial/interpolated/budget_\b.csv};
            \noexpand\addplot[fill opacity=0.5,pc2] table [x expr=-1*\noexpand\thisrow{mu_o_1}, y expr=-1*\noexpand\thisrow{mu_o_5}, col sep=comma] {data/cs/ari/binomial/interpolated/budget_\b.csv};
            % std plots
            \noexpand\addplot[name path=binomialupper\b, draw=none] table [x expr=-1*\noexpand\thisrow{mu_o_1}, y expr=-1*\noexpand\thisrow{mu_o_5}+\noexpand\thisrow{std_o_5}, col sep=comma] {data/cs/arh/binomial/interpolated/budget_\b.csv};
            \noexpand\addplot[name path=binomiallower\b, draw=none] table [x expr=-1*\noexpand\thisrow{mu_o_1}, y expr=-1*\noexpand\thisrow{mu_o_5}-\noexpand\thisrow{std_o_5}, col sep=comma] {data/cs/arh/binomial/interpolated/budget_\b.csv};
            \noexpand\addplot[pc1!15] fill between[of={binomialupper\b} and {binomiallower\b}];

            \noexpand\addplot[name path=odeupper\b, draw=none] table [x expr=-1*\noexpand\thisrow{mu_o_1}, y expr=-1*\noexpand\thisrow{mu_o_5}+\noexpand\thisrow{std_o_5}, col sep=comma] {data/cs/ari/binomial/interpolated/budget_\b.csv};
            \noexpand\addplot[name path=odelower\b, draw=none] table [x expr=-1*\noexpand\thisrow{mu_o_1}, y expr=-1*\noexpand\thisrow{mu_o_5}-\noexpand\thisrow{std_o_5}, col sep=comma] {data/cs/ari/binomial/interpolated/budget_\b.csv};
            \noexpand\addplot[pc2!15] fill between[of={odeupper\b} and {odelower\b}];
        }
        \temp
    }
    \end{groupplot}
\end{tikzpicture}

### 2023-10-19 paper/fig/arh_binomial_both.tex
update figures

@@
         ylabel={total lost contacts},
-        width=0.45\textwidth
+        width=0.45\textwidth,
+        x dir= reverse,
+        y dir= reverse,
     }
@@
         \edef\temp{
-            \noexpand\addplot[fill opacity=0.5,\c] table [x=mu_o_1, y=mu_o_5, col sep=comma] {data/cs/arh/binomial/interpolated/budget_\b.csv};
+            \noexpand\addplot[fill opacity=0.5,\c] table [x expr=-1*\noexpand\thisrow{mu_o_1}, y expr=-1*\noexpand\thisrow{mu_o_5}, col sep=comma] {data/cs/arh/binomial/interpolated/budget_\b.csv};
         }
@@
         \edef\temp{
-            \noexpand\addplot[name path=upper\b, draw=none] table [x=mu_o_1, y expr=\noexpand\thisrow{mu_o_5}+\noexpand\thisrow{std_o_5}, col sep=comma] {data/cs/arh/binomial/interpolated/budget_\b.csv};
-            \noexpand\addplot[name path=lower\b, draw=none] table [x=mu_o_1, y expr=\noexpand\thisrow{mu_o_5}-\noexpand\thisrow{std_o_5}, col sep=comma] {data/cs/arh/binomial/interpolated/budget_\b.csv};
+            \noexpand\addplot[name path=upper\b, draw=none] table [x expr=-1*\noexpand\thisrow{mu_o_1}, y expr=-1*\noexpand\thisrow{mu_o_5}+\noexpand\thisrow{std_o_5}, col sep=comma] {data/cs/arh/binomial/interpolated/budget_\b.csv};
+            \noexpand\addplot[name path=lower\b, draw=none] table [x expr=-1*\noexpand\thisrow{mu_o_1}, y expr=-1*\noexpand\thisrow{mu_o_5}-\noexpand\thisrow{std_o_5}, col sep=comma] {data/cs/arh/binomial/interpolated/budget_\b.csv};
             \noexpand\addplot[\c!15] fill between[of={upper\b} and {lower\b}];
@@
     }
-    \addplot[only marks,mark=+] table [x=o_0, y=o_1, col sep=comma] {data/cs/cs_fixed.csv};
+    \addplot[only marks,mark=+] table [x expr=-1*\thisrow{o_0}, y expr=-1*\thisrow{o_1}, col sep=comma] {data/cs/cs_fixed.csv};
     % best performing run
@@
         \edef\temp{
-            \noexpand\addplot[only marks,mark=o,fill opacity=0.5,\c] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/binomial/budget_\b.csv};
+            \noexpand\addplot[only marks,mark=o,fill opacity=0.5,\c] table [x expr=-1*\noexpand\thisrow{o_1}, y expr=-1*\noexpand\thisrow{o_5}, col sep=comma] {data/cs/arh/binomial/budget_\b.csv};
         }
@@
     }
-    \addplot[only marks,mark=+] table [x=o_0, y=o_1, col sep=comma] {data/cs/cs_fixed.csv};
+    \addplot[only marks,mark=+] table [x expr=-1*\thisrow{o_0}, y expr=-1*\thisrow{o_1}, col sep=comma] {data/cs/cs_fixed.csv};
     \end{groupplot}

### 2023-10-19 paper/fig/arh_binomial_interpolated.tex
update figures

\begin{tikzpicture}
    \pgfplotsset{
        scale only axis,
        % height=6cm,
        xlabel={total number of daily new hospitalizations},
        ylabel={total lost contacts}
    }
    \begin{axis}[
    legend entries={{$\text{PCN}_{b=2}$}, {$\text{PCN}_{b=3}$}, {$\text{PCN}_{b=4}$}, {$\text{PCN}_{b=5}$}, {$\text{PCN}$}, {Fixed}},
    % width=9.5cm,
    mark size=1pt,
    ]
    \foreach \b/\c in {2/pc1, 3/pc2, 4/pc3, 5/pc4, {None}/pc5}
    {
        \edef\temp{
            \noexpand\addplot[draw opacity=1,fill opacity=0.5,\c] table [x=mu_o_1, y=mu_o_5, col sep=comma] {data/cs/arh/binomial/interpolated/budget_\b.csv};
        }
        \temp
    }
    \foreach \b/\c in {2/pc1, 3/pc2, 4/pc3, 5/pc4, {None}/pc5}
    {
        \edef\temp{
            \noexpand\addplot[name path=upper\b, draw=none] table [x=mu_o_1, y expr=\noexpand\thisrow{mu_o_5}+\noexpand\thisrow{std_o_5}, col sep=comma] {data/cs/arh/binomial/interpolated/budget_\b.csv};
            \noexpand\addplot[name path=lower\b, draw=none] table [x=mu_o_1, y expr=\noexpand\thisrow{mu_o_5}-\noexpand\thisrow{std_o_5}, col sep=comma] {data/cs/arh/binomial/interpolated/budget_\b.csv};
            \noexpand\addplot[\c!20] fill between[of={upper\b} and {lower\b}];
        }
        \temp
    }
    \addplot[only marks,mark=+] table [x=o_0, y=o_1, col sep=comma] {data/cs/cs_fixed.csv};
    \end{axis}
\end{tikzpicture}

### 2023-10-19 paper/fig/arh_inflection_point.tex
update figures

\tikzsetnextfilename{arh_inflection_point}
\begin{tikzpicture}
    \pgfplotsset{
          scale only axis,
          date coordinates in=x,
          xticklabel=\day/\month,
          height=2cm,
          every axis title/.append style={at={(0.07,0.0)}},
        }
        \begin{groupplot}[
        group style={
                % set how the plots should be organized
                group size=1 by 5,
                % only show ticklabels and axis labels on the bottom
                y descriptions at=edge left,
                x descriptions at=edge bottom,
                % set the `vertical sep' to zero
                horizontal sep=0pt,
                vertical sep=0pt,
            },
          width=9.5cm,
          axis y line*=left,
          xlabel={time},
          date ZERO=2020-05-04,
          ylabel={individuals},
          ymin=-10,
          ymax=120,
          ytick={0,25,50,75,100},
          % scaled y ticks={base 10:-3},
          % every y tick scale label/.style={
          %     at={(0,1)},yshift=-1pt,anchor=north west,inner sep=0pt
          % }
        ]
        % GROUP WITH LEFT Y-AXIS: PEOPLE (HOSP, ICU, DEATHS)
        \nextgroupplot[scaled y ticks=false, title=$\text{PCN}_{b=2}$]
        \addplot[mark=none, cyan] table [x=dates, y=i_hosp_new, col sep=comma] {data/transitions/arh/binomial/budget_2/policies-transitions/136/run_0.csv};
        \label{i_hosp_new}
        \addplot[mark=none, orange] table [x=dates, y=i_icu_new, col sep=comma] {data/transitions/arh/binomial/budget_2/policies-transitions/136/run_0.csv};
        \label{i_icu_new}
        \addplot[mark=none, red] table [x=dates, y=d_new, col sep=comma] {data/transitions/arh/binomial/budget_2/policies-transitions/136/run_0.csv};
        \label{d_new}
        \nextgroupplot[scaled y ticks=false, title=$\text{PCN}_{b=3}$]
        \addplot[mark=none, cyan] table [x=dates, y=i_hosp_new, col sep=comma] {data/transitions/arh/binomial/budget_3/policies-transitions/136/run_0.csv};
        \addplot[mark=none, orange] table [x=dates, y=i_icu_new, col sep=comma] {data/transitions/arh/binomial/budget_3/policies-transitions/136/run_0.csv};
        \addplot[mark=none, red] table [x=dates, y=d_new, col sep=comma] {data/transitions/arh/binomial/budget_3/policies-transitions/136/run_0.csv};
        \nextgroupplot[scaled y ticks=false, title=$\text{PCN}_{b=4}$]
        \addplot[mark=none, cyan] table [x=dates, y=i_hosp_new, col sep=comma] {data/transitions/arh/binomial/budget_4/policies-transitions/73/run_0.csv};
        \addplot[mark=none, orange] table [x=dates, y=i_icu_new, col sep=comma] {data/transitions/arh/binomial/budget_4/policies-transitions/73/run_0.csv};
        \addplot[mark=none, red] table [x=dates, y=d_new, col sep=comma] {data/transitions/arh/binomial/budget_4/policies-transitions/73/run_0.csv};
        \nextgroupplot[scaled y ticks=false, title=$\text{PCN}_{b=5}$]
        \addplot[mark=none, cyan] table [x=dates, y=i_hosp_new, col sep=comma] {data/transitions/arh/binomial/budget_5/policies-transitions/86/run_0.csv};
        \addplot[mark=none, orange] table [x=dates, y=i_icu_new, col sep=comma] {data/transitions/arh/binomial/budget_5/policies-transitions/86/run_0.csv};
        \addplot[mark=none, red] table [x=dates, y=d_new, col sep=comma] {data/transitions/arh/binomial/budget_5/policies-transitions/86/run_0.csv};
        % \nextgroupplot[scaled y ticks=false, title=$\text{PCN}_{b=\infty}$]
        % \addplot[mark=none, cyan] table [x=dates, y=i_hosp_new, col sep=comma] {data/transitions/arh/binomial/budget_None/policies-transitions/48/run_0.csv};
        % \addplot[mark=none, orange] table [x=dates, y=i_icu_new, col sep=comma] {data/transitions/arh/binomial/budget_None/policies-transitions/48/run_0.csv};
        % \addplot[mark=none, red] table [x=dates, y=d_new, col sep=comma] {data/transitions/arh/binomial/budget_None/policies-transitions/48/run_0.csv};
        \end{groupplot}
        % GROUP WITH RIGHT Y-AXIS: ACTIONS (P_W, P_S, P_L)
        \begin{groupplot}[
        group style={
                % set how the plots should be organized
                group size=1 by 5,
                % only show ticklabels and axis labels on the bottom
                y descriptions at=edge right,
                x descriptions at=edge bottom,
                % set the `vertical sep' to zero
                horizontal sep=0pt,
                vertical sep=0pt
            },
          width=9.5cm,
          axis y line*=right,
          axis x line=none,
          ymax=1.1,
          ylabel={proportion},
          legend style={at={(0.5,1)},anchor=south},
        %   legend pos=outer north east,
          legend columns=3, 
          legend cell align=left,
        ]
        \nextgroupplot[scaled y ticks=false]
        \addplot[mark=none, dashed, blue] table [x=dates, y=p_w, col sep=comma] {data/transitions/arh/binomial/budget_2/policies-transitions/136/run_0.csv};
        \label{p_w}
        \addplot[mark=none, dashed, magenta] table [x=dates, y=p_s, col sep=comma] {data/transitions/arh/binomial/budget_2/policies-transitions/136/run_0.csv};
        \label{p_s}
        \addplot[mark=none, dashed, darkgray] table [x=dates, y=p_l, col sep=comma] {data/transitions/arh/binomial/budget_2/policies-transitions/136/run_0.csv};
        \label{p_l}
        \addlegendimage{mark=none, cyan}\addlegendentry{$p_w$}
        \addlegendimage{mark=none, orange}\addlegendentry{$p_s$}
        \addlegendimage{mark=none, red}\addlegendentry{$p_l$}
        \addlegendimage{/pgfplots/refstyle=p_w}\addlegendentry{{daily new hosp}}
        \addlegendimage{/pgfplots/refstyle=p_s}\addlegendentry{{daily new ICU}}
        \addlegendimage{/pgfplots/refstyle=p_l}\addlegendentry{{daily deaths}}
        \nextgroupplot[scaled y ticks=false]
        \addplot[mark=none, dashed, blue] table [x=dates, y=p_w, col sep=comma] {data/transitions/arh/binomial/budget_3/policies-transitions/136/run_0.csv};
        \addplot[mark=none, dashed, magenta] table [x=dates, y=p_s, col sep=comma] {data/transitions/arh/binomial/budget_3/policies-transitions/136/run_0.csv};
        \addplot[mark=none, dashed, darkgray] table [x=dates, y=p_l, col sep=comma] {data/transitions/arh/binomial/budget_3/policies-transitions/136/run_0.csv};
        \nextgroupplot[scaled y ticks=false]
        \addplot[mark=none, dashed, blue] table [x=dates, y=p_w, col sep=comma] {data/transitions/arh/binomial/budget_4/policies-transitions/73/run_0.csv};
        \addplot[mark=none, dashed, magenta] table [x=dates, y=p_s, col sep=comma] {data/transitions/arh/binomial/budget_4/policies-transitions/73/run_0.csv};
        \addplot[mark=none, dashed, darkgray] table [x=dates, y=p_l, col sep=comma] {data/transitions/arh/binomial/budget_4/policies-transitions/73/run_0.csv};
        \nextgroupplot[scaled y ticks=false]
        \addplot[mark=none, dashed, blue] table [x=dates, y=p_w, col sep=comma] {data/transitions/arh/binomial/budget_5/policies-transitions/86/run_0.csv};
        \addplot[mark=none, dashed, magenta] table [x=dates, y=p_s, col sep=comma] {data/transitions/arh/binomial/budget_5/policies-transitions/86/run_0.csv};
        \addplot[mark=none, dashed, darkgray] table [x=dates, y=p_l, col sep=comma] {data/transitions/arh/binomial/budget_5/policies-transitions/86/run_0.csv};
        \end{groupplot}
      \end{tikzpicture}

### 2023-10-19 paper/fig/arh_ode.tex
update figures

--- 2023-01-31 paper/fig/arh.tex
@@
         % height=6cm,
-    xlabel={total daily new hospitalizations},
-    ylabel={total lost contacts},
-    width=0.45\textwidth
+        xlabel={total number of daily new hospitalizations},
+        ylabel={total lost contacts}
     }
-\begin{groupplot}[
-    group style={
-        group size=2 by 1,
-        y descriptions at=edge left,
-        vertical sep=0pt,
-        horizontal sep=0pt,
-    },
+    \begin{axis}[
     legend entries={{$\text{PCN}_{b=2}$}, {$\text{PCN}_{b=3}$}, {$\text{PCN}_{b=4}$}, {$\text{PCN}_{b=5}$}, {$\text{PCN}$}, {Fixed}},
+    % width=9.5cm,
+    mark size=1pt,
+    draw opacity=1
     ]
-\nextgroupplot[mark size=1pt,draw opacity=1]
     \addplot[only marks,mark=o,pc1] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/ode/budget_2.csv};
@@
     \addplot[only marks,mark=+] table [x=o_0, y=o_1, col sep=comma] {data/cs/cs_fixed.csv};
-\nextgroupplot[mark size=1pt]
-        \addplot[only marks,draw opacity=1,fill opacity=0.5,mark=o,pc1] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/binomial/budget_2.csv};
-        \addplot[only marks,draw opacity=1,fill opacity=0.5,mark=o,pc2] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/binomial/budget_3.csv};
-        \addplot[only marks,draw opacity=1,fill opacity=0.5,mark=o,pc3] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/binomial/budget_4.csv};
-        \addplot[only marks,draw opacity=1,fill opacity=0.5,mark=o,pc4] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/binomial/budget_5.csv};
-        \addplot[only marks,draw opacity=1,fill opacity=0.5,mark=o,pc5] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/binomial/budget_None.csv};
-        \addplot[only marks,mark=+] table [x=o_0, y=o_1, col sep=comma] {data/cs/cs_fixed.csv};
-
-\end{groupplot}
+    \end{axis}
 \end{tikzpicture}

### 2023-10-19 paper/main.tex
update figures

@@
 Infectious disease outbreaks can have a disruptive impact on public health and societal processes.
-As decision making in the context of epidemic mitigation is hard, reinforcement learning provides a methodology to automatically learn prevention strategies in combination with complex epidemic models.
+As decision-making in the context of epidemic mitigation is hard, reinforcement learning provides a methodology to automatically learn prevention strategies in combination with complex epidemic models.
 Current research focuses on optimizing policies with respect to a single objective, such as the pathogen's attack rate.
@@
 We consider the first wave of the Belgian COVID-19 epidemic, which was mitigated by a lockdown, and study different deconfinement strategies, aiming to minimize both COVID-19 cases (i.e., infections and hospitalizations) and the societal burden that is induced by the applied mitigation measures.
-We contribute a multi-objective Markov decision process that encapsulates the stochastic compartment model that was used to inform policy makers during the COVID-19 epidemic.
+We contribute a multi-objective Markov decision process that encapsulates the stochastic compartment model that was used to inform policymakers during the COVID-19 epidemic.
 As these social mitigation measures are implemented in a continuous action space that modulates the contact matrix of the age-structured epidemic model, we extend PCN to this setting.
@@
 To address these challenges, reinforcement learning provides a methodology to automatically learn mitigation strategies in combination with complex epidemic models~\cite{libin2020}.
-Previous research focused on optimising policies with respect to a single objective, such as the pathogen's attack rate, while the mitigation of epidemics is a problem that inherently covers distinct  and possibly conflicting criteria (i.a., prevalence, mortality, morbidity, cost).
+Previous research focused on optimizing policies with respect to a single objective, such as the pathogen's attack rate, while the mitigation of epidemics is a problem that inherently covers distinct  and possibly conflicting criteria (i.a., prevalence, mortality, morbidity, cost).
 Therefore, optimizing on a single objective requires that  these distinct criteria are somehow aggregated into a single metric.
@@
 However, this is a challenging process, as the decision maker is generally not aware of all possible trade-offs, and as such this may result in a metric that exhibits suboptimal performance.
-Furthermore, manually designing such metrics is time-consuming, costly and error-prone, as this non-intuitive process requires repetitive and tedious tuning to achieve the desired behaviour~\cite{roijers2013survey}.
+Furthermore, manually designing such metrics is time-consuming, costly and error-prone, as this non-intuitive process requires repetitive and tedious tuning to achieve the desired behavior~\cite{roijers2013survey}.
 Moreover, taking a single objective approach has several other limiting factors.
@@
 When the incidence of confirmed cases was steadily decreasing, epidemiological experts were tasked to investigate deconfinement strategies, to reduce the severe social contact and mobility restrictions.
-Here, we consider an epidemiological model that was constructed to describe the Belgian COVID-19 epidemic and was fitted to hospitalisation incidence data and serial sero-prevalence data~\cite{abrams2021modelling}.
+Here, we consider an epidemiological model that was constructed to describe the Belgian COVID-19 epidemic and was fitted to hospitalization incidence data and serial sero-prevalence data~\cite{abrams2021modelling}.
 This model constitutes a stochastic discrete-time age-structured compartmental model that simulates mitigation strategies by varying social distancing parameters concerning school, work and leisure contacts.
-Based on this model, we contribute a novel mutli-objective epidemiological reinforcement learning environment~(\momdpname), in the form of a multi-objective Markov decision process (MOMDP) \cite{roijers2013survey}.
+Based on this model, we contribute a novel multi-objective epidemiological reinforcement learning environment~(\momdpname), in the form of a multi-objective Markov decision process (MOMDP) \cite{roijers2013survey}.
 \momdpname\ encapsulates the epidemiological model developed by \citet{abrams2021modelling} to implement state transitions, with an action space that combines a proportional reduction of school, work and leisure contacts at each time step, Furthermore, it defines a reward function based on two objectives: the attack rate (i.e., proportion of the population affected by the pathogen) and the social burden that is induced by the mitigation measures.
@@
 
-By evaluating the solution set of mitigation policies learned by PCN, we observe that PCN minimises the social burden in scenarios where hospitalization rates are sufficiently low.
+By evaluating the solution set of mitigation policies learned by PCN, we observe that PCN minimizes the social burden in scenarios where hospitalization rates are sufficiently low.
 Therefore, in this work we illustrate that multi-objective reinforcement learning can provide important insights concerning the trade-offs between complex mitigation polices in real-world epidemiological models.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Background}
+\section{COVID-19 model and \momdpname\  MOMDP}
+\label{sec:sars-cov2-momdp}
+
+\emph{Reinforcement learning} (RL) is a set of machine learning techniques to learn the optimal behavior for sequential decision problems.
+Through trial and error, the RL agent interacts with its environment and observes the consequences of its decisions on the environment.
+Decisions that led to desirable consequences are reinforced, such that this behavior can be repeated.
+Otherwise, the agent explores the environment by trying different decisions.
+Developing mitigation strategies for epidemic control is a sequential decision problem on which we can apply RL, which will learn a desirable strategy.
+
+In this Section, we first introduce the compartment model that we use to simulate the epidemic (Sec.~\ref{sec:sars-cov2-momdp_model}).
+Then, we explain the intervention strategies that can be made to control the evolution of the epidemic (Sec.~\ref{sec:interventions}).
+Afterwards, we introduce the multi-objective reinforcement learning (MORL) framework (Sec.~\ref{sec:reinforcement-learning}).
+Finally, we propose \momdpname, a MORL environment that encompasses the epidemiological model and intervention strategies, so we can learn optimal mitigation strategies (Sec.~\ref{sec:covid-momdp}).
+
 In the following, we use bold notation to denote variables with a vectorial value, while non-bold notation for its scalar counterpart.
+\subsection{Stochastic compartment model for SARS-CoV-2}
+\label{sec:sars-cov2-momdp_model}
+To evaluate non-pharmaceutical interventions, we consider the compartmental model presented by Abrams et al., that was used to investigate exit strategies in Belgium after the first epidemic wave of SARS-CoV-2~\cite{abrams2021modelling}.
+This model concerns a discrete-time stochastic model, that considers an age-structured population.
+The model generalizes a standard SEIR model\footnote{A standard SEIR model divides the population into four different states, i.e., susceptible, exposed, infectious and recovered individuals.}, extended to capture the different stages of disease spread and history that are associated with SARS-CoV-2 (i.e., pre-symptomatic, asymptomatic, symptomatic with mild symptoms and symptomatic with severe symptoms) and to represent the stages associated with severe disease, i.e., hospitalization, admission to the intensive care unit (ICU) and death.
+
+A visual representation of the model is depicted in Fig.~\ref{fig:compartment_model}. It shows the different compartments, as well as the flow rates at which individuals move between compartments.
+
+\begin{figure}[t!]
+    \centering
+    \begin{tikzpicture}
+     \draw (0.0pt, 0.0pt)node[fill=green!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](0){$\textbf{S}$};
+     \draw (60.0pt, 0.0pt)node[fill=orange!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](1){$\textbf{E}$};
+     \draw (120.0pt, 0.0pt)node[fill=red!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](2){$\textbf{I}^{presym}$};
+     \draw (180.0pt, 10.0pt)node[fill=red!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](3){$\textbf{I}^{asym}$};
+     \draw (170.0pt, -30.0pt)node[fill=red!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](4){$\textbf{I}^{mild}$};
+     \draw (170.0pt, -70.0pt)node[fill=red!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](5){$\textbf{I}^{sev}$};
+     \draw (230.0pt, -50.0pt)node[fill=blue!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](6){$\textbf{I}^{hosp}$};
+      \draw (75.0pt, -70.0pt)node[fill=gray!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](10){$\textbf{H}^{new}$};
+     \draw (230.0pt, -90.0pt)node[fill=blue!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](7){$\textbf{I}^{icu}$};
+     \draw (300.0pt, -50.0pt)node[fill=orange!50, thick, minimum height=0.6cm,minimum width=0.6cm, draw](8){$\textbf{R}$};
+     \draw (300.0pt, -90.0pt)node[fill=purple!50, thick, minimum height=0.6cm,minimum width=0.6cm, draw](9){$\textbf{D}$};
+
+     %\path[->] (6) edge[dotted, bend right = 50]
+     \path[->] (0) edge[] node[above]{$\boldsymbol{\lambda}(t)$} (1);
+     \path[->] (1) edge[] node[above]{$\gamma$} (2);
+     \path[->] (2) edge[] node[below]{$\textbf{p}\theta$} (3);
+     \path[->] (2) edge[] node[near end, below left]{$(1-\textbf{p})\theta$} (4);
+     \path[->] (4) edge[] node[left]{$\boldsymbol{\psi}$} (5);
+     \path[->] (5) edge[] node[near end, below left]{$(1-\phi_{1})\boldsymbol{\omega}$} (7);
+     \path[->] (5) edge[] node[above]{$\phi_{1}\boldsymbol{\omega}$} (6);
+     \path[->] (7) edge[] node[near end, below right]{$\boldsymbol{\delta}_{4}$} (8);
+     \path[->] (7) edge[] node[above]{$\boldsymbol{\tau}_{2}$} (9);
+     \path[->] (6) edge[] node[above]{$\boldsymbol{\delta}_{3}$} (8);
+     \path[->] (6) edge[] node[near start, below left]{$\boldsymbol{\tau}_{1}$} (9);
+     \path[->] (3) edge[bend left = 30] node[above right]{$\delta_{1}$} (8);
+     \path[->] (4) edge[bend left = 30] node[above left]{$\boldsymbol{\delta}_{2}$} (8);
+      \path[->] (5) edge[] node[below]{$\textbf{I}^{sev} - \textbf{H}^{new}$} (10);
+    \end{tikzpicture}
+
+    \caption{Schematic diagram of the compartmental model for SARS-CoV-2 presented by \citet{abrams2021modelling} which is used to derive the MOMDP.}
+    \label{fig:compartment_model}
+\end{figure}
+
+These flow rates are defined by a set of ordinary differential equations, which are outlined as follows:
+
+\begin{align*}
+\frac{d\textbf{S}(t)}{dt} & = -\lambda(t)(\textbf{S})(t), \\
+%
+\frac{d\textbf{E}(t)}{dt} & = \lambda(t)(\textbf{S})(t) - \gamma \textbf{E}(t), \\
+%
+\frac{d\textbf{I}^{presym}(t)}{dt} & = \gamma \textbf{E}(t) - \theta \textbf{I}^{presym}(t), \\
+%
+\frac{d\textbf{I}^{asym}(t)}{dt} & = \theta p \textbf{I}^{presym}(t) - \delta_{1} \textbf{I}^{asym}(t), \\
+%
+\frac{d\textbf{I}^{mild}(t)}{dt} & = \theta (1 - p) \textbf{I}^{presym}(t) - \{ \psi + \omega_{2} \} \textbf{I}^{mild}(t), \\
+%
+\frac{d\textbf{I}^{sev}(t)}{dt} & = \psi \textbf{I}^{mild}(t) - \omega \textbf{I}^{sev}(t), \\
+%
+\frac{d\textbf{I}^{hosp}(t)}{dt} & = \phi_{1} \omega \textbf{I}^{sev}(t) - \{ \delta_{3} + \tau_{1} \} \textbf{I}^{hosp}(t), \\
+%
+\frac{d\textbf{I}^{icu}(t)}{dt} & = (1 - \phi_1) \omega \textbf{I}^{sev}(t) - \{ \delta_{4} + \tau_{2} \} \textbf{I}^{icu}(t), \\
+%
+\frac{d\textbf{D}(t)}{dt} & = \tau_{1} \textbf{I}^{hosp}(t) - \tau_{2} \textbf{I}^{icu}(t), \\
+%
+\frac{d\textbf{R}(t)}{dt} & = \delta_{1} \textbf{I}^{asym}(t) + \delta_{2} \textbf{I}^{mild}(t) + \delta_{3} \textbf{I}^{hosp}(t) + \delta_{4} \textbf{I}^{icu}(t)
+\end{align*}
+%
+
+In this set of ordinary differential equations, each state variable represents a vector over all age groups for a particular compartment at time $t$.
+For example, $\textbf{S} = (S_{1}(t), S_{2}(t), ..., S_{\agegroup}(t))^{T}$ is the vector representing the susceptible members of the population of each age group $\agegroup$ at time $t$.
+Infection dynamics are governed by an age-specific force of infection $\foi$:
+
+\begin{equation}
+    \foi(\agegroup,t) = \sum_{\agegroup'=1}^{\agegroups}\beta(\agegroup,\agegroup')I_{\agegroup'}(t),
+\end{equation}
+%
+where $\agegroups$ is the total number of age groups, and $\beta(\agegroup,\agegroup')$ is the time-invariant transmission rate that encodes the average per capita rate at which an infectious individual in age group $\agegroup$ makes an effective contact with a susceptible individual in age group $\agegroup'$, per unit of time.
+
+As we consider an age-structured population, we consider this extended SEIR structure for $\agegroups = 10$ age groups,  i.e., $[0-10), [10-20), [20-30), [30-40), [40-50), [50-60), [60-70), [70-80), [80-90), [90,100+)$.
+Contacts of the different age-groups, which impact the propagation rate of the epidemic, are modeled using social contact matrices.
+We define a social contact matrix for 6 different social environments: $\chome, \cwork, \ctransport, \cschool, \cleisure, \cother$, for the home, work, transport, school, leisure, other environments respectively.
+The social contact matrix across all social environments is defined as:
+
+\begin{equation}
+    \ctotal = \chome + \cwork + \ctransport + \cschool + \cleisure + \cother
+\end{equation}
+
+Under the social contact hypothesis \cite{wallinga2006using}, we have that:
+%
+\begin{equation}
+\beta(\agegroup,\agegroup') = \proportionalityfactor \cdot \ctotal(\agegroup,\agegroup'),
+\end{equation}
+%
+where $\proportionalityfactor$ is a proportionality factor.
+
+Following \citet{abrams2021modelling}, we rely on distinct social contact matrices for symptomatic and asymptomatic individuals, respectively $\ctotal_{s}$ and $\ctotal_{a}$.
+Therefore, we define the transmission rates for both symptomatic and asymptomatic individuals as follows:
+
+\begin{equation}
+\boldsymbol{\beta}_{s}(\agegroup, \agegroup) = \proportionalityfactor_{s} \cdot C_{s}(\agegroup, \agegroup'),    
+\end{equation}
+and
+\begin{equation}
+\boldsymbol{\beta}_{a}(\agegroup, \agegroup') = \proportionalityfactor_{a} \cdot C_{a}(\agegroup, \agegroup').  
+\end{equation}
+
+The age-dependent force of infection can be defined as follows:
+
+\begin{equation}
+    \boldsymbol{\lambda}(t) = \boldsymbol{\beta}_{a} \times \{ \textbf{I}^{presym}(t) + \textbf{I}^{asym}(t) \} +
+    \boldsymbol{\beta}_{s} \times \{ \textbf{I}^{mild}(t) + \textbf{I}^{sev}(t) \},
+\end{equation}
+%
+where $\boldsymbol{\lambda}(t) = (\lambda(1, t), \lambda(2, t), ..., \lambda(\agegroups, t))$.
+For all further information about the different compartments and parameters we refer the reader to the work of \citet{abrams2021modelling}.
+
+\todo{epi friends, please check} Time- and age-dependent behavioral changes introduce substantial uncertainty regarding the outcome of an outbreak.
+To evaluate intervention strategies that modulate such behavioral changes, the use of stochastic epidemiological models is warranted \cite{abrams2021modelling}.
+Moreover, the effect of stochasticity is most pronounced when studying the initial growth of an epidemic \cite{britton2009epidemic}, be it during its emergence, or when implementing deconfinement strategies.
+
+By formulating the set of differential equations defined above, as a chain-binomial, we can obtain stochastic trajectories from this model \cite{bailey1975mathematical}.
+A chain-binomial model assumes a stochastic model where infected individuals are generated by some underlying probability distribution.
+For this stochastic model we consider a time interval $(t, t +h]$, where $h$ is defined as the length between two consecutive time points.
+In this work we set $h = \frac{1}{240}$.
+
+\subsection{Interventions strategies}
+\label{sec:interventions}
+To model different types of interventions, we follow the contact reduction scheme presented by \citet{abrams2021modelling}.
+Firstly, to consider distinct exit scenarios, we modulate the social contact matrices to reflect a contact reduction in a particular age group.
+
+We consider a contact reduction function that imposes a proportional reduction of work (including transport) $p_w$, school $p_s$ and leisure $p_l$ contacts, which is implemented as a linear combination of social contact matrices:
+\begin{equation}
+\label{eq:scm}
+    \hat{C}(p_w, p_s, p_l) = \chome + p_w (\cwork + \ctransport) + p_s \cschool + p_l (\cleisure + \cother)
+\end{equation}
+
+We denote $\hat{C}_t$ the social contact matrix at timestep $t$, resulting from the reduction function $\hat{C}$.
+Secondly, we assume that compliance to the interventions is gradual and model this using a logistic compliance function (see details in Sec.~\ref{sec:compliance-function} of the Appendix).
 
 \subsection{Multi-Objective Reinforcement Learning}
+\label{sec:reinforcement-learning}
 Real-world decisions problems typically consider multiple and possibly conflicting objectives.
@@
 
-\subsection{Multi-Objective Metrics}
+\subsubsection{Multi-Objective Metrics}
 Comparing the learned coverage sets produced by different algorithms is a non-trivial task, as one algorithm's output might dominate the other in some region of the objective-space, but be dominated in another.
@@
     \begin{tikzpicture}
-      \begin{axis}[legend pos=south west, xmin=-1, xmax=4.5, ymin=-0.5, ymax=4.5,ytick={0,1,2,3,4,5},scale only axis,width=0.4\textwidth]
+      \begin{axis}[
+        legend pos=south west,
+        xmin=-1,
+        xmax=4.5,
+        ymin=-0.5,
+        ymax=4.5,
+        ytick={0,1,2,3,4,5},
+        scale only axis,
+        width=0.4\textwidth,
+        xlabel={objective $1$},
+        ylabel={objective $2$}]
       \addplot[only marks,mark=x,mark size=3] coordinates {(-0.5,0)};
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{COVID-19 model and \momdpname\  MOMDP}
-\label{sec:sars-cov2-momdp}
-\subsection{Stochastic compartment model for SARS-CoV-2}
-\label{sec:sars-cov2-momdp_model}
-To evaluate non-pharmaceutical interventions, we consider the compartmental model presented by Abrams et al., that was used to investigate exit strategies in Belgium after the first epidemic wave of SARS-CoV-2~\cite{abrams2021modelling}.
-This model concerns a discrete-time stochastic model, that considers an age-structured population.
-The model generalises a standard SEIR model\footnote{A standard SEIR model divides the population into four different states, i.e., susceptible, exposed, infectious and recovered individuals.}, extended to capture the different stages of disease spread and history that are associated with SARS-CoV-2 (i.e., pre-symptomatic, asymptomatic, symptomatic with mild symptoms and symptomatic with severe symptoms) and to represent the stages associated with severe disease, i.e., hospitalisation, admission to the intensive care unit (ICU) and death.
-
-A visual representation of the model is depicted in Fig.~\ref{fig:compartment_model}. It shows the different compartments, as well as the flow rates at which individuals move between compartments.
-
-These flow rates are defined by a set of ordinary differential equations, which are outlined as follows:
-
-\begin{align*}
-\frac{d\textbf{S}(t)}{dt} & = -\lambda(t)(\textbf{S})(t), \\
-%
-\frac{d\textbf{E}(t)}{dt} & = \lambda(t)(\textbf{S})(t) - \gamma \textbf{E}(t), \\
-%
-\frac{d\textbf{I}^{presym}(t)}{dt} & = \gamma \textbf{E}(t) - \theta \textbf{I}^{presym}(t), \\
-%
-\frac{d\textbf{I}^{asym}(t)}{dt} & = \theta p \textbf{I}^{presym}(t) - \delta_{1} \textbf{I}^{asym}(t), \\
-%
-\frac{d\textbf{I}^{mild}(t)}{dt} & = \theta (1 - p) \textbf{I}^{presym}(t) - \{ \psi + \omega_{2} \} \textbf{I}^{mild}(t), \\
-%
-\frac{d\textbf{I}^{sev}(t)}{dt} & = \psi \textbf{I}^{mild}(t) - \omega \textbf{I}^{sev}(t), \\
-%
-\frac{d\textbf{I}^{hosp}(t)}{dt} & = \phi_{1} \omega \textbf{I}^{sev}(t) - \{ \delta_{3} + \tau_{1} \} \textbf{I}^{hosp}(t), \\
-%
-\frac{d\textbf{I}^{icu}(t)}{dt} & = (1 - \phi_1) \omega \textbf{I}^{sev}(t) - \{ \delta_{4} + \tau_{2} \} \textbf{I}^{icu}(t), \\
-%
-\frac{d\textbf{D}(t)}{dt} & = \tau_{1} \textbf{I}^{hosp}(t) - \tau_{2} \textbf{I}^{icu}(t), \\
-%
-\frac{d\textbf{R}(t)}{dt} & = \delta_{1} \textbf{I}^{asym}(t) + \delta_{2} \textbf{I}^{mild}(t) + \delta_{3} \textbf{I}^{hosp}(t) + \delta_{4} \textbf{I}^{icu}(t)
-\end{align*}
-%
-
-In this set of ordinary differential equations, each state variable represents a vector over all age groups for a particular compartment at time $t$.
-For example, $\textbf{S} = (S_{1}(t), S_{2}(t), ..., S_{\agegroup}(t))^{T}$ is the vector representing the susceptible members of the population of each age group $\agegroup$ at time $t$.
-Infection dynamics are governed by an age-specific force of infection $\foi$:
-
-\begin{equation}
-    \foi(\agegroup,t) = \sum_{\agegroup'=1}^{\agegroups}\beta(\agegroup,\agegroup')I_{\agegroup'}(t),
-\end{equation}
-%
-where $\agegroups$ is the total number of age groups, and $\beta(\agegroup,\agegroup')$ is the time-invariant transmission rate that encodes the average per capita rate at which an infectious individual in age group $\agegroup$ makes an effective contact with a susceptible individual in age group $\agegroup'$, per unit of time.
-
-As we consider an age-structured population, we consider this extended SEIR structure for $\agegroups = 10$ age groups,  i.e., $[0-10), [10-20), [20-30), [30-40), [40-50), [50-60), [60-70), [70-80), [80-90), [90,100+)$.
-Contacts of the different age-groups, which impact the propagation rate of the epidemic, are modeled using social contact matrices.
-We define a social contact matrix for 6 different social environments: $\chome, \cwork, \ctransport, \cschool, \cleisure, \cother$, for the home, work, transport, school, leisure, other environments respectively.
-The social contact matrix across all social environments is defined as:
-
-\begin{equation}
-    \ctotal = \chome + \cwork + \ctransport + \cschool + \cleisure + \cother
-\end{equation}
-
-Under the social contact hypothesis \cite{wallinga2006using}, we have that:
-%
-\begin{equation}
-\beta(\agegroup,\agegroup') = \proportionalityfactor \cdot \ctotal(\agegroup,\agegroup'),
-\end{equation}
-%
-where $\proportionalityfactor$ is a proportionality factor.
-
-Following \citet{abrams2021modelling}, we rely on distinct social contact matrices for symptomatic and asymptomatic individuals, respectively $\ctotal_{s}$ and $\ctotal_{a}$.
-Therefore, we define the transmission rates for both symptomatic and asymptomatic individuals as follows:
-
-\begin{equation}
-\boldsymbol{\beta}_{s}(\agegroup, \agegroup) = \proportionalityfactor_{s} \cdot C_{s}(\agegroup, \agegroup'),    
-\end{equation}
-and
-\begin{equation}
-\boldsymbol{\beta}_{a}(\agegroup, \agegroup') = \proportionalityfactor_{a} \cdot C_{a}(\agegroup, \agegroup').  
-\end{equation}
-
-The age-dependent force of infection can be defined as follows:
-
-\begin{equation}
-    \boldsymbol{\lambda}(t) = \boldsymbol{\beta}_{a} \times \{ \textbf{I}^{presym}(t) + \textbf{I}^{asym}(t) \} +
-    \boldsymbol{\beta}_{s} \times \{ \textbf{I}^{mild}(t) + \textbf{I}^{sev}(t) \},
-\end{equation}
-%
-where $\boldsymbol{\lambda}(t) = (\lambda(1, t), \lambda(2, t), ..., \lambda(\agegroups, t))$.
-For all further information about the different compartments and parameters we refer the reader to the work of \citet{abrams2021modelling}.
-
-\todo{epi friends, please check} Time- and age-dependent behavioral changes introduce substantial uncertainty regarding the outcome of an outbreak.
-To evaluate intervention strategies that modulate such behavioral changes, the use of stochastic epidemiological models is warranted \cite{abrams2021modelling}.
-Moreover, the effect of stochasticity is most pronounced when studying the initial growth of an epidemic \cite{britton2009epidemic}, be it during its emergence, or when implementing deconfinement strategies.
-
-By formulating the set of differential equations defined above, as a chain-binomial, we can obtain stochastic trajectories from this model \cite{bailey1975mathematical}.
-A chain-binomial model assumes a stochastic model where infected individuals are generated by some underlying probability distribution.
-For this stochastic model we consider a time interval $(t, t +h]$, where $h$ is defined as the length between two consecutive time points.
-In this work we set $h = \frac{1}{240}$.
-
-\subsection{Interventions strategies}
-\label{sec:interventions}
-To model different types of interventions, we follow the contact reduction scheme presented by \citet{abrams2021modelling}.
-Firstly, to consider distinct exit scenarios, we modulate the social contact matrices to reflect a contact reduction in a particular age group.
-
-We consider a contact reduction function that imposes a proportional reduction of work (including transport) $p_w$, school $p_s$ and leisure $p_l$ contacts, which is implemented as a linear combination of social contact matrices:
-\begin{equation}
-\label{eq:scm}
-    \hat{C}(p_w, p_s, p_l) = \chome + p_w (\cwork + \ctransport) + p_s \cschool + p_l (\cleisure + \cother)
-\end{equation}
-
-We denote $\hat{C}_t$ the social contact matrix at timestep $t$, resulting from the reduction function $\hat{C}$.
-Secondly, we assume that compliance to the interventions is gradual and model this using a logistic compliance function (see details in Sec.~\ref{sec:compliance-function} of the Appendix).
-
 \subsection{The \momdpname\ Environment}
-\begin{figure}[t!]
-    \centering
-    \begin{tikzpicture}
-     \draw (0.0pt, 0.0pt)node[fill=green!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](0){$\textbf{S}$};
-     \draw (60.0pt, 0.0pt)node[fill=orange!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](1){$\textbf{E}$};
-     \draw (120.0pt, 0.0pt)node[fill=red!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](2){$\textbf{I}^{presym}$};
-     \draw (180.0pt, 10.0pt)node[fill=red!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](3){$\textbf{I}^{asym}$};
-     \draw (170.0pt, -30.0pt)node[fill=red!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](4){$\textbf{I}^{mild}$};
-     \draw (170.0pt, -70.0pt)node[fill=red!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](5){$\textbf{I}^{sev}$};
-     \draw (230.0pt, -50.0pt)node[fill=blue!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](6){$\textbf{I}^{hosp}$};
-      \draw (75.0pt, -70.0pt)node[fill=gray!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](10){$\textbf{H}^{new}$};
-     \draw (230.0pt, -90.0pt)node[fill=blue!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](7){$\textbf{I}^{icu}$};
-     \draw (300.0pt, -50.0pt)node[fill=orange!50, thick, minimum height=0.6cm,minimum width=0.6cm, draw](8){$\textbf{R}$};
-     \draw (300.0pt, -90.0pt)node[fill=purple!50, thick, minimum height=0.6cm,minimum width=0.6cm, draw](9){$\textbf{D}$};
-
-     %\path[->] (6) edge[dotted, bend right = 50]
-     \path[->] (0) edge[] node[above]{$\boldsymbol{\lambda}(t)$} (1);
-     \path[->] (1) edge[] node[above]{$\gamma$} (2);
-     \path[->] (2) edge[] node[below]{$\textbf{p}\theta$} (3);
-     \path[->] (2) edge[] node[near end, below left]{$(1-\textbf{p})\theta$} (4);
-     \path[->] (4) edge[] node[left]{$\boldsymbol{\psi}$} (5);
-     \path[->] (5) edge[] node[near end, below left]{$(1-\phi_{1})\boldsymbol{\omega}$} (7);
-     \path[->] (5) edge[] node[above]{$\phi_{1}\boldsymbol{\omega}$} (6);
-     \path[->] (7) edge[] node[near end, below right]{$\boldsymbol{\delta}_{4}$} (8);
-     \path[->] (7) edge[] node[above]{$\boldsymbol{\tau}_{2}$} (9);
-     \path[->] (6) edge[] node[above]{$\boldsymbol{\delta}_{3}$} (8);
-     \path[->] (6) edge[] node[near start, below left]{$\boldsymbol{\tau}_{1}$} (9);
-     \path[->] (3) edge[bend left = 30] node[above right]{$\delta_{1}$} (8);
-     \path[->] (4) edge[bend left = 30] node[above left]{$\boldsymbol{\delta}_{2}$} (8);
-      \path[->] (5) edge[] node[below]{$\textbf{I}^{sev} - \textbf{H}^{new}$} (10);
-    \end{tikzpicture}
-
-    \caption{Schematic diagram of the compartmental model for SARS-CoV-2 presented by \citet{abrams2021modelling} which is used to derive the MOMDP.}
-    \label{fig:compartment_model}
-\end{figure}
+\label{sec:covid-momdp}
 In order to apply multi-objective reinforcement learning, we construct the \momdpname\ MOMDP based on the epidemiological model introduced in Sec.~\ref{sec:sars-cov2-momdp_model} and graphically depicted in Fig.~\ref{fig:compartment_model}.
@@
 
-\paragraph{Transition function:} The model defined by \citet{abrams2021modelling} utilises a model transition probability $M$ (see Sec.~\ref{sec:binomial-model} of the Appendix for details on $M$), that progresses the epidemiological model in one timestep based on the currently installed social contact matrix $\hat{C}(p_{w}, p_{s}, p_{l})$.
+\paragraph{Transition function:} The model defined by \citet{abrams2021modelling} utilizes a model transition probability $M$ (see Sec.~\ref{sec:binomial-model} of the Appendix for details on $M$), that progresses the epidemiological model in one timestep based on the currently installed social contact matrix $\hat{C}(p_{w}, p_{s}, p_{l})$.
 We use this function as the transition function in \momdpname.
@@
 for each age group $k \in \{1, \ldots, \agegroups\}$, where $S$ encodes the members of the population who are susceptible to infection and $E$ encodes the members of the population who have been exposed to COVID-19.
-Moreover, $I^{presym}$, $I^{asym}$, $I^{mild}$, $I^{hosp}$, $I^{icu}$ denote the members of the population infected with COVID-19 and are, respectively, presymptomatic, asymptomatic, mildly symptomatic, hospitalised, or in the ICU.
-Finally, $H^{new}_k$ represents the number of newly hospitalised individuals in age group $k$.
+Moreover, $I^{presym}$, $I^{asym}$, $I^{mild}$, $I^{hosp}$, $I^{icu}$ denote the members of the population infected with COVID-19 and are, respectively, pre-symptomatic, asymptomatic, mildly symptomatic, hospitalized, or in the ICU.
+Finally, $H^{new}_k$ represents the number of newly hospitalized individuals in age group $k$.
 
@@
 
-\paragraph{Reward function:} We define a vectorial reward function which considers multiple objectives: attack rate (i.e., infections and hospitalisations) and the social burden imposed by the interventions on the population.
+\paragraph{Reward function:} We define a vectorial reward function which considers multiple objectives: attack rate (i.e., infections and hospitalizations) and the social burden imposed by the interventions on the population.
 
@@
 
-The reward function to reduce the attack rate in terms of hospitalisations is defined as the negative number of new hospitalizations:
+The reward function to reduce the attack rate in terms of hospitalizations is defined as the negative number of new hospitalizations:
 \begin{equation}
@@
 Transitions $x= \langle \mdpstate_t, h_t, \momdpreturn_t \rangle, y=\action_t$ are sampled from the dataset, and the ground-truth output $y$ is compared with the predicted output $\hat{y}=\pi(\mdpstate_t, h_t, \momdpreturn_t)$.
-The predictor (i.e., the policy) is then updated using the cross-entropy loss function \cite{shore1980axiomatic}:
+The predictor (i.e., the policy) is then updated using the cross-entropy loss function~\cite{shore1980axiomatic}:
 \begin{equation}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Analysing COVID-19 deconfinement policies}
+\section{Results}
 \label{sec:experiments}
@@
 This allows us to obtain a strong baseline for potential exit strategies over the objective space.
-We note that while such fixed policies are a feasible approach, they do not scale well in terms of action and objective spaces and they will not be able to provide an adaptive restriction level, which is what we aim to provide using PCN.
+We note that while such fixed policies are a feasible approach, they do not scale well in terms of action and objective spaces and they will not be able to provide an adaptive restriction level, which is our aim using PCN.
 
 All experiments are averaged over 10 runs.
-The hyper-parameters and the neural network architecture can be found in Sec.s~\ref{sec:pcn-hyperparameters} and Sec.~\ref{sec:pcn-architecture} of the Appendix, respectively.
+The hyper-parameters and the neural network architecture can be found in Sec.~\ref{sec:pcn-hyperparameters} and Sec.~\ref{sec:pcn-architecture} of the Appendix, respectively.
 
@@
     \input{fig/arh_binomial_both}
+    \caption{The Pareto front of policies discovered by PCN using \momdpname, showing the different compromises between the number of hospitalizations and the number of lost contacts.
+On the right, we show, for each budget setting (colored, subscript indicates budget) the coverage set learned by the best performing run.
+On the left, we show an interpolated average of the coverage sets learned by the different runs, with the shaded regions corresponding to the standard deviation.
+For comparison, the basline is displayed on both plots (in black).
+As the budget increases, so does the size of the coverage set learnt by PCN.
+Changes are most noticeable in the less restrictive trade-offs in terms of social burden.}
+    \label{fig:pf-binomial-arh-budgets}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \input{fig/arh_inflection_point}
     \caption{The Pareto front of policies discovered by PCN using \momdpname, showing the different compromises between the number of hospitalizations and the number of lost contacts.
@@
 Still, the coverage set of when trained on $\mathbf{R}_{\text{ARI}}$ is systematically dominated by the one when trained on $\mathbf{R}_{\text{ARH}}$.
-\todo{Please check: is this actually true?} While hospitalisations and infections are highly correlated, they differ in terms of age-groups.
+\todo{Please check: is the following actually true?} While hospitalisations and infections are highly correlated, they differ in terms of age-groups.
 Older age-groups are more susceptible to be hospitalised after being infected, but they do not form the majority of the population.
@@
 \end{figure}
-
-\subsection{Policy executions}
-
-Depending on the budget, PCN learns a coverage set containing more than 150 different policies.
-To gain a better insight about their behavior, and how they differ from each other, we plot executions of each learned policy in Fig.s~\ref{fig:policy-executions-b2-0}-\ref{fig:policy-executions-b2-140}.
-The plots are displayed from the least restrictive policy in terms of social burden to the most restrictive one.
-Since the \momdpname\ model is stochastic, we show 10 executions of the same policy, on each plot.
-
-\foreach \i/\j in {0/19, 20/39, 40/59, 60/79, 80/99, 100/119, 120/139, 140/153}{
-    \begin{figure}
-        \centering
-        \plotPolicies{\i}{\j}{2}
-        \caption{Execution of policies \i\ to \j, with a budget of 2.}
-        \label{fig:policy-executions-b2-\i}
-    \end{figure}
-}
-
-\foreach \i/\j in {0/19, 20/39, 40/59, 60/79, 80/99, 100/119, 120/139, 140/159, 160/160}{
-    \begin{figure}
-        \centering
-        \plotPolicies{\i}{\j}{3}
-        \caption{Execution of policies \i\ to \j, with a budget of 3.}
-        \label{fig:policy-executions-b3-\i}
-    \end{figure}
-}
-
-\foreach \i/\j in {0/19, 20/39, 40/59, 60/79, 80/93}{
-    \begin{figure}
-        \centering
-        \plotPolicies{\i}{\j}{4}
-        \caption{Execution of policies \i\ to \j, with a budget of 4.}
-        \label{fig:policy-executions-b4-\i}
-    \end{figure}
-}
-
-\foreach \i/\j in {0/19, 20/39, 40/59, 60/79, 80/99, 100/114}{
-    \begin{figure}
-        \centering
-        \plotPolicies{\i}{\j}{5}
-        \caption{Execution of policies \i\ to \j, with a budget of 5.}
-        \label{fig:policy-executions-b5-\i}
-    \end{figure}
-}
-
-\foreach \i/\j in {0/19, 20/39, 40/59, 60/79, 80/99, 100/116}{
-    \begin{figure}
-        \centering
-        \plotPolicies{\i}{\j}{None}
-        \caption{Execution of policies \i\ to \j, with no budget restriction ($b=\infty$).}
-        \label{fig:policy-executions-binf-\i}
-    \end{figure}
-}
 

### 2023-10-19 mones/mones.py
mones baseline

@@ -1,0 +1,250 @@
+import torch
+import torch.nn as nn
+import torch.nn.functional as F
+import numpy as np
+import copy
+from mones.metrics import non_dominated, compute_hypervolume
+from logger import Logger
+import torch.multiprocessing as mp
+
+
+class Multiprocessor(object):
+
+    def __init__(self):
+        self.processes = []
+        self.queue = mp.Manager().Queue()
+
+    @staticmethod
+    def _wrapper(func, queue, args, kwargs):
+        torch.set_num_threads(2)
+        ret = func(*args, **kwargs)
+        queue.put(ret)
+
+    def run(self, func, *args, **kwargs):
+        args2 = [func, self.queue, args, kwargs]
+        p = mp.Process(target=self._wrapper, args=args2)
+        self.processes.append(p)
+        p.start()
+
+    def wait(self):
+        rets = []
+        for p in self.processes:
+            ret = self.queue.get()
+            rets.append(ret)
+        for p in self.processes:
+            p.join()
+        return rets
+
+
+
+def n_parameters(model):
+    return np.sum([torch.prod(torch.tensor(p.shape)) for p in model.parameters()])
+
+
+def set_parameters(model, z):
+    assert len(z) == n_parameters(model), 'not providing correct amount of parameters'
+    s = 0
+    for p in model.parameters():
+        n_p = torch.prod(torch.tensor(p.shape))
+        p.data = z[s:s+n_p].view_as(p)
+        s += n_p
+
+
+def run_episode(env, model):
+
+    e_r = 0; done = False
+    o = env.reset()
+    actions = []
+    while not done:
+        with torch.no_grad():
+            obs = [torch.tensor([o_]) for o_ in o] if type(o) == tuple else torch.tensor([o])
+            actor_out = model(obs)
+            action = actor_out
+            action = torch.argmax(actor_out, 1)
+
+            action = action.numpy().flatten()[0]
+            actions.append(action)
+
+        n_o, r, done, _ = env.step(action)
+        e_r += r
+        o = n_o
+
+    return torch.from_numpy(e_r).float()
+
+
+def run_n_episodes(env, model, n):
+    returns = [run_episode(env, model) for _ in range(n)]
+    return torch.mean(torch.stack(returns, dim=0), dim=0)
+
+
+def indicator_hypervolume(points, ref, nd_penalty=0.):
+    # compute hypervolume of dataset
+    nd = non_dominated(points)
[…]

### 2023-10-21 mones/mones.py
mones continuous action bugfix

@@ -59,10 +59,11 @@ def run_episode(env, model):
         with torch.no_grad():
             obs = [torch.tensor([o_]) for o_ in o] if type(o) == tuple else torch.tensor([o])
             actor_out = model(obs)
-            action = actor_out
-            action = torch.argmax(actor_out, 1)
+            action = actor_out.numpy().flatten()
+            # TODO hacky
+            if model.__class__.__name__ != 'ContinuousHead':
+                action = np.argmax(action)
 
-            action = action.numpy().flatten()[0]
             actions.append(action)
 
         n_o, r, done, _ = env.step(action)

### 2024-05-28 paper/main.tex
main

@@
 \subsection{Impact of budgets on the coverage set}
+\label{sec:pf-binomial-arh-budgets}
 
@@
     \label{fig:pf-binomial-arh-ari}
+\end{figure}
+
+\subsection{Comparison of PCN with a dynamic baseline}
+\label{sec:cs-pcn-vs-mones}
+
+In this work, we focus on the application of multi-objective reinforcement learning for epidemic outbreaks.
+As a reference, we use fixed policies to show that the dynamic policies we learn are able to properly adapt to the evolution of the epidemic.
+We extend PCN towards this application, and assess the quality of the learned policies (see Fig.~\ref{fig:policy-execution},\ref{fig:pe-binomial-arh-budgets}).
+Nonetheless, the theoretical Pareto front of optimal policies is unknown.
+To confirm the quality of the policies learned by PCN, we compare it to  Multi-Objective Natural Evolution Strategies (MONES)~\cite{parisi2017manifold}, a dynamic baseline which also learns a coverage set of policies, and makes no assumptions on the shape of the Pareto front.
+
+Given a parametrized policy, MONES learns a distribution over the policy parameters such that sampling from this distribution produces a Pareto-efficient policy.
+Each sample produces a different policy, leading to a different non-dominated return.
+This poses the advantage that we can sample infinitely many different policies.
+However, as a drawback, we do not know the return the sampled policy will achieve without executing it first.
+Another considerable drawback is that MONES computes the Fisher information matrix, which is squared in the number of policy-parameters.
+Due to the size of \momdpname's state-space and, consequently, the size of the neural network that represents the policy, it becomes a considerable challenge to compute this matrix.
+To avoid this matrix to be intractable, we had to make assumptions on the space of policy-parameters, i.e., zero-correlation between parameters.
+
+We train MONES on both the Binomial and ODE version of \momdpname, for which the results can be seen on Fig.~\ref{fig:pf-binomial-pcn-mones},~\ref{fig:pf-ode-pcn-mones}, respectively.
+In both cases, MONES outperforms the fixed baseline, especially in the less restrictive policies in terms of social burden.
+However, PCN systematically learns policies that dominate the ones learned by MONES.
+Moreover, as explained in Sec.~\ref{sec:pf-binomial-arh-budgets}, given a higher budget, PCN learns increasingly better policies.
+This seems to be less the case for MONES.On the Binomial environment, the difference in budget barely impacts the learned policies.
+On the ODE environment, this difference is only visible on a limited region of the coverage set.
+Thus, MONES makes a suboptimal use of its budget.
+This confirms that anticipating the evolution of the epidemic to make best use of the leftover budget is a challenging problem, for which the use of adequate MORL techniques could greatly benefit decision-makers.
+
+\begin{figure}[t!]
+    \centering
+    \input{fig/arh_binomial_mones}
+    \caption{Comparison of PCN and MONES on the Binomial environment}
+    \label{fig:pf-binomial-pcn-mones}
+\end{figure}
+
+\begin{figure}[t!]
+    \centering
+    \input{fig/arh_ode_mones}
+    \caption{Comparison of PCN and MONES on the ODE environment}
+    \label{fig:pf-ode-pcn-mones}
 \end{figure}

### 2024-05-28 paper/policy-executions.tex
main

--- 2023-03-31 paper/main.tex
@@
 \pgfplotsset{compat=newest}
-\tikzexternalize[prefix=cache/]
+\tikzexternalize[prefix=cache-policies/]
+
+\newcommand{\momdpname}{MOBelCov}
 
@@
 
-% default \cite to use parentheses
-\renewcommand{\cite}[1]{\citep{#1}}
-
-\newcommand{\foi}{\lambda}
-\newcommand\todo[1]{\textcolor{red}{(TODO: #1)}}
-
-%MDP
-%state space
-\newcommand{\mdpstatespace}{\mathcal{S}}
-%action space
-\newcommand{\mdpactionspace}{\mathcal{A}}
-%mdp discount factor
-\newcommand{\mdpdiscount}{\gamma}
-%mdp policy
-\newcommand{\mdppolicy}{\pi}
-%mdpvaluefunction
-\newcommand{\mdpvaluefunction}{V}
-%vectorial value function
-\newcommand{\momdpvaluefunction}{\mathbf{V}}
-%mdpqfunction
-\newcommand{\mdpqfunction}{Q}
-%mdp q function estimate
-\newcommand{\mdpqfunctionestimate}{\hat{Q}}
-%mdp optimal q function
-\newcommand{\mdpoptimalqfunction}{Q^{*}}
-%q step size
-\newcommand{\qstepsize}{\alpha}
-%\advantageestimate
-\newcommand{\advantageestimate}{\hat{A}}
-%transition function
-\newcommand{\mdptransition}{\mathcal{T}}
-%reward fn
-\newcommand{\mdprewardfn}{\mathcal{R}}
-%vectorial reward fn
-\newcommand{\momdprewardfn}{\bm{\mathcal{R}}}
-%mdp state
-\newcommand{\mdpstate}{\mathbf{s}}
-%mdp reward
-\newcommand{\mdpreward}{r}
-\newcommand{\momdpreward}{\mathbf{r}}
-\newcommand{\momdpreturn}{\mathbf{R}}
-%mdp action
-\newcommand{\action}{a}
-%vectorial action
-\newcommand{\mdpaction}{\mathbf{a}}
-% the 6 social contact matrices
-\newcommand{\ctotal}{C}
-\newcommand{\chome}{C_{\text{home}}}
-\newcommand{\cwork}{C_{\text{work}}}
-\newcommand{\ctransport}{C_{\text{transport}}}
-\newcommand{\cschool}{C_{\text{school}}}
-\newcommand{\cleisure}{C_{\text{leisure}}}
-\newcommand{\cother}{C_{\text{other}}}
-% tentative momdp env name
-\newcommand{\momdpname}{MOBelCov}
-%other params
-\newcommand{\proportionalityfactor}{q}
-\newcommand{\agegroup}{k}
-\newcommand{\agegroups}{K}
-\newcommand{\budget}{\bm{b}}
-
-\newcommand{\ltuple}{\langle}
-\newcommand{\rtuple}{\rangle}
-
-% color scheme for plots: plot-color-i
-\definecolor{pc1}{RGB}{76, 120, 168}
-\definecolor{pc2}{RGB}{245, 133, 24}
-\definecolor{pc3}{RGB}{228, 87, 88}
-\definecolor{pc4}{RGB}{114, 183, 178}
-\definecolor{pc5}{RGB}{84, 162, 75}
-\definecolor{pc6}{RGB}{238, 202, 59}
-\definecolor{pc7}{RGB}{178, 121, 162}
-
-\title{Exploring the Pareto front of multi-objective COVID-19 mitigation policies using reinforcement learning}
-
-% Here you can change the date presented in the paper title
-%\date{September 9, 1985}
-% Or remove it
-\date{}
-
-\author{Mathieu Reymond \\
-	Vrije Universiteit Brussel\\
-	Brussels, Belgium \\
-	\texttt{mathieu.reymond@vub.be} \\
-	\And
-	Conor F. Hayes \\
-	National University of Ireland Galway\\
-	Galway, Ireland\\
-% 	\texttt{c.hayes13@nuigalway.ie} \\
-	\And
-	Lander Willem \\
-	University of Antwerp \\
-	Antwerp, Belgium \\
-	\And
-	Roxana R\u{a}dulescu \\
-	Vrije Universiteit Brussel\\
-	Brussels, Belgium \\
-	\And
-	Steven Abrams \\
-	Hasselt University \\
-	Hasselt, Belgium \\
-	\And
-	Diederik M.\ Roijers \\
-	HU University of Applied Sciences Utrecht \\
-	Utrecht, the Netherlands \\
-	\And
-	Enda Howley \\
-	National University of Ireland Galway\\
-	Galway, Ireland\\
-	\And
-	Patrick Mannion \\
-	National University of Ireland Galway\\
-	Galway, Ireland\\
-	\And
-	Niel Hens \\
-	Hasselt University \\
-	Hasselt, Belgium \\
-	\And
-	Ann Now\'{e} \\
-	Vrije Universiteit Brussel\\
-	Brussels, Belgium \\
-	\And
-	Pieter Libin \\
-	Vrije Universiteit Brussel\\
-	Brussels, Belgium \\
-    \texttt{pieter.libin@vub.be} \\
-% 	\AND
-% 	Coauthor \\
-% 	Affiliation \\
-% 	Address \\
-% 	\texttt{email} \\
-}
-
-% \author{Mathieu Reymond\inst{1}{\Letter} \and Conor F.
-Hayes \inst{4} \and Lander Willem  \inst{6}\and Roxana R\u{a}dulescu \inst{1} \and Steven Abrams \inst{3,7} \and Diederik M.\ Roijers\inst{1,5}\and Enda Howley\inst{4} \and Patrick Mannion\inst{4} \and Niel Hens \inst{3,6} \and Ann Now\'{e}\inst{1} \and Pieter Libin\inst{1,2,3}{\Letter}}
-
-% \authorrunning{M. Reymond et al.}
-
-% \institute{ 
-% Artificial intelligence lab, Vrije Universiteit Brussel, Brussels, Belgium \email{\{pieter.libin,mathieu.reymond,diederik.roijers,ann.nowe\}@vub.be} 
-% \and
-% Rega institute, KU Leuven, Leuven, Belgium 
-% \and  
-% Data science institute, Hasselt University, Hasselt, Belgium
-% \and
-% National University of Ireland Galway, Ireland \email{\{c.hayes13\}@nuigalway.ie}
-% \and
-% HU University of Applied Sciences Utrecht, the Netherlands
-% \and
-% Centre for Health Economic Research and Modelling Infectious Diseases
-% CHERMID, University of Antwerp, Belgium
-% \and
-% Global Health Institute, University of Antwerp, Belgium
-% }
-
-% Uncomment to override  the `A preprint' in the header
-%\renewcommand{\headeright}{Technical Report}
-%\renewcommand{\undertitle}{Technical Report}
-\renewcommand{\shorttitle}{Exploring the Pareto front of multi-objective COVID-19 mitigation policies}
-
-%%% Add PDF metadata to help others organize their library
-%%% Once the PDF is generated, you can check the metadata with
-%%% $ pdfinfo template.pdf
-\hypersetup{
-pdftitle={Exploring the Pareto front of multi-objective COVID-19 mitigation policies using reinforcement learning},
-pdfauthor={Mathieu Reymond, Conor F.~Hayes, Lander Willem, Roxana R\u{a}dulescu, Steven Abrams , Diederik M.\ Roijers, Enda Howley, Patrick Mannion, Niel Hens, Ann Now\'{e}, Pieter Libin},
-pdfkeywords={Multi-objective reinforcement learning, Epidemic control, COVID-19 epidemic models},
-}
-
+\pagestyle{empty}
 \begin{document}
-\maketitle
-
-\begin{abstract}
-Infectious disease outbreaks can have a disruptive impact on public health and societal processes.
-As decision making in the context of epidemic mitigation is hard, reinforcement learning provides a methodology to automatically learn prevention strategies in combination with complex epidemic models.
-Current research focuses on optimizing policies with respect to a single objective, such as the pathogen's attack rate.
-However, as the mitigation of epidemics involves distinct, and possibly conflicting, criteria (i.a., prevalence, mortality, morbidity, cost), a multi-objective decision approach is warranted to learn balanced policies.
-To lift this decision-making process to real-world epidemic models, we apply deep multi-objective reinforcement learning and build upon a state-of-the-art algorithm, Pareto Conditioned Networks (PCN), to learn a set of solutions that approximates the Pareto front of the decision problem.
-We consider the first wave of the Belgian COVID-19 epidemic, which was mitigated by a lockdown, and study different deconfinement strategies, aiming to minimize both COVID-19 cases (i.e., infections and hospitalizations) and the societal burden that is induced by the applied mitigation measures.
-We contribute a multi-objective Markov decision process that encapsulates the stochastic compartment model that was used to inform policy makers during the COVID-19 epidemic.
-As these social mitigation measures are implemented in a continuous action space that modulates the contact matrix of the age-structured epidemic model, we extend PCN to this setting.
-We evaluate the solution set that PCN returns, and observe that it correctly learns to reduce the social burden whenever the hospitalization rates are sufficiently low.
-In this work, we thus demonstrate that multi-objective reinforcement learning is attainable in complex epidemiological models and provides essential insights to balance complex mitigation policies.
-\end{abstract}
-
-% keywords can be removed
-\keywords{Multi-objective reinforcement learning  \and Epidemic control \and COVID-19 epidemic models}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction}
-Infectious disease outbreaks represent a major global challenge \cite{miranda2022tale}.
-To this end, understanding the complex dynamics that underlie these epidemics is essential.
-Epidemiological transmission models allow us to capture and understand such dynamics and facilitate the study of prevention strategies through simulation.
-However, developing efficient mitigation strategies remains a challenging process, given the non-linear and complex nature of epidemics.
-
-To address these challenges, reinforcement learning provides a methodology to automatically learn mitigation strategies in combination with complex epidemic models~\cite{libin2020}.
-Previous research focused on optimising policies with respect to a single objective, such as the pathogen's attack rate, while the mitigation of epidemics is a problem that inherently covers distinct  and possibly conflicting criteria (i.a., prevalence, mortality, morbidity, cost).
-Therefore, optimizing on a single objective requires that  these distinct criteria are somehow aggregated into a single metric.
-Typically, a decision maker will rely on an expert's assistance to manually design such a metric.
-However, this is a challenging process, as the decision maker is generally not aware of all possible trade-offs, and as such this may result in a metric that exhibits suboptimal performance.
-Furthermore, manually designing such metrics is time-consuming, costly and error-prone, as this non-intuitive process requires repetitive and tedious tuning to achieve the desired behaviour~\cite{roijers2013survey}.
-Moreover, taking a single objective approach has several other limiting factors.
-For example, it reduces the explainability of the learned solution, as we cannot compare the learned behavior with alternatives.
-Moreover, the preferences of the decision maker might change over time, in which case the optimization process needs to be done again~\cite{hayes2021practical}.
-
-This challenging process can be circumvented by taking an explicit multi-objective approach that aims to learn the different trade-offs regarding the considered criteria.
-By assuming that a decision maker will always prefer solutions for which at least one objective improves, it is possible to learn a set of optimal solutions referred to as the \emph{Pareto front}~\cite{hayes2021practical}.
-This enables decision makers to review each solution on the Pareto front before making a decision, thereby being aware of the trade-offs that each solution implies.
-
-In this work, we investigate the use of \emph{multi-objective reinforcement learning} (MORL) to learn a set of solutions that approximate the Pareto front of multi-objective epidemic mitigation strategies.
-We consider the first wave of the Belgian COVID-19 epidemic, which was mitigated by a strict lockdown~\cite{willem2021impact}.
-When the incidence of confirmed cases was steadily decreasing, epidemiological experts were tasked to investigate deconfinement strategies, to reduce the severe social contact and mobility restrictions.
-Here, we consider an epidemiological model that was constructed to describe the Belgian COVID-19 epidemic and was fitted to hospitalisation incidence data and serial sero-prevalence data~\cite{abrams2021modelling}.
-This model constitutes a stochastic discrete-time age-structured compartmental model that simulates mitigation strategies by varying social distancing parameters concerning school, work and leisure contacts.
-Based on this model, we contribute a novel mutli-objective epidemiological reinforcement learning environment~(\momdpname), in the form of a multi-objective Markov decision process (MOMDP) \cite{roijers2013survey}.
-\momdpname\ encapsulates the epidemiological model developed by \citet{abrams2021modelling} to implement state transitions, with an action space that combines a proportional reduction of school, work and leisure contacts at each time step, Furthermore, it defines a reward function based on two objectives: the attack rate (i.e., proportion of the population affected by the pathogen) and the social burden that is induced by the mitigation measures.
-
-To learn and explore the trade-offs between the attack rate and social burden we use a state-of-the-art MORL approach based on Pareto Conditioned Networks (PCN)~\cite{reymond2022pcn}.
-PCN uses a single neural network to learn the policies that belong to the Pareto front.
-As PCN is an algorithm designed for discrete action-spaces, we extend it towards continuous action-spaces to accommodate \momdpname's action-space.
-With this continuous action variant of PCN, we explore the Pareto front of multi-objective COVID-19 mitigation policies.
-
-By evaluating the solution set of mitigation policies learned by PCN, we observe that PCN minimises the social burden in scenarios where hospitalization rates are sufficiently low.
-Therefore, in this work we illustrate that multi-objective reinforcement learning can provide important insights concerning the trade-offs between complex mitigation polices in real-world epidemiological models.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Background}
-In the following, we use bold notation to denote variables with a vectorial value, while non-bold notation for its scalar counterpart.
-
-\subsection{Multi-Objective Reinforcement Learning}
-Real-world decisions problems typically consider multiple and possibly conflicting objectives.
-Multi-objective reinforcement learning (MORL) can be used to find optimal solutions for sequential decision making problems with multiple objectives~\cite{hayes2021practical}.
-Multi-objective sequential problems are typically modelled as a multi-objective Markov decision process (MOMDP), i.e., a tuple, $\mathcal{M} = \ltuple\mdpstatespace, \mdpactionspace, \mdptransition, \mdpdiscount, \momdprewardfn\rtuple$, where $\mdpstatespace$, $\mdpactionspace$ denote the state and action spaces respectively, $\mdptransition \colon \mdpstatespace \times \mdpactionspace \times \mdpstatespace  \to \left[ 0, 1 \right]$ denotes a probabilistic transition function, $\mdpdiscount$ is a discount factor determining the importance of future rewards and $\momdprewardfn \colon \mdpstatespace \times \mdpactionspace \times \mdpstatespace \to \mathbb{R}^n$ is an $n$-dimensional vector-valued immediate reward function, where $n$ corresponds to the number of objectives.
-
-While MOMDPs define the actions an agent can take, it does not define the agent's behavior, i.e., which actions are taken in each state.
-Given this MOMDP, an agent follows a policy $\mdppolicy$, that expresses the probability to take action $\action \in \mdpactionspace$ when in state $\mdpstate \in \mdpstatespace$: $\mdpstatespace \times \mdpactionspace \to \left[ 0, 1 \right]$.
-We measure the performance of $\mdppolicy$ through the expected sum of discounted rewards (denoted the \emph{Value}) it achieves from the initial state $\mdpstate_0$ until the end of the decision problem:
-
-\begin{equation}
-    \momdpvaluefunction^\mdppolicy = \mathbb{E} \left [ \sum_{t=0}^h \mdpdiscount^t \momdpreward_t |\ \mdppolicy, \mdpstate_0 \right ],
-\end{equation}
-%
-where $\momdpreward_t$ corresponds to the multi-objective reward observed at time $t$, by following policy $\mdppolicy$, given by the reward function $\mdprewardfn(\mdpstate_t, \action_t, \mdpstate_{t+1})$, where $\mdpstate_t$ is the current state, $\action_t$ corresponds to the action that was taken, and $\mdpstate_{t+1}$ signifies the state reached by taking action $\action_t$.
-
-On the one hand, for single-objective RL, where $n = 1$, the goal is to find the policy $\mdppolicy^{*}$ that maximizes the $\mdpvaluefunction$-value:
-\begin{equation}
-\mdppolicy^*= \arg\max_\mdppolicy \mdpvaluefunction^\mdppolicy.
-\end{equation}
-
-On the other hand, for MORL, $n > 1$ which leads to vectorial returns.
-In this case, there can be policies for which, without any additional information, it is impossible to know if one is better than the other.
-For example, it is impossible to decide which policy between $\mdppolicy_1, \mdppolicy_2$ is optimal if both policies lead to expected returns $\momdpvaluefunction^{\mdppolicy_1}=(0,1), \momdpvaluefunction^{\mdppolicy_2}=(1,0)$ respectively.
-We call these solutions \emph{non-dominated}, i.e., solutions for which it is impossible to improve an objective without hampering another.
-The set that contains all the non-dominated solutions of the decision problem is called the \emph{Pareto front} $\mathcal{F}$.
-Our goal is to find the set of policies that lead to all the $\mdpvaluefunction$-values contained in the Pareto front $\Pi^* = \{\ \pi | \momdpvaluefunction^{\pi} \in \mathcal{F}\}$.
-In general, we call any set of $\mdpvaluefunction$-values a \emph{solution set}.
-When a solution set contains only  non-dominated $\mdpvaluefunction$-values, it is referred to as a \emph{coverage set}.
-In the case that no $\mdppolicy$ exists that has a $\momdpvaluefunction^{\pi}$ dominating any of the solutions in a coverage set, then this coverage set is the Pareto front.
-
-\subsection{Multi-Objective Metrics}
-Comparing the learned coverage sets produced by different algorithms is a non-trivial task, as one algorithm's output might dominate the other in some region of the objective-space, but be dominated in another.
-Intuitively, one would generally prefer the algorithm that covers a wider range of decision maker preferences.
-%In this work we use several metrics to evaluate our algorithm's performance.
-
-A widely used metric in the literature is called the \emph{hypervolume}~\cite{zitzler2003}.
-This metric evaluates the learned coverage set by computing its volume with respect to a fixed reference point.
-The reference point is taken as a lower bound on the achievable returns so that the volumes are always positive.
-By definition, the solutions contained in the Pareto front dominate all other possible solutions.
-Thus, no other solution can further increase the volume under the Pareto front.
-This means that the hypervolume is the highest for the Pareto front.
-One drawback of the hypervolume is that it can be difficult to interpret.
-For example, when working in high-dimensional objective-spaces, adding or removing a single point can drastically change hypervolume values, especially if the point lies close to an extremum of said space.
-
-To counterpoise these shortcomings, we evaluate our work using an additional metric called the $\varepsilon$-indicator $I_\varepsilon$~\cite{zitzler2003}, which measures how close a coverage set is to the Pareto front $\mathcal{F}$.
-Intuitively, $I_\varepsilon$ shows that any solution of $\mathcal{F}$ is \emph{at most} $\varepsilon$ better with respect to each objective $o$ than the closest solution of the evaluated coverage set:
-
-\begin{equation}
-\label{eq:epsilon-metric}
-    I_{\varepsilon} = 
-    \inf_{\varepsilon\in\mathbb{R}} 
-    \{ 
-    \forall\ \momdpvaluefunction^\pi\!{\in}\ \mathcal{F}, ~
-    \exists\ \momdpvaluefunction^{\pi'}\!{\in}\ \hat{\Pi} :\ 
-    || \momdpvaluefunction^\pi - \momdpvaluefunction^{\pi'} ||_{\infty} \le \varepsilon
-    \},
-\end{equation}
-%
-where $||.||_{\infty}$, the L-infinity norm, is defined as the magnitude of the largest entry of a vector.
-
-The main disadvantage of this metric is that we need the true Pareto front to compute it, which is unknown for our MOMDP (see Sec.~\ref{sec:sars-cov2-momdp}).
-To still gain insights from our learned policies, we approximate the true Pareto front using the non-dominated policies across all runs.
-
-We note that the $\varepsilon$-indicator metric is quite pessimistic, as it measures worst-case performance~\cite{zintgraf2015}, i.e., it will still report low performance as long as a single point of the Pareto front is incorrectly modeled, even if all the other points are covered.
-As such, we also use the $I_{\varepsilon-mean}$ which measures the \emph{average} $\varepsilon$ distance of the solutions in $\mathcal{F}$ with respect to the evaluated coverage set~\cite{reymond2022pcn}.
-
-Fig.~\ref{fig:paretofrontmetrics} shows a visual representation of the hypervolume and $\varepsilon$ metrics in two dimensions.
-
-\begin{figure}[t!]
-    \centering
-    \begin{tikzpicture}
-      \begin{axis}[legend pos=south west, xmin=-1, xmax=4.5, ymin=-0.5, ymax=4.5,ytick={0,1,2,3,4,5},scale only axis,width=0.4\textwidth]
-      \addplot[only marks,mark=x,mark size=3] coordinates {(-0.5,0)};
-      \addplot[only marks,mark=*] coordinates {(1,4)(2,2)(4,1)};
-      \addplot[only marks,mark=o] coordinates {(0.5,3)(0.75,2.3)(2.3,1)(3.3,0.7)};
-      % Hypervolume
-      \fill [cyan,opacity=0.2] (-0.5,0) -- (-0.5,3) -- (0.5,3) -- (0.5,2.3) -- (0.75,2.3) --
-      (0.75,1) -- (2.3,1) -- (2.3,0.7) -- (3.3,0.7) -- (3.3,0) -- cycle;
-      % epsilons
-      \draw [dashed, -|] (0.5, 3) -- node[right]{$\varepsilon_1$} (0.5,4);
-      %\draw [dashed, -|] (0.75,2.3) -- node[below]{x} (2,2.3);
-      \draw [dashed, -|] (2.3,1) -- node[right]{$\varepsilon_2$} (2.3,2);
-      \draw [dashed, -|] (3.3,0.7) -- node[below]{$\varepsilon_3$} (4,0.7);
-      \draw [opacity=0.2,line width=2,orange, dotted] (-0.5,0) -- (-0.5,4) -- (1,4) -- (1,2) --
-      (2,2) -- (2,1) -- (4,1) -- (4,0) -- (-0.5,0);
-      \end{axis}
-    \end{tikzpicture}
-    \caption{Example of a Pareto front (black dots) and a coverage set (white dots) in a 2-objective environment.
-The hypervolume metric (in light blue) measures the volume of all dominated solutions with respect to a reference point (cross).
-The reference point is taken as a lower bound on the achievable returns, which is why it can differ from the origin.
-The $\varepsilon$ metrics first compute the maximum distance between each point in the Pareto front and its closest point in the coverage set ($\varepsilon_i$).
-We can then take their maximum value to compute the $I_\varepsilon$ metric, or their mean value to obtain the $I_{\varepsilon-mean}$ metric of the coverage set.}
-    \label{fig:paretofrontmetrics}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{COVID-19 model and \momdpname\  MOMDP}
-\label{sec:sars-cov2-momdp}
-\subsection{Stochastic compartment model for SARS-CoV-2}
-\label{sec:sars-cov2-momdp_model}
-To evaluate non-pharmaceutical interventions, we consider the compartmental model presented by Abrams et al., that was used to investigate exit strategies in Belgium after the first epidemic wave of SARS-CoV-2~\cite{abrams2021modelling}.
-This model concerns a discrete-time stochastic model, that considers an age-structured population.
-The model generalises a standard SEIR model\footnote{A standard SEIR model divides the population into four different states, i.e., susceptible, exposed, infectious and recovered individuals.}, extended to capture the different stages of disease spread and history that are associated with SARS-CoV-2 (i.e., pre-symptomatic, asymptomatic, symptomatic with mild symptoms and symptomatic with severe symptoms) and to represent the stages associated with severe disease, i.e., hospitalisation, admission to the intensive care unit (ICU) and death.
-
-A visual representation of the model is depicted in Fig.~\ref{fig:compartment_model}. It shows the different compartments, as well as the flow rates at which individuals move between compartments.
-
-These flow rates are defined by a set of ordinary differential equations, which are outlined as follows:
-
-\begin{align*}
-\frac{d\textbf{S}(t)}{dt} & = -\lambda(t)(\textbf{S})(t), \\
-%
-\frac{d\textbf{E}(t)}{dt} & = \lambda(t)(\textbf{S})(t) - \gamma \textbf{E}(t), \\
-%
-\frac{d\textbf{I}^{presym}(t)}{dt} & = \gamma \textbf{E}(t) - \theta \textbf{I}^{presym}(t), \\
-%
-\frac{d\textbf{I}^{asym}(t)}{dt} & = \theta p \textbf{I}^{presym}(t) - \delta_{1} \textbf{I}^{asym}(t), \\
-%
-\frac{d\textbf{I}^{mild}(t)}{dt} & = \theta (1 - p) \textbf{I}^{presym}(t) - \{ \psi + \omega_{2} \} \textbf{I}^{mild}(t), \\
-%
-\frac{d\textbf{I}^{sev}(t)}{dt} & = \psi \textbf{I}^{mild}(t) - \omega \textbf{I}^{sev}(t), \\
-%
-\frac{d\textbf{I}^{hosp}(t)}{dt} & = \phi_{1} \omega \textbf{I}^{sev}(t) - \{ \delta_{3} + \tau_{1} \} \textbf{I}^{hosp}(t), \\
-%
-\frac{d\textbf{I}^{icu}(t)}{dt} & = (1 - \phi_1) \omega \textbf{I}^{sev}(t) - \{ \delta_{4} + \tau_{2} \} \textbf{I}^{icu}(t), \\
-%
-\frac{d\textbf{D}(t)}{dt} & = \tau_{1} \textbf{I}^{hosp}(t) - \tau_{2} \textbf{I}^{icu}(t), \\
-%
-\frac{d\textbf{R}(t)}{dt} & = \delta_{1} \textbf{I}^{asym}(t) + \delta_{2} \textbf{I}^{mild}(t) + \delta_{3} \textbf{I}^{hosp}(t) + \delta_{4} \textbf{I}^{icu}(t)
-\end{align*}
-%
-
-In this set of ordinary differential equations, each state variable represents a vector over all age groups for a particular compartment at time $t$.
-For example, $\textbf{S} = (S_{1}(t), S_{2}(t), ..., S_{\agegroup}(t))^{T}$ is the vector representing the susceptible members of the population of each age group $\agegroup$ at time $t$.
-Infection dynamics are governed by an age-specific force of infection $\foi$:
-
-\begin{equation}
-    \foi(\agegroup,t) = \sum_{\agegroup'=1}^{\agegroups}\beta(\agegroup,\agegroup')I_{\agegroup'}(t),
-\end{equation}
-%
-where $\agegroups$ is the total number of age groups, and $\beta(\agegroup,\agegroup')$ is the time-invariant transmission rate that encodes the average per capita rate at which an infectious individual in age group $\agegroup$ makes an effective contact with a susceptible individual in age group $\agegroup'$, per unit of time.
-
-As we consider an age-structured population, we consider this extended SEIR structure for $\agegroups = 10$ age groups,  i.e., $[0-10), [10-20), [20-30), [30-40), [40-50), [50-60), [60-70), [70-80), [80-90), [90,100+)$.
-Contacts of the different age-groups, which impact the propagation rate of the epidemic, are modeled using social contact matrices.
-We define a social contact matrix for 6 different social environments: $\chome, \cwork, \ctransport, \cschool, \cleisure, \cother$, for the home, work, transport, school, leisure, other environments respectively.
-The social contact matrix across all social environments is defined as:
-
-\begin{equation}
-    \ctotal = \chome + \cwork + \ctransport + \cschool + \cleisure + \cother
-\end{equation}
-
-Under the social contact hypothesis \cite{wallinga2006using}, we have that:
-%
-\begin{equation}
-\beta(\agegroup,\agegroup') = \proportionalityfactor \cdot \ctotal(\agegroup,\agegroup'),
-\end{equation}
-%
-where $\proportionalityfactor$ is a proportionality factor.
-
-Following \citet{abrams2021modelling}, we rely on distinct social contact matrices for symptomatic and asymptomatic individuals, respectively $\ctotal_{s}$ and $\ctotal_{a}$.
-Therefore, we define the transmission rates for both symptomatic and asymptomatic individuals as follows:
-
-\begin{equation}
-\boldsymbol{\beta}_{s}(\agegroup, \agegroup) = \proportionalityfactor_{s} \cdot C_{s}(\agegroup, \agegroup'),    
-\end{equation}
-and
-\begin{equation}
-\boldsymbol{\beta}_{a}(\agegroup, \agegroup') = \proportionalityfactor_{a} \cdot C_{a}(\agegroup, \agegroup').  
-\end{equation}
-
-The age-dependent force of infection can be defined as follows:
-
-\begin{equation}
-    \boldsymbol{\lambda}(t) = \boldsymbol{\beta}_{a} \times \{ \textbf{I}^{presym}(t) + \textbf{I}^{asym}(t) \} +
-    \boldsymbol{\beta}_{s} \times \{ \textbf{I}^{mild}(t) + \textbf{I}^{sev}(t) \},
-\end{equation}
-%
-where $\boldsymbol{\lambda}(t) = (\lambda(1, t), \lambda(2, t), ..., \lambda(\agegroups, t))$.
-For all further information about the different compartments and parameters we refer the reader to the work of \citet{abrams2021modelling}.
-
-\todo{epi friends, please check} Time- and age-dependent behavioral changes introduce substantial uncertainty regarding the outcome of an outbreak.
-To evaluate intervention strategies that modulate such behavioral changes, the use of stochastic epidemiological models is warranted \cite{abrams2021modelling}.
-Moreover, the effect of stochasticity is most pronounced when studying the initial growth of an epidemic \cite{britton2009epidemic}, be it during its emergence, or when implementing deconfinement strategies.
-
-By formulating the set of differential equations defined above, as a chain-binomial, we can obtain stochastic trajectories from this model \cite{bailey1975mathematical}.
-A chain-binomial model assumes a stochastic model where infected individuals are generated by some underlying probability distribution.
-For this stochastic model we consider a time interval $(t, t +h]$, where $h$ is defined as the length between two consecutive time points.
-In this work we set $h = \frac{1}{240}$.
-
-\subsection{Interventions strategies}
-\label{sec:interventions}
-To model different types of interventions, we follow the contact reduction scheme presented by \citet{abrams2021modelling}.
-Firstly, to consider distinct exit scenarios, we modulate the social contact matrices to reflect a contact reduction in a particular age group.
-
-We consider a contact reduction function that imposes a proportional reduction of work (including transport) $p_w$, school $p_s$ and leisure $p_l$ contacts, which is implemented as a linear combination of social contact matrices:
-\begin{equation}
-\label{eq:scm}
-    \hat{C}(p_w, p_s, p_l) = \chome + p_w (\cwork + \ctransport) + p_s \cschool + p_l (\cleisure + \cother)
-\end{equation}
-
-We denote $\hat{C}_t$ the social contact matrix at timestep $t$, resulting from the reduction function $\hat{C}$.
-Secondly, we assume that compliance to the interventions is gradual and model this using a logistic compliance function (see details in Sec.~\ref{sec:compliance-function} of the Appendix).
-
-\subsection{The \momdpname\ Environment}
-\begin{figure}[t!]
-    \centering
-    \begin{tikzpicture}
-     \draw (0.0pt, 0.0pt)node[fill=green!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](0){$\textbf{S}$};
-     \draw (60.0pt, 0.0pt)node[fill=orange!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](1){$\textbf{E}$};
-     \draw (120.0pt, 0.0pt)node[fill=red!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](2){$\textbf{I}^{presym}$};
-     \draw (180.0pt, 10.0pt)node[fill=red!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](3){$\textbf{I}^{asym}$};
-     \draw (170.0pt, -30.0pt)node[fill=red!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](4){$\textbf{I}^{mild}$};
-     \draw (170.0pt, -70.0pt)node[fill=red!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](5){$\textbf{I}^{sev}$};
-     \draw (230.0pt, -50.0pt)node[fill=blue!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](6){$\textbf{I}^{hosp}$};
-      \draw (75.0pt, -70.0pt)node[fill=gray!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](10){$\textbf{H}^{new}$};
-     \draw (230.0pt, -90.0pt)node[fill=blue!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](7){$\textbf{I}^{icu}$};
-     \draw (300.0pt, -50.0pt)node[fill=orange!50, thick, minimum height=0.6cm,minimum width=0.6cm, draw](8){$\textbf{R}$};
-     \draw (300.0pt, -90.0pt)node[fill=purple!50, thick, minimum height=0.6cm,minimum width=0.6cm, draw](9){$\textbf{D}$};
-
-     %\path[->] (6) edge[dotted, bend right = 50]
-     \path[->] (0) edge[] node[above]{$\boldsymbol{\lambda}(t)$} (1);
-     \path[->] (1) edge[] node[above]{$\gamma$} (2);
-     \path[->] (2) edge[] node[below]{$\textbf{p}\theta$} (3);
-     \path[->] (2) edge[] node[near end, below left]{$(1-\textbf{p})\theta$} (4);
-     \path[->] (4) edge[] node[left]{$\boldsymbol{\psi}$} (5);
-     \path[->] (5) edge[] node[near end, below left]{$(1-\phi_{1})\boldsymbol{\omega}$} (7);
-     \path[->] (5) edge[] node[above]{$\phi_{1}\boldsymbol{\omega}$} (6);
-     \path[->] (7) edge[] node[near end, below right]{$\boldsymbol{\delta}_{4}$} (8);
-     \path[->] (7) edge[] node[above]{$\boldsymbol{\tau}_{2}$} (9);
-     \path[->] (6) edge[] node[above]{$\boldsymbol{\delta}_{3}$} (8);
-     \path[->] (6) edge[] node[near start, below left]{$\boldsymbol{\tau}_{1}$} (9);
-     \path[->] (3) edge[bend left = 30] node[above right]{$\delta_{1}$} (8);
-     \path[->] (4) edge[bend left = 30] node[above left]{$\boldsymbol{\delta}_{2}$} (8);
-      \path[->] (5) edge[] node[below]{$\textbf{I}^{sev} - \textbf{H}^{new}$} (10);
-    \end{tikzpicture}
-
-    \caption{Schematic diagram of the compartmental model for SARS-CoV-2 presented by \citet{abrams2021modelling} which is used to derive the MOMDP.}
-    \label{fig:compartment_model}
-\end{figure}
-In order to apply multi-objective reinforcement learning, we construct the \momdpname\ MOMDP based on the epidemiological model introduced in Sec.~\ref{sec:sars-cov2-momdp_model} and graphically depicted in Fig.~\ref{fig:compartment_model}.
-Moreover, we consider a finite-horizon setting where we simulate the compartment model for a fixed number of weeks.
-
-\paragraph{Action-space:} Our actions concern the installment of a social contact matrix with a particular reduction resulting from the reduction function $\hat{C}$ (see Sec.~\ref{sec:interventions}).
-To this end, we use the proportional reduction parameters $p_w, p_s, p_l$ defined in Sec.~\ref{sec:interventions}.
-Thus, each $\mdpaction \in \mdpactionspace$ is a 3-dimensional continuous vector in $[0,1]^3$ (i.e., $\mdpaction = [p_{w}, p_{s}, p_{l}]$) which impacts the social contact matrix according to Eq.~\ref{eq:scm}.
-
-\paragraph{Transition function:} The model defined by \citet{abrams2021modelling} utilises a model transition probability $M$ (see Sec.~\ref{sec:binomial-model} of the Appendix for details on $M$), that progresses the epidemiological model in one timestep based on the currently installed social contact matrix $\hat{C}(p_{w}, p_{s}, p_{l})$.
-We use this function as the transition function in \momdpname.
-
-In a classical MDP, executing an action $\mdpaction_t$ in any state $\mdpstate_t$ leads to a next state $\mdpstate_{t+1}$ according to the transition function $\mdptransition$.
-At every timestep $t$, the agent is free to choose the action to perform.
-In our case, this potentially results in a different restriction $[p_{w}, p_{s}, p_{l}]$ every week.
-However, we argue that in the context of mitigation policies, consistency is important and policies that impose changes too frequently will be hard to adhere to.
-
-In order to learn realistic and consistent mitigation policies, we introduce a \emph{budget} regarding the number of times a policy can change its actions until the terminal state of the MOMDP is reached.
-Concretely, when the action changes, i.e., if the social restriction proposed by the policy is different from the one that is currently in place, we reduce the budget by one.
-We only allow action changes as long as there is budget left.
-We note that we can mimic a no-limit budget setting by choosing a budget that corresponds to the horizon of the environment.
-
-Finally, for each timestep $t$, our transition function $\mdptransition$ uses the model transition probability $M$ to simulate the model for one week, using $\hat{C}_t$ obtained from $\mdpaction_t$.
-
-% As mentioned above and detailed in the SI, the compartment model is described by a set of ODEs.
-To obtain a stochastic version, the ODEs can be formulated as a chain-binomial process.
-Based on these cases, we also create and experiment on two versions of the \momdpname\ environment: the (deterministic) \emph{ODE model} (i.e., a MOMDP with a deterministic transition function) and the (stochastic) \emph{Binomial model} (i.e., a MOMDP with a stochastic transition function).
-
-\paragraph{State-space:} The state of the MOMDP concerns a 3-tuple.
-The first element, $\mdpstate_m$, directly corresponds to the aggregation of the state variables in the epidemiological model, i.e., a tuple,
-\begin{equation}
-    \ltuple S_k, E_k, I_k^{presym}, I_k^{asym}, I_k^{mild}, I_k^{sev}, I_k^{hosp}, I_k^{icu}, H^{new}_k, D_k, R_k \rtuple ,
-\end{equation}
-%
-for each age group $k \in \{1, \ldots, \agegroups\}$, where $S$ encodes the members of the population who are susceptible to infection and $E$ encodes the members of the population who have been exposed to COVID-19.
-Moreover, $I^{presym}$, $I^{asym}$, $I^{mild}$, $I^{hosp}$, $I^{icu}$ denote the members of the population infected with COVID-19 and are, respectively, presymptomatic, asymptomatic, mildly symptomatic, hospitalised, or in the ICU.
-Finally, $H^{new}_k$ represents the number of newly hospitalised individuals in age group $k$.
-
-We parameterize the epidemiological model using the mean of the posteriors as specified by \citet{abrams2021modelling} (details in Sec.~\ref{sec:compartment-model-parameters} of the Appendix).
-
-The second element of the tuple consists of the social contact matrix $\hat{C}_t$ that is currently in place.
-The reason to incorporate it in the state-space is two-fold.
-First, \citet{abrams2021modelling} define a compliance function, simulating the time people need to get used to the new rules set in place.
-As such, during the simulated week, there is a gradual shift from the current $\hat{C}_t$ to the new social contact matrix, $\hat{C}_{t+1}$.
-Thus, we need to include the current $\hat{C}_t$, to maintain a Markovian environment.
-Secondly, we require the current $\hat{C}_t$ to determine whether the action changes the social restrictions in place, and thus consume part of the budget.
-
-The third element of the tuple concerns the budget $\budget$.
-We incorporate a distinct budget per action-dimension, so $p_{w}, p_{s}$ and $p_{l}$ each have their own budget, resulting in a vector $\budget = [b_w, b_s, b_l]$.
-As such, it is possible that, at timestep $t$, the budget for one of the proportional reductions is reduced but not the others.
-
-Therefore, we define a state in \momdpname\ as follows:
-\begin{equation}
-    \mdpstate = \mdpstate_m \cup \hat{C} \cup \budget
-\end{equation}
-
-\paragraph{Reward function:} We define a vectorial reward function which considers multiple objectives: attack rate (i.e., infections and hospitalisations) and the social burden imposed by the interventions on the population.
-
-The attack rate in terms of infections is defined as the difference in susceptibles from the current state to the next state~\cite{libin2020}.
-Since this is a cost that needs to be minimized, we defined the corresponding reward function as the negative attack rate:
-\begin{equation}
-\mdprewardfn_{\text{ARI}}(\mdpstate,\mdpaction,\mdpstate') = -(\sum_{k=1}^{\agegroups}S_k(\mdpstate)-\sum_{k=1}^{\agegroups}S_k(\mdpstate')).
-\end{equation}
-
-The reward function to reduce the attack rate in terms of hospitalisations is defined as the negative number of new hospitalizations:
-\begin{equation}
-\mdprewardfn_{\text{ARH}}(\mdpstate,\mdpaction,\mdpstate') = -\sum_{k=1}^{\agegroups}H^\text{new}_k(\mdpstate).
-\label{eqn:attack-rate-hosp}
-\end{equation}
-
-Finally, we use the missed contacts resulting from the intervention measures as a proxy for societal burden.
-To quantify missed contacts, we consider the original social contact matrix $C$ and the installed social contact matrix $\hat{C}$ to compute the difference $\hat{C} - C$.
-The resulting difference matrix quantifies the average frequency of contacts missed.
-To determine missed contacts for the entire population, we apply the difference matrix to the population sizes of the respective age groups that are currently uninfected (i.e., susceptible and recovered individuals).
-Formally, we define the social burden reward function $\mdprewardfn_{\text{SB}}$, as follows:
-
-\begin{equation}
-    \mdprewardfn_{\text{SB}}(\mdpstate,\mdpaction,\mdpstate') = \sum_{i=1}^{\agegroups}\sum_{j=1}^{\agegroups}(\hat{C}-C)_{ij}S_i(\mdpstate)S_j(\mdpstate) + \sum_{i=1}^{\agegroups}\sum_{j=1}^{\agegroups}(\hat{C}-C)_{ij}R_i(\mdpstate)R_j(\mdpstate),
-\end{equation}
-where $S_k(\mdpstate)$ represents the number of susceptible individuals in age group $k$ in state $\mdpstate$ and $R_k$ represents the number of recovered individuals in age group $k$ in state $\mdpstate$.
-In Sec.~\ref{sec:experiments}, we optimize PCN on two different variants for the multi-objective reward function: $[\mdprewardfn_\text{ARH}, \mdprewardfn_\text{SB}]$ and $[\mdprewardfn_\text{ARI}, \mdprewardfn_\text{SB}]$, to study the impact of these distinct attack rate quantities.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Pareto Conditioned Networks}
-In multi-objective optimization, the set of optimal policies can grow exponentially in the number of objectives.
-Thus, recovering them all is a computationally expensive process and requires an exhaustive exploration of the complete state space.
-To address this problem, we use Pareto Conditioned Networks~(PCN), a method that uses a single neural network to encompass all non-dominated policies \cite{reymond2022pcn}.
-The key idea behind PCN is to use supervised learning techniques to improve the policy instead of resorting to temporal-difference learning.
-% This eliminates the moving-target problem \cite{schmidhuber2019}, resulting in stable learning algorithm.
-
-One major challenge of using neural networks as function approximators in RL is that the target (e.g., the optimal action of the policy) is not known in advance, as opposed to classical supervised learning where the ground-truth target is provided.
-As the behavior of the agent improves over time, the action used as target can change, often leading to hard-to-tune and brittle learners \cite{mnih2015,fu2019}.
-
-Instead of trying to continuously improve the policy by learning actions that should lead to the highest cumulative reward, PCN flips the problem, by learning actions that should lead to any \emph{desired} cumulative reward (be it high or low).
-In this way, all past trajectories can be reused for supervision, since their returns are known, as well as the actions needed to reach said returns.
-We can thus train a policy that, conditioned on a desired return, provides the optimal action to reach said return.
-By leveraging the generalization properties of neural networks, we can accumulate increasingly better experience by conditioning on increasingly higher reward-goals.
-
-Instead of using a neural network that takes a state $\mdpstate$ as input, PCN uses a single neural network that takes a tuple $\ltuple \mdpstate, \hat{h}, \mathbf{\hat{R}} \rtuple$ as input.
-$\mathbf{\hat{R}}$ represents the \emph{desired return} of the decision maker, i.e., the return PCN should reach at the end of the episode.
-$\hat{h}$ denotes the \emph{desired horizon} that expresses the number of timesteps that should be executed before reaching $\mathbf{\hat{R}}$.
-At execution time, both $\hat{h}$ and $\mathbf{\hat{R}}$ are chosen by the decision maker at the start of the episode.
-Then, at every timestep, the desired horizon is updated according to the perceived reward $\momdpreward_t$, $\mathbf{\hat{R}} \leftarrow \mathbf{\hat{R}} - \momdpreward_t$ and the desired horizon is decreased by one, $\hat{h} \leftarrow \hat{h}-1$.
-
-PCN's neural network has an output for each action $\action_i \in \mathcal{A}$.
-Each output represents the confidence the network has that, by taking the corresponding action in $\mdpstate$, $\mathbf{\hat{R}}$ will be reached in $\hat{h}$ timesteps.
-We can draw an analogy with a classification problem where the network should learn to classify $\ltuple \mdpstate, \hat{h}, \mathbf{\hat{R}}\rtuple$ to its corresponding label $\action_i$.
-
-Similar to classification, PCN requires a labeled dataset with training examples to learn a mapping from input to label.
-However, contrary to classification, the data present in the dataset is not fixed.
-PCN collects data from the trajectories experienced while exploring the environment.
-Thus, the dataset improves over time, as we collect increasingly better trajectories.
-
-Concretely, the PCN algorithm implements an iterative process that repeatedly executes 3 main steps:
-\begin{enumerate}
-    \item Use the current policy to interact with the environment and generate trajectories (see Sec.~\ref{sec:pcn-policy-exploration}).
-Initially, a random policy is used to fill the dataset with random trajectories.
-    \item Update the dataset to incorporate these new trajectories and prune the least interesting trajectories in terms of returns (see Sec.~\ref{sec:pcn-building-dataset}).
-    \item Update the policy by training it on the updated dataset (see Sec.~\ref{sec:pcn-continuous}).
-\end{enumerate}
-
-\subsection{Building and updating the dataset}
-\label{sec:pcn-building-dataset}
-
-Given a trajectory composed of $T$ transitions $\ltuple \mdpstate_0, \action_0, \bm{r}_0, \mdpstate_1\rtuple, \dots, \ltuple \mdpstate_{T-1}, \action_{T-1}, \bm{r}_{T-1}, \mdpstate_T\rtuple$ we can compute, for each transition at timestep $t$, the future discounted return $\bm{R}_t = \sum_{i=t}^T \gamma^i\bm{r}_i$ and the leftover horizon $h_t = T-t$.
-Since for this trajectory executing action $a_t$ in state $s_t$ resulted in return $\mathbf{R}_t$ in $h_t$ timesteps, we add a datapoint with input $\langle s, \hat{h}, \mathbf{\hat{R}}\rangle = \langle s_t, h_t, \mathbf{R}_t \rangle$ and output $a=a_t$ to the dataset.
-In other words, when the observed return corresponds to the desired return in that state, then $a_t$ is the optimal action to take.
-
-So that our neural network is trained on relevant transitions only, we keep a fix-sized dataset.
-Thus, adding a trajectory means we need to remove another one.
-Our aim is to keep trajectories that span different parts of the objective space, such that we can explore as many types of trade-offs as possible.
-However, to improve the trade-offs we currently have in our dataset, we need to only keep solutions that are close to our current estimate of the Pareto front.
-Thus, we assign a priority for each trajectory in our dataset that combines two metrics: the \emph{crowding distance}~\cite{deb2000}, which measures the distance of a solution with its closest neighbors, and the \emph{$\ell^2$-norm} with the closest non-dominated solution, which indicates how close the solution is from our current estimate of the Pareto front.
-
-The dataset is then updated by only keeping the trajectories with the top-$N$ priorities, where $N$ is the size of the dataset.
-
-\subsection{Generating trajectories with the current policy}
-\label{sec:pcn-policy-exploration}
-
-In the previous section, we explain how PCN updates the dataset given a trajectory. In this section, we explain how PCN produces said trajectory.
-
-It is unrealistic to expect PCN to reliably produce trajectories with high-valued desired returns when it has only been trained on datapoints originating from random trajectories.
-Rather, we expect PCN to produce trajectories with returns in the range of the ones from the current training data.
-Therefore, if we obtain trajectories reaching high returns, PCN will be able to confidently return high-return policies.
-
-PCN leverages the fact that, due to the generalization capabilities of neural networks, the policies obtained from the network will still be reliable even if the desired return is marginally higher than what is present in the training data.
-In fact, they will perform similar actions to those in the training data, but lead to a higher return.
-Thus, we incrementally condition the network on increasingly higher returns, to obtain trajectories that extend the boundaries of PCN's current coverage set.
-
-More precisely, we randomly select a non-dominated return $\mathbf{R}_{nd}$ and its corresponding horizon $\hat{h}$ from the dataset.
-By randomly picking a non-dominated return from the entire coverage set we ensure equal probability to improve each part of the objective space.
-To generate trajectories that improve upon $\mathbf{R}_{nd}$, we randomly select an objective $o$ and increase $R_{nd,o}$ by a sample from the uniform distribution $\mathcal{U}(0, \sigma_o)$, where $\sigma_o$ is the standard deviation for the selected objective, using all non-dominated returns from the trajectories in the dataset.
-This then becomes our desired return $\bm{\hat{R}}$.
-By restricting the improvement to at most $\sigma_o$, $\mathbf{\hat{R}}$ stays in the range of achievable returns.
-By only modifying one objective at a time, the changes to the network's input compared to the training data are kept at a minimum.
-
-\subsection{Training the network for continuous actions}
-\label{sec:pcn-continuous}
-
-PCN trains the network as a classification problem, where each class represents a different action.
-Transitions $x= \langle \mdpstate_t, h_t, \momdpreturn_t \rangle, y=\action_t$ are sampled from the dataset, and the ground-truth output $y$ is compared with the predicted output $\hat{y}=\pi(\mdpstate_t, h_t, \momdpreturn_t)$.
-The predictor (i.e., the policy) is then updated using the cross-entropy loss function \cite{shore1980axiomatic}:
-\begin{equation}
-    H = -\sum_{\action \in \mathcal{A}}{y_{\action} \log \pi(\action|\mdpstate_t,h_t,\momdpreturn_t)}
-\end{equation}
-where $y_\action = 1$ if $\action = \action_t$ and $y_\action = 0$ otherwise.
-
-While the original PCN algorithm is designed for MOMDPs with discrete actions-spaces, the problem we tackle (see Sec.~\ref{sec:sars-cov2-momdp}) is defined in terms of a continuous action-space.
-We thus extend PCN to the continuous action-space setting.
-We change the output of the neural network such that there is a single output value for each dimension of the action-space.
-Since the actions should be bound in the domain of possible actions ($[0,1]$ in the case of \momdpname, see Sec.~\ref{sec:sars-cov2-momdp}), we apply a sigmoid non-linearity function on each output, as the output of the sigmoid function is bound in $[0,1]$.
-As such, we have a regression problem instead of a classification problem, as the labeled dataset now uses continuous labels $y = \mdpaction_t$ instead of categories.
-We thus use a Mean Squared Error (MSE) loss to update the policy:
-
-\begin{equation}
-    MSE = \frac{1}{|\mathcal{A}|}\sum_{a \in \mathcal{A}}{(\hat{y}_a - y_a)^2}
-\end{equation}
-
-Learning the full set of Pareto-efficient policies $\Pi^*$ requires that the policies $\pi^* \in \Pi^*$ are deterministic stationary policies.
-If stochastic policies are permitted, the set of optimal policies corresponds to the convex part of the Pareto front~\cite{roijers2013survey}.
-However, we argue that, in the context of mitigation policies, deterministic policies are required, as the population needs to be informed in advance of imposed measures.
-Thus, we use the output $\mathbf{\hat{y}}$, which is deterministic, as action at execution time.
-However, PCN improves its policy through exploration, by continuously updating its dataset with better trajectories.
-Thus, at training time, we use a stochastic policy by adding random noise to the action \cite{lillicrap2015continuous}:
-\begin{equation}
-    \mdpaction_t = \pi(\mdpstate_t, h_t, \momdpreturn_t) + \eta s \text { with } s \sim \mathcal{N},
-\end{equation}
-
-where $\mathcal{N}$ is the standard Normal distribution and $\eta$ is a hyper-parameter defining the magnitude of noise to be added.
-
-\subsection{Coping with stochastic transitions}
-\label{sec:pcn-stochastic}
-
-PCN trains its policy on a dataset that is collected by executing trajectories.
-It assumes that reenacting a transition from the dataset leads to the same episodic return.
-When the transition function $\mathcal{T}$ of the MOMDP is deterministic, the whole trajectory can be faithfully reenacted, which guarantees the same return.
-Combined with the fact that PCN's policy is deterministic at execution time, conditioning the policy on a target episodic return is equivalent to conditioning it on the V-value $\mathbf{V}$.
-
-However, when $\mdptransition$ is stochastic this can no longer be guaranteed.
-To mitigate this, we add a limited amount of random noise to $\mathbf{R}_t$ when performing gradient descent, which reduces the risk of overfitting~\cite{zur2009noise}.
-Moreover, while the \momdpname\ model is stochastic, the variation is entirely due to the sampling of the binomial distributions in the binomial-chain.
-While this variation accumulates over time, the time window we consider for each timestep (i.e., one week) is short enough that the accumulation remains bounded.
-Thus, the possible next-states resulting from a state-action pair are similar to each other.
-This allows PCN to compensate if $\momdpreward_t = \momdprewardfn(\mdpstate_t, \mdpaction_t, \mdpstate_{t+1})$ is worse than expected.
-
-Although we use a stochastic model to cope with the uncertainty of the outcome of the outbreak, it is possible to deterministically evaluate the set of ordinary differential equations that define the model.
-We assess the validity of our approach by executing PCN on this deterministic variant, and observing similar performance as with the \momdpname\ model.
-We show the results in Sec.~\ref{sec:cs-ode-vs-binomial} of the Appendix.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Analysing COVID-19 deconfinement policies}
-\label{sec:experiments}
-Our goal is to use PCN to learn deconfinment strategies in the \momdpname\ environment.
-We aim to learn policies that balance between the epidemiological objective of minimising the attack rate (i.e., $\mdprewardfn_\text{ARH}$ for hospitalisation and $\mdprewardfn_\text{ARI}$ for infection) and the social burden (i.e., $\mdprewardfn_\text{SB}$) experienced by the population population due to the implement mitigation measures.
-To this end, we consider two cases for the vectorial reward functions $[\mdprewardfn_\text{ARH}, \mdprewardfn_\text{SB}]$ and $[\mdprewardfn_\text{ARI}, \mdprewardfn_\text{SB}]$, to learn and analyse policies under different targets with respect to the considered attack rate.
-
-To conduct this analysis, we apply our extension of PCN for continuous action-spaces on the \momdpname\ model.
-As explained in Sec.~\ref{sec:pcn-stochastic}, we extend PCN for environments with stochastic transitions.
-
-Conform to \citet{abrams2021modelling}, the simulation starts on the 1st of March 2020, by seeding a number of infections in the population.
-Two weeks later, on the 14th of March, the Belgian government initiated a full lockdown.
-This is implemented by fixing the actions $p_w, p_s, p_l$ to $0.2, 0, 0.1$ respectively.
-This lockdown ended on the 4th of May 2020, at which point the government decided on a multi-phase exit strategy to incrementally reduce teleworking, reopen schools and allow leisure activities, such as the re-opening of bars and the cultural sector.
-It is from this day onward that PCN aims to learn policies for diverse exit strategies, compromising between the total number of daily hospitalizations and the total number of contacts lost as a proxy for social burden.
-The simulation lasts throughout the summer school holidays, from 01/07/2020 to 31/08/2020.
-Schools are closed during the school holidays, which is simulated by setting $p_s = 0$, regardless of the corresponding value outputted by the policy, i.e., during periods of school closure $p_s$ is ignored.
-
-% https://nl.wikipedia.org/wiki/Maatregelen_tijdens_de_coronacrisis_in_Belgi%C3%AB
-We draw the analogy with the multi-phase exit strategy established by the Belgian government and the restriction on the number of action-changes imposed by the \momdpname's budget $\budget$.
-Indeed, on the 11th of May 2020, exactly one week after the end of the lockdown, stores and certain companies were allowed to reopen, under strict conditions.
-This corresponds to altering $p_w$ in our MOMDP.
-One week later, on 18th of May, primary and secondary schools reopened for limited sized class-groups, and the cultural sector reopened partially.
-This is equivalent to increasing $p_s$ and $p_l$.
-Further changes of restrictions occured on the 8th of June, 1st, 9th and 25th of July.
-Thus, we argue that, with a limited budget, we can achieve realistic policies.
-In our experiments, we consider budgets of 2 to 5, as these closely relate to the number of changes that occured until the end of the summer holidays of 2020.
-As an upper bound, we also consider a no-limit budget setting.
-
-To evaluate the quality of the policies learned by PCN, we compare PCN to a baseline.
-This baseline consists of a set of 100 fixed policies, that iterate over all the possible social restriction levels, with values ranging between 0 and 1.
-Concretely, each policy uses a fixed proportional reduction $p_w=p_l=p_s=u, u\sim\mathcal{U}(0, 1)$ throughout the episode.
-In other words, the fixed policies directly operate in a fine-grained manner on the whole contact reduction function $\hat{C}$.
-This allows us to obtain a strong baseline for potential exit strategies over the objective space.
-We note that while such fixed policies are a feasible approach, they do not scale well in terms of action and objective spaces and they will not be able to provide an adaptive restriction level, which is what we aim to provide using PCN.
-
-All experiments are averaged over 10 runs.
-The hyper-parameters and the neural network architecture can be found in Sec.s~\ref{sec:pcn-hyperparameters} and Sec.~\ref{sec:pcn-architecture} of the Appendix, respectively.
-
-\subsection{Learned coverage set}
-
-\begin{figure}
-    \centering
-    \input{fig/arh_binomial_both}
-    \caption{The Pareto front of policies discovered by PCN using \momdpname, showing the different compromises between the number of hospitalizations and the number of lost contacts.
-On the right, we show, for each budget setting (colored, subscript indicates budget) the coverage set learned by the best performing run.
-On the left, we show an interpolated average of the coverage sets learned by the different runs, with the shaded regions corresponding to the standard deviation.
-For comparison, the basline is displayed on both plots (in black).
-As the budget increases, so does the size of the coverage set learnt by PCN.
-Changes are most noticeable in the less restrictive trade-offs in terms of social burden.}
-    \label{fig:pf-binomial-arh-budgets}
-\end{figure}
-
-\begin{table}[t]
-    \centering
-    \setlength{\tabcolsep}{0.5em} % for the horizontal padding
-    {\renewcommand{\arraystretch}{1.2}% for the vertical padding
-    \input{tab/arh/binomial}
-    }
-    \caption{Evaluation metrics for the coverage sets comparing hospitalizations with social burden.
-In general, an increase of budget results in a better coverage set.
-Training on infections (ARI) still provides a competitive coverage set in terms of hospitalizations.
-All PCN coverage sets outperform the baseline.}
-    \label{tab:coverage-set}
-  \end{table}
-
-We learn a coverage set (see Fig.~\ref{fig:pf-binomial-arh-budgets}) that ranges from imposing minimal restrictions to enforcing many restrictions.
-In Fig.~\ref{fig:pf-binomial-arh-budgets}, we display on the right the coverage set of the best-performing run in terms of hypervolume, for each budget setting.
-On the left, we show an interpolated average of the coverage sets learned by the different runs.
-
-Regardless of the imposed budget, we notice that the coverage sets discovered by PCN almost completely dominate the coverage set of the baseline, demonstrating that there are better alternatives to the fixed policies.
-This is most evident in the compromising policies, where one has to carefully choose when to remove social restrictions while at the same time minimizing the impact on daily new hospitalizations.
-In these scenarios, PCN learns policies that drastically reduce the total number of new hospitalizations (e.g., more than $20000$) for the same social burden.
-We analyse the executions of such policies in Fig.~\ref{fig:policy-execution} (middle plot), that shows a flattened hospitalization curve, with a gradual increase of social freedom during the school holidays such that the curve of the epidemic is flattened and gradually decreases over time.
-
-Interestingly, we notice that the most restrictive policy (i.e., the one that prioritizes hospitalizations over social burden, see Fig.~\ref{fig:policy-execution}, bottom plot) still starts to gradually increase $p_w$ and $p_l$ from the end of July onward.
-This is because by then, the epidemic has mostly faded out, and it is safe to reduce social restrictions.
-The timing of this reduction is important as reducing restrictions too soon can lead to a new wave.
-PCN learns the impact of its decisions over time, and correctly infers the timing at which restrictions can be safely lifted.
-
-Finally, the top plot shows that, without imposing social restrictions, the number of hospitalisations peaks on the 15th of June.
-By the beginning of July, the epidemic has spread out over the majority of the population \todo{epi friends: can we mention here that we reached herd immunity?}, and the number of admissions at the hospital has been reduced to a fraction of the number of hospitalisations at the peak.
-Thus, without social restrictions, we do not take advantage of the natural decrease of social contacts due to school holidays, as a significant proportion of the population has already been infected before the start of the holidays.
-
-\subsection{Impact of budgets on the coverage set}
-
-\begin{figure}
-    \centering
-    \input{fig/budget_5}
-    \caption{Selection of policies learned by $\text{PCN}_{b=5}$, from most restrictive in terms of social burden (top) to least restrictive (bottom).
-The x-axis represents the time, starting from the end of the lockdown, on the 4th of May, until the end of the school holidays, on the 1st of September.
-Since the lockdown is simulated before the start of the exit strategy, the start-state differs for each episode (i.e., the hospital already contains infected individuals).
-The left y-axis represents the number of individuals affected by the epidemic.
-The full-lined plots represent the number of individuals admitted into the hospital, ICU and deceased between the last timestep and the current one.
-The plot showing the newly deceased individuals closely relates to the ICU admissions.
-The right y-axis represents the proportional reduction in effect, with 1 meaning a business-as-usual policy, and 0 meaning a complete suppression of social contacts.
-The dotted-lined plots represent the proportional reductions for the work, leasure and school environments.
-We note that the school reduction automatically goes to 0 at the start of the school holidays.
-    }
-    \label{fig:policy-execution}
-\end{figure}
-
-Fig.~\ref{fig:pf-binomial-arh-budgets} demonstrates that the budget impacts the learnt coverage set.
-In general, an increase of budget is associated with an increasingly better coverage set, as policies learned using a higher budget dominate the ones learned with a lower budget.
-This is to be expected, as a higher budget gives the agent more freedom to change its actions as the epidemic progresses.
-%However, we also notice that this increase of coverage sets is not necessarily gradual.
-% Between the run with a budget of 2 and budget of 3, the difference in coverage set seems marginal.
-\todo{insight}.
-
-Moreover, we observe that the difference is concentrated around the less restrictive policies in terms of social burden.
-We postulate that this region contains the most complex policies, as these try to maintain as much social freedom as possible, while containing the number of hospitalisations.
-In these cases, the timing of the actions coincide with the timing and duration of the peak of the epidemic, and a higher budget allows for more fine-grained control to manage this timing.
-To confirm this, we select, for each budget setting, the solution where the difference in performance is most noticeable, corresponding to the solutions with a total number of hospitalisations around $80000$ (which is in the middle of the range of possible hospitalisations, as can be seen in Fig.~\ref{fig:pf-binomial-arh-budgets}).
-We plot the execution of the corresponding policies in Fig.~\ref{fig:pe-binomial-arh-budgets} and analyse their impact in terms of social burden.
-First, we observe that the lower-budget policies are unable to reduce the social restrictions past the peak of the epidemic.
-In contrast, the setting with no budget restrictions meticulously controls the restrictions as the epidemic progresses, completely removing restrictions by the end of the wave.
-Second, we note that the policy with a budget of 5 resembles the execution of the one without restrictions.
-However, due to its budget, the policy is unable to progressively reduce the restrictions and instead resorts to a halfway compromise.
-Compared to this specific region of the coverage set, the difference in performance between different budget settings seems marginal around the extremas.
-At the extremas, the policies are less complex (e.g., business-as-usual, resulting in the same action executed throughout the episode) and are thus less impacted by the budget restrictions.
-
-Finally, we observe that, while the extremas deliver similar trade-offs for any of the chosen budgets, these trade-offs differ for the setting without budget restrictions.
-Indeed, in this setting, PCN does not learn the most extreme policies with respect to restrictions, even though there are no constraints on the action-set.
-As explained in Sec.~\ref{sec:pcn-policy-exploration}, PCN searches for increasingly better solutions using a stochastic policy.
-Thus, at every timestep, the action can change compared to the previous one.
-Continuously outputting the same action (e.g., no social restrictions) becomes a complicated task.
-In comparison, for the settings with a limited budget, the action stays the same as the previous one once the budget has been spent.
-As such, it is easier to learn the most extreme policies.
-Thus, in the specific case where we have an unlimited budget, the freedom of action actually hinders PCN's search, for certain regions of the reward-space.
-
-\begin{figure}
-    \centering
-    \input{fig/all_budgets}
-    \caption{Execution of the policies attaining a number of hospitalisations around $80000$, for different budgets.
-From top to bottom we display the policy executions with budget 2,3,4,5 and no-limit, respectively.
-We notice that the lower-budget policies are unable to reduce the social restrictions past the peak.
-The setting without budget restrictions finely controls the restrictions as the epidemic progresses, completely removing restrictions by the end of the wave.
-Finally, there is no consensus on which social environment to restrict most: certain policies provide similar restrictions for $p_w$ and $p_l$, while others impose harsher restrictions on one social environment than the other.
-    }
-    \label{fig:pe-binomial-arh-budgets}
-\end{figure}
-
-\subsection{$R_\text{ARH}$ versus on $R_\text{ARI}$}
-\label{sec:arh-vs-ari}
-
-Next, we assess the difference in coverage sets when optimizing on $\mdprewardfn_\text{ARH}$ versus $\mdprewardfn_\text{ARI}$.
-Although these reward functions have a different scale (there are more infected persons than hospitalised ones), our experiments show that infections and hospitalizations are tightly correlated.
-This is expected, as during the initial phase of the epidemic, limited immunity was present in the population (i.e., limited natural immunity and no vaccines), which induces a tight coupling between infection and hospitalisation cases.
-This is confirmed in Table~\ref{tab:coverage-set}.
-In this table, we show the different performance metrics (hypervolume, $I_\varepsilon$, $I_{\varepsilon-mean}$) with respect to the objectives $[\mdprewardfn_\text{ARH}, \mdprewardfn_\text{SB}]$.
-The table is split in two parts.
-The left-side shows the different performance metrics, for PCN using $[\mdprewardfn_\text{ARH}, \mdprewardfn_\text{SB}]$ as optimization criterions.
-The right-side shows the same performance metrics, but with PCN using $[\mdprewardfn_\text{ARI}, \mdprewardfn_\text{SB}]$ as optimization criterions.
-
-Even with the $\mdprewardfn_\text{ARI}$, the increased budget shows an increase in hypervolume in terms of hospitalisations.
-Moreover, those hypervolumes are close to the ones trained on $\mdprewardfn_\text{ARH}$, indicating that their coverage sets are similar.
-However, regardless of the imposed budget, the hypervolumes are slightly worse.
-This is to be expected, since those experiments are not directly optimised on $\mdprewardfn_\text{ARH}$.
-We draw a similar conclusion for $I_\varepsilon$: for budgets 2, 3 and 5, the difference between the worst-performing policy for the $\mdprewardfn_\text{ARI}$ variant and the $\mdprewardfn_\text{ARH}$ is less than $0.01$, indicating less than $1\%$ difference in return values between the two variants when comparing their worst-performing policy.
-As an exception, we notice that PCN without budget restrictions results in better performance across every metric for the $\mdprewardfn_\text{ARI}$ variant.
-Still, due to the high standard deviation of the unlimited budget, $\mdprewardfn_\text{ARH}$ setting, we do not believe this difference is meaningful.
-Thus, we could optimise on the attack rate of hospitalisations with $\mdprewardfn_\text{ARI}$.
-As there is a 2 week delay for hospitalisations, this would facilitate learning policies to react to unexpected changes earlier than using $\mdprewardfn_\text{ARH}$, given that a good proxy to the actual number of infections was available (e.g., due to a scale up of PCR testing, as was the case after the first lockdown).
-
-\subsection{Robustness of policy executions}
-\label{sec:pcn-robustness}
-
-\begin{table}
-    \centering
-    \setlength{\tabcolsep}{0.5em} % for the horizontal padding
-    {\renewcommand{\arraystretch}{1.2}% for the vertical padding
-    \input{tab/arh/arh_binomial_robustness}
-    }
-    \caption{Comparing the difference in the desired return provided to PCN and the actual return PCN obtained when executing its policy.
-We see that, regardless of the setting, the learned policy faithfully receives a return similar to its desired return.}
-    \label{tab:pcn-robustness}
-\end{table}
-
-The dataset of trajectories that PCN is trained on is pruned over time to keep only the most relevant trajectories.
-The returns of these trajectories are used in Fig.~\ref{fig:pf-binomial-arh-budgets} to visualize the learned coverage set.
-Each of these returns can be used as desired return for policy execution.
-We now assess the robustness of the executed policies, by comparing the return obtained after executing the policy with the corresponding target return.
-For each run, we execute the policy $10$ times and compute the $I_\varepsilon$ and $I_{\varepsilon-mean}$ metrics with respect to the coverage set learned during the run.
-We show that the executed policies reliably obtain returns that are similar to the desired return used to condition PCN.
-
-Results are shown in Table~\ref{tab:pcn-robustness}.
-The $I_\varepsilon$ indicators shows that, regardless of the budget, the decision maker will lose at worst a $0.050$ normalized return in any of the objectives.
-On average, it will lose $0.010$ normalized returns, i.e., on average, the return obtained by executing a policy will either result in an additional $1441$ hospitalisations than expected, or result in $12$ additional social contacts lost.
-Moreover, we emphasize that the learned coverage set contains the non-dominated returns encountered over the whole training procedure.
-Since the \momdpname\ model is stochastic, for multiple executions of the same policy, the executions kept in the coverage sets are the ones for which the samples from the binomial-chain resulted in a better progression of the epidemic than average.
-Thus, we expect our policy-executions to be close to the target selected from the coverage set, but not exactly on target.
-Based on this analysis, we conclude that the policies trained by PCN are robust and produce returns as close as possible from their chosen target.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Related work}
-Reinforcement learning (RL) has been used in conjunction with epidemiological models to learn policies to limit the spread of diseases and predict the effects of possible mitigation strategies \cite{probert2019context,libin2018bayesian,libin2020}.
-
-RL and Deep RL have been used extensively as a decision making aid to reduce the spread of COVID-19.
-For example, to learn effective mitigation strategies \cite{ohi2020exploring}, to learn efficacy of lockdown and travel restrictions \cite{kwak2021covid} and to limit the influx of asymptomatic travellers \cite{bastani2021efficient}.
-
-Multi-objective methods have also been deployed to learn optimal strategies to mitigate the spread of COVID-19.
-Wan et al.
-\cite{wan2021multi} implement a model-based multi-objective policy search method and demonstrate their method on COVID-19 data from China.
-Given that this method is model-based, a model of the transition function must be learned by sampling from the environment.
-The method proposed by Wan et al.
-\cite{wan2021multi} only considers a discrete action space which limits the application of their algorithm.
-Wan et al.
-\cite{wan2021multi} use linear weights to compute a set of Pareto optimal policies.
-However, methods which use linear weights can only learn policies on the convex-hull of the Pareto front \cite{vamplew2008limitations}, therefore the full Pareto front cannot be learned.
-We note that the method proposed by Kompella et al.
-\cite{kompella2020reinforcement} considers multiple objectives.
-However, the objectives are combined using a weighted sum with hand-tuned weights which are determined by the authors.
-The weighted sum is applied by the reward function and a single objective RL method is used to learn a single optimal policy.
-In contrast to previous work, our approach makes no assumptions regarding the scalarisation function of the user and is able to discover Pareto fronts of arbitrary shape.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusion and discussion}
-Making decisions on how to mitigate epidemics has important ethical implications with respect to public health and societal burden.
-In this regard, it is crucial to approach this decision making from a balanced perspective, to which end we argue that multi-objective decision making is essential.
-In this work, we establish a novel approach, i.e., an expert system, to study multi-faceted policies, and this approach shows great potential to study future epidemic mitigation policies.
-We are aware of the ethical implications that expert systems have on the decision process and we make the disclaimer that all results based on the expert system that we propose should be carefully interpreted by experts in the field of public health, and in a broader context that encompasses health economics, well-being and education \todo{epi friends, other things to add here?
-To make sure we are exhaustive}.
-We note that the work in this manuscript was conducted by a interdisciplinary consortium that includes computer scientists and scientists with a background public health, epidemiology and bio-statistics.
-
-In this work, we focus on the clinical outcomes of intervention strategies and use the reduced contacts as proxy for social burden.
-This could be extended into more formal health economic evaluations, by designing reward functions that explicitely consider distinct health economic principles.
-The COVID-19 pandemic demonstrates the broad impact of infectious diseases on sectors other than health care.
-This stresses the need to capture a societal and thus multi-objective perspective in the decision making process on public health and health care interventions.
-Our learned policies confirm this, showing that focusing solely on reducing the number of hospitalizations results in taking drastic measures -- more than a thousand social interactions lost per person over the span of 4 months -- that may have a long-lasting impact on the population.
-
-Although we use an age-structured compartment model, with social contact matrices to model social interactions, this remains a model that evaluates the progression of the pandemic as an aggregated process over the population.
-Individual-based models could provide more detailed and localized policies, potentially further improving the quality of possible trade-offs, and would provide an interesting avenue for future work.
-However, due to the computational cost of simulating such models, and the number of interactions required by reinforcement learning in general, this remains a challenging problem, that will require fundamental research to improve the sample efficiency of  multi-objective reinforcement learning algorithms.
-
-While we are able to interpret and analyse the obtained policies and their corresponding trade-offs, as we can plot the Pareto front for two objectives, this approach cannot be used for problems with more objectives, which will be necessary to cover reward functions that cover distinct health economic principles.
-To facilitate this kind of research, new algorithms are necessary to enable reinforcement learning in many-objective contexts and to interprete the learnt policies.
-
-In this work, PCN is able to cope with model stochasticity, as the stochasticity is limited.
-In settings where the stochasticity is more pronounced, e.g., when designing policies to control the initial outbreak of an epidemic, further methodological extensions are warranted.
-
-Finally, in this work we studied policies that aim to balance social burden and hospitalizations.
-Yet, the methodology that we propose shows promise to address a wide variety of public health challenges, such as balancing the number of lost schooldays with respect to the attack rate of infections in schools \cite{torneri2021controlling}, contact tracing effort compared to the impact of such policies \cite{willem2021impact}, the impact of antivirals on the epidemic while balancing the likelihood for resistance mutations to emerge \cite{torneri2020prospect}, and to balance the cost of universal testing and its impact on an emerging epidemic \cite{libin2021assessing}.
-
-To conclude, we show that multi-objective reinforcement learning can be used to learn a wide set of high-quality policies on real-world problems, providing the decision maker with insightful and diverse alternatives, and showing the impact of extreme measures.
-Furthermore, we show that action budgets can act as a regulariser that facilitates learning realistic policies that can be easily conveyed to decision makers.
-
-\bibliographystyle{unsrtnat}
-\bibliography{references}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\appendix
-\counterwithin{figure}{section}
-\counterwithin{table}{section}
-\counterwithin{equation}{section}
-
-\section{Stochastic Compartmental Model}
-\label{sec:binomial-model}
-In this work we utilise the compartmental model proposed by \citet{abrams2021modelling} and extend the model to a multi-objective Markov decision process (MOMDP).
-Fig.~\ref{fig:compartment_model} shows a visual depiction of the compartment model.
-The flows of the deterministic model are defined by a set of ordienary differential equations, outlined in Sec.~\ref{sec:sars-cov2-momdp_model}.
-
-Intervening in the spread of the virus by, for example, reducing social contacts or government interventions introduces uncertainty in the further course of the outbreak.
-Therefore, to understand how this uncertainty affect the spread of the disease we introduce a stochastic component model which can model the uncertainty generated by interventions in social contacts.
-
-By formulating the set of differential equations defined in Sec.~\ref{sec:sars-cov2-momdp_model}, as a chain-binomial, we can obtain stochastic trajectories from this model \cite{bailey1975mathematical}.
-A chain-binomial model assumes a stochastic model where infected individuals are generated by some underlying probability distribution.
-For the stochastic model we consider a time interval $(t, t +h]$, where $h$ is defined as the length between two consecutive time points.
-Similar to \citet{abrams2021modelling}, in this work we set $h = \frac{1}{24}$.
-\citet{abrams2021modelling} define the set of differential equations as a chain binomial as follows:
-
-\begin{align*}
-\textbf{S}_{t+h} (k) & = \textbf{S}_{t} (k) - \textbf{E}_{new, t+h}(k),\\
-%
-\textbf{E}_{t+h}(k) & = \textbf{E}_{t}(k) + \textbf{E}_{new, t+h}(k) - \textbf{I}^{presym}_{new, t + h} (k),\\
-%
-\textbf{I}^{presym}_{t+h}(k) & = \textbf{I}^{presym}_{t}(k) + \textbf{I}^{presym}_{new, t + h}(k) - \textbf{I}^{asym}_{new, t + h}(k) - \textbf{I}^{mild}_{new, t + h}(k),\\
-%
-\textbf{I}^{asym}_{t+h}(k) & = \textbf{I}^{asym}_{t}(k) + \textbf{I}^{asym}_{new, t + h}(k) - \textbf{R}^{asym}_{new, t + h}(k),\\
-%
-\textbf{I}^{mild}_{t+h}(k) & = \textbf{I}^{mild}_{t}(k) + \textbf{I}^{mild}_{new, t + h}(k) - \textbf{I}^{sev}_{new, t + h}(k) - \textbf{R}^{mild}_{new, t + h}(k),\\
-%
-\textbf{I}^{sev}_{t+h}(k) & = \textbf{I}^{sev}_{t}(k) + \textbf{I}^{sev}_{new, t + h}(k) - \textbf{I}^{hosp}_{new, t + h}(k) - \textbf{I}^{icu}_{new, t + h}(k),\\
-%
-\textbf{I}^{hosp}_{t+h}(k) & = \textbf{I}^{hosp}_{t}(k) + \textbf{I}^{hosp}_{new, t + h}(k) - \textbf{D}^{hosp}_{new, t + h}(k) - \textbf{R}^{hosp}_{new, t + h}(k),\\
-%
-\textbf{I}^{icu}_{t+h}(k) & = \textbf{I}^{icu}_{t}(k) + \textbf{I}^{icu}_{new, t + h}(k) - \textbf{D}^{icu}_{new, t + h}(k) - \textbf{R}^{icu}_{new, t + h}(k),\\
-%
-\textbf{D}_{t+h}(k) & = \textbf{D}_{t}(k) + \textbf{D}^{hosp}_{new, t + h}(k) + \textbf{D}^{icu}_{new, t + h}(k),\\
-%
-\textbf{R}_{t+h}(k) & = \textbf{R}_{t}(k) + \textbf{R}^{asym}_{new, t + h}(k) + \textbf{R}^{mild}_{new, t + h}(k) + \textbf{R}^{hosp}_{new, t + h}(k) + \textbf{R}^{icu}_{new, t + h}(k)
-\end{align*}
-
-where,
-
-\begin{align*}
-\textbf{E}_{new, t+h} & \sim \textit{Binomial} \left( \textbf{S}_{t} (k), p^{*}_t (k) = 1 - \{ 1 -  p^{*}_t (k)\}^{\textbf{I}_{t}} \right), \\
-%
-p^{*}_t (k) & = 1 - exp \left[ -h \sum^{K}_{k' = 1} \beta_{asym}(k, k')\{ \textbf{I}^{asym}_{t}(k')\} + \beta_{sym}(k, k') \{\textbf{I}^{mild}_{t}(k')+\textbf{I}^{sev}_{t}(k')\} \right],\\
-%
-\textbf{I}^{presym}_{new, t+h} (k)&  \sim Binomial\left( \textbf{I}^{presym}_{t}(k), 1 - exp(-hp(k)\theta) \right),\\
-%
-\textbf{I}^{mild}_{new, t+h} (k) & \sim Binomial\left( \textbf{I}^{presym}_{t}(k), 1 - exp\left[-h \{1-p(k)\}\theta \right]) \right),\\
-%
-\textbf{I}^{sev}_{new, t+h} (k) & \sim Binomial\left( \textbf{I}^{mild}_{t}(k), 1 - exp\{-h\psi(k)\}) \right),\\
-%
-\textbf{I}^{hosp}_{new, t+h} (k) & \sim Binomial\left( \textbf{I}^{sev}_{t}(k), 1 - exp\{-h\phi_{1}(k)\omega(k)\}) \right),\\
-%
-\textbf{I}^{icu}_{new, t+h} (k) & \sim Binomial\left( \textbf{I}^{sev}_{t}(k), 1 - exp\left[-h \{1-\phi_{1}(k)\}\omega(k) \right] \right),\\
-%
-\textbf{D}^{hosp}_{new, t+h} (k) & \sim Binomial\left( \textbf{I}^{hosp}_{t}(k), 1 - exp\{-h\tau_{1}(k) \} \right),\\
-%
-\textbf{D}^{icu}_{new, t+h} (k) & \sim Binomial\left( \textbf{I}^{icu}_{t}(k), 1 - exp\{-h\tau_{2}(k) \} \right),\\
-%
-\textbf{R}^{asym}_{new, t+h} (k) & \sim Binomial\left( \textbf{I}^{asym}_{t}(k), 1 - exp\left(-h\delta_{2}(k) \right) \right), \\
-%
-\textbf{R}^{hosp}_{new, t+h} (k) & \sim Binomial\left( \textbf{I}^{hosp}_{t}(k), 1 - exp\{-h\delta_{3}(k) \} \right),\\
-%
-\textbf{R}^{icu}_{new, t+h} (k) & \sim Binomial\left( \textbf{I}^{icu}_{t}(k), 1 - exp\{-h\delta_{4}(k) \} \right).
-\end{align*}
-
-Given \momdpname\ also calculates new hospitalisations, $\textbf{H}^{new}$, we define $\textbf{H}^{new}$ for the stochastic compartmental model as follows:
-\[
-\textbf{H}^{new}_{t+h}(k) = \textbf{H}^{new}_{t}(k) + \textbf{I}^{hosp}_{new, t+h} (k).
-\]
-For more details on this model and the chain-binomial representation of the differential equations, we refer the reader to the work of \citet{abrams2021modelling}.
-
-To create a version of \momdpname\ with a stochastic transition function, $M$, we utilise the stochastic compartmental model outlined above.
-Given the transitions within the compartmental model are derived by an underlying probability distribution it is possible to utilise the stochastic compartmental model transitions for \momdpname.
-As previously outlined in Sec.~\ref{sec:sars-cov2-momdp_model} the contact matrix $\hat{C}$ applied the model state $s_{m}$ progresses the model and returns a new model state $s'_{m}$.
-Given the underlying model dynamics are governed in a probabilistic manner, the model returns $s'_{m}$ stochastically.
-Therefore, it is possible to use this process as a stochastic transition function, $M$, for \momdpname.
-
-\subsection{A Note on Model Parameters}
-\label{sec:compartment-model-parameters}
-
-The model is parameterised using the mean of the posteriors as reported by \citet{abrams2021modelling}.
-
-The population size for each of the considered age groups was taken from the Belgian statistical agency STATBEL\footnote{\url{https://statbel.fgov.be/nl/themas/bevolking/structuur-van-de-bevolking\#figures}}.
-To initialise the model, we used the number of confirmed cases until 13 March 2020 \cite{abrams2021modelling}, as reported by the Belgian agency for public health Sciensano\footnote{\url{https://epistat.wiv-isp.be/covid/}}.
-
-\subsection{Modelling interventions}
-\label{sec:compliance-function}
-In order to model different types of interventions, we follow \citet{abrams2021modelling}.
-Firstly, to consider distinct exit scenarios, we alter the social contact matrices to reflect a contact reduction in a particular age group.
-Secondly, we assume that compliance to the interventions is gradual and model this using a logistic compliance function.
-We use the logistic compliance function in function of time $t$,
-%
-\begin{equation}
-c(t,t_I) = \frac{\exp(\beta^*_0 + \beta^*_1(t-t_I))}{1+\exp(\beta^*_0 + \beta^*_1(t-t_I))},
-\end{equation}
-%
-where $t_I$ indicates the time the intervention started.
-We initialise $\beta^*_1$ to the value estimated in by Abrams et al.
-and choose $\beta^*_0=-5$, as an intercept to have $c(t)=0$ for $t=0$, in correspondence with Fig.~F2 in the Supplementary Information of \citet{abrams2021modelling}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Additional results}
-
-\subsection{Comparison of coverage sets learned on ODE and Binomial models}
-\label{sec:cs-ode-vs-binomial}
-
-\begin{figure}[t!]
-    \centering
-    \input{fig/arh_binomial_ode}
-    \caption{Comparison of learned coverage sets when PCN interacts with the ODE (i.e., deterministic) variant of the compartment model.
-We observe that, regardless of the chosen budget, PCN learns a similar coverage set on both variants of the model, indicating that it is able to cope with the stochasticity present in the Binomial variant.}
-    \label{fig:pf-binomial-ode}
-\end{figure}
-
-The coverage sets displayed in Fig.~\ref{fig:pf-binomial-arh-budgets} correspond to PCN trained on the \momdpname\ model, which is stochastic.
-To show that PCN copes with the stochasticity of this model, we compare these coverage sets with the ones learned on the ODE model, which is deterministic.
-
-Results are shown in Fig.~\ref{fig:pf-binomial-ode}.
-First, in a similar fashion as Fig.~\ref{fig:pf-binomial-arh-budgets}, we show the interpolated average coverage set for different budget setting, with PCN trained on the ODE model.
-We observe similar trends as for the Binomial model.
-Next, for each budget setting, we compare the interpolated average coverage set of the Binomial setting with the ODE setting.
-We observe similar coverage set, regardless of the budget setting, indicating that PCN is able to cope with stochastic transitions.
-
-\subsection{Comparison of coverage sets learned on $\mathbf{R}_{\text{ARH}}$ and $\mathbf{R}_{\text{ARI}}$}
-\label{sec:cs-arh-vs-ari}
-
-In Sec.~\ref{sec:arh-vs-ari}, we show that we can learn all trade-offs between social burden and hospitalisations, while using the attack rate over infections as reward function.
-
-Results are shown in Fig.~\ref{fig:pf-binomial-arh-ari}.
-We observe that the learned coverage sets are similar, regardless of the budget setting.
-Still, the coverage set of when trained on $\mathbf{R}_{\text{ARI}}$ is systematically dominated by the one when trained on $\mathbf{R}_{\text{ARH}}$.
-\todo{Please check: is this actually true?} While hospitalisations and infections are highly correlated, they differ in terms of age-groups.
-Older age-groups are more susceptible to be hospitalised after being infected, but they do not form the majority of the population.
-For trade-offs where infections and social burden need to be balanced, the proportional reductions target different social environments than for trade-offs balancing hospitalisations and social burden.
-For example, the work environment is majorly comprised of individuals with a more robust immune system, reducing the social contact in this environment greatly affects the number of infections, but has a lesser impact on the number of hospitalisations.
-
-\begin{figure}[t!]
-    \centering
-    \input{fig/arh_binomial_ari}
-    \caption{Comparison of learned coverage sets when PCN learns using the attack rate of infections $\mathbf{R}_{\text{ARI}}$.}
-    \label{fig:pf-binomial-arh-ari}
-\end{figure}
-
-\subsection{Policy executions}
+\section{Policy executions}
 
 Depending on the budget, PCN learns a coverage set containing more than 150 different policies.
-To gain a better insight about their behavior, and how they differ from each other, we plot executions of each learned policy in Fig.s~\ref{fig:policy-executions-b2-0}-\ref{fig:policy-executions-b2-140}.
+To gain a better insight about their behavior, and how they differ from each other, we plot executions of each learned policy in Fig.~\ref{fig:policy-executions-b2-0}-\ref{fig:policy-executions-b2-140}.
 The plots are displayed from the least restrictive policy in terms of social burden to the most restrictive one.
@@
 
-\subsection{Experiment parameters}
-\label{sec:pcn-hyperparameters}
-
-We used the same hyper-parameters across all experiments.
-Each experiment resulted in 10 independent trials.
-Finally, we performed a grid-search over possible hyper-parameter values.
-All hyper-parameters used and their possible values explored during grid-search are displayed in Table~\ref{tab:pcn-hyperparameters}.
-
-\begin{table}[t]
-    \centering
-    \begin{tabular}{|c|c|c|}
-        \hline
-        hyper-parameter & value & grid-search \\
-        \hline
-        learning rate & $0.001$ &  \\
-        total training timesteps & $300000$  & \\
-        batch size & $256$ & $256, 1024$ \\
-        model updates & $50$ & \\
-        episodes between updates & $10$ & \\
-        ER size (in episodes) & $1000$ & $400, 500, 1000$ \\
-        initial random episodes & $200$ & $50, 200$ \\
-        exploration noise & $0.1$ & $0, 0.1, 0.2$ \\
-        desired return noise & $0.05$ & $0, 0.05, 0.1, 0.2$ \\
-        reward scaling & $[10000, 100]$ & \\
-        \hline
-    \end{tabular}
-    \caption{The different hyper-parameters used by our extension of PCN. The right-most column also shows, when applicable, the different values tried during grid-search.}
-    \label{tab:pcn-hyperparameters}
-\end{table}
-
-\begin{table}[h]
-    \centering
-    \input{tab_nn-architecture}
-    \caption{The 4 different neural network architectures explored for our experiments. All the displayed results use the \texttt{dense-big} variant.}
-    \label{tab:nn-architecture}
-\end{table}
-
-\subsection{Neural network architecture}
-\label{sec:pcn-architecture}
-
-Next to the hyper-parameter search, we also performed a grid search over 4 different neural network architectures.
-All the architectures have the same structure.
-We use a compartment embedding $sc_{emb}$, a social contact matrix embedding $sm_{emb}$ and a school-holidays embedding $sh_{emb}$ that take as inputs the compartment, the previous $p_w, p_s, p_l$ values (as they fully define the SCM $\hat{C}$) and a boolean flag for school holidays, respectively.
-All these embeddings have a same-sized embedding of 64, which are multiplied together to form the full state embedding.
-This state-embedding is used as input for another network, $s_{emb}$.
-Additionally, we use a common embedding $c_{emb}$ for the concatenation of the desired return and horizon.
-Finally, the results of $s_{emb}$ and $c_{emb}$ are multiplied together, before passing through a fully connected network $fc$ that has 3 outputs, one for $p_w, p_s, p_l$ respectively.
-
-All the architectures of the different components are displayed in Table~\ref{tab:nn-architecture}. The variant used in all experiments is \texttt{dense-big}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \end{document}

### 2024-05-28 paper/fig/arh_binomial_mones.tex
mones

--- 2023-10-19 paper/fig/arh_binomial_ari.tex
@@
     ]
-    % interpolated plot for ARI
-    \nextgroupplot[mark size=1pt,draw opacity=1,legend entries={{$\text{PCN}_{b=2}$}, {$\text{PCN}_{b=3}$}, {$\text{PCN}_{b=4}$}, {$\text{PCN}_{b=5}$}, {$\text{PCN}$}, {Fixed}},title={$\mdprewardfn_\text{ARI}$, different budgets}]
+    % interpolated plot for MONES
+    \nextgroupplot[mark size=1pt,draw opacity=1,legend entries={{$\text{MONES}_{b=2}$}, {$\text{MONES}_{b=3}$}, {$\text{MONES}_{b=4}$}, {$\text{MONES}_{b=5}$}, {$\text{MONES}$}, {Fixed}},title={MONES algorithm, different budgets}]
     % mean line
@@
         \edef\temp{
-            \noexpand\addplot[fill opacity=0.5,\c] table [x expr=-1*\noexpand\thisrow{mu_o_1}, y expr=-1*\noexpand\thisrow{mu_o_5}, col sep=comma] {data/cs/ari/binomial/interpolated/budget_\b.csv};
+            \noexpand\addplot[fill opacity=0.5,\c] table [x expr=-1*\noexpand\thisrow{mu_o_1}, y expr=-1*\noexpand\thisrow{mu_o_5}, col sep=comma] {data/mones/arh/binomial/interpolated/budget_\b.csv};
         }
@@
         \edef\temp{
-            \noexpand\addplot[name path=upper\b, draw=none] table [x expr=-1*\noexpand\thisrow{mu_o_1}, y expr=-1*\noexpand\thisrow{mu_o_5}+\noexpand\thisrow{std_o_5}, col sep=comma] {data/cs/ari/binomial/interpolated/budget_\b.csv};
-            \noexpand\addplot[name path=lower\b, draw=none] table [x expr=-1*\noexpand\thisrow{mu_o_1}, y expr=-1*\noexpand\thisrow{mu_o_5}-\noexpand\thisrow{std_o_5}, col sep=comma] {data/cs/ari/binomial/interpolated/budget_\b.csv};
+            \noexpand\addplot[name path=upper\b, draw=none] table [x expr=-1*\noexpand\thisrow{mu_o_1}, y expr=-1*\noexpand\thisrow{mu_o_5}+\noexpand\thisrow{std_o_5}, col sep=comma] {data/mones/arh/binomial/interpolated/budget_\b.csv};
+            \noexpand\addplot[name path=lower\b, draw=none] table [x expr=-1*\noexpand\thisrow{mu_o_1}, y expr=-1*\noexpand\thisrow{mu_o_5}-\noexpand\thisrow{std_o_5}, col sep=comma] {data/mones/arh/binomial/interpolated/budget_\b.csv};
             \noexpand\addplot[\c!15] fill between[of={upper\b} and {lower\b}];
@@
         \edef\temp{
-            \noexpand\nextgroupplot[draw opacity=1,legend entries={$R_{\text{ARH}}$, $R_{\text{ARI}}$},title={$\text{PCN}_{b=\b}$}]
+            \noexpand\nextgroupplot[draw opacity=1,legend entries={$\text{PCN}_{b=\b}$, $\text{MONES}_{b=\b}$},title={budget \b}]
             \noexpand\addplot[fill opacity=0.5,pc1] table [x expr=-1*\noexpand\thisrow{mu_o_1}, y expr=-1*\noexpand\thisrow{mu_o_5}, col sep=comma] {data/cs/arh/binomial/interpolated/budget_\b.csv};
-            \noexpand\addplot[fill opacity=0.5,pc2] table [x expr=-1*\noexpand\thisrow{mu_o_1}, y expr=-1*\noexpand\thisrow{mu_o_5}, col sep=comma] {data/cs/ari/binomial/interpolated/budget_\b.csv};
+            \noexpand\addplot[fill opacity=0.5,pc2] table [x expr=-1*\noexpand\thisrow{mu_o_1}, y expr=-1*\noexpand\thisrow{mu_o_5}, col sep=comma] {data/mones/arh/binomial/interpolated/budget_\b.csv};
             % std plots
@@
 
-            \noexpand\addplot[name path=odeupper\b, draw=none] table [x expr=-1*\noexpand\thisrow{mu_o_1}, y expr=-1*\noexpand\thisrow{mu_o_5}+\noexpand\thisrow{std_o_5}, col sep=comma] {data/cs/ari/binomial/interpolated/budget_\b.csv};
-            \noexpand\addplot[name path=odelower\b, draw=none] table [x expr=-1*\noexpand\thisrow{mu_o_1}, y expr=-1*\noexpand\thisrow{mu_o_5}-\noexpand\thisrow{std_o_5}, col sep=comma] {data/cs/ari/binomial/interpolated/budget_\b.csv};
+            \noexpand\addplot[name path=odeupper\b, draw=none] table [x expr=-1*\noexpand\thisrow{mu_o_1}, y expr=-1*\noexpand\thisrow{mu_o_5}+\noexpand\thisrow{std_o_5}, col sep=comma] {data/mones/arh/binomial/interpolated/budget_\b.csv};
+            \noexpand\addplot[name path=odelower\b, draw=none] table [x expr=-1*\noexpand\thisrow{mu_o_1}, y expr=-1*\noexpand\thisrow{mu_o_5}-\noexpand\thisrow{std_o_5}, col sep=comma] {data/mones/arh/binomial/interpolated/budget_\b.csv};
             \noexpand\addplot[pc2!15] fill between[of={odeupper\b} and {odelower\b}];

### 2024-05-28 paper/fig/arh_ode_mones.tex
mones

--- 2024-05-28 paper/fig/arh_binomial_mones.tex
@@
         \edef\temp{
-            \noexpand\addplot[fill opacity=0.5,\c] table [x expr=-1*\noexpand\thisrow{mu_o_1}, y expr=-1*\noexpand\thisrow{mu_o_5}, col sep=comma] {data/mones/arh/binomial/interpolated/budget_\b.csv};
+            \noexpand\addplot[fill opacity=0.5,\c] table [x expr=-1*\noexpand\thisrow{mu_o_1}, y expr=-1*\noexpand\thisrow{mu_o_5}, col sep=comma] {data/mones/arh/ode/interpolated/budget_\b.csv};
         }
@@
         \edef\temp{
-            \noexpand\addplot[name path=upper\b, draw=none] table [x expr=-1*\noexpand\thisrow{mu_o_1}, y expr=-1*\noexpand\thisrow{mu_o_5}+\noexpand\thisrow{std_o_5}, col sep=comma] {data/mones/arh/binomial/interpolated/budget_\b.csv};
-            \noexpand\addplot[name path=lower\b, draw=none] table [x expr=-1*\noexpand\thisrow{mu_o_1}, y expr=-1*\noexpand\thisrow{mu_o_5}-\noexpand\thisrow{std_o_5}, col sep=comma] {data/mones/arh/binomial/interpolated/budget_\b.csv};
+            \noexpand\addplot[name path=upper\b, draw=none] table [x expr=-1*\noexpand\thisrow{mu_o_1}, y expr=-1*\noexpand\thisrow{mu_o_5}+\noexpand\thisrow{std_o_5}, col sep=comma] {data/mones/arh/ode/interpolated/budget_\b.csv};
+            \noexpand\addplot[name path=lower\b, draw=none] table [x expr=-1*\noexpand\thisrow{mu_o_1}, y expr=-1*\noexpand\thisrow{mu_o_5}-\noexpand\thisrow{std_o_5}, col sep=comma] {data/mones/arh/ode/interpolated/budget_\b.csv};
             \noexpand\addplot[\c!15] fill between[of={upper\b} and {lower\b}];
@@
         \edef\temp{
-            \noexpand\nextgroupplot[draw opacity=1,legend entries={$\text{PCN}_{b=\b}$, $\text{MONES}_{b=\b}$},title={budget \b}]
-            \noexpand\addplot[fill opacity=0.5,pc1] table [x expr=-1*\noexpand\thisrow{mu_o_1}, y expr=-1*\noexpand\thisrow{mu_o_5}, col sep=comma] {data/cs/arh/binomial/interpolated/budget_\b.csv};
-            \noexpand\addplot[fill opacity=0.5,pc2] table [x expr=-1*\noexpand\thisrow{mu_o_1}, y expr=-1*\noexpand\thisrow{mu_o_5}, col sep=comma] {data/mones/arh/binomial/interpolated/budget_\b.csv};
+            \noexpand\nextgroupplot[draw opacity=1,legend entries={$\text{PCN}_{b=\b}$, $\text{MONES}_{b=\b}$},title={budget $\b$}]
+            \noexpand\addplot[fill opacity=0.5,pc1] table [x expr=-1*\noexpand\thisrow{mu_o_1}, y expr=-1*\noexpand\thisrow{mu_o_5}, col sep=comma] {data/cs/arh/ode/interpolated/budget_\b.csv};
+            \noexpand\addplot[fill opacity=0.5,pc2] table [x expr=-1*\noexpand\thisrow{mu_o_1}, y expr=-1*\noexpand\thisrow{mu_o_5}, col sep=comma] {data/mones/arh/ode/interpolated/budget_\b.csv};
             % std plots
-            \noexpand\addplot[name path=binomialupper\b, draw=none] table [x expr=-1*\noexpand\thisrow{mu_o_1}, y expr=-1*\noexpand\thisrow{mu_o_5}+\noexpand\thisrow{std_o_5}, col sep=comma] {data/cs/arh/binomial/interpolated/budget_\b.csv};
-            \noexpand\addplot[name path=binomiallower\b, draw=none] table [x expr=-1*\noexpand\thisrow{mu_o_1}, y expr=-1*\noexpand\thisrow{mu_o_5}-\noexpand\thisrow{std_o_5}, col sep=comma] {data/cs/arh/binomial/interpolated/budget_\b.csv};
+            \noexpand\addplot[name path=binomialupper\b, draw=none] table [x expr=-1*\noexpand\thisrow{mu_o_1}, y expr=-1*\noexpand\thisrow{mu_o_5}+\noexpand\thisrow{std_o_5}, col sep=comma] {data/cs/arh/ode/interpolated/budget_\b.csv};
+            \noexpand\addplot[name path=binomiallower\b, draw=none] table [x expr=-1*\noexpand\thisrow{mu_o_1}, y expr=-1*\noexpand\thisrow{mu_o_5}-\noexpand\thisrow{std_o_5}, col sep=comma] {data/cs/arh/ode/interpolated/budget_\b.csv};
             \noexpand\addplot[pc1!15] fill between[of={binomialupper\b} and {binomiallower\b}];
 
-            \noexpand\addplot[name path=odeupper\b, draw=none] table [x expr=-1*\noexpand\thisrow{mu_o_1}, y expr=-1*\noexpand\thisrow{mu_o_5}+\noexpand\thisrow{std_o_5}, col sep=comma] {data/mones/arh/binomial/interpolated/budget_\b.csv};
-            \noexpand\addplot[name path=odelower\b, draw=none] table [x expr=-1*\noexpand\thisrow{mu_o_1}, y expr=-1*\noexpand\thisrow{mu_o_5}-\noexpand\thisrow{std_o_5}, col sep=comma] {data/mones/arh/binomial/interpolated/budget_\b.csv};
+            \noexpand\addplot[name path=odeupper\b, draw=none] table [x expr=-1*\noexpand\thisrow{mu_o_1}, y expr=-1*\noexpand\thisrow{mu_o_5}+\noexpand\thisrow{std_o_5}, col sep=comma] {data/mones/arh/ode/interpolated/budget_\b.csv};
+            \noexpand\addplot[name path=odelower\b, draw=none] table [x expr=-1*\noexpand\thisrow{mu_o_1}, y expr=-1*\noexpand\thisrow{mu_o_5}-\noexpand\thisrow{std_o_5}, col sep=comma] {data/mones/arh/ode/interpolated/budget_\b.csv};
             \noexpand\addplot[pc2!15] fill between[of={odeupper\b} and {odelower\b}];

### 2024-05-28 main_pcn.py
default hp from paper

@@ -254,9 +254,9 @@ if __name__ == '__main__':
     parser = argparse.ArgumentParser(description='PCN')
     parser.add_argument('--objectives', default=[1, 5], type=int, nargs='+', help='index for ari, arh, pw, ps, pl, ptot')
     parser.add_argument('--env', default='ode', type=str, help='ode or binomial')
-    parser.add_argument('--action', default='discrete', type=str, help='discrete, multidiscrete or continuous')
+    parser.add_argument('--action', default='continuous', type=str, help='discrete, multidiscrete or continuous')
     parser.add_argument('--lr', default=1e-3, type=float, help='learning rate')
-    parser.add_argument('--steps', default=3e5, type=float, help='total timesteps')
+    parser.add_argument('--steps', default=5e5, type=float, help='total timesteps')
     parser.add_argument('--batch', default=256, type=int, help='batch size')
     parser.add_argument('--model-updates', default=50, type=int,
         help='number of times the model is updated at every training iteration')
@@ -265,12 +265,12 @@ if __name__ == '__main__':
               Initially fill ER with n random episodes')
     parser.add_argument('--n-episodes', default=10, type=int,
         help='number of episodes to run between each training iteration')
-    parser.add_argument('--er-size', default=400, type=int,
+    parser.add_argument('--er-size', default=1000, type=int,
         help='max size (in episodes) of the ER buffer')
     parser.add_argument('--threshold', default=0.02, type=float, help='crowding distance threshold before penalty')
-    parser.add_argument('--noise', default=0.0, type=float, help='noise applied on target-return on batch-update')
-    parser.add_argument('--model', default='conv1dsmall', type=str, help='conv1d(big|small), dense(big|small)')
-    parser.add_argument('--clip_grad_norm', default=None, type=float, help='clip gradient norm during pcn update')
+    parser.add_argument('--noise', default=0.05, type=float, help='noise applied on target-return on batch-update')
+    parser.add_argument('--model', default='densebig_silu', type=str, help='conv1d(big|small), dense(big|small)')
+    parser.add_argument('--clip_grad_norm', default=5, type=float, help='clip gradient norm during pcn update')
     parser.add_argument('--budget', default=None, type=int, help='number of times each action is allowed to change')
     args = parser.parse_args()
     print(args)
[…]
